### 2015-05-11 O3CycleScheme.tex
Troposheric ozone production chemical scheme (CH4 only)

\documentclass[]{standalone}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}

\usepackage{graphicx}
%\usepackage[left=2cm,right=2cm,top=2cm,bottom=2cm]{geometry}

% Figures
\usepackage{tikz}
\usetikzlibrary{shapes,arrows}
\usetikzlibrary{positioning}
\usetikzlibrary{calc}
%\usepackage{chemfig}

% Tikz styles
\tikzstyle{reg} = [rounded rectangle, text width=2cm, fill=black!5, align=center, anchor=west,font=\sffamily\Large\bfseries, minimum height=1cm, inner sep=0]
\tikzstyle{rad} = [rounded rectangle, draw=none, text width=2cm, fill=red!10, align=center, anchor=west,font=\sffamily\Large\bfseries, minimum height=1cm, inner sep=0]
\tikzstyle{long} = [rounded rectangle, draw=none, text width=3.5cm, fill=black!5, align=center, anchor=west,font=\sffamily\Large\bfseries, minimum height=1cm, inner sep=0]   
\tikzstyle{add} = [scale=1.5, draw=none,fill=none,align=center]
\tikzstyle{emp} = [draw=none,fill=none]
\tikzstyle{ar} = [->,draw=black!70,line width=2]
\tikzstyle{ln} = [-,draw=black!70,line width=2]

\begin{document}
\begin{tikzpicture}[node distance = 4cm, auto]

% Nodes: big cycle: CH4-CH302-CH3NO3-CH3O-HCHO-HO2-OH-O3
\node[reg] (ch4) {$CH_4$};
\node[add](foo1_1) at (3,-2) {};
\node[add](foo1_2) at (4,-2) {$O_2$\\$OH$};
\node[reg] (ch3o2) at (5,-4){$CH_3 O_2$};
\node[reg] (oh) at (5,4){$OH$};
\node[long] (ch3ooh) at (0,-8){$CH_3OOH$};
\node[reg] (oz1) at (0,8){$O_3$};
\node[emp] (foo2) at ($(oz1)!.5!(oh)$){};
\node[reg] (ho2) at (15,4){$HO_2$};
\node[add] (oz_up) at ($(ho2)!.5!(oh) + (0,0.5)$){$O_3$};
\node[reg] (h2o2) at (17.5,7.5){$H_2O_2$};
\node[add] (ho2_) at (16.5,6){$HO_2$};
\node[emp] (nox_top) at ($(ho2)!.5!(oh) + (0,1.5)$){};
\node[reg] (ch3o) at (15,-4){$CH_3 O$};
\node[reg] (hcho) at (20,0){$HCHO$};
\node[emp] (foo5) at (19,-2){};
\node[long] (ch3no3) at (10,-10){$CH_3 NO_3$};
\node[add] (foo1_2) at (7,-9) {$NO$};
%\node[emp] (dep1) at (11,-12){};
%\node[add] (dep1_) at (10.3,-11.3){deposition};

%Nodes: Bottom NOx cycle + misc
\node[add] (no_1) at (7,-6){$NO$};
\node[add] (no2_1) at (13,-6){$NO_2$};
\node[emp] (nox_bottom) at ($(no_1)!.5!(no2_1)+(0,2)$){};
\node[emp] (foo7) at (11.5,-7.5){};
\node[add] (hv_11) at (12,-7){$h\nu$};
\node[reg] (o3_1) at (9.25,-8.75){$O_3$};
\node[add] (o2_1) at (11.2,-7.9){$O_2$};
\node[emp] (foo3) at (16,-8){};
\node[add] (hv_12) at (15,-8.75){$h\nu$};
\node[emp] (foo4) at (19,-8){};
\node[add] (oh_) at (20,-6.5){$OH$};
\node[add] (o2_) at (18,-3){$O_2$};

%Nodes: Top NOx cycle + misc
\node[add] (no2_2) at ($(ho2)!.5!(oh)+(-3,3.5)$){$NO_2$};
\node[add] (no_2) at ($(ho2)!.5!(oh)+(3,3.5)$){$NO$};
\node[add] (hv_2) at (10,9){$h\nu$};
\node[reg] (o3_2) at (13,10.5){$O_3$};
\node[add] (o2_2) at (12.75,9.75){$O_2$};
\node[emp] (foo6) at (11.5,9.5){};

%Nodes: CO & CO2
\node[reg] (co) at ($(ho2)!.5!(oh)+(-4,-3)$){$CO$};
\node[reg] (co2) at ($(ho2)!.5!(oh)+(2,-3)$){$CO_2$};
\node[emp] (cc_o2) at ($(ho2)!.5!(oh)+(0,-2.5)$){};
\node[emp] (cc_1) at ($(ho2)!.5!(oh)+(-0.5,-1.5)$){};
\node[emp] (cc_2) at ($(ho2)!.5!(oh)+(0.5,-1.5)$){};
\node[add] (cc_o2) at ($(ho2)!.5!(oh)+(0,-2)$){$O_2$};

%Nodes: HCHO decompostion
\node[long] (co_2ho2) at (25,3){$CO~+~2HO_2$};
\node[add] (hcho_hv1) at ($(hcho)!.5!(co_2ho2)$){$h\nu$};
\node[long] (co_h2) at (25,0){$CO~+~H_2$};
\node[add] (hcho_hv2) at ($(hcho)!.5!(co_h2)+(0.25,0.25)$){$h\nu$};
\node[long] (co_ho2) at (25,-3){$CO~+~HO_2$};
\node[add] (hcho_oh) at ($(hcho)!.5!(co_ho2)+(0.25,0)$){$OH$};
%\node[long] (hno3dep) at (25,-6){\small $CO~+~HO_2~+~HNO_3$\\ deposition?};
%\node[add] (hcho_no3) at ($(hcho)!.5!(hno3dep)$){$NO_3$};

%Nodes: after decomposition HO_2 + CO_2
\node[emp] (decomp) at ($(co_2ho2)!.5!(co_ho2)+(3,0)$){};
\node[long] (decomp2) at ($(decomp)+(2,0)$){$HO_2~+~CO_2$};
\node[add] (decomp_oh) at ($(decomp)+(1,0.5)$){$OH$};

% Connections: big cycle
\draw[ln] (oh.west) to [out=180,in=90](foo1_1.center);
\draw[ln] (ch4.east) to [out=0,in=90](foo1_1.center);
\draw[ar] (foo1_1.center) to [out=270,in=180](ch3o2.west);
\draw[ar] (ch3o2) to (ch3ooh);
\node[add](foo1) at ($(ch3o2)!.5!(ch3ooh) + (-1,0)$) {$HO_2$};
%\draw[ar] (oz1) to node[right]{$h\nu$}(foo2.north west);
\draw[ar] (oz1) to (oh);
\node[add] (o3_to_OH) at ($(oz1)!.5!(oh) + (0.5,0.5)$){$h\nu$\\ $H_2O$};
%\draw[ar] (foo2.south east) to node[right]{$H_2O$}(oh);
\draw[ar] (ho2) to (h2o2);
\draw[ln] (ho2.north west) to [out=135,in=0](nox_top.center);
\draw[ar] (nox_top.center) to [out=180,in=45](oh.north east);
\draw[ar] (ho2) to (oh);
\draw[ln] (ch3o2) to (nox_bottom.center);
\draw[ar] (nox_bottom.center) to (ch3o);
\draw[ar] (ch3o2.south) to [out=270,in=180](ch3no3.west);
\draw[ln] (ch3o.east) to [out=0,in=270](foo5.center);
\draw[ar] (foo5.center) to [out=90,in=0](ho2.east);
\draw[ar] (foo5.center) to [out=90,in=180](hcho.west);

% Connections: bottom NOx cycle + misc
\draw[ln] (no_1.north) to [out=90,in=180](nox_bottom.center);
\draw[ar] (nox_bottom.center) to [out=0,in=90](no2_1.north);
\draw[ln] (no2_1.south) to [out=270,in=0](foo7.center);
\draw[ar] (foo7.center) to [out=180,in=90](o3_1.north);
\draw[ar] (foo7.center) to [out=180,in=270](no_1.south);
\draw[ln] (ch3no3.east) to [out=0,in=270](foo3.center);
\draw[ar] (foo3.center) to [out=90,in=270](ch3o.south);
\draw[ar] (foo3.center) to [out=90,in=0](no2_1.east);
\draw[ln] (ch3no3.east) to [out=0,in=270](foo4.center);
\draw[ar] (foo4.center) to [out=90,in=270](hcho.south);
\draw[ar] (foo4.center) to [out=90,in=0](no2_1.east);
%\draw[ar] (ch3no3.south) -- (dep1.center);

% Connections: top NOx cycle
\draw[ln] (no_2.south) to [out=270,in=0](nox_top.center);
\draw[ar] (nox_top.center) to [out=180,in=270](no2_2.south);
\draw[ln] (no2_2.north) to [out=90,in=180](foo6.center);
\draw[ar] (foo6.center) to [out=0,in=90](no_2.north);
\draw[ar] (foo6.center) to [out=0,in=270](o3_2.south);

%Connections: CO & CO2
\draw[ln] (oh.south east) to [out=315,in=180](cc_1.center);
\draw[ln] (co.north east) to [out=45,in=180](cc_1.center);
\draw[ln] (cc_1.center) to [out=0,in=180](cc_2.center);
\draw[ar] (cc_2.center) to [out=0,in=225](ho2.south west);
\draw[ar] (cc_2.center) to [out=0,in=135](co2.north west);

%Connections: HCHO decomp
\draw[ar] (hcho.east) to [out=0,in=180](co_2ho2.west);
\draw[ar] (hcho.east) to [out=0,in=180](co_h2.west);
\draw[ar] (hcho.east) to [out=0,in=180](co_ho2.west);
%\draw[ar] (hcho.east) to [out=0,in=180](hno3dep.west);

\draw[ln] (co_2ho2.east) to [out=0,in=180](decomp.center);
\draw[ln] (co_h2.east) to [out=0,in=180](decomp.center);
\draw[ln] (co_ho2.east) to [out=0,in=180](decomp.center);
\draw[ar] (decomp.center) to [out=0,in=180](decomp2.west);

\end{tikzpicture}

\end{document}

### 2015-07-18 MSc_diss.tex
.latex file with thesis text, started introduction and literature review

\documentclass[11pt,a4paper]{article}
% Encoding and language
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
% Colors
\usepackage[usenames,dvipsnames]{color}
\usepackage{colortbl}
% Paper size and margins
%\usepackage{rotating}
\usepackage{pdflscape}
%\usepackage[a4paper,showframe]{geometry}
\usepackage{vmargin}
\setmarginsrb{2cm}{2cm}{2cm}{2cm}{0cm}{0cm}{0cm}{1.5cm}
\usepackage{pdflscape}
\usepackage{afterpage}
% Indenting
\usepackage{indentfirst}
%\setlength{\parindent}{1cm}
\setlength{\parskip}{0.5cm}
% Symbols
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
% Equations
\usepackage{eulervm} % upright font in eqns (Palatino)
\usepackage{cool}
\usepackage{mathtools} % for arrows
\renewcommand{\theequation}{R\arabic{equation}}
% References
\usepackage{natbib}
\bibliographystyle{apalike}
% Hyperlinks and labels
%\usepackage{showkeys}
\usepackage[linktocpage=true,plainpages=false,pdfpagelabels=false]{hyperref}
\definecolor{citecolor}{rgb}{0.2,0.7,0.9}
\definecolor{urlcolor}{rgb}{0,0,1}
\hypersetup{
    colorlinks, linkcolor={Black},
    citecolor={Black}, urlcolor={Black}
}
% Figures
\usepackage{subcaption}
\usepackage{alphalph}
\usepackage{tikz}
\usetikzlibrary{shapes,arrows}
\usetikzlibrary{positioning}
\usetikzlibrary{calc}
% Tikz styles
\tikzstyle{reg} = [rounded rectangle, text width=2cm, fill=black!5, align=center, anchor=west,font=\sffamily\Large\bfseries, minimum height=1cm, inner sep=0]
\tikzstyle{rad} = [rounded rectangle, draw=none, text width=2cm, fill=red!10, align=center, anchor=west,font=\sffamily\Large\bfseries, minimum height=1cm, inner sep=0]
\tikzstyle{long} = [rounded rectangle, draw=none, text width=3.5cm, fill=black!5, align=center, anchor=west,font=\sffamily\Large\bfseries, minimum height=1cm, inner sep=0]   
\tikzstyle{add} = [scale=1.5, draw=none,fill=none,align=center]
\tikzstyle{emp} = [draw=none,fill=none]
\tikzstyle{ar} = [->,draw=black!70,line width=2]
\tikzstyle{ln} = [-,draw=black!70,line width=2]
% Title etc
\newcommand{\mytitle}{Investigation of the relationship between tropospheric ozone production efficiency and carbon bond emissions}
\newcommand{\authorname}{Maria Zamyatina}
\newcommand{\authornumber}{Student No: 100106685}
\newcommand{\degname}{Master of Science}
\newcommand{\univname}{University of East Anglia}
\newcommand{\univaddr}{University Plain \\ Norwich NR47TJ UK}
\newcommand{\depname}{School of Environmental Sciences}
\newcommand{\mydate}{\the\year}
%% Remove author and date from the maketitle command
%\makeatletter
%\renewcommand{\@maketitle}{
%\newpage
% \null
% \vskip 2em%
% \begin{center}%
%  {\LARGE \@title \par}%
% \end{center}%
% \par} \makeatother

% PDF meta-data
\hypersetup{pdftitle={\mytitle}}
\hypersetup{pdfauthor=\authorname}

\title{\mytitle}
% =========================================================================
\begin{document}

\input{MSc_diss_title}

\begin{abstract}
This work is based on alkyl nitrate measurements from four field campaigns, from ground-based sites across the British Isles and aircraft measurements over the North Atlantic. The ratios between the alkyl nitrates and their parent hydrocarbons are compared with those expected from the photochemical theory. The sensitivity analysis to the theoretical assumptions and uncertainties in measurements revealed a good general agreement between observed and theoretical data.
\end{abstract}

\section{Introduction} \label{sec:intro}
The Earth’s atmosphere is the recipient of a vast range of chemical compounds emitted by natural and anthropogenic sources. The oxidation of these compounds is the main process taking place in the atmosphere that keeps their concentrations at levels that do not distort the chemical balance of the atmosphere \citep{Prinn2003}. In other words, oxidation is a removal mechanism for these compounds (various pollutants and greenhouse gases), and its rate is often referred to as the “cleansing” capacity of the atmosphere. This capacity depends mainly on two factors: the global burden of the principal oxidants such as ozone ($O_3$), the hydroxyl radical ($OH$) and hydrogen peroxide ($H_2O_2$), and climate \citep{Prinn2003,Thompson1992}.

Tropospheric ozone ($O_3$) plays an important role in the chemical and thermal balance of the atmosphere. Although it accounts only for about 10\% of the total ozone, it is a primary precursor to the formation of hydroxyl radicals ($OH$) which in association with hydrogen peroxide ($H_2O_2$) determines the ‘cleansing’ capacity of the atmosphere \citep{Prinn2003,Tarasick2008,Thompson1992}. On the other hand, being a source of the main pollutant removal agents, ozone is an air pollutant itself. Long exposure to its high concentrations causes respiratory problems in humans and damage to sensitive plant species \citep{Fowler2008}. Last but not least characteristic of tropospheric ozone is that it is an important greenhouse gas with a current estimated radiative forcing of  $0.40\pm 0.20~Wm^{–2}$ which is approximately one fifth of the $CO_2$ radiative forcing \citep{Hartmann2013,Myhre2013}. Due to these reasons a considerable interest in quantifying tropospheric ozone concentrations, budget and trends exists for more than a 100 years \citep{Becker2004}.

Unlike many other air pollutants, tropospheric ozone is not directly emitted, but produced following the oxidation of carbon monoxide ($CO$), methane ($CH_4$), and nonmethane volatile compounds (nmVOCs) in the presence of nitrogen oxides (NOx) \citep{Crutzen1973,Myhre2013}. The emissions of these precursors arise from both natural and anthropogenic sources which have increased/intensified since pre-industrial times \citep{Parrish2014,Volz1988} (Figure 1). Presumably it has led to a roughly doubling in background tropospheric ozone concentration \citep{Guicherit2000,Hartmann2013,Tarasick2008,Vingarzan2004}, although contribution from another major source, transport from the stratosphere, may also took place \citep{Fowler2008}. Climate is also believed to be a key driver of ozone production and destruction since the chemical reactions affecting ozone concentration are strongly dependent on such climatic factors as temperature, rainfall and humidity. Therefore in a changing climate tropospheric ozone is not anymore a local air quality issue, but is a global pollution problem \citep{Fowler2008}.

\begin{figure}[h]
\includegraphics[width=\linewidth]{{./pics/Isaksen2009_O3}.png}
\caption{Observed surface ozone at different Northern Hemispheric surface stations \citep{Isaksen2009}.}
\label{fig:ratio_curves}
\end{figure}

According to the International Panel on Climate Change (IPCC) Fifth Assessment report (AR5) for present (about 2000) conditions the global mean tropospheric ozone budget is approximately 331 Tg \citep{Myhre2013}. However, despite general agreement on how the drivers impact ozone concentrations at global and regional scales, its budget varies considerably between different modelling and observational studies \citep{Stevenson2006,Wild2007,Young2012}. Table 1 demonstrates that the range of the global tropospheric ozone burden estimates is approximately 60 Tg which is 18\% of the mean. In terms of ozone production the variation is even higher and account for 28\% of the corresponding value. In that regard, two questions arise: why does the tropospheric ozone budget differ between models and observations? And secondly, why does it vary to this extent, especially in terms of production? The only way to answer these questions is to perform a rigorous investigation of the factors that drive tropospheric ozone and attribute those differences to respective factors \citep{Young2012}.

\begin{table}[h] % O3 budget
\centering
\caption{Summary of tropospheric ozone global budget model and observation estimates for present (about 2000) conditions. All uncertainties quoted as 1 standard deviation (68\% confidence interval). Adapted from \citep{Myhre2013}}
\begin{tabular}{ccl}
\hline
Burden, Tg & Production, Tg yr–1 & Reference \\
\hline
\multicolumn{3}{c}{Modelling studies} \\
$337\pm23$  & $4877\pm853$ & Young et al. (2013); ACCMIP \\
$323$	    & -	           & Archibald et al. (2011) \\
$330$	    & $4876$       & Kawase et al. (2011) \\
$312$       & $4289$       & Huijnen et al. (2010) \\
$334$       & $3826$       & Zeng et al. (2010) \\
$324$       & $4870$       & Wild and Palmer (2008) \\
$314$       & -	           & Zeng et al. (2008) \\
$319$	    & $4487$	   & Wu et al. (2007) \\
$372$	    & $5042$       & Horowitz (2006) \\
$349$	    & $4384$	   & Liao et al. (2006) \\
$344\pm39$	& $5110\pm606$ & Stevenson et al. (2006); ACCENT \\
$314\pm33$	& $4465\pm514$ & Wild (2007) (post-2000 studies) \\
\hline
\multicolumn{3}{c}{Observational studies} \\
$333$   	& -	           & Fortuin and Kelder (1998) \\
$327$	    & -	           & Logan (1999) \\
$325$	    & -	           & Ziemke et al. (2011); 60S–60N \\
$319–351$	& -	           & Osterman et al. (2008); 60S–60N \\
\hline
\end{tabular}
\end{table}

A considerable effort has been made to tackle these problems, and generally there are two possible ways of doing it. The first one implies the direct analysis of tropospheric chemical schemes used in climate-chemistry models. For example, Emmerson and Evans (2009) compared the state of the art Master Chemical Mechanism which contains approximately 5600 species and 13500 reactions \citep{Jenkin2002} with other six chemical schemes and found four significant variations between them. This approach is linked with big challenges since a modern global model is a very sophisticated representation of the climate-chemistry interactions, and therefore it is often difficult to separate influence of one reaction from another. However, even being so developed climate-chemistry models do not take into account all known chemical reactions taking place in the real troposphere. A complete representation requires many thousands of species and tens of thousands of reactions, which is beyond the numerical capabilities currently available. This significantly limits the practicable size of chemical schemes and typically involves a reduction of the number of VOCs considered and lumping the carbon from the discarded species into representative surrogates.

The second way to understanding the discrepancies between models and observations is to conduct an idealised theoretical study using a conceptual model. The advantage of this approach is that such kind of models can be quite easily constructed, and more importantly can be focused on representation of a certain chemical reactions which presumably could explain the discrepancy between models and observations. An example of such set of reactions is alkyl nitrate chemistry.

Alkyl nitrates are important tropospheric trace gases which are formed following the same chemistry that leads to the production of ozone \citep{Reeves2007}. The simplest example, in case of $CH_4$ oxidation, can be described by the set of reactions below (also shown in Figure 2).

%\input{O3CycleScheme}

In the presence of UV light shorter than 320 nm and water vapour $O_3$ photolysis is the main source of OH radicals in the troposphere:
\begin{equation} \label{eq:O3hv}
O_3 + hv \rightarrow O(^1D) + O_2
\end{equation}
\begin{equation} \label{eq:O1D+H2O}
O(^1D) + H_2O \rightarrow 2OH
\end{equation}
Once formed the OH radicals react mainly with $CO$ and $CH_4$ to initiate the $O_3$ formation or removal cycles by producing peroxy radicals ($HO_2$, $CH_3O_2$):
\begin{equation} \label{eq:OH+CO}
OH + CO \rightarrow HO_2 + CO_2
\end{equation}
\begin{equation} \label{eq:OH+CH4}
OH + CH_4 \rightarrow CH_3O_2 + H_2O
\end{equation}

The fate of the peroxy radicals depends on the presence of NOx. At low NOx levels (remote regions), the $HO_2$ radical reacts with another HO2 to form hydrogen peroxide, H2O2, or with the methyl peroxy radical, $CH_3O_2$, to form methyl hydrogen peroxide ($CH_3OOH$):
\begin{equation} \label{eq:HO2+HO2}
HO_2 + HO_2 \rightarrow H_2O_2 + O_2
\end{equation}
\begin{equation} \label{eq:HO2+CH3O2}
HO_2 + CH_3O_2 \rightarrow CH_3OOH + O_2
\end{equation}
On a whole, at low NOx levels a net $O_3$ loss takes place, because the reaction sequence is initiated by $O_3$ photolysis. Plus under the same conditions, some additional $O_3$ removal occurs due to the re-action of $HO_2$ radicals with $O_3$,
\begin{equation} \label{eq:HO2+O3}
HO_2 + O_3 \rightarrow OH + 2O_2
\end{equation}
leading to the regeneration of OH radicals as part of an $O_3$ depleting HOx catalytic cycle.

At intermediate NOx levels (rural areas of most industrialised countries), the peroxy radicals react with nitrogen monoxide ($NO$) to form nitrogen dioxide ($NO_2$) with the subsequent photolysis of $NO_2$ generating $O_3$:
\begin{equation} \label{eq:HO2+NO}
HO_2 + NO \rightarrow NO_2 + OH
\end{equation}
\begin{equation} \label{eq:CH_3O_2+NO}
CH_3O_2 + NO \rightarrow NO_2 + CH_3O
\end{equation}
\begin{equation} \label{eq:NO2hv}
NO_2 + hv \rightarrow NO + O(^3P)
\end{equation}
\begin{equation} \label{eq:O3P+O2}
O(^3P) + O_2 + M \rightarrow O_3 + M
\end{equation}
As it is shown in Figure 2 reactions (8) and (9) form part of the free-radical propagated $O_3$ forming cycles, which may occur a number of times before being halted by a radical termination reactions (not shown). However, the reaction of $CH_3O_2$ with NO is a two-channel reaction \citep{Day2003}, and its second branch can terminate the $O_3$ forming cycle earlier by producing an alkyl nitrate, in this case methyl nitrate ($CH_3ONO_2$):
\begin{equation} \label{eq:CH3NO3form}
CH_3O_2 + NO \rightarrow CH_3ONO_2
\end{equation}

Being a reservoir for the short lived NOx and having atmospheric lifetimes ranging from several days to about a month \citep{Reeves2007} alkyl nitrates can be transported over long distances, and then lost through photolysis and reaction with $OH$:
\begin{equation} \label{eq:CH3NO3hv}
CH_3ONO_2 + hv \rightarrow CH_3O + NO_2
\end{equation}
\begin{equation} \label{eq:CH3NO3+OH}
CH_3ONO_2 + OH \rightarrow HCHO + NO_2
\end{equation}
It means that alkyl nitrates can ‘release’ $NO_2$ back to the atmosphere after a while, and in association with reactions (10) and (11) start a new $O_3$ formation cycle being quite away from the source.

The free-radical propagated oxidation of hydrocarbons leads to the generation of carbonyl compounds, in this case formaldehyde ($HCHO$). Its further oxidation and photolysis contributes to radical generation:
\begin{equation} \label{eq:HCHOhv1}
HCHO + hv \rightarrow 2HO_2 + CO
\end{equation}
\begin{equation} \label{eq:HCHOhv2}
HCHO + hv \rightarrow CO + H_2
\end{equation}
\begin{equation} \label{eq:HCHO+OH}
HCHO + OH \rightarrow HO_2 + CO
\end{equation}
As a result, the formation of $HCHO$ has an impact on the rates of the oxidation cycles described above, through secondary radical generation \citep{Fowler2008}.
At high NOx levels (urban environment) reactions (8) and (9) dominate for $CH_3O_2$ and $HO_2$, but $O_3$ formation becomes inhibited by further increases in NOx. In other words, at high NOx regime the $O_3$ formation rate becomes sensitive to the concentration of $CH_4$ or $CO$ and to inputs of other VOCs \citep{Fowler2008}.

To sum up, alkyl nitrates are produced as a minor branch of the reaction of peroxy radicals with NO \citep{Day2003}. Since these compounds are removed from the troposphere by photolysis and OH oxidation they can provide information on photochemical evolution of an air mass through relationships with their parent hydrocarbons \citep{Bertman1995}. Acting as reservoir species for reactive nitrogen (NOy) and having a relatively long atmospheric lifetimes alkyl nitrates play an important role in long range transport of NOx. Furthermore, alkyl nitrates link to ozone and carbonyl production making them useful tracers for both species \citep{Worton2010}. Despite their relative importance alkyl nitrates are usually not included in the climate-chemistry models, for example in UKCA (Reeves, personal communication), which potentially could lead to discrepancies between models and observations described before.

\subsection{Objectives}
The overall purpose of this research is to investigate the impact of alkyl nitrate chemistry on ozone production efficiency. Specific aims are:
\begin{itemize}
\item to investigate of the relationship between ozone production and carbon bond emissions (see Data and Methods);
\item to analyse the impact of inclusion of alkyl nitrates chemistry on this relationship;
\item to analyse everything mentioned above in various NOx regimes.
\end{itemize}

\section{Literature review} \label{sec:data}
Alkyl nitrates ($RONO_2$) are organic tropospheric trace gases. They are produced photochemically through the oxidation of their parent hydrocarbons \citep{Roberts1990} or emitted directly from equatorial oceans \citep{Blake2003} and biomass burning \citep{Simpson2002}. Being involved in the ozone production chemistry, these species can act as reservoirs of nitrogen oxides which availability largely affects the production of ozone \citep{Reeves2007}. Since the alkyl nitrates have atmospheric lifetimes ranging from several days to about a month, they can redistribute nitrogen oxides presented over polluted continental regions to clean remote marine environments, and therefore are considered as one of the tracers for determining the anthropogenic influence \citep{Atherton1989, Reeves2007, Worton2005}.

Alkyl nitrates are removed from the atmosphere by photolysis \citep{Turberg1990} or through oxidation by hydroxyl radical ($OH$) \citep{Talukdar1997}. Photolysis is the dominant loss mechanism for the compounds with short carbon chains and becomes more important with decreasing carbon number. Reaction with $OH$ follows the opposite trend and is most important for the pentyl nitrates and higher.

Owing to the existence of the relationship between alkyl nitrates and their parent hydrocarbons, measurements of their concentrations can provide information on the extent of photochemical processing in the air mass. \cite{Bertman1995} developed a kinetic approach to describe this relationship through the evolution of the ratio of alkyl nitrate to its parent hydrocarbon as a function of time. This approach has been extensively used to study air mass photochemical ageing at several sites, in North America \citep{Bertman1995, Roberts1998}, China \citep{Simpson2006}, the British Isles \citep{Worton2010} and over the Pacific \citep{Simpson2003} and North Atlantic Ocean \citep{Reeves2007}.

\cite{Bertman1995} found that the theoretical ratios of  2-butyl nitrate to n-butane and 2- and 3-pentyl nitrates to pentane strongly agree with those measured from the ambient air showing no evidence for primary sources of these nitrates. On the contrary, measurements of the other ratios, ethyl/ethane, n-propyl nitrate/propane, and 2-propyl nitrate/propane, appeared to be significantly higher than predicted, relative to the 2-butyl nitrate/n-butane ratio, meaning that there should be an additional source of these alkyl nitrates. \cite{Roberts1998} generally approved the theory by comparing Bertman's conclusions with the results from a new observational dataset.

Using methyl nitrate as a tracer of marine $RONO_2$ \cite{Simpson2006} revealed that the South China Sea in not a region of strong $RONO_2$ emissions suggesting that photochemical rather marine production of alkyl nitrates is dominant at the site.

\cite{Worton2010} found that the photochemical source of the short-chain alkyl nitrates originates mostly from the photochemical oxidation and decomposition of longer-chain compounds rather than from the oxidation of the parent hydrocarbons as it is thought in the previous studies.

The brief description of the measurement sites and data is presented in section \ref{sec:data}. In section \ref{sec:method} the methodology used in this work is defined. Section \ref{sec:res} contains the results. Section \ref{sec:discuss} discusses the results, particularly focusing attention on the analysis of the discrepancy between the observational data and the results of theoretical modelling. Conclusions constitute section \ref{sec:conclusion}.

\section{Methodology} \label{sec:method}
Below is the simplified scheme of atmospheric hydrocarbon oxidation leading to the formation of alkyl nitrates (see also Fig. \ref{fig:scheme_bertman}). It begins from the process of $H$ abstraction by the $OH$ radical where the fraction of the primary or secondary radical is denoted by $\alpha_1$.
\begin{equation} \label{eq:alkane_oh}
RH + \bullet OH \xrightarrow{k_1, \alpha_1} R\bullet + H_2O
\end{equation}
The resulting radical rapidly reacts with $O_2$ and forms an alkyl peroxy radical ($R\bullet$):
\begin{equation} \label{eq:rad_o2}
R\bullet + O_2 \xrightarrow{k_2} ROO\bullet
\end{equation}
Then the alkyl peroxy radical reacts with $NO$, when it can either lose an oxygen atom to form alkoxy radical ($RO\bullet$) or bond with $NO$ to form an alkyl nitrate. The former reaction is a crucial in the net production of tropospheric ozone, since photolysis of $NO_2$ is the main photochemical source of $O_3$ in the troposphere.
\begin{subequations} \label{eq:peroxy_no0}
\begin{align}
ROO\bullet + NO &\xrightarrow{k_3, (1-\alpha_3)} RO\bullet + NO_2 \label{eq:peroxy_no1}\\
ROO\bullet + NO &\xrightarrow{k_3, \alpha_3} RONO_2 \label{eq:peroxy_no2}
\end{align}
\end{subequations}
In other words, a branching ratio $\alpha_3$ of the reactions between alkyl peroxy radicals and $NO$ leads to formation of nitrate products via rearrangement of a high-energy intermediate \citep{Bertman1995}. The branching ratio $\alpha_3$ is larger when the parent alkane is larger and is a function of temperature and pressure.

\begin{figure}[h]
\centering
\begin{tikzpicture}[node distance = 4cm, auto]
\node[reg] (rh) {$RH$};
\node[rad] (roo) at (4,0){$ROO\bullet$};
\node[add] (oh) at (3,0.3) {$\bullet OH$};
\node[rad] (roor) at (4,-2){$ROOR'$};
\node[add] (r_oo) at (3.7,-1) {$R'OO\bullet$};
\node[reg] (rono2) at (2,3){$RONO_2$};
\node[rad] (ro) at (6,3){$RO\bullet$};
\node[add] (alpha) at ($(roo)!.5!(rono2) + (-1,-0.2)$) {$\alpha_3$};
\node[add] (alpha2) at ($(roo)!.5!(ro) + (1,-0.2)$) {$1-\alpha_3$};
\node[add] (no) at ($(roo)+(0,1.5)$) {$NO$};
\node[reg] (rcho) at (9,3){$RCHO$};
\node[add] (prod) at (0,3){\textit{products}};

\draw[ar] (rh.east) to [out=0,in=180](roo.west);
\draw[ar] (roo.south) to [out=270,in=90](roor.north);
\draw[ar] (roo.north) to [out=90,in=270](rono2.south);
\draw[ar] (roo.north) to [out=90,in=270](ro.south);
\draw[ar] (rono2.west) to [out=180,in=0](prod.east);
\draw[ar] (ro.east) to [out=0,in=180](rcho.west);
\end{tikzpicture}
\caption{Schematic representation of the alkyl nitrate chemistry and their formation through the photooxydation of alkanes. Adapted from \citep{Bertman1995}.}
\label{fig:scheme_bertman}
\end{figure}

Apart from $NO$, alkyl peroxy radicals can react with each other and give peroxides, without formation of nitrates. As the eqs. \eqref{eq:peroxy_no0} outline, whether $ROO\bullet$ will bond with nitrates or with other peroxy radicals depends partially on $NO$ concentration. The reaction between $ROO\bullet$ and $NO$ is more likely when nitrate concentration is high.

The alkyl nitrates are removed from the atmosphere through photolysis and reaction with hydroxyl radical.
\begin{equation} \label{eq:an_photo}
RONO_2 + h\nu \xrightarrow{j_4} RO\bullet + NO_2
\end{equation}
\begin{equation} \label{eq:an_oh}
RONO_2 + OH \xrightarrow{k_5} \mathit{products}
\end{equation}
The relative importance of these two loss reactions depends generally on the net UV intensity and the size of the carbon chain in alkyl nitrates. Photolysis is relatively more important for nitrates with $< C_4$, while nitrates with longer chains ($\geq C_5$) tend to react faster with $OH$. In case of butyl nitrates both mechanisms are of similar significance \citep{Worton2010}.

Assuming that the only source of alkyl nitrates is the $OH$ reaction with alkane, and that reactions take place in $NO$-rich atmosphere (so no $ROO\bullet$ self-reactions occur), the alkyl nitrate part in this generalised scheme can be reduced to the two equations of formation and loss:
\begin{subequations} \label{eq:an_form_loss0}
\begin{align}
RH &\xrightarrow{\beta k_A} RONO2 \label{eq:an_form_loss1}\\
RONO_2 &\xrightarrow{k_B} \mathit{products} \label{eq:an_form_loss2}
\end{align}
\end{subequations}
where $k_A = k_1[OH]$ and $k_B=j_4 + k_5[OH]$. The factor $\beta=\alpha_1\alpha_3$ takes into account the fraction of $H$ atom abstraction from the particular hydrocarbon (with the reaction rate $k_1$) and the branching ratios leading to the nitrate. The photolytic loss rate of $RONO_2$ is denoted by $j_4$, and $k_5$ is the rate of alkyl nitrate destruction by $OH$.

These kinetic equations lead to the ordinary differential equations which can be integrated in time ($t$) giving a relationship between the alkyl nitrates and their parent alkanes concentrations:
\begin{equation} \label{eq:integral}
\frac{[RONO_2]}{[RH]} = \frac{\beta k_A}{k_B - k_A}\left(1-e^{(k_A - k_B)t}\right)+\frac{[RONO_2]_0}{[RH]_0}e^{(k_A - k_B)t}
\end{equation}
where subscript $0$ denotes the initial concentrations.

A further good approximation is that there is no alkyl nitrate initially (\textit{i.e.} no direct emissions), hence, the second term in eq. (\eqref{eq:integral}) becomes zero. Finally, the evolution of alkyl nitrates in an air mass can be modelled using the following formula.
\begin{equation} \label{eq:model}
\frac{[RONO_2]}{[RH]} = \frac{\beta k_A}{k_B - k_A}\left(1-e^{(k_A - k_B)t}\right)
\end{equation}
It has to be noted that eq. \eqref{eq:model} assumes no mixing with surrounding air. This simplification is reasonable only when mixing affects both the parent alkane and the corresponding alkyl nitrate similarly, keeping the ratios in eqs. \eqref{eq:integral} and \eqref{eq:model} the same \citep{Reeves2007}.

\section{Results} \label{sec:res}

\section{Discussion} \label{sec:discuss}
The experimental results presented in the previous section provide an opportunity to speculate about the behaviour of our simple model of alkyl nitrate production and loss with time in comparison with the observational data.

\subsection{Short-chained alkyl nitrates} \label{sec:short_an}
\subsubsection{Short photochemical processing times} \label{sec:short_an_short_time}

\section{Conclusion} \label{sec:conclusion}

\newpage
\clearpage
\thispagestyle{empty}% empty page style (?)
\begin{landscape}% Landscape page
\centering
	%\input{subfig_28_land}
\end{landscape}

\bibliography{diss_refs}

\end{document}

### 2015-07-18 MSc_diss.tex
Started to describe the model

@@
 
+\tableofcontents
+\newpage
+
 \begin{abstract}
-This work is based on alkyl nitrate measurements from four field campaigns, from ground-based sites across the British Isles and aircraft measurements over the North Atlantic.
-The ratios between the alkyl nitrates and their parent hydrocarbons are compared with those expected from the photochemical theory.
-The sensitivity analysis to the theoretical assumptions and uncertainties in measurements revealed a good general agreement between observed and theoretical data.
+Abstract.
 \end{abstract}
@@
 The oxidation of these compounds is the main process taking place in the atmosphere that keeps their concentrations at levels that do not distort the chemical balance of the atmosphere \citep{Prinn2003}.
-In other words, oxidation is a removal mechanism for these compounds (various pollutants and greenhouse gases), and its rate is often referred to as the “cleansing” capacity of the atmosphere.
+In other words, oxidation is a removal “cleansing” mechanism for these compounds (various pollutants and greenhouse gases), and its rate is often referred to as the oxidizing capacity of the atmosphere.
 This capacity depends mainly on two factors: the global burden of the principal oxidants such as ozone ($O_3$), the hydroxyl radical ($OH$) and hydrogen peroxide ($H_2O_2$), and climate \citep{Prinn2003,Thompson1992}.
 
+\subsection{Tropospheric ozone}
 Tropospheric ozone ($O_3$) plays an important role in the chemical and thermal balance of the atmosphere.
@@
 
+\subsection{Alkyl nitrates}
 Alkyl nitrates are important tropospheric trace gases which are formed following the same chemistry that leads to the production of ozone \citep{Reeves2007}.
@@
 
-\subsection{Objectives}
-The overall purpose of this research is to investigate the impact of alkyl nitrate chemistry on ozone production efficiency. Specific aims are:
+\subsection{Motivation}
+lalala
+
+\section{Overall objective and specific aims}
+\subsection{Overall objective}
+The overall purpose of this research is to investigate the impact of alkyl nitrate chemistry on ozone production efficiency.
+\subsection{Specific aims}
+Specific aims are:
 \begin{itemize}
@@
 \section{Methodology} \label{sec:method}
+\subsection{Model development}
+A one box model was developed to investigate the impact of alkyl nitrate chemistry on ozone production efficiency. It will be written using the programming language FACSIMILE.
+
+\subsection{Photolysis rate coefficients}
+From MCMv3.3.
+
+\subsection{Design of experiments}
+Hypothesis: ANs have an impact on O3.
+
+To test the hypothesis box model runs were performed at a series of different NOx and VOCs initial concentrations, so that an $O_3$ isopleth plot could be constructed, similar to those found in \citep{Dodge1977} and \citep{Sillman1999}.
+The following initial concentrations were used:
+
+\begin{table} % model setup: O3, CO
+\caption{Model setup}
+\centering
+\begin{tabular}{ccc}
+\hline
+Compound       & Initial concentration, ppb \\
+\hline
+$O_3$          & 40 \\
+$CO$           & 100 \\
+\hline
+\end{tabular}
+\end{table}
+
+\begin{table} % model setup: NOx
+\caption{Model setup: NOx initial concentrations (in ppb)}
+\centering
+\begin{tabular}{ccc}
+\hline
+$NO$      & $NO_2$      & NOx  ($=NO+NO_2$) \\
+\hline
+0.0015    & 0.0035      & 0.005 \\
+0.007     & 0.018       & 0.025 \\
+0.015     & 0.035       & 0.05  \\
+0.03      & 0.07        & 0.1   \\
+0.07      & 0.18        & 0.25  \\
+0.15      & 0.35        & 0.5   \\
+0.25      & 0.5         & 0.75  \\
+0.3       & 0.7         & 1     \\
+0.7       & 1.8         & 2.5   \\
+1.5       & 3.5         & 5     \\
+3         & 7           & 10    \\
+\hline
+\end{tabular}
+\end{table}
+
+\begin{table} % model setup: VOCs
+\caption{Model setup: VOCs initial concentrations (ppb)}
+\centering
+\begin{tabular}{ccccccccc}
+\hline
+Experiment number & $CH_4$ & $C_2H_6$ & $C_3H_8$ & $nC_4H_{10}$ & $iC_4H_{10}$ & $nC_5H_{12}$ & $iC_5H_{12}$ \\
+\hline
+1  & 1780 &	0.620 &	0.100 &	0.050 &	0.250 &	0.250 &	0.250 \\
+2  & 1807 &	1.258 &	0.390 &	0.245 &	0.325 &	0.275 &	0.425 \\
+3  & 1834 &	1.896 &	0.680 &	0.440 &	0.400 &	0.300 &	0.600 \\
+4  & 1861 &	2.534 &	0.970 &	0.635 &	0.475 &	0.325 &	0.775 \\
+5  & 1888 &	3.172 &	1.260 &	0.830 &	0.550 &	0.350 &	0.950 \\
+6  & 1915 &	3.810 &	1.550 &	1.025 &	0.625 &	0.375 &	1.125 \\
+7  & 1942 &	4.448 &	1.840 &	1.220 &	0.700 &	0.400 &	1.300 \\
+8  & 1969 &	5.086 &	2.130 &	1.415 &	0.775 &	0.425 &	1.475 \\
+9  & 1996 &	5.724 &	2.420 &	1.610 &	0.850 &	0.450 &	1.650 \\
+10 & 2023 &	6.362 &	2.710 &	1.805 &	0.925 &	0.475 &	1.825 \\
+11 & 2050 &	7.000 &	3.000 &	2.000 &	1.000 &	0.500 &	2.000 \\
+\hline
+\end{tabular}
+\end{table}
+
+\subsection{Bertman}
 Below is the simplified scheme of atmospheric hydrocarbon oxidation leading to the formation of alkyl nitrates (see also Fig.

### 2015-07-18 MSc_diss.tex
Minor changes

@@
 \hypersetup{
-    colorlinks, linkcolor={Black},
-    citecolor={Black}, urlcolor={Black}
+    colorlinks, linkcolor={Blue},
+    citecolor={Blue}, urlcolor={Black}
 }
@@
 \section{Methodology} \label{sec:method}
-\subsection{Model development}
-A one box model was developed to investigate the impact of alkyl nitrate chemistry on ozone production efficiency. It will be written using the programming language FACSIMILE.
+\subsection{Model development and description}
+A one (0D) box model is developed to investigate the impact of alkyl nitrate chemistry on ozone production efficiency.
+It is written from scratch using FACSIMILE for Windows.
+This special high-level programming language is a powerful modelling tool designed to efficiently solve differential equations including those resulting from a set of chemical reactions of interest.
 
-\subsection{Photolysis rate coefficients}
-From MCMv3.3.
+The model is intended to represent the steady state of the troposphere under mean summer daytime meteorology and clear skies.
+To achieve that the air temperature is fixed at $25^{\circ}C$ and relative humidity at 70\%.
+There is no transport and mixing in the model.
+
+\subsection{Chemical mechanism}
+The chemical mechanism of the model is based on recommendations from \citep{Atkinson2004} and the Master Chemical Mechanism (http://mcm.leeds.ac.uk/MCM/home.htt).
+On a whole it includes 90 species and 165 reactions.
+It is focused on representation of the chemistry related to $O_3$ production, namely:
+\begin{itemize}
+\item $O_3$ null cycle, HOx and NOx cycles;
+\item oxidation of methane ($CH_4$) and other alkanes, namely ethane ($C_2H_6$), propane ($C_3H_8$), n-butane ($nC_4H_{10}$), i-butane ($iC_4H_{10}$), n-pentane ($nC_5H_{12}$) and i-pentane ($nC_5H_{12}$)
+\item alkyl nitrate chemistry.
+\end{itemize}
+
+Photolysis rates are calculated externally using parameters from MCMv3.3.
+
+The only sink for NOx is alkyl nitrates.
+
+\subsection{Emissions and deposition}
+There are no emissions and deposition is the model.
 

### 2015-07-19 MSc_diss.tex
Continued to work on section about model development

@@
 Although it accounts only for about 10\% of the total ozone, it is a primary precursor to the formation of hydroxyl radicals ($OH$) which in association with hydrogen peroxide ($H_2O_2$) determines the ‘cleansing’ capacity of the atmosphere \citep{Prinn2003,Tarasick2008,Thompson1992}.
-On the other hand, being a source of the main pollutant removal agents, ozone is an air pollutant itself.
+On the other hand, being a source of main pollutant removal agents, ozone is an air pollutant itself.
 Long exposure to its high concentrations causes respiratory problems in humans and damage to sensitive plant species \citep{Fowler2008}.
@@
 Unlike many other air pollutants, tropospheric ozone is not directly emitted, but produced following the oxidation of carbon monoxide ($CO$), methane ($CH_4$), and nonmethane volatile compounds (nmVOCs) in the presence of nitrogen oxides (NOx) \citep{Crutzen1973,Myhre2013}.
-The emissions of these precursors arise from both natural and anthropogenic sources which have increased/intensified since pre-industrial times \citep{Parrish2014,Volz1988} (Figure 1).
-Presumably it has led to a roughly doubling in background tropospheric ozone concentration \citep{Guicherit2000,Hartmann2013,Tarasick2008,Vingarzan2004}, although contribution from another major source, transport from the stratosphere, may also took place \citep{Fowler2008}.
+The emissions of these precursors arise from both natural and anthropogenic sources which have intensified since pre-industrial times \citep{Parrish2014,Volz1988} (Figure \ref{fig:O3observations}).
+Presumably it has led to roughly a doubling in background tropospheric ozone concentration \citep{Guicherit2000,Hartmann2013,Tarasick2008,Vingarzan2004}, although contribution from another major source, transport from the stratosphere, may also took place \citep{Fowler2008}.
 Climate is also believed to be a key driver of ozone production and destruction since the chemical reactions affecting ozone concentration are strongly dependent on such climatic factors as temperature, rainfall and humidity.
@@
 \caption{Observed surface ozone at different Northern Hemispheric surface stations \citep{Isaksen2009}.}
-\label{fig:ratio_curves}
+\label{fig:O3observations}
 \end{figure}
 
-According to the International Panel on Climate Change (IPCC) Fifth Assessment report (AR5) for present (about 2000) conditions the global mean tropospheric ozone budget is approximately 331 Tg \citep{Myhre2013}.
+According to the Intergovernmental Panel on Climate Change (IPCC) Fifth Assessment report (AR5) for present conditions (around year 2000) the global mean tropospheric ozone budget is approximately 331 Tg \citep{Myhre2013}.
 However, despite general agreement on how the drivers impact ozone concentrations at global and regional scales, its budget varies considerably between different modelling and observational studies \citep{Stevenson2006,Wild2007,Young2012}.
-Table 1 demonstrates that the range of the global tropospheric ozone burden estimates is approximately 60 Tg which is 18\% of the mean.
+Table \ref{tab:O3budget} demonstrates that the range of the global tropospheric ozone budget estimates is approximately 60 Tg which is 18\% of the mean.
 In terms of ozone production the variation is even higher and account for 28\% of the corresponding value.
@@
 Adapted from \citep{Myhre2013}}
+\label{tab:O3budget}
 \begin{tabular}{ccl}
@@
 \subsection{Chemical mechanism}
-The chemical mechanism of the model is based on recommendations from \citep{Atkinson2004} and the Master Chemical Mechanism (http://mcm.leeds.ac.uk/MCM/home.htt).
-On a whole it includes 90 species and 165 reactions.
-It is focused on representation of the chemistry related to $O_3$ production, namely:
+The chemical mechanism of the model is based on recommendations from \citep{Atkinson2004} and information from the Master Chemical Mechanism, MCM v3.3 \citep{Jenkin1997,Saunders2003}, via website: http://mcm.leeds.ac.uk/MCM.
+(On a whole it includes 90 species and 165 reactions).
+The mechanism is developed to represent the chemistry that governs the production and loss of ozone, hydrocarbons and alkyl nitrates.
+These include:
 \begin{itemize}
-\item $O_3$ null cycle, HOx and NOx cycles;
+\item the formation of ozone as a result of the photolysis of $NO_2$ followed by the reaction of ground-state oxygen atom ($O(^3P)$) with molecular oxygen and a third body, M ($N_2$ or $O_2$);
+\item the loss of ozone through its photolysis and reactions with $O(^3P)$, $OH$, $HO_2$ and $NO$, among which the last is of particular interest because it gives back $NO_2$;
+\item the formation of alkylperoxy radicals ($RO_2$) from oxidation of the hydrocarbons;
+\item the competition of $NO$ and $HO_2$ for reaction with $HO_2$ and $RO_2$;
+\item the formation of alkyl nitrates which adds up to the competition of $NO$ and $HO_2$ for reaction with the peroxy radicals;
+\item the loss of alkyl nitrates through their photolysis and the reaction with $OH$;
+\item the subsequent reactions of hydroperoxides and aldehydes formed from the organic reactions mentioned above.
 \item oxidation of methane ($CH_4$) and other alkanes, namely ethane ($C_2H_6$), propane ($C_3H_8$), n-butane ($nC_4H_{10}$), i-butane ($iC_4H_{10}$), n-pentane ($nC_5H_{12}$) and i-pentane ($nC_5H_{12}$)
-\item alkyl nitrate chemistry.
 \end{itemize}
 
-Photolysis rates are calculated externally using parameters from MCMv3.3.
+Since the most alkoxy radicals ($RO$) react rapidly with $O_2$ under atmospheric conditions and most large (>=C4) alkoxy radicals also rapidly decompose or isomerize, alkyl nitrate formation from reaction RO+NO2 = RONO2 is relatively unimportant for most organics under atmospheric conditions while observations showed that the reaction RO2+NO = RONO2 gives a higher yield.
+
+Since one of the main mechanisms of the alkyl nitrate production (>C3!!) is the gas gas phase oxidation of alkanes by OH in the presence of NOx.
+
+Apart from methane in the model there are other nonmethane 
+It is worth mentioning that only alkanes  
+Due to purely theoretical purpose of this investigation the only sink for NOx considered in the mechanism is their loss due to formation of alkyl nitrates.
+It means that reactions involving formation of nitric acid ($HNO_3$) and peroxyacetyl nitrates (PAN) are not considered here, which could lead to discrepancies between our modelling results and those from other research studies.
+
+\subsection{Photolysis}
+To estimate photolysis rates for a given geographical location, one must take into account the latitude and season, as well as the time of the day.
+
+24-hour average photolysis rates are calculated externally (outside FACSIMILE program) for 1 July using the formula and parametrisations used in MCM v3.3.
+The formula that describes variation of photolysis rates with solar zenith angle for the core photolysis reactions was determined by the MCM's authors by using a two stream isotopic scattering model and by calculating rates for clear sky conditions at an altitude of 0.5 km on 1 July at a latitude of 45$^{\circ}$N.
+has the following form:
+\begin{equation} \label{eq:MCMphotolysis}
+\renewcommand{\theequation}{\arabic{equation}}
+J = l(cos\chi)^mexp(-nsec\chi)
+\end{equation}
+where $J$ - photolysis rate coefficient, $\chi$ - solar zenith angle, $l$, $m$ and $n$ - optimised parameters.
+and it is a result of determination from the available data on absorption cross section and quantum data for the core reactions using two stream  Hayman.
+
+The values of solar zenith angle are calculated using formulas from NOAA Solar Calculator (http://www.esrl.noaa.gov/gmd/grad/solcalc/).
 
@@
 \begin{table} % model setup: O3, CO
-\caption{Model setup}
+\caption{Fixed and variable parameters used in box model runs}
 \centering
@@
 \hline
-Compound       & Initial concentration, ppb \\
+Parameter & Initial mixing ratio & Type \\
 \hline
-$O_3$          & 40 \\
-$CO$           & 100 \\
+$H_2O$    & 2\%                  & Fixed    \\
+$O_2$     & 21\%     ???         & Fixed    \\
+$N_2$     & 79\%        ???      & Fixed    \\
+$H_2$     & 525 ppb  ???         & Variable \\
+$O_3$     & 40 ppb               & Variable \\
+$CO$      & 100 ppb              & Variable \\
 \hline
@@
 \begin{table} % model setup: NOx
-\caption{Model setup: NOx initial concentrations (in ppb)}
+\caption{Model setup: NOx initial concentrations (ppb)}
 \centering
@@
 \section{Results} \label{sec:res}
+Dependence of alkyl nitrate yield on carbon number (paper about n-alkanes): bigger yield of ANs from larger carbon number means that the potential for contributing to photochemical air pollution may be less for the larger (C>6) n-alkanes that for the smaller ones.
 

### 2015-07-23 MSc_diss.tex
Minor additions in methodology

@@
 
+!! from \citep{Finlayson-Pitts2000}
+The relationship between the peroxy radical concentration and the ozone photolysis rate constant for these higher NO conditions can be again approximated using steady-state analysis [..].
+While OH is recycled in its reactions with CO and CH4 via HO2, it is pernanetly removed at higher NOx concentrations by the reaction of OH with NO2, forming nitric acid:
+OH + NO2 = M = HONO2 (113).
+This reaction is primarily a daytime reaction because most OH sources are photolytic in nature. As a result, the NO2 reaction with OH competes!!!! with NO2 photolysis (page 290/993)!!
+         Rate coeff (OH+NO2 = HNO3)  J(NO2 = NO+O)
+Book     2.47e-30          7.00e-3
+My model 1.15e-11          4.39e-03
+Thus, the reaction with OH is not usually a dominant loss process for NO2, but it is still sufficiently fact to from significant amounts of HNO3 during the day, particularly in polluted regions with relatively large NO2 concentrations.
+Nitric acid undergoes both wet and dry deposition rapidly and can be neutralized by ammonia, the major gaseous base found in the atmosphere.
+
 \subsection{Motivation}
-lalala
+lalalaep
 
@@
 \subsection{Model development and description}
-A one (0D) box model is developed to investigate the impact of alkyl nitrate chemistry on ozone production efficiency.
-It is written from scratch using FACSIMILE for Windows.
-This special high-level programming language is a powerful modelling tool designed to efficiently solve differential equations including those resulting from a set of chemical reactions of interest.
+A zero-dimensional (box) model is developed to investigate the impact of alkyl nitrate chemistry on ozone production efficiency.
+Its code is written using FACSIMILE for Windows.
+FACSIMILE is a special high-level programming language which is a powerful modelling tool designed to efficiently solve differential equations including those resulting from a set of chemical reactions.
 
 The model is intended to represent the steady state of the troposphere under mean summer daytime meteorology and clear skies.
-To achieve that the air temperature is fixed at $25^{\circ}C$ and relative humidity at 70\%.
-There is no transport and mixing in the model.
+To achieve that a number of assumptions have been made.
+First, the model assumes a uniform mixing of individual constituents of the troposphere, second, there are no atmospheric transport (no 'winds') and no emissions of species into the box meaning that the time evolution of each species is controlled by the chemical interactions alone and determined by their initial concentrations.
+The air temperature and relative humidity are assumed to be $25^{\circ}C$ and 70\% respectively and are fixed in all model runs.
+Other assumptions concerning the chemical mechanism can be found in the next section.
+
+!! For long-lived species, averaging may be reasonable.
 
 \subsection{Chemical mechanism}
-The chemical mechanism of the model is based on recommendations from \citep{Atkinson2004} and information from the Master Chemical Mechanism, MCM v3.3 \citep{Jenkin1997,Saunders2003}, via website: http://mcm.leeds.ac.uk/MCM.
-(On a whole it includes 90 species and 165 reactions).
+The chemical mechanism of the model is based on recommendations from \citep{Atkinson2004} and information from the Master Chemical Mechanism, MCM v3.3 \citep{Jenkin1997,Saunders2003}, retrieved via website: http://mcm.leeds.ac.uk/MCM.
+On a whole it includes 90 species and 165 reactions.
 The mechanism is developed to represent the chemistry that governs the production and loss of ozone, hydrocarbons and alkyl nitrates.
@@
 \item the formation of alkyl nitrates which adds up to the competition of $NO$ and $HO_2$ for reaction with the peroxy radicals;
-\item the loss of alkyl nitrates through their photolysis and the reaction with $OH$;
+\item the loss of alkyl nitrates through their photolysis and reaction with $OH$;
 \item the subsequent reactions of hydroperoxides and aldehydes formed from the organic reactions mentioned above.
-\item oxidation of methane ($CH_4$) and other alkanes, namely ethane ($C_2H_6$), propane ($C_3H_8$), n-butane ($nC_4H_{10}$), i-butane ($iC_4H_{10}$), n-pentane ($nC_5H_{12}$) and i-pentane ($nC_5H_{12}$)
+\end{itemize}
+The full list of reactions can be found in Appendix ??
+\ref{sec:appendix1}.
+It is important to mention that due to purely theoretical purpose of this research and its focus on investigation of the impact of alkyl nitrate chemistry on ozone production some of the reactions that lead to formation of species that serve as a sink for $NO_x$ are not included into the chemical mechanism.
+However, these reactions play an important role in the chemistry of the troposphere, and below here is a description of their role and anticipated effect of their exclusion.
+\begin{itemize}
+\item The formation of the nitric acid ($HNO_3$) is a major loss process for $NO_x$ during daytime in the atmosphere (e.g.
+\citep{Holloway2010}).
+Being a 'sticky' molecule $HNO_3$ readily adsorbs to surfaces, particularly if there is water on the surface, and deposits either in precipitation (wet deposition) or in dry form (dry deposition)/undergoes dry and wet deposition.
+Because of the high deposition velocity, dry deposition of $HNO_3$ can be responsible for much of the removal of inorganic nitrogen from the troposphere \citep{Finlayson-Pitts2000}, which means that exclusion of $HNO_3$ formation from the chemical mechanism potentially leads to overproduction of ozone, (because the competition between $NO_2$ photolysis and its oxidation by $OH$ radical ($NO_2 + OH \xrightarrow{M} HNO_3$) is ignored and other reactions).
+\item the formation of nitrous acid ($HONO$) which photolysis during morning hours could be one of the sources of $OH$ radicals in the troposphere;
+(\citep{Finlayson-Pitts2000})
+!!
+HONO is formed whenever NO2 and water are present, its contribution to the total acidity has become of interest and concern.
+But since it photolyses rapidly, its concentrations and contribution to the total acidity during the day are not significant compared to sulfuric and nitric acids.
+\item the formation of the nitrate radical ($NO_3$);
+(\citep{Finlayson-Pitts2000})
+NO2 also reacts with O3, forming the nitrate radical,
+NO2 + O3 = NO3 + O2.
+NO3 is recognized to be a major contributor to the chemistry of organics in the troposphere at night.
+Because it absorbs in the red region of the spectrum, it photolyzes rapidly during the day so that its chemistry other than photolysis is essentially restricted to the dark hours.
+In addition to reacting with organics, NO3 also reacts with NO2, forming dinitrogen pentoxide in a reversible, equilibrium reaction:
+NO3 + NO2 <=> M N2O5.
+In any case, because of this equilibrium, sink for N2O5 such as hydrolysis are, in essence, also sinks for NO3 as well.
+\item the formation of dinitrogen pentoxide ($N_2O_5$)
+In the troposphere N2O5 is an important nighttime source of nitric acid through its rapid hydrolysis on wet surfaces and aerosol particles:
+N2O5 + H2O = surfaces, aerosol particles 2HNO3.
+The NO2-NO3 reaction may also have a small contribution from a two-body reaction:
+NO3+NO2 = NO+NO2+O2
+however, it appears to be minor.
+\item the formation of peroxyacetyl nitrate ($RC(O)OONO_2$, PAN)
 \end{itemize}
 
-Since the most alkoxy radicals ($RO$) react rapidly with $O_2$ under atmospheric conditions and most large (>=C4) alkoxy radicals also rapidly decompose or isomerize, alkyl nitrate formation from reaction RO+NO2 = RONO2 is relatively unimportant for most organics under atmospheric conditions while observations showed that the reaction RO2+NO = RONO2 gives a higher yield.
+!! from \citep{Finlayson-Pitts2000}
+!
+In urban areas: air pollution is characterized more by the formation of ozone and other oxidants rather than by SO2, particles, and sulfuric acid.
+In these regions, the primary pollutants are NOx (mainly NO) and volatile organic compounds (VOC), which undergo photochemical reactions in sunlight to form a host of secondary pollutants, the most of which is O3.
+Some of these are criteria pollutants for which air quality standards have been set, such as O3, SO2, CO, NO2, PM10 and PM2.5.
+Other are so-called "trace" noncritetia pollutants, e.g., gaseous peroxyacetyl nitrate (CH3C(O)OONO2,,PAN), nitric acid (HNO3 or HONO2), formaldehyde (HCHO), and formic acid (HCOOH).
+The overall reaction is now written as:
+VOC + NOx + hv = O3 + PAN + HNO3 + ... + Particles, etc. (6)
 
-Since one of the main mechanisms of the alkyl nitrate production (>C3!!) is the gas gas phase oxidation of alkanes by OH in the presence of NOx.
+! oxidation of methane ($CH_4$) and other alkanes, namely ethane ($C_2H_6$), propane ($C_3H_8$), n-butane ($nC_4H_{10}$), i-butane ($iC_4H_{10}$), n-pentane ($nC_5H_{12}$) and i-pentane ($nC_5H_{12}$)
 
-Apart from methane in the model there are other nonmethane 
+!
+Since the most alkoxy radicals ($RO$) react rapidly with $O_2$ under atmospheric conditions and most large (>=C4) alkoxy radicals also rapidly decompose or isomerize, alkyl nitrate formation from reaction RO+NO2 = RONO2 is relatively unimportant for most organics under atmospheric conditions while observations showed that the reaction RO2+NO = RONO2 gives a higher yield (from \citep{Atkinson2000}, actually from 1982).
+
+! Since one of the main mechanisms of the alkyl nitrate production (>C3!!) is the gas gas phase oxidation of alkanes by OH in the presence of NOx (\citep{Newland2013}).
+
+! Apart from methane in the model there are other nonmethane 
 It is worth mentioning that only alkanes  
@@
 $H_2O$    & 2\%                  & Fixed    \\
-$O_2$     & 21\%     ???         & Fixed    \\
-$N_2$     & 79\%        ???      & Fixed    \\
+$O_2$     & 21\%                 & Fixed    \\
+$N_2$     & 78\%                 & Fixed    \\
 $H_2$     & 525 ppb  ???         & Variable \\
@@
 
+\subsection{Justification of model performance}
+!! from \citep{Finlayson-Pitts2000}
+Urban:
+Certain reproducible features of time-concentration profiles for  pollutants are observed in "smoggy" ambient air.
+Figure 1.3, a classic example of historical interest, shows such profiles for NO, NO2, and total oxidant (mainly O3) in Pasadena, California, during a  severe photochemical air pollution episode in July 1973.
+Reproducible  features  include  the  following:
+\begin{itemize}
+\item In the early morning, the concentration of NO rises and  reaches a maximum at a time that approximately coincides with the  maximum emissions of NO, in this case, peak  automobile  traffic; 
+\item Subsequently, NO2 rises to a maximum;
+\item Oxidant (e.g., O3) levels, which are relatively low in the  early morning, increase significantly about noon when the NO concentration drops to a low value.
+\end{itemize}
+
 \section{Discussion} \label{sec:discuss}
@@
 
+\section{Appendix I} \label{sec:appendix1}
+
 \end{document}

### 2015-07-24 MSc_diss.tex
Minor changes in section about the chemical mechanism

@@
 \citep{Holloway2010}).
-Being a 'sticky' molecule $HNO_3$ readily adsorbs to surfaces, particularly if there is water on the surface, and deposits either in precipitation (wet deposition) or in dry form (dry deposition)/undergoes dry and wet deposition.
-Because of the high deposition velocity, dry deposition of $HNO_3$ can be responsible for much of the removal of inorganic nitrogen from the troposphere \citep{Finlayson-Pitts2000}, which means that exclusion of $HNO_3$ formation from the chemical mechanism potentially leads to overproduction of ozone, (because the competition between $NO_2$ photolysis and its oxidation by $OH$ radical ($NO_2 + OH \xrightarrow{M} HNO_3$) is ignored and other reactions).
-\item the formation of nitrous acid ($HONO$) which photolysis during morning hours could be one of the sources of $OH$ radicals in the troposphere;
-(\citep{Finlayson-Pitts2000})
-!!
-HONO is formed whenever NO2 and water are present, its contribution to the total acidity has become of interest and concern.
-But since it photolyses rapidly, its concentrations and contribution to the total acidity during the day are not significant compared to sulfuric and nitric acids.
-\item the formation of the nitrate radical ($NO_3$);
-(\citep{Finlayson-Pitts2000})
-NO2 also reacts with O3, forming the nitrate radical,
-NO2 + O3 = NO3 + O2.
-NO3 is recognized to be a major contributor to the chemistry of organics in the troposphere at night.
-Because it absorbs in the red region of the spectrum, it photolyzes rapidly during the day so that its chemistry other than photolysis is essentially restricted to the dark hours.
-In addition to reacting with organics, NO3 also reacts with NO2, forming dinitrogen pentoxide in a reversible, equilibrium reaction:
-NO3 + NO2 <=> M N2O5.
-In any case, because of this equilibrium, sink for N2O5 such as hydrolysis are, in essence, also sinks for NO3 as well.
-\item the formation of dinitrogen pentoxide ($N_2O_5$)
-In the troposphere N2O5 is an important nighttime source of nitric acid through its rapid hydrolysis on wet surfaces and aerosol particles:
-N2O5 + H2O = surfaces, aerosol particles 2HNO3.
-The NO2-NO3 reaction may also have a small contribution from a two-body reaction:
-NO3+NO2 = NO+NO2+O2
-however, it appears to be minor.
-\item the formation of peroxyacetyl nitrate ($RC(O)OONO_2$, PAN)
+Being a 'sticky' molecule $HNO_3$ readily adsorbs to surfaces, particularly if there is water on the surface, and (deposits either in precipitation (wet deposition) or in dry form (dry deposition))/(undergoes dry and wet deposition).
+Because of the high deposition velocity, dry deposition of $HNO_3$ can be responsible for much of the removal of inorganic nitrogen from the troposphere \citep{Finlayson-Pitts2000}, which means that exclusion of $HNO_3$ (formation and removal)/(chemistry) from the chemical mechanism (potentially) leads to overproduction of ozone in the model, especially at high $NO_2$ concentrations because the competition between $NO_2$ photolysis and its reaction with $OH$ radical ($NO_2 + OH \xrightarrow{M} HNO_3$) is ignored.
+\item Since this research is focused on the daytime chemistry some reactions that are believed to be significant in nighttime are not included into the chemical mechanism of the model.
+Among them are reactions that involve the nitrate radical ($NO_3$) and dinitrogen pentoxide ($N_2O_5$).
+\begin{itemize}
+\item $NO_3$ radical rapidly photolyzes during the daylight hours, but in nighttime its concentration can increase to measurable levels \citep{Atkinson2008}.
+This increase in concentration and a lesser availability of the OH radical (which sources are mostly photolytic in nature) makes $NO_3$ radical a major contributor to the night chemistry of organics in the troposphere \citep{Finlayson-Pitts2000}.
+However, exclusion of $NO_3$ radical reactions from our chemical mechanism should not have a huge impact on the results anyway since the mechanism considers mainly decomposition of alkanes with which $NO_3$ radical reacts to a generally minor extent \citep{Atkinson2008}???.
+\item ! the nighttime NO3 radical reaction is of minor significance in term of the overall removal of the alkanes (Atkinson, 1990a)
+\item In addition to reacting with organics, $NO_3$ also reacts with $NO_2$, forming dinitrogen pentoxide ($NO_3 + NO_2 \xleftrightarrow{M} N_2O_5$) which in turn is an important nighttime source of nitric acid through its rapid hydrolysis on wet surfaces and aerosol particles ($N_2O_5 + H_2O \xrightarrow{surf., aerosol} 2HNO_3$) \citep{Finlayson-Pitts2000}.
+Exclusion of $N_2O_5$ from the chemical mechanism assumes no cycling of nitrogen oxides within the troposphere through $NO_3$ and $N_2O_5$, eventually forming $HNO_3$, which are not included into the mechanism anyway.
+\item ! In any case, because of this equilibrium, sink for N2O5 such as hydrolysis are, in essence, also sinks for NO3 as well.
+\end{itemize}
+\item Another species involved in $NO_x$ tropospheric chemistry is the nitrous acid ($HONO$).
+The major source of nitrous acid is believed to be heterogeneous 'dark' reactions of $NO_2$, including hydrolysis (reaction with water vapour) of $NO_2$ on aerosol and particulate matter surfaces in nighttime.
+During the daytime, $HONO$ can be formed by the reaction of $OH$ radical with $NO$ ($OH + NO \xrightarrow{M} HONO$).
+However, since $HONO$ strongly absorbs UV radiation, it rapidly photolyzes, and under particular conditions (when the nighttime $HONO$ concentration reaches almost 15 ppb \citep{Finlayson-Pitts2000}), photolysis of nighttime-generated $HONO$ can become a major source of $OH$ radicals in the early morning hours ($HONO + hv \rightarrow OH + NO$) \citep{Lammel1996}.
+Therefore exclusion of $HONO$ from the chemical mechanism potentially leads to underestimation of $OH$ concentrations, but having neither formation nor removal of $HONO$ should balance this effect out???.
+\item Peroxyacetyl nitrates ($RC(O)OONO_2$, PAN)
+is important as a means of transporting NOx over large distances.
+It is quite stable at low temperatures but decomposes at higher temperatures to release NO2.
+Hence whether photolysis can compete with the thermal decomposition is of great interest.
+
+The rate constant for PAN decomposition is strongly temperature dependent.
+As a result, this is important in the atmosphere because PAN acts as a reservoir for NO2.
+At low temperatures, much of the total NOx may be tied up in PAN; as the atmosphere warms, due to either duirnal temperature variations or the air mass being transported into warmer regions.
+the rate of PAN decomposition increases.
+This releases NO2, which can then form other secondary pollutants such as O3 and HNO3.
+
+PAN's thermal decomposition releases both NO2 and an organic free radical (acetylperoxy radical), so that it can act as an NOx reservoir and ultimately as a source of OH in the dark.
+The oxidation of organics produces a variety of free radicals, some of which can form peroxyacyl nitrates.
+For example, acetaldehyde is a classic precursor to the PAN through the following reaction sequence:
+CH3CHO + OH = CH3CO + H2O; 
+CH3CO + O2 + M = CH3C(O)OO + M;
+CH3C(O)OO + NO2 + M = CH3C(O)OONO2 + M.
+The key feature of the last reaction is that it is reversible.
+Of all the possible fates for PANs in the atmosphere, thermal decomposition is usually the most important.
+The strong dependence has important implications for the role of PAN.
+When it is formed at lower temperatures or is transported into colder regions, it is stabilized and acts as an NOx reservoir.
+When an air mass containing PAN is transported into warmer regions, however, the PAN decomposes, releasing both the NO2and the acetylperoxy free radical.
+If sufficient NO is present, the organic free radical reacts int the following manner:
+CH3C(O)OO + NO = CH3C(O)O + NO2;
+CH3C(O)O = CH3 + CO2
+followed by the reaction of CH3 with O2, etc. to form HCHO and HO2, or HCHO, CH3OH and CH3OOH under low-NOx conditions. As a result, PAN should act as an accelerator for photochemical smog formation.
+At lower temperatures such as those found in the Arctic, the majority  of NOy is often present in the form of PAN because of its thermal stability.
 \end{itemize}
@@
 The values of solar zenith angle are calculated using formulas from NOAA Solar Calculator (http://www.esrl.noaa.gov/gmd/grad/solcalc/).
-
-The only sink for NOx is alkyl nitrates.
-
-\subsection{Emissions and deposition}
-There are no emissions and deposition is the model.
 

### 2015-07-24 MSc_diss.tex
Additions to desctiption of the chemical mechanism

@@
 \begin{itemize}
-\item The formation of the nitric acid ($HNO_3$) is a major loss process for $NO_x$ during daytime in the atmosphere (e.g.
-\citep{Holloway2010}).
-Being a 'sticky' molecule $HNO_3$ readily adsorbs to surfaces, particularly if there is water on the surface, and (deposits either in precipitation (wet deposition) or in dry form (dry deposition))/(undergoes dry and wet deposition).
-Because of the high deposition velocity, dry deposition of $HNO_3$ can be responsible for much of the removal of inorganic nitrogen from the troposphere \citep{Finlayson-Pitts2000}, which means that exclusion of $HNO_3$ (formation and removal)/(chemistry) from the chemical mechanism (potentially) leads to overproduction of ozone in the model, especially at high $NO_2$ concentrations because the competition between $NO_2$ photolysis and its reaction with $OH$ radical ($NO_2 + OH \xrightarrow{M} HNO_3$) is ignored.
+\item The formation of the nitric acid ($HNO_3$) is a major loss process for $NO_x$ during daytime in the atmosphere.
+Being a 'sticky' molecule $HNO_3$ readily adsorbs to surfaces, particularly if there is water on the surface, and deposits either in precipitation (wet deposition) or in dry form (dry deposition).
+Because of the high deposition velocity, dry deposition of $HNO_3$ can be responsible for much of the removal of inorganic nitrogen from the troposphere \citep{Finlayson-Pitts2000}, which means that exclusion of $HNO_3$ (formation and removal)/(chemistry) from the chemical mechanism potentially leads to overproduction of ozone in the model, especially at high $NO_2$ concentrations because the competition between $NO_2$ photolysis and its reaction with $OH$ radical ($NO_2 + OH \xrightarrow{M} HNO_3$) is ignored.
 \item Since this research is focused on the daytime chemistry some reactions that are believed to be significant in nighttime are not included into the chemical mechanism of the model.
@@
 \item $NO_3$ radical rapidly photolyzes during the daylight hours, but in nighttime its concentration can increase to measurable levels \citep{Atkinson2008}.
-This increase in concentration and a lesser availability of the OH radical (which sources are mostly photolytic in nature) makes $NO_3$ radical a major contributor to the night chemistry of organics in the troposphere \citep{Finlayson-Pitts2000}.
+This increase in concentration and a lesser availability of the OH radical (which sources are mostly photolytic in nature) make $NO_3$ radical a major contributor to the night chemistry of organics in the troposphere \citep{Finlayson-Pitts2000}.
 However, exclusion of $NO_3$ radical reactions from our chemical mechanism should not have a huge impact on the results anyway since the mechanism considers mainly decomposition of alkanes with which $NO_3$ radical reacts to a generally minor extent \citep{Atkinson2008}???.
@@
 Therefore exclusion of $HONO$ from the chemical mechanism potentially leads to underestimation of $OH$ concentrations, but having neither formation nor removal of $HONO$ should balance this effect out???.
-\item Peroxyacetyl nitrates ($RC(O)OONO_2$, PAN)
-is important as a means of transporting NOx over large distances.
-It is quite stable at low temperatures but decomposes at higher temperatures to release NO2.
-Hence whether photolysis can compete with the thermal decomposition is of great interest.
+\item Peroxyacetyl nitrate ($CH_3C(O)OONO_2$, PAN) is the most abundant organic nitrate in the troposphere \citep{Hewitt1994}.
+A classic precursor to the PAN is acetaldehyde ($CH_3CHO$) which oxidation produces PAN through the following reaction sequence:
+\begin{itemize}
+\item[] $CH_3CHO + OH \rightarrow CH_3CO + H_2O$
+\item[] $CH_3CO + O_2 \xrightarrow{M} CH_3C(O)OO$
+\item[] $CH_3C(O)OO + NO_2 \xrightarrow{M} CH_3C(O)OONO_2$.
+\end{itemize}
+Of all the possible fates for PAN in the atmosphere, thermal decomposition is usually the most important because the rate constant for PAN decomposition strongly depends on temperature.
+At low temperatures it is quite stable, but at higher temperatures it decomposes giving $NO_2$ and the acetylperoxy radical ($CH_3C(O)OO$)
+\begin{itemize}
+\item[] $CH_3C(O)OONO_2 \xleftrightarrow{M} NO_2 + CH_3C(O)OO$.
+\end{itemize}
+This temperature dependence has important implications for the role of PAN.
+When it is formed at lower temperatures (e.g., at high altitude) or transported into colder regions, PAN can act as an $NO_x$ reservoir.
+Observations of PAN and the sum of reactive nitrogen ($NO_y$) confirm it revealing that PAN can constitute up to 90\% of the total $NO_y$ budget at higher northern latitudes or higher altitudes (e.g., \citep{Jacobi2000}).
+When an air mass containing PAN is transported into warmer regions, $NO_2$ tied up in PAN can be released back and cause the formation of $O_3$ and $HNO_3$ far away from the original $NO_x$ source.
+Moreover, if sufficient $NO$ is present, $CH_3C(O)OO$ released from PAN can react in the following manner:
+\begin{itemize}
+\item[] $CH_3C(O)OO + NO \rightarrow CH_3C(O)O + NO_2$
+\item[] $CH_3C(O)O \rightarrow CH_3 + CO_2$
+\end{itemize}
+followed by the reaction of $CH_3$ with $O_2$, etc.
+to form $HCHO$ and $HO_2$, or $HCHO$, $CH_3OH$ and $CH_3OOH$ under low-$NO_x$ conditions.
+As a result of this chemistry, PAN can act as an accelerator for photochemical smog formation \citep{Finlayson-Pitts2000}.
+Exclusion of PAN from the chemical mechanism of the model means that we omit the global transport of $NO_x$ provided by PAN, and therefore do not fully estimate tropospheric ozone production.
+However, since the focus of this research is on the impact of alkyl nitrate chemistry on ozone such a miscount is acceptable due to purely theoretical purpose of the research.
+\item Inclusion of all chemistry listed above into the model is a subject for future work.
+\end{itemize}
 
-The rate constant for PAN decomposition is strongly temperature dependent.
-As a result, this is important in the atmosphere because PAN acts as a reservoir for NO2.
-At low temperatures, much of the total NOx may be tied up in PAN; as the atmosphere warms, due to either duirnal temperature variations or the air mass being transported into warmer regions.
-the rate of PAN decomposition increases.
-This releases NO2, which can then form other secondary pollutants such as O3 and HNO3.
+The constructed chemical mechanism has a limited representation of organic chemistry.
+It includes (oxidation schemes)/(degradation reactions) of only seven alkanes: methane ($CH_4$), ethane ($C_2H_6$), propane ($C_3H_8$), n-butane ($nC_4H_{10}$), i-butane ($iC_4H_{10}$), n-pentane ($nC_5H_{12}$) and i-pentane ($iC_5H_{12}$).
+While the oxidation scheme of $CH_4$ is full, oxidation of other six alkanes is not full, but terminated at different oxidation steps.
 
-PAN's thermal decomposition releases both NO2 and an organic free radical (acetylperoxy radical), so that it can act as an NOx reservoir and ultimately as a source of OH in the dark.
-The oxidation of organics produces a variety of free radicals, some of which can form peroxyacyl nitrates.
-For example, acetaldehyde is a classic precursor to the PAN through the following reaction sequence:
-CH3CHO + OH = CH3CO + H2O; 
-CH3CO + O2 + M = CH3C(O)OO + M;
-CH3C(O)OO + NO2 + M = CH3C(O)OONO2 + M.
-The key feature of the last reaction is that it is reversible.
-Of all the possible fates for PANs in the atmosphere, thermal decomposition is usually the most important.
-The strong dependence has important implications for the role of PAN.
-When it is formed at lower temperatures or is transported into colder regions, it is stabilized and acts as an NOx reservoir.
-When an air mass containing PAN is transported into warmer regions, however, the PAN decomposes, releasing both the NO2and the acetylperoxy free radical.
-If sufficient NO is present, the organic free radical reacts int the following manner:
-CH3C(O)OO + NO = CH3C(O)O + NO2;
-CH3C(O)O = CH3 + CO2
-followed by the reaction of CH3 with O2, etc. to form HCHO and HO2, or HCHO, CH3OH and CH3OOH under low-NOx conditions. As a result, PAN should act as an accelerator for photochemical smog formation.
-At lower temperatures such as those found in the Arctic, the majority  of NOy is often present in the form of PAN because of its thermal stability.
-\end{itemize}
+It has been done due to computational restrictions and  because otherwise the mechanism would 
+Another limitation of the model's chemical mechanism is that it does not consider all possible organic reactions (which otherwise could lead to a too large number of reactions but is focused on representing the oxidation of alkanes
+full oxidation scheme for methane ($CH_4$) only .
 
@@
 
-! oxidation of methane ($CH_4$) and other alkanes, namely ethane ($C_2H_6$), propane ($C_3H_8$), n-butane ($nC_4H_{10}$), i-butane ($iC_4H_{10}$), n-pentane ($nC_5H_{12}$) and i-pentane ($nC_5H_{12}$)
-
 !
@@
 ! Since one of the main mechanisms of the alkyl nitrate production (>C3!!) is the gas gas phase oxidation of alkanes by OH in the presence of NOx (\citep{Newland2013}).
-
-! Apart from methane in the model there are other nonmethane 
-It is worth mentioning that only alkanes  
-Due to purely theoretical purpose of this investigation the only sink for NOx considered in the mechanism is their loss due to formation of alkyl nitrates.
-It means that reactions involving formation of nitric acid ($HNO_3$) and peroxyacetyl nitrates (PAN) are not considered here, which could lead to discrepancies between our modelling results and those from other research studies.
 

### 2015-07-25 MSc_diss.tex
Minor changes in description of the chemical mechanism

@@
 The constructed chemical mechanism has a limited representation of organic chemistry.
-It includes (oxidation schemes)/(degradation reactions) of only seven alkanes: methane ($CH_4$), ethane ($C_2H_6$), propane ($C_3H_8$), n-butane ($nC_4H_{10}$), i-butane ($iC_4H_{10}$), n-pentane ($nC_5H_{12}$) and i-pentane ($iC_5H_{12}$).
-While the oxidation scheme of $CH_4$ is full, oxidation of other six alkanes is not full, but terminated at different oxidation steps.
+It includes oxidation of only seven alkanes: methane ($CH_4$), ethane ($C_2H_6$), propane ($C_3H_8$), n-butane ($nC_4H_{10}$), i-butane ($iC_4H_{10}$), n-pentane ($nC_5H_{12}$) and i-pentane ($iC_5H_{12}$).
+$CH_4$ undergoes a full oxidation in the model, but oxidation of other six alkanes follows the reaction sequence:
+\begin{itemize}
+\item[] $RH + OH \rightarrow RO_2 + H_2O$
+\item[] $RO_2 + HO_2 \rightarrow ROOH + O2$
+\item[] $RO_2 + NO \rightarrow RO + NO_2$
+\item[] $ROOH + hv \rightarrow RO + OH$
+\item[] $ROOH + OH \rightarrow products$
+\item[] $RO + O_2 \rightarrow products$
+\end{itemize}
+and is 'terminated' at different oxidation steps.
+Ethane oxidation is terminated after the formation of the acetylperoxyl radicals ($C_2H_3CO_3$), propane - after the formation of the propyldioxy radicals ($C_3H_5O_3$), n-butane - after the formation of ethanal ($CH_3CHO$), butanal ($C_3H_7CHO$), methyl ethyl ketone ($C_2H_5COCH_3$), and 1-($\lambda^1$-oxidanylmethoxy)-2-methoxy-ethane ($C_4H_9O_3$), i-butane - after the formation of 2-methylpropanal ($iso-C_3H_7CHO$) and acetone ($CH_3COCH_3$), n-pentane - after the formation of pentanal ($C_4H_9CHO$), methyl propyl ketone ($n-C_3H_7COCH_3$), and diethyl ketone ($(C_2H_5)_2CO$), i-pentane - after the formation of 2-methylbutanal ($sec-C_4H_9CHO$) and isopropyl methyl ketone ($iso-C_3H_7COCH_3$).
+It has been done to limit the number of reactions in the mechanism, which otherwise would include 1678 reactions describing evolution of 539 species, and therefore would be computationally too expensive.
+For the same reason all self-reactions of organic peroxy radicals ($RO_2$) are omitted as well.
+To avoid a huge impact of incomplete oxidation of alkanes on the results the model's simulation time has been restricted to 24 hours.
 
-It has been done due to computational restrictions and  because otherwise the mechanism would 
-Another limitation of the model's chemical mechanism is that it does not consider all possible organic reactions (which otherwise could lead to a too large number of reactions but is focused on representing the oxidation of alkanes
-full oxidation scheme for methane ($CH_4$) only .
+No halon chemistry.
 
@@
 \end{table}
+
+\begin{figure}[h]
+\includegraphics[width=0.5\linewidth]{{./pics/Sillman1999_Fig1}.png}
+%\caption{chem_allAN_netO3_rate}
+\label{fig:O3observations}
+\end{figure}
 
@@
 
+\begin{figure}[h]
+\includegraphics[width=\linewidth]{D:/FACSIMILE/ANsCBmodel/matlab/ANsCB_pics/ANsNOxVOC/allAN/chem_allAN_netO3_rate.png}
+%\caption{chem_allAN_netO3_rate}
+\label{fig:O3observations}
+\end{figure}
+
 \section{Discussion} \label{sec:discuss}

### 2015-07-26 MSc_diss.tex
Finished section about model development

@@
 
+!
+Since the most alkoxy radicals ($RO$) react rapidly with $O_2$ under atmospheric conditions and most large (>=C4) alkoxy radicals also rapidly decompose or isomerize, alkyl nitrate formation from reaction RO+NO2 = RONO2 is relatively unimportant for most organics under atmospheric conditions while observations showed that the reaction RO2+NO = RONO2 gives a higher yield (from \citep{Atkinson2000}, actually from 1982).
+
+! Since one of the main mechanisms of the alkyl nitrate production (>C3!!) is the gas gas phase oxidation of alkanes by OH in the presence of NOx (\citep{Newland2013}).
+
 \section{Methodology} \label{sec:method}
@@
 The air temperature and relative humidity are assumed to be $25^{\circ}C$ and 70\% respectively and are fixed in all model runs.
-Other assumptions concerning the chemical mechanism can be found in the next section.
-
-!! For long-lived species, averaging may be reasonable.
+Other assumptions concerning the chemical mechanism and photolysis rates can be found in the following sections.
 
@@
 \end{itemize}
-and is 'terminated' at different oxidation steps.
-Ethane oxidation is terminated after the formation of the acetylperoxyl radicals ($C_2H_3CO_3$), propane - after the formation of the propyldioxy radicals ($C_3H_5O_3$), n-butane - after the formation of ethanal ($CH_3CHO$), butanal ($C_3H_7CHO$), methyl ethyl ketone ($C_2H_5COCH_3$), and 1-($\lambda^1$-oxidanylmethoxy)-2-methoxy-ethane ($C_4H_9O_3$), i-butane - after the formation of 2-methylpropanal ($iso-C_3H_7CHO$) and acetone ($CH_3COCH_3$), n-pentane - after the formation of pentanal ($C_4H_9CHO$), methyl propyl ketone ($n-C_3H_7COCH_3$), and diethyl ketone ($(C_2H_5)_2CO$), i-pentane - after the formation of 2-methylbutanal ($sec-C_4H_9CHO$) and isopropyl methyl ketone ($iso-C_3H_7COCH_3$).
+and then is 'terminated' after formation of certain species.
+Ethane oxidation is terminated after formation of the acetylperoxyl radicals ($C_2H_3CO_3$), propane - after formation of the propyldioxy radicals ($C_3H_5O_3$), n-butane - after formation of ethanal ($CH_3CHO$), butanal ($C_3H_7CHO$), methyl ethyl ketone ($C_2H_5COCH_3$), and 1-($\lambda^1$-oxidanylmethoxy)-2-methoxy-ethane ($C_4H_9O_3$), i-butane - after formation of 2-methylpropanal ($iso-C_3H_7CHO$) and acetone ($CH_3COCH_3$), n-pentane - after formation of pentanal ($C_4H_9CHO$), methyl propyl ketone ($n-C_3H_7COCH_3$), and diethyl ketone ($(C_2H_5)_2CO$), i-pentane - after formation of 2-methylbutanal ($sec-C_4H_9CHO$) and isopropyl methyl ketone ($iso-C_3H_7COCH_3$).
 It has been done to limit the number of reactions in the mechanism, which otherwise would include 1678 reactions describing evolution of 539 species, and therefore would be computationally too expensive.
@@
 
-No halon chemistry.
-
-!! from \citep{Finlayson-Pitts2000}
-!
-In urban areas: air pollution is characterized more by the formation of ozone and other oxidants rather than by SO2, particles, and sulfuric acid.
-In these regions, the primary pollutants are NOx (mainly NO) and volatile organic compounds (VOC), which undergo photochemical reactions in sunlight to form a host of secondary pollutants, the most of which is O3.
-Some of these are criteria pollutants for which air quality standards have been set, such as O3, SO2, CO, NO2, PM10 and PM2.5.
-Other are so-called "trace" noncritetia pollutants, e.g., gaseous peroxyacetyl nitrate (CH3C(O)OONO2,,PAN), nitric acid (HNO3 or HONO2), formaldehyde (HCHO), and formic acid (HCOOH).
-The overall reaction is now written as:
-VOC + NOx + hv = O3 + PAN + HNO3 + ... + Particles, etc. (6)
-
-!
-Since the most alkoxy radicals ($RO$) react rapidly with $O_2$ under atmospheric conditions and most large (>=C4) alkoxy radicals also rapidly decompose or isomerize, alkyl nitrate formation from reaction RO+NO2 = RONO2 is relatively unimportant for most organics under atmospheric conditions while observations showed that the reaction RO2+NO = RONO2 gives a higher yield (from \citep{Atkinson2000}, actually from 1982).
-
-! Since one of the main mechanisms of the alkyl nitrate production (>C3!!) is the gas gas phase oxidation of alkanes by OH in the presence of NOx (\citep{Newland2013}).
-
+Lastly, the chemical mechanism of the model does not include halogen chemistry.
+Despite that halogens (chlorine ($Cl$), bromine ($Br$) and iodine ($I$)) are very reactive chemical compounds mostly known by their role in anthropogenic stratospheric ozone-depletion chemistry, and that they are directly (for example, reaction of the $Cl$ radical with hydrocarbons (e.g., $CH_4$)) and indirectly (coupling with the cycles of sulfur and nitrogen that eventually influence $O_3$, $OH$ and $NO_x$ concentrations) involved in the tropospheric chemistry \citep{VonGlasow2014,Platt2003}, they are excluded from the chemical mechanism to keep it more transparent.
 \subsection{Photolysis}
 To estimate photolysis rates for a given geographical location, one must take into account the latitude and season, as well as the time of the day.
-
-24-hour average photolysis rates are calculated externally (outside FACSIMILE program) for 1 July using the formula and parametrisations used in MCM v3.3.
-The formula that describes variation of photolysis rates with solar zenith angle for the core photolysis reactions was determined by the MCM's authors by using a two stream isotopic scattering model and by calculating rates for clear sky conditions at an altitude of 0.5 km on 1 July at a latitude of 45$^{\circ}$N.
-has the following form:
+For the purpose of this research 24-hour average photolysis rates are calculated externally (outside FACSIMILE program) for 1 July using the formula and parametrisations from the MCM v3.3.
+The formula that describes variation of photolysis rate with solar zenith angle for the core photolysis reactions in the MCM was obtained by its authors using cross-section and quantum yield data and two stream isotopic scattering model described in \citep{Hayman1997}.
+Their calculations were performed for clear sky conditions at an altitude of 0.5 km for 1 July at a latitude of 45$^{\circ}$N, resulting in the following formula:
 \begin{equation} \label{eq:MCMphotolysis}
@@
 \end{equation}
-where $J$ - photolysis rate coefficient, $\chi$ - solar zenith angle, $l$, $m$ and $n$ - optimised parameters.
-and it is a result of determination from the available data on absorption cross section and quantum data for the core reactions using two stream  Hayman.
+where $J$ - photolysis rate coefficient, $\chi$ - solar zenith angle, $l$, $m$ and $n$ - optimising parameters \citep{Jenkin1997,Saunders2003}.
 
-The values of solar zenith angle are calculated using formulas from NOAA Solar Calculator (http://www.esrl.noaa.gov/gmd/grad/solcalc/).
+The values of solar zenith angle are calculated using NOAA Solar Calculator (http://www.esrl.noaa.gov/gmd/grad/solcalc/).
 
 \subsection{Design of experiments}
-Hypothesis: ANs have an impact on O3.
+To investigate the influence of alkyl nitrate chemistry on (potential changes in) tropospheric $O_3$ production, box model runs were performed at a series of different $NO_x$ and $VOCs$ initial concentrations, so that an $O_3$ isopleth plot could be constructed, similar to those found in \citep{Dodge1977} and \citep{Sillman1999}.
+Table \ref{tab:setupO3CO} shows initial concentrations of 'independent'/global parameters and indicates which of them are fixed or allowed to vary during the model run.
+Table \ref{tab:setupNOx} and Table \ref{tab:setupVOCs} show the initial concentrations of $NO_x$ and $VOCs$ respectively that are used at a series of experiments.
 
-To test the hypothesis box model runs were performed at a series of different NOx and VOCs initial concentrations, so that an $O_3$ isopleth plot could be constructed, similar to those found in \citep{Dodge1977} and \citep{Sillman1999}.
-The following initial concentrations were used:
+The lower limit of the range of $NO_x$ and $VOCs$ initial concentrations is based on observational data from Mauna Loa station (Hawaii, US) which is believed to be representative for conditions of remote troposphere.
+The upper limit is based on data from
+Marylebone Road station (London, UK) which is a kerbside monitoring site located in central London, in close proximity to a 6 lane road, with very high traffic flow and frequent congestion \citep{VonSchneidemesser2010}, therefore being representative for conditions of urban troposphere.
+
+Two experiments have been performed.
+The aim of the first one is to reveal the total impact of alkyl nitrate chemistry on $O_3$ production, the aim of the second one is to estimate the contribution of individual alkyl nitrates to changes in $O_3$ production.
+
+For the first experiment the model was run under two configurations, (1) with all alkyl nitrate chemistry switched on, and (2) with all alkyl nitrate chemistry switched off.
+In every model run $VOCs$ concentrations were (changed according to the data from Table \ref{tab:setupVOCs} and) were fixed throughout the run, while the initial concentrations of $NO_x$ were (changed according to the data from Table \ref{tab:setupNOx} and were) allowed to vary throughout the run.
+As a result, the first experiment consists of 253 model runs including 11 runs (varying $NO_x$) with inorganic chemistry only.
+
+For the second experiment the model was run using the same approach concerning variations in $NO_x$ and $VOCs$ concentrations, but changes in alkyl nitrate chemistry were treated differently.
+In every model run reactions of individual alkyl nitrates or their groups (defined in Table\ref{tab:RHandANS} were either switched on or switched off, resulting in other additional 847 model runs.
+
+\begin{table} % RH and ANs
+\caption{Alkyl nitrates and their parent alkanes}\label{tab:RHandANS}
+\centering
+\begin{tabular}{llll}
+\hline
+Parent alkane& Structural formula          & Name                     & Group\\
+\hline
+$CH_4$       & $CH_3ONO_2$                 & methyl nitrate           & $CH_3ONO_2$ \\
+\hline
+$C_2H_6$     & $CH_3CH_2ONO_2$             & ethyl nitrate            & $C_2H_5ONO_2$ \\
+\hline
+$C_3H_8$     & $(CH_3)_2CHONO_2$           & iso-propyl nitrate       & $\Sigma C_3H_7ONO_2$\\
+             & $CH_3(CH_2)_2ONO_2$         & n-propyl nitrate         & \\
+\hline
+$nC_4H_{10}$ & $CH_3(CH_2)_3ONO_2$         & n-butyl nitrate          & $\Sigma nC_4H_9ONO_2$\\
+             & $CH_3CH_2CH(CH_3)ONO_2$     & sec-butyl nitrate        & \\
+\hline
+$iC_4H_{10}$ & $CH_3CH(CH_3)CH_2ONO_2$     & iso-butyl nitrate        & $\Sigma iC_4H_9ONO_2$\\
+             & $CH_2C(CH_3)CH_2ONO_2$      & tert-butyl nitrate       & \\
+\hline
+$nC_5H_{12}$ & $CH_3(CH_2)_4ONO_2$         & n-pentyl nitrate         & $\Sigma nC_5H_{11}ONO_2$\\
+             & $CH_3(CH_2)_2CH(CH_3)ONO_2$ & 2-pentyl nitrate         & \\
+             & $???$                       & 3-pentyl nitrate         & \\
+\hline
+$iC_5H_{12}$ & $???$                       & 2-methylbutyl nitrate    & $\Sigma iC_5H_{11}ONO_2$\\
+             & $(CH_3)_2CH(CH_2)_2ONO_2$   & 3-methyl-2-butyl nitrate & \\
+             & $(CH_3)_2C(CH_2CH_3)ONO_2$  & 2-methyl-2-butyl nitrate & \\
+\hline
+\end{tabular}
+\end{table}
+
+!! For long-lived species, averaging may be reasonable.
 
 \begin{table} % model setup: O3, CO
-\caption{Fixed and variable parameters used in box model runs}
+\caption{Fixed and variable parameters used in box model runs}\label{tab:setupO3CO}
 \centering
@@
 $N_2$     & 78\%                 & Fixed    \\
-$H_2$     & 525 ppb  ???         & Variable \\
+$H_2$     & 525 ppb              & Variable \\
 $O_3$     & 40 ppb               & Variable \\
@@
 \begin{table} % model setup: NOx
-\caption{Model setup: NOx initial concentrations (ppb)}
+\caption{Model setup: NOx initial concentrations (ppb)}\label{tab:setupNOx}
 \centering
@@
 \begin{table} % model setup: VOCs
-\caption{Model setup: VOCs initial concentrations (ppb)}
+\caption{Model setup: VOCs concentrations (ppb)}\label{tab:setupVOCs}
 \centering

### 2015-07-27 MSc_diss.tex
Continued subsection about model validation

@@
 
-\section{Methodology} \label{sec:method}
-\subsection{Model development and description}
-A zero-dimensional (box) model is developed to investigate the impact of alkyl nitrate chemistry on ozone production efficiency.
-Its code is written using FACSIMILE for Windows.
-FACSIMILE is a special high-level programming language which is a powerful modelling tool designed to efficiently solve differential equations including those resulting from a set of chemical reactions.
-
-The model is intended to represent the steady state of the troposphere under mean summer daytime meteorology and clear skies.
-To achieve that a number of assumptions have been made.
-First, the model assumes a uniform mixing of individual constituents of the troposphere, second, there are no atmospheric transport (no 'winds') and no emissions of species into the box meaning that the time evolution of each species is controlled by the chemical interactions alone and determined by their initial concentrations.
-The air temperature and relative humidity are assumed to be $25^{\circ}C$ and 70\% respectively and are fixed in all model runs.
-Other assumptions concerning the chemical mechanism and photolysis rates can be found in the following sections.
-
-\subsection{Chemical mechanism}
-The chemical mechanism of the model is based on recommendations from \citep{Atkinson2004} and information from the Master Chemical Mechanism, MCM v3.3 \citep{Jenkin1997,Saunders2003}, retrieved via website: http://mcm.leeds.ac.uk/MCM.
-On a whole it includes 90 species and 165 reactions.
-The mechanism is developed to represent the chemistry that governs the production and loss of ozone, hydrocarbons and alkyl nitrates.
-These include:
-\begin{itemize}
-\item the formation of ozone as a result of the photolysis of $NO_2$ followed by the reaction of ground-state oxygen atom ($O(^3P)$) with molecular oxygen and a third body, M ($N_2$ or $O_2$);
-\item the loss of ozone through its photolysis and reactions with $O(^3P)$, $OH$, $HO_2$ and $NO$, among which the last is of particular interest because it gives back $NO_2$;
-\item the formation of alkylperoxy radicals ($RO_2$) from oxidation of the hydrocarbons;
-\item the competition of $NO$ and $HO_2$ for reaction with $HO_2$ and $RO_2$;
-\item the formation of alkyl nitrates which adds up to the competition of $NO$ and $HO_2$ for reaction with the peroxy radicals;
-\item the loss of alkyl nitrates through their photolysis and reaction with $OH$;
-\item the subsequent reactions of hydroperoxides and aldehydes formed from the organic reactions mentioned above.
-\end{itemize}
-The full list of reactions can be found in Appendix ??
-\ref{sec:appendix1}.
-It is important to mention that due to purely theoretical purpose of this research and its focus on investigation of the impact of alkyl nitrate chemistry on ozone production some of the reactions that lead to formation of species that serve as a sink for $NO_x$ are not included into the chemical mechanism.
-However, these reactions play an important role in the chemistry of the troposphere, and below here is a description of their role and anticipated effect of their exclusion.
-\begin{itemize}
-\item The formation of the nitric acid ($HNO_3$) is a major loss process for $NO_x$ during daytime in the atmosphere.
-Being a 'sticky' molecule $HNO_3$ readily adsorbs to surfaces, particularly if there is water on the surface, and deposits either in precipitation (wet deposition) or in dry form (dry deposition).
-Because of the high deposition velocity, dry deposition of $HNO_3$ can be responsible for much of the removal of inorganic nitrogen from the troposphere \citep{Finlayson-Pitts2000}, which means that exclusion of $HNO_3$ (formation and removal)/(chemistry) from the chemical mechanism potentially leads to overproduction of ozone in the model, especially at high $NO_2$ concentrations because the competition between $NO_2$ photolysis and its reaction with $OH$ radical ($NO_2 + OH \xrightarrow{M} HNO_3$) is ignored.
-\item Since this research is focused on the daytime chemistry some reactions that are believed to be significant in nighttime are not included into the chemical mechanism of the model.
-Among them are reactions that involve the nitrate radical ($NO_3$) and dinitrogen pentoxide ($N_2O_5$).
-\begin{itemize}
-\item $NO_3$ radical rapidly photolyzes during the daylight hours, but in nighttime its concentration can increase to measurable levels \citep{Atkinson2008}.
-This increase in concentration and a lesser availability of the OH radical (which sources are mostly photolytic in nature) make $NO_3$ radical a major contributor to the night chemistry of organics in the troposphere \citep{Finlayson-Pitts2000}.
-However, exclusion of $NO_3$ radical reactions from our chemical mechanism should not have a huge impact on the results anyway since the mechanism considers mainly decomposition of alkanes with which $NO_3$ radical reacts to a generally minor extent \citep{Atkinson2008}???.
-\item ! the nighttime NO3 radical reaction is of minor significance in term of the overall removal of the alkanes (Atkinson, 1990a)
-\item In addition to reacting with organics, $NO_3$ also reacts with $NO_2$, forming dinitrogen pentoxide ($NO_3 + NO_2 \xleftrightarrow{M} N_2O_5$) which in turn is an important nighttime source of nitric acid through its rapid hydrolysis on wet surfaces and aerosol particles ($N_2O_5 + H_2O \xrightarrow{surf., aerosol} 2HNO_3$) \citep{Finlayson-Pitts2000}.
-Exclusion of $N_2O_5$ from the chemical mechanism assumes no cycling of nitrogen oxides within the troposphere through $NO_3$ and $N_2O_5$, eventually forming $HNO_3$, which are not included into the mechanism anyway.
-\item ! In any case, because of this equilibrium, sink for N2O5 such as hydrolysis are, in essence, also sinks for NO3 as well.
-\end{itemize}
-\item Another species involved in $NO_x$ tropospheric chemistry is the nitrous acid ($HONO$).
-The major source of nitrous acid is believed to be heterogeneous 'dark' reactions of $NO_2$, including hydrolysis (reaction with water vapour) of $NO_2$ on aerosol and particulate matter surfaces in nighttime.
-During the daytime, $HONO$ can be formed by the reaction of $OH$ radical with $NO$ ($OH + NO \xrightarrow{M} HONO$).
-However, since $HONO$ strongly absorbs UV radiation, it rapidly photolyzes, and under particular conditions (when the nighttime $HONO$ concentration reaches almost 15 ppb \citep{Finlayson-Pitts2000}), photolysis of nighttime-generated $HONO$ can become a major source of $OH$ radicals in the early morning hours ($HONO + hv \rightarrow OH + NO$) \citep{Lammel1996}.
-Therefore exclusion of $HONO$ from the chemical mechanism potentially leads to underestimation of $OH$ concentrations, but having neither formation nor removal of $HONO$ should balance this effect out???.
-\item Peroxyacetyl nitrate ($CH_3C(O)OONO_2$, PAN) is the most abundant organic nitrate in the troposphere \citep{Hewitt1994}.
-A classic precursor to the PAN is acetaldehyde ($CH_3CHO$) which oxidation produces PAN through the following reaction sequence:
-\begin{itemize}
-\item[] $CH_3CHO + OH \rightarrow CH_3CO + H_2O$
-\item[] $CH_3CO + O_2 \xrightarrow{M} CH_3C(O)OO$
-\item[] $CH_3C(O)OO + NO_2 \xrightarrow{M} CH_3C(O)OONO_2$.
-\end{itemize}
-Of all the possible fates for PAN in the atmosphere, thermal decomposition is usually the most important because the rate constant for PAN decomposition strongly depends on temperature.
-At low temperatures it is quite stable, but at higher temperatures it decomposes giving $NO_2$ and the acetylperoxy radical ($CH_3C(O)OO$)
-\begin{itemize}
-\item[] $CH_3C(O)OONO_2 \xleftrightarrow{M} NO_2 + CH_3C(O)OO$.
-\end{itemize}
-This temperature dependence has important implications for the role of PAN.
-When it is formed at lower temperatures (e.g., at high altitude) or transported into colder regions, PAN can act as an $NO_x$ reservoir.
-Observations of PAN and the sum of reactive nitrogen ($NO_y$) confirm it revealing that PAN can constitute up to 90\% of the total $NO_y$ budget at higher northern latitudes or higher altitudes (e.g., \citep{Jacobi2000}).
-When an air mass containing PAN is transported into warmer regions, $NO_2$ tied up in PAN can be released back and cause the formation of $O_3$ and $HNO_3$ far away from the original $NO_x$ source.
-Moreover, if sufficient $NO$ is present, $CH_3C(O)OO$ released from PAN can react in the following manner:
-\begin{itemize}
-\item[] $CH_3C(O)OO + NO \rightarrow CH_3C(O)O + NO_2$
-\item[] $CH_3C(O)O \rightarrow CH_3 + CO_2$
-\end{itemize}
-followed by the reaction of $CH_3$ with $O_2$, etc.
-to form $HCHO$ and $HO_2$, or $HCHO$, $CH_3OH$ and $CH_3OOH$ under low-$NO_x$ conditions.
-As a result of this chemistry, PAN can act as an accelerator for photochemical smog formation \citep{Finlayson-Pitts2000}.
-Exclusion of PAN from the chemical mechanism of the model means that we omit the global transport of $NO_x$ provided by PAN, and therefore do not fully estimate tropospheric ozone production.
-However, since the focus of this research is on the impact of alkyl nitrate chemistry on ozone such a miscount is acceptable due to purely theoretical purpose of the research.
-\item Inclusion of all chemistry listed above into the model is a subject for future work.
-\end{itemize}
-
-The constructed chemical mechanism has a limited representation of organic chemistry.
-It includes oxidation of only seven alkanes: methane ($CH_4$), ethane ($C_2H_6$), propane ($C_3H_8$), n-butane ($nC_4H_{10}$), i-butane ($iC_4H_{10}$), n-pentane ($nC_5H_{12}$) and i-pentane ($iC_5H_{12}$).
-$CH_4$ undergoes a full oxidation in the model, but oxidation of other six alkanes follows the reaction sequence:
-\begin{itemize}
-\item[] $RH + OH \rightarrow RO_2 + H_2O$
-\item[] $RO_2 + HO_2 \rightarrow ROOH + O2$
-\item[] $RO_2 + NO \rightarrow RO + NO_2$
-\item[] $ROOH + hv \rightarrow RO + OH$
-\item[] $ROOH + OH \rightarrow products$
-\item[] $RO + O_2 \rightarrow products$
-\end{itemize}
-and then is 'terminated' after formation of certain species.
-Ethane oxidation is terminated after formation of the acetylperoxyl radicals ($C_2H_3CO_3$), propane - after formation of the propyldioxy radicals ($C_3H_5O_3$), n-butane - after formation of ethanal ($CH_3CHO$), butanal ($C_3H_7CHO$), methyl ethyl ketone ($C_2H_5COCH_3$), and 1-($\lambda^1$-oxidanylmethoxy)-2-methoxy-ethane ($C_4H_9O_3$), i-butane - after formation of 2-methylpropanal ($iso-C_3H_7CHO$) and acetone ($CH_3COCH_3$), n-pentane - after formation of pentanal ($C_4H_9CHO$), methyl propyl ketone ($n-C_3H_7COCH_3$), and diethyl ketone ($(C_2H_5)_2CO$), i-pentane - after formation of 2-methylbutanal ($sec-C_4H_9CHO$) and isopropyl methyl ketone ($iso-C_3H_7COCH_3$).
-It has been done to limit the number of reactions in the mechanism, which otherwise would include 1678 reactions describing evolution of 539 species, and therefore would be computationally too expensive.
-For the same reason all self-reactions of organic peroxy radicals ($RO_2$) are omitted as well.
-To avoid a huge impact of incomplete oxidation of alkanes on the results the model's simulation time has been restricted to 24 hours.
-
-Lastly, the chemical mechanism of the model does not include halogen chemistry.
-Despite that halogens (chlorine ($Cl$), bromine ($Br$) and iodine ($I$)) are very reactive chemical compounds mostly known by their role in anthropogenic stratospheric ozone-depletion chemistry, and that they are directly (for example, reaction of the $Cl$ radical with hydrocarbons (e.g., $CH_4$)) and indirectly (coupling with the cycles of sulfur and nitrogen that eventually influence $O_3$, $OH$ and $NO_x$ concentrations) involved in the tropospheric chemistry \citep{VonGlasow2014,Platt2003}, they are excluded from the chemical mechanism to keep it more transparent.
-\subsection{Photolysis}
-To estimate photolysis rates for a given geographical location, one must take into account the latitude and season, as well as the time of the day.
-For the purpose of this research 24-hour average photolysis rates are calculated externally (outside FACSIMILE program) for 1 July using the formula and parametrisations from the MCM v3.3.
-The formula that describes variation of photolysis rate with solar zenith angle for the core photolysis reactions in the MCM was obtained by its authors using cross-section and quantum yield data and two stream isotopic scattering model described in \citep{Hayman1997}.
-Their calculations were performed for clear sky conditions at an altitude of 0.5 km for 1 July at a latitude of 45$^{\circ}$N, resulting in the following formula:
-\begin{equation} \label{eq:MCMphotolysis}
-\renewcommand{\theequation}{\arabic{equation}}
-J = l(cos\chi)^mexp(-nsec\chi)
-\end{equation}
-where $J$ - photolysis rate coefficient, $\chi$ - solar zenith angle, $l$, $m$ and $n$ - optimising parameters \citep{Jenkin1997,Saunders2003}.
-
-The values of solar zenith angle are calculated using NOAA Solar Calculator (http://www.esrl.noaa.gov/gmd/grad/solcalc/).
-
-\subsection{Design of experiments}
-To investigate the influence of alkyl nitrate chemistry on (potential changes in) tropospheric $O_3$ production, box model runs were performed at a series of different $NO_x$ and $VOCs$ initial concentrations, so that an $O_3$ isopleth plot could be constructed, similar to those found in \citep{Dodge1977} and \citep{Sillman1999}.
-Table \ref{tab:setupO3CO} shows initial concentrations of 'independent'/global parameters and indicates which of them are fixed or allowed to vary during the model run.
-Table \ref{tab:setupNOx} and Table \ref{tab:setupVOCs} show the initial concentrations of $NO_x$ and $VOCs$ respectively that are used at a series of experiments.
-
-The lower limit of the range of $NO_x$ and $VOCs$ initial concentrations is based on observational data from Mauna Loa station (Hawaii, US) which is believed to be representative for conditions of remote troposphere.
-The upper limit is based on data from
-Marylebone Road station (London, UK) which is a kerbside monitoring site located in central London, in close proximity to a 6 lane road, with very high traffic flow and frequent congestion \citep{VonSchneidemesser2010}, therefore being representative for conditions of urban troposphere.
-
-Two experiments have been performed.
-The aim of the first one is to reveal the total impact of alkyl nitrate chemistry on $O_3$ production, the aim of the second one is to estimate the contribution of individual alkyl nitrates to changes in $O_3$ production.
-
-For the first experiment the model was run under two configurations, (1) with all alkyl nitrate chemistry switched on, and (2) with all alkyl nitrate chemistry switched off.
-In every model run $VOCs$ concentrations were (changed according to the data from Table \ref{tab:setupVOCs} and) were fixed throughout the run, while the initial concentrations of $NO_x$ were (changed according to the data from Table \ref{tab:setupNOx} and were) allowed to vary throughout the run.
-As a result, the first experiment consists of 253 model runs including 11 runs (varying $NO_x$) with inorganic chemistry only.
-
-For the second experiment the model was run using the same approach concerning variations in $NO_x$ and $VOCs$ concentrations, but changes in alkyl nitrate chemistry were treated differently.
-In every model run reactions of individual alkyl nitrates or their groups (defined in Table\ref{tab:RHandANS} were either switched on or switched off, resulting in other additional 847 model runs.
-
-\begin{table} % RH and ANs
-\caption{Alkyl nitrates and their parent alkanes}\label{tab:RHandANS}
-\centering
-\begin{tabular}{llll}
-\hline
-Parent alkane& Structural formula          & Name                     & Group\\
-\hline
-$CH_4$       & $CH_3ONO_2$                 & methyl nitrate           & $CH_3ONO_2$ \\
-\hline
-$C_2H_6$     & $CH_3CH_2ONO_2$             & ethyl nitrate            & $C_2H_5ONO_2$ \\
-\hline
-$C_3H_8$     & $(CH_3)_2CHONO_2$           & iso-propyl nitrate       & $\Sigma C_3H_7ONO_2$\\
-             & $CH_3(CH_2)_2ONO_2$         & n-propyl nitrate         & \\
-\hline
-$nC_4H_{10}$ & $CH_3(CH_2)_3ONO_2$         & n-butyl nitrate          & $\Sigma nC_4H_9ONO_2$\\
-             & $CH_3CH_2CH(CH_3)ONO_2$     & sec-butyl nitrate        & \\
-\hline
-$iC_4H_{10}$ & $CH_3CH(CH_3)CH_2ONO_2$     & iso-butyl nitrate        & $\Sigma iC_4H_9ONO_2$\\
-             & $CH_2C(CH_3)CH_2ONO_2$      & tert-butyl nitrate       & \\
-\hline
-$nC_5H_{12}$ & $CH_3(CH_2)_4ONO_2$         & n-pentyl nitrate         & $\Sigma nC_5H_{11}ONO_2$\\
-             & $CH_3(CH_2)_2CH(CH_3)ONO_2$ & 2-pentyl nitrate         & \\
-             & $???$                       & 3-pentyl nitrate         & \\
-\hline
-$iC_5H_{12}$ & $???$                       & 2-methylbutyl nitrate    & $\Sigma iC_5H_{11}ONO_2$\\
-             & $(CH_3)_2CH(CH_2)_2ONO_2$   & 3-methyl-2-butyl nitrate & \\
-             & $(CH_3)_2C(CH_2CH_3)ONO_2$  & 2-methyl-2-butyl nitrate & \\
-\hline
-\end{tabular}
-\end{table}
-
-!! For long-lived species, averaging may be reasonable.
-
-\begin{table} % model setup: O3, CO
-\caption{Fixed and variable parameters used in box model runs}\label{tab:setupO3CO}
-\centering
-\begin{tabular}{ccc}
-\hline
-Parameter & Initial mixing ratio & Type \\
-\hline
-$H_2O$    & 2\%                  & Fixed    \\
-$O_2$     & 21\%                 & Fixed    \\
-$N_2$     & 78\%                 & Fixed    \\
-$H_2$     & 525 ppb              & Variable \\
-$O_3$     & 40 ppb               & Variable \\
-$CO$      & 100 ppb              & Variable \\
-\hline
-\end{tabular}
-\end{table}
-
-\begin{table} % model setup: NOx
-\caption{Model setup: NOx initial concentrations (ppb)}\label{tab:setupNOx}
-\centering
-\begin{tabular}{ccc}
-\hline
-$NO$      & $NO_2$      & NOx  ($=NO+NO_2$) \\
-\hline
-0.0015    & 0.0035      & 0.005 \\
-0.007     & 0.018       & 0.025 \\
-0.015     & 0.035       & 0.05  \\
-0.03      & 0.07        & 0.1   \\
-0.07      & 0.18        & 0.25  \\
-0.15      & 0.35        & 0.5   \\
-0.25      & 0.5         & 0.75  \\
-0.3       & 0.7         & 1     \\
-0.7       & 1.8         & 2.5   \\
-1.5       & 3.5         & 5     \\
-3         & 7           & 10    \\
-\hline
-\end{tabular}
-\end{table}
-
-\begin{table} % model setup: VOCs
-\caption{Model setup: VOCs concentrations (ppb)}\label{tab:setupVOCs}
-\centering
-\begin{tabular}{ccccccccc}
-\hline
-Experiment number & $CH_4$ & $C_2H_6$ & $C_3H_8$ & $nC_4H_{10}$ & $iC_4H_{10}$ & $nC_5H_{12}$ & $iC_5H_{12}$ \\
-\hline
-1  & 1780 &	0.620 &	0.100 &	0.050 &	0.250 &	0.250 &	0.250 \\
-2  & 1807 &	1.258 &	0.390 &	0.245 &	0.325 &	0.275 &	0.425 \\
-3  & 1834 &	1.896 &	0.680 &	0.440 &	0.400 &	0.300 &	0.600 \\
-4  & 1861 &	2.534 &	0.970 &	0.635 &	0.475 &	0.325 &	0.775 \\
-5  & 1888 &	3.172 &	1.260 &	0.830 &	0.550 &	0.350 &	0.950 \\
-6  & 1915 &	3.810 &	1.550 &	1.025 &	0.625 &	0.375 &	1.125 \\
-7  & 1942 &	4.448 &	1.840 &	1.220 &	0.700 &	0.400 &	1.300 \\
-8  & 1969 &	5.086 &	2.130 &	1.415 &	0.775 &	0.425 &	1.475 \\
-9  & 1996 &	5.724 &	2.420 &	1.610 &	0.850 &	0.450 &	1.650 \\
-10 & 2023 &	6.362 &	2.710 &	1.805 &	0.925 &	0.475 &	1.825 \\
-11 & 2050 &	7.000 &	3.000 &	2.000 &	1.000 &	0.500 &	2.000 \\
-\hline
-\end{tabular}
-\end{table}
-
-\begin{figure}[h]
-\includegraphics[width=0.5\linewidth]{{./pics/Sillman1999_Fig1}.png}
-%\caption{chem_allAN_netO3_rate}
-\label{fig:O3observations}
-\end{figure}
-
 \subsection{Bertman}
@@
 
-\section{Results} \label{sec:res}
-Dependence of alkyl nitrate yield on carbon number (paper about n-alkanes): bigger yield of ANs from larger carbon number means that the potential for contributing to photochemical air pollution may be less for the larger (C>6) n-alkanes that for the smaller ones.
+\section{Methodology} \label{sec:method}
+\subsection{Model development and description}
+A zero-dimensional (box) model is developed to investigate the impact of alkyl nitrate chemistry on ozone production efficiency.
+Its code is written using FACSIMILE for Windows.
+FACSIMILE is a special high-level programming language which is a powerful modelling tool designed to efficiently solve differential equations including those resulting from a set of chemical reactions.
 
-\subsection{Justification of model performance}
+! \citep{Newland2013}
+[The model is written using the programming language FACSIMILE (Curtis and Sweetenham, 1987) and the differential equations are assembled and integrated using the Gear’s method (1971).
+The mixing ratios of the input species in the model are driven entirely by the emissions, the loss processes and the transport with no redistribution of the burden to ensure that mixing ratios agree with observations.]
+
+The model is intended to represent the steady state of the troposphere under mean summer daytime meteorology and clear skies.
+To achieve that a number of assumptions have been made.
+First, the model assumes a uniform mixing of individual constituents of the troposphere, second, there are no atmospheric transport (no 'winds') and no emissions of species into the box meaning that the time evolution of each species is controlled by the chemical interactions alone and determined by their initial concentrations.
+The air temperature and relative humidity are assumed to be $25^{\circ}C$ and 70\% respectively and are fixed in all model runs.
+Other assumptions concerning the chemical mechanism and photolysis rates can be found in the following sections.
+
+\subsection{Chemical mechanism}
+The chemical mechanism of the model is based on recommendations from \citep{Atkinson2004} and information from the Master Chemical Mechanism, MCM v3.3 \citep{Jenkin1997,Saunders2003}, retrieved via website: http://mcm.leeds.ac.uk/MCM.
+On a whole it includes 90 species and 165 reactions.
+The mechanism is developed to represent the chemistry that governs the production and loss of ozone, hydrocarbons and alkyl nitrates.
+These include:
+
+! \citep{Newland2013}
+[The majority of the kinetic reaction rates and photochemical data used in the full chemistry model were taken from the IUPAC Subcommittee for Gas Kinetic Data Evaluation (http://www.iupac-kinetic.ch.cam.ac.uk; Atkinson et al., 2004; Atkinson et al., 2006; Atkinson et al., 2008).
+If the required data were not provided by IUPAC then data were taken from the NASA / JPL ‘Chemical Kinetics and Photochemical Data for Use in Atmospheric Studies: Evaluation No.
+17’ (Sander et al., 2011).
+Both these sources consider the full range of literature available on each reaction and provide a recommendation based on a synthesis of this data.
+There are sometimes minor differences between the recommendations of each organisation.]
+
+\begin{itemize}
+\item the formation of ozone as a result of the photolysis of $NO_2$ followed by the reaction of ground-state oxygen atom ($O(^3P)$) with molecular oxygen and a third body, M ($N_2$ or $O_2$);
+\item the loss of ozone through its photolysis and reactions with $O(^3P)$, $OH$, $HO_2$ and $NO$, among which the last is of particular interest because it gives back $NO_2$;
+\item the formation of alkylperoxy radicals ($RO_2$) from oxidation of the hydrocarbons;
+\item the competition of $NO$ and $HO_2$ for reaction with $HO_2$ and $RO_2$;
+\item the formation of alkyl nitrates which adds up to the competition of $NO$ and $HO_2$ for reaction with the peroxy radicals;
+\item the loss of alkyl nitrates through their photolysis and reaction with $OH$;
+\item the subsequent reactions of hydroperoxides and aldehydes formed from the organic reactions mentioned above.
+\end{itemize}
+The full list of reactions can be found in Appendix ??
+\ref{sec:appendix1}.
+It is important to mention that due to purely theoretical purpose of this research and its focus on investigation of the impact of alkyl nitrate chemistry on ozone production some of the reactions that lead to formation of species that serve as a sink for $NO_x$ are not included into the chemical mechanism.
+However, these reactions play an important role in the chemistry of the troposphere, and below here is a description of their role and anticipated effect of their exclusion.
+\begin{itemize}
+\item The formation of the nitric acid ($HNO_3$) is a major loss process for $NO_x$ during daytime in the atmosphere.
+Being a 'sticky' molecule $HNO_3$ readily adsorbs to surfaces, particularly if there is water on the surface, and deposits either in precipitation (wet deposition) or in dry form (dry deposition).
+Because of the high deposition velocity, dry deposition of $HNO_3$ can be responsible for much of the removal of inorganic nitrogen from the troposphere \citep{Finlayson-Pitts2000}, which means that exclusion of $HNO_3$ (formation and removal)/(chemistry) from the chemical mechanism potentially leads to overproduction of ozone in the model, especially at high $NO_2$ concentrations because the competition between $NO_2$ photolysis and its reaction with $OH$ radical ($NO_2 + OH \xrightarrow{M} HNO_3$) is ignored.
+\item Since this research is focused on the daytime chemistry some reactions that are believed to be significant in nighttime are not included into the chemical mechanism of the model.
+Among them are reactions that involve the nitrate radical ($NO_3$) and dinitrogen pentoxide ($N_2O_5$).
+\begin{itemize}
+\item $NO_3$ radical rapidly photolyzes during the daylight hours, but in nighttime its concentration can increase to measurable levels \citep{Atkinson2008}.
+This increase in concentration and a lesser availability of the OH radical (which sources are mostly photolytic in nature) make $NO_3$ radical a major contributor to the night chemistry of organics in the troposphere \citep{Finlayson-Pitts2000}.
+However, exclusion of $NO_3$ radical reactions from our chemical mechanism should not have a huge impact on the results anyway since the mechanism considers mainly decomposition of alkanes with which $NO_3$ radical reacts to a generally minor extent \citep{Atkinson2008}???.
+\item ! the nighttime NO3 radical reaction is of minor significance in term of the overall removal of the alkanes (Atkinson, 1990a)
+\item In addition to reacting with organics, $NO_3$ also reacts with $NO_2$, forming dinitrogen pentoxide ($NO_3 + NO_2 \xleftrightarrow{M} N_2O_5$) which in turn is an important nighttime source of nitric acid through its rapid hydrolysis on wet surfaces and aerosol particles ($N_2O_5 + H_2O \xrightarrow{surf., aerosol} 2HNO_3$) \citep{Finlayson-Pitts2000}.
+Exclusion of $N_2O_5$ from the chemical mechanism assumes no cycling of nitrogen oxides within the troposphere through $NO_3$ and $N_2O_5$, eventually forming $HNO_3$, which are not included into the mechanism anyway.
+\item ! In any case, because of this equilibrium, sink for N2O5 such as hydrolysis are, in essence, also sinks for NO3 as well.
+\end{itemize}
+\item Another species involved in $NO_x$ tropospheric chemistry is the nitrous acid ($HONO$).
+The major source of nitrous acid is believed to be heterogeneous 'dark' reactions of $NO_2$, including hydrolysis (reaction with water vapour) of $NO_2$ on aerosol and particulate matter surfaces in nighttime.
+During the daytime, $HONO$ can be formed by the reaction of $OH$ radical with $NO$ ($OH + NO \xrightarrow{M} HONO$).
+However, since $HONO$ strongly absorbs UV radiation, it rapidly photolyzes, and under particular conditions (when the nighttime $HONO$ concentration reaches almost 15 ppb \citep{Finlayson-Pitts2000}), photolysis of nighttime-generated $HONO$ can become a major source of $OH$ radicals in the early morning hours ($HONO + hv \rightarrow OH + NO$) \citep{Lammel1996}.
+Therefore exclusion of $HONO$ from the chemical mechanism potentially leads to underestimation of $OH$ concentrations, but having neither formation nor removal of $HONO$ should balance this effect out???.
+\item Peroxyacetyl nitrate ($CH_3C(O)OONO_2$, PAN) is the most abundant organic nitrate in the troposphere \citep{Hewitt1994}.
+A classic precursor to the PAN is acetaldehyde ($CH_3CHO$) which oxidation produces PAN through the following reaction sequence:
+\begin{itemize}
+\item[] $CH_3CHO + OH \rightarrow CH_3CO + H_2O$
+\item[] $CH_3CO + O_2 \xrightarrow{M} CH_3C(O)OO$
+\item[] $CH_3C(O)OO + NO_2 \xrightarrow{M} CH_3C(O)OONO_2$.
+\end{itemize}
+Of all the possible fates for PAN in the atmosphere, thermal decomposition is usually the most important because the rate constant for PAN decomposition strongly depends on temperature.
+At low temperatures it is quite stable, but at higher temperatures it decomposes giving $NO_2$ and the acetylperoxy radical ($CH_3C(O)OO$)
+\begin{itemize}
+\item[] $CH_3C(O)OONO_2 \xleftrightarrow{M} NO_2 + CH_3C(O)OO$.
+\end{itemize}
+This temperature dependence has important implications for the role of PAN.
+When it is formed at lower temperatures (e.g., at high altitude) or transported into colder regions, PAN can act as an $NO_x$ reservoir.
+Observations of PAN and the sum of reactive nitrogen ($NO_y$) confirm it revealing that PAN can constitute up to 90\% of the total $NO_y$ budget at higher northern latitudes or higher altitudes (e.g., \citep{Jacobi2000}).
+When an air mass containing PAN is transported into warmer regions, $NO_2$ tied up in PAN can be released back and cause the formation of $O_3$ and $HNO_3$ far away from the original $NO_x$ source.
+Moreover, if sufficient $NO$ is present, $CH_3C(O)OO$ released from PAN can react in the following manner:
+\begin{itemize}
+\item[] $CH_3C(O)OO + NO \rightarrow CH_3C(O)O + NO_2$
+\item[] $CH_3C(O)O \rightarrow CH_3 + CO_2$
+\end{itemize}
+followed by the reaction of $CH_3$ with $O_2$, etc.
+to form $HCHO$ and $HO_2$, or $HCHO$, $CH_3OH$ and $CH_3OOH$ under low-$NO_x$ conditions.
+As a result of this chemistry, PAN can act as an accelerator for photochemical smog formation \citep{Finlayson-Pitts2000}.
+Exclusion of PAN from the chemical mechanism of the model means that we omit the global transport of $NO_x$ provided by PAN, and therefore do not fully estimate tropospheric ozone production.
+However, since the focus of this research is on the impact of alkyl nitrate chemistry on ozone such a miscount is acceptable due to purely theoretical purpose of the research.
+\item Inclusion of all chemistry listed above into the model is a subject for future work.
+\end{itemize}
+
+The constructed chemical mechanism has a limited representation of organic chemistry.
+It includes oxidation of only seven alkanes: methane ($CH_4$), ethane ($C_2H_6$), propane ($C_3H_8$), n-butane ($nC_4H_{10}$), i-butane ($iC_4H_{10}$), n-pentane ($nC_5H_{12}$) and i-pentane ($iC_5H_{12}$).
+$CH_4$ undergoes a full oxidation in the model, but oxidation of other six alkanes follows the reaction sequence:
+\begin{itemize}
+\item[] $RH + OH \rightarrow RO_2 + H_2O$
+\item[] $RO_2 + HO_2 \rightarrow ROOH + O2$
+\item[] $RO_2 + NO \rightarrow RO + NO_2$
+\item[] $ROOH + hv \rightarrow RO + OH$
+\item[] $ROOH + OH \rightarrow products$
+\item[] $RO + O_2 \rightarrow products$
+\end{itemize}
+and then is 'terminated' after formation of certain species.
+Ethane oxidation is terminated after formation of the acetylperoxyl radicals ($C_2H_3CO_3$), propane - after formation of the propyldioxy radicals ($C_3H_5O_3$), n-butane - after formation of ethanal ($CH_3CHO$), butanal ($C_3H_7CHO$), methyl ethyl ketone ($C_2H_5COCH_3$), and 1-($\lambda^1$-oxidanylmethoxy)-2-methoxy-ethane ($C_4H_9O_3$), i-butane - after formation of 2-methylpropanal ($iso-C_3H_7CHO$) and acetone ($CH_3COCH_3$), n-pentane - after formation of pentanal ($C_4H_9CHO$), methyl propyl ketone ($n-C_3H_7COCH_3$), and diethyl ketone ($(C_2H_5)_2CO$), i-pentane - after formation of 2-methylbutanal ($sec-C_4H_9CHO$) and isopropyl methyl ketone ($iso-C_3H_7COCH_3$).
+It has been done to limit the number of reactions in the mechanism, which otherwise would include 1678 reactions describing evolution of 539 species, and therefore would be computationally too expensive.
+For the same reason all self-reactions of organic peroxy radicals ($RO_2$) are omitted as well.
+To avoid a huge impact of incomplete oxidation of alkanes on the results the model's simulation time has been restricted to 24 hours.
+
+\citep{Newland2013}
+[The alkanes and alkyl nitrates studied in this work were all explicitly described in the model with oxidation of the alkane giving a peroxy radical and reaction of the peroxy radical with NO giving an alkyl nitrate.]
+
+Lastly, the chemical mechanism of the model does not include halogen chemistry.
+Despite that halogens (chlorine ($Cl$), bromine ($Br$) and iodine ($I$)) are very reactive chemical compounds mostly known by their role in anthropogenic stratospheric ozone-depletion chemistry, and that they are directly (for example, reaction of the $Cl$ radical with hydrocarbons (e.g., $CH_4$)) and indirectly (coupling with the cycles of sulfur and nitrogen that eventually influence $O_3$, $OH$ and $NO_x$ concentrations) involved in the tropospheric chemistry \citep{VonGlasow2014,Platt2003}, they are excluded from the chemical mechanism to keep it more transparent.
+\subsection{Photolysis}
+To estimate photolysis rates for a given geographical location, one must take into account the latitude and season, as well as the time of the day.
+For the purpose of this research 24-hour average photolysis rates are calculated externally (outside FACSIMILE program) for 1 July using the formula and parametrisations from the MCM v3.3.
+The formula that describes variation of photolysis rate with solar zenith angle for the core photolysis reactions in the MCM was obtained by its authors using cross-section and quantum yield data and two stream isotopic scattering model described in \citep{Hayman1997}.
+Their calculations were performed for clear sky conditions at an altitude of 0.5 km for 1 July at a latitude of 45$^{\circ}$N, resulting in the following formula:
+\begin{equation} \label{eq:MCMphotolysis}
+\renewcommand{\theequation}{\arabic{equation}}
+J = l(cos\chi)^mexp(-nsec\chi)
+\end{equation}
+where $J$ - photolysis rate coefficient, $\chi$ - solar zenith angle, $l$, $m$ and $n$ - optimising parameters \citep{Jenkin1997,Saunders2003}.
+
+The values of solar zenith angle are calculated using NOAA Solar Calculator (http://www.esrl.noaa.gov/gmd/grad/solcalc/).
+
+\subsection{Design of experiments}
+To investigate the influence of alkyl nitrate chemistry on (potential changes in) tropospheric $O_3$ production, box model runs were performed at a series of different $NO_x$ and $VOCs$ initial concentrations, so that an $O_3$ isopleth plot could be constructed, similar to those found in \citep{Dodge1977} and \citep{Sillman1999}.
+Table \ref{tab:setupO3CO} shows initial concentrations of 'independent'/global parameters and indicates which of them are fixed or allowed to vary during the model run.
+Table \ref{tab:setupNOx} and Table \ref{tab:setupVOCs} show the initial concentrations of $NO_x$ and $VOCs$ respectively that are used at a series of experiments.
+
+The lower limit of the range of $NO_x$ and $VOCs$ initial concentrations is based on observational data from Mauna Loa station (Hawaii, US) which is believed to be representative for conditions of remote troposphere.
+The upper limit is based on data from
+Marylebone Road station (London, UK) which is a kerbside monitoring site located in central London, in close proximity to a 6 lane road, with very high traffic flow and frequent congestion \citep{VonSchneidemesser2010}, therefore being representative for conditions of urban troposphere.
+
+Two experiments have been performed.
+The aim of the first one is to reveal the total impact of alkyl nitrate chemistry on $O_3$ production, the aim of the second one is to estimate the contribution of individual alkyl nitrates to changes in $O_3$ production (sensitivity!).
+
+For the first experiment the model was run under two configurations, (1) with all alkyl nitrate chemistry switched on, and (2) with all alkyl nitrate chemistry switched off.
+In every model run $VOCs$ concentrations were (changed according to the data from Table \ref{tab:setupVOCs} and) were fixed throughout the run, while the initial concentrations of $NO_x$ were (changed according to the data from Table \ref{tab:setupNOx} and were) allowed to vary throughout the run.
+As a result, the first experiment consists of 253 model runs including 11 runs (varying $NO_x$) with inorganic chemistry only.
+
+For the second experiment the model was run using the same approach concerning variations in $NO_x$ and $VOCs$ concentrations, but changes in alkyl nitrate chemistry were treated differently.
+In every model run reactions of individual alkyl nitrates or their groups (defined in Table \ref{tab:RHandANS}) were either switched on or switched off, resulting in other 847 additional model runs.
+
+\begin{table} % RH and ANs
+\caption{Alkyl nitrates and their parent alkanes}\label{tab:RHandANS}
+\centering
+\begin{tabular}{llll}
+\hline
+Parent alkane& Structural formula          & Name                     & Group\\
+\hline
+$CH_4$       & $CH_3ONO_2$                 & methyl nitrate           & $CH_3ONO_2$ \\
+\hline
+$C_2H_6$     & $CH_3CH_2ONO_2$             & ethyl nitrate            & $C_2H_5ONO_2$ \\
+\hline
+$C_3H_8$     & $(CH_3)_2CHONO_2$           & iso-propyl nitrate       & $\Sigma C_3H_7ONO_2$\\
+             & $CH_3(CH_2)_2ONO_2$         & n-propyl nitrate         & \\
+\hline
+$nC_4H_{10}$ & $CH_3(CH_2)_3ONO_2$         & n-butyl nitrate          & $\Sigma nC_4H_9ONO_2$\\
+             & $CH_3CH_2CH(CH_3)ONO_2$     & sec-butyl nitrate        & \\
+\hline
+$iC_4H_{10}$ & $CH_3CH(CH_3)CH_2ONO_2$     & iso-butyl nitrate        & $\Sigma iC_4H_9ONO_2$\\
+             & $CH_2C(CH_3)CH_2ONO_2$      & tert-butyl nitrate       & \\
+\hline
+$nC_5H_{12}$ & $CH_3(CH_2)_4ONO_2$         & n-pentyl nitrate         & $\Sigma nC_5H_{11}ONO_2$\\
+             & $CH_3(CH_2)_2CH(CH_3)ONO_2$ & 2-pentyl nitrate         & \\
+             & $???$                       & 3-pentyl nitrate         & \\
+\hline
+$iC_5H_{12}$ & $???$                       & 2-methylbutyl nitrate    & $\Sigma iC_5H_{11}ONO_2$\\
+             & $(CH_3)_2CH(CH_2)_2ONO_2$   & 3-methyl-2-butyl nitrate & \\
+             & $(CH_3)_2C(CH_2CH_3)ONO_2$  & 2-methyl-2-butyl nitrate & \\
+\hline
+\end{tabular}
+\end{table}
+
+!! For long-lived species, averaging may be reasonable.
+
+\begin{table} % model setup: O3, CO
+\caption{Fixed and variable 'independent' parameters used in box model runs}\label{tab:setupO3CO}
+\centering
+\begin{tabular}{ccc}
+\hline
+Parameter & Initial mixing ratio & Type \\
+\hline
+$H_2O$    & 2\%                  & Fixed    \\
+$O_2$     & 21\%                 & Fixed    \\
+$N_2$     & 78\%                 & Fixed    \\
+$H_2$     & 525 ppb              & Variable \\
+$O_3$     & 40 ppb               & Variable \\
+$CO$      & 100 ppb              & Variable \\
+\hline
+\end{tabular}
+\end{table}
+
+\begin{table} % model setup: NOx
+\caption{Model setup: NOx initial concentrations (ppb)}\label{tab:setupNOx}
+\centering
+\begin{tabular}{ccc}
+\hline
+$NO$      & $NO_2$      & NOx  ($=NO+NO_2$) \\
+\hline
+0.0015    & 0.0035      & 0.005 \\
+0.007     & 0.018       & 0.025 \\
+0.015     & 0.035       & 0.05  \\
+0.03      & 0.07        & 0.1   \\
+0.07      & 0.18        & 0.25  \\
+0.15      & 0.35        & 0.5   \\
+0.25      & 0.5         & 0.75  \\
+0.3       & 0.7         & 1     \\
+0.7       & 1.8         & 2.5   \\
+1.5       & 3.5         & 5     \\
+3         & 7           & 10    \\
+\hline
+\end{tabular}
+\end{table}
+
+\begin{table} % model setup: VOCs
+\caption{Model setup: VOCs concentrations (ppb)}\label{tab:setupVOCs}
+\centering
+\begin{tabular}{ccccccccc}
+\hline
+Experiment number & $CH_4$ & $C_2H_6$ & $C_3H_8$ & $nC_4H_{10}$ & $iC_4H_{10}$ & $nC_5H_{12}$ & $iC_5H_{12}$ \\
+\hline
+A  & 1780 &	0.620 &	0.100 &	0.050 &	0.250 &	0.250 &	0.250 \\
+B  & 1807 &	1.258 &	0.390 &	0.245 &	0.325 &	0.275 &	0.425 \\
+C  & 1834 &	1.896 &	0.680 &	0.440 &	0.400 &	0.300 &	0.600 \\
+D  & 1861 &	2.534 &	0.970 &	0.635 &	0.475 &	0.325 &	0.775 \\
+E  & 1888 &	3.172 &	1.260 &	0.830 &	0.550 &	0.350 &	0.950 \\
+F  & 1915 &	3.810 &	1.550 &	1.025 &	0.625 &	0.375 &	1.125 \\
+G  & 1942 &	4.448 &	1.840 &	1.220 &	0.700 &	0.400 &	1.300 \\
+H  & 1969 &	5.086 &	2.130 &	1.415 &	0.775 &	0.425 &	1.475 \\
+I  & 1996 &	5.724 &	2.420 &	1.610 &	0.850 &	0.450 &	1.650 \\
+J  & 2023 &	6.362 &	2.710 &	1.805 &	0.925 &	0.475 &	1.825 \\
+K  & 2050 &	7.000 &	3.000 &	2.000 &	1.000 &	0.500 &	2.000 \\
+\hline
+\end{tabular}
+\end{table}
+\subsection{Model validation}
+Time evolution of species concentrations in the model agrees reasonably well with photochemical theory as it can be seen, for example, from Figure ???.
+At low $NO_x$ levels (50 ppt) $O_3$ is being removed from the system because there is not enough $NO$ that could react with the peroxy radicals ($HO_2$ and $RO_2$) to produce $NO_2$ (which photolysis in conjunction with the reaction $O(^3P)+O_2\xrightarrow{M} O_3$ is the main source of $O_3$ in the troposphere).
+Some additional $O_3$ removal occurs because $HO_2$ radicals can react with $O_3$ leading to the regeneration of $OH$ radicals as part of an $O_3$ depleting $OH$-$HO_2$ ($HO_x$) cycle.
+In other words, the following reaction sequence dominates:
+\begin{itemize}
+\item[] $O_3 + hv \rightarrow O(^1D) + O_2$
+\item[] $O(^1D) + H_2O \rightarrow OH + OH$
+\item[] $O(^1D) \xrightarrow{M} O(^3P)$
+\item[] $O(^3P) + O_2 \xrightarrow{M} O_3$
+\item[] $OH + CO \rightarrow HO_2 + CO_2$
+\item[] $OH + RH \rightarrow RO_2 + H_2O$
+\item[] $HO_2 + HO_2 \rightarrow H_2O_2 + O_2$
+\item[] $RO_2 + HO_2 \rightarrow ROOH + O_2$
+\item[] $O_3 + HO_2 \rightarrow OH + 2O_2$.
+\end{itemize}
+A more pronounced $O_3$ loss takes place under lower $VOC$ concentrations because in this case more $OH$ radicals are left in the system after oxidation of $CO$ and alkanes leading to a greater concentration of $HO_2$ radicals, and therefore to a greater chance of the last reaction in the sequence.
+
+The concentration of NOx is the main factor which determines whether O3 is produced or removed in the atmosphere.
+changing balance of the free-radical reactions, and their impact on O3, as the level of NOx increases.
+
+NO and NO2 exist in a state equilibrium with rapid interconversion between the two molecules in the troposphere.
+
 !! from \citep{Finlayson-Pitts2000}
@@
 
-\begin{figure}[h]
+%\newpage
+%\clearpage
+%\thispagestyle{empty}% empty page style (?)
+%\begin{landscape}% Landscape page
+%\centering
+%	\input{diss_valid_vs_time}
+%\end{landscape}
+
+\section{Results} \label{sec:res}
+\begin{figure}[h] % Sillman1999
+\includegraphics[width=0.5\linewidth]{{./pics/Sillman1999_Fig1_adapted}.png}
+\caption{Isopleths giving net rate of ozone production (ppb/h, solid lines) as a function of $VOC$ (ppbC) and $NO_x$ (ppb) for mean summer daytime meteorology and clear skies.
+The solid lines represent production rates of 1, 2.5, 5, 10, 15, 20 and 30 ppb/h.
+Adapted from \citep{Sillman1999}.}\label{fig:Sillman1999}
+\end{figure}
+\begin{figure}[h] % allAN_netO3_rate
 \includegraphics[width=\linewidth]{D:/FACSIMILE/ANsCBmodel/matlab/ANsCB_pics/ANsNOxVOC/allAN/chem_allAN_netO3_rate.png}
-%\caption{chem_allAN_netO3_rate}
-\label{fig:O3observations}
+\caption{allAN, netO3 rate}\label{fig:allAN_netO3_rate}
 \end{figure}
+
+!
+Dependence of alkyl nitrate yield on carbon number (paper about n-alkanes): bigger yield of ANs from larger carbon number means that the potential for contributing to photochemical air pollution may be less for the larger (C>6) n-alkanes that for the smaller ones.
 
@@
 
-\subsection{Short-chained alkyl nitrates} \label{sec:short_an}
-\subsubsection{Short photochemical processing times} \label{sec:short_an_short_time}
-
 \section{Conclusion} \label{sec:conclusion}
-
-\newpage
-\clearpage
-\thispagestyle{empty}% empty page style (?)
-\begin{landscape}% Landscape page
-\centering
-	%\input{subfig_28_land}
-\end{landscape}
-
 \bibliography{diss_refs}

### 2015-07-27 MSc_diss.tex
Finished figures with time series for section about model validation

@@
 \usetikzlibrary{calc}
+\usepackage{epstopdf}
 % Tikz styles
@@
 \end{itemize}
-A more pronounced $O_3$ loss takes place under lower $VOC$ concentrations because in this case more $OH$ radicals are left in the system after oxidation of $CO$ and alkanes leading to a greater concentration of $HO_2$ radicals, and therefore to a greater chance of the last reaction in the sequence.
+A more pronounced $O_3$ loss takes place under lower $VOC$ concentrations because in this case more $OH$ radicals are left in the system after oxidation of $CO$ and alkanes leading to a greater concentration of $HO_2$ radicals, and therefore to a greater recurrence of the last reaction in the sequence.
+
+%\begin{figure}[h] % ModelValidation
+%\includegraphics[width=0.5\linewidth]{{D:/FACSIMILE/ANsCBmodel/matlab/ANsCB_pics/ANsNOxVOC/noAN/chem_noAN_valid_50ppt5000ppt_O3_vs_time}.png}
+%%\caption{out.}\label{fig:ModelValidationO3}
+%\end{figure}
+%\begin{figure}[h] % ModelValidation
+%\includegraphics[width=0.5\linewidth]{{D:/FACSIMILE/ANsCBmodel/matlab/ANsCB_pics/ANsNOxVOC/noAN/chem_noAN_valid_50ppt5000ppt_O1D_vs_time}.png}
+%%\caption{out.}\label{fig:ModelValidationO1D}
+%\end{figure}
 
@@
 
-%\newpage
-%\clearpage
-%\thispagestyle{empty}% empty page style (?)
-%\begin{landscape}% Landscape page
-%\centering
-%	\input{diss_valid_vs_time}
-%\end{landscape}
+\newpage
+\clearpage
+\thispagestyle{empty}% empty page style (?)
+\input{diss_valid_vs_time}
 

### 2015-07-27 diss_valid_vs_time.tex
Figures with mixing ratio time series for section about model validation

\graphicspath{{d:/FACSIMILE/ANsCBmodel/matlab/ANsCB_pics/ANsNOxVOC/noAN/}}

\def \expid {chem_noAN_valid_50ppt5000ppt_}
\def\exppostfix {_vs_time}

%\renewcommand\thesubfigure{\alphalph{\value{subfigure}}}
\newlength{\subfigsize} \setlength{\subfigsize}{0.3\linewidth}
\newlength{\subfigwidth} \setlength{\subfigwidth}{\subfigsize}
%\newlength{\subfigheight} \setlength{\subfigheight}{2.5cm} %keepaspectratio=false,height=\subfigheight
% 0.3 makes 3x3 matrix

\begin{figure}[h!]
\centering

\begin{subfigure}[t]{\subfigsize}\includegraphics[width=\subfigwidth]{{\expid O3\exppostfix}.eps}\caption{$O_3$}\label{fig:O3valid}\end{subfigure}\hfill
\begin{subfigure}[t]{\subfigsize}\includegraphics[width=\subfigwidth]{{\expid O1D\exppostfix}.eps}\caption{$O(^1D)$}\label{fig:O1Dvalid}\end{subfigure}\hfill
\begin{subfigure}[t]{\subfigsize}\includegraphics[width=\subfigwidth]{{\expid OH\exppostfix}.eps}\caption{$OH$}\label{fig:OHvalid}\end{subfigure}\hfill
\begin{subfigure}[t]{\subfigsize}\includegraphics[width=\subfigwidth]{{\expid NO\exppostfix}.eps}\caption{$NO$}\label{fig:NOvalid}\end{subfigure}\hfill
\begin{subfigure}[t]{\subfigsize}\includegraphics[width=\subfigwidth]{{\expid NO2\exppostfix}.eps}\caption{$NO_2$}\label{fig:NO2valid}\end{subfigure}\hfill
\begin{subfigure}[t]{\subfigsize}\includegraphics[width=\subfigwidth]{{\expid HO2\exppostfix}.eps}\caption{$HO_2$}\label{fig:HO2valid}\end{subfigure}\hfill
\begin{subfigure}[t]{\subfigsize}\includegraphics[width=\subfigwidth]{{\expid H2O2\exppostfix}.eps}\caption{$H_2O_2$}\label{fig:H2O2valid}\end{subfigure}\hfill
\begin{subfigure}[t]{\subfigsize}\includegraphics[width=\subfigwidth]{{\expid RO2\exppostfix}.eps}\caption{$\Sigma RO_2$}\label{fig:RO2valid}\end{subfigure}\hfill
\begin{subfigure}[t]{\subfigsize}\includegraphics[width=\subfigwidth]{{\expid CO\exppostfix}.eps}\caption{$CO$}\label{fig:COvalid}\end{subfigure}\hfill

\caption{Changes in mixing ratios of $O_3$, $O(^1D)$, $OH$, $NO$, $NO_2$, $HO_2$, $H_2O_2$, total $RO_2$ and $CO$ with time (in ppb) under different $NO_x$ and $VOC$ conditions: 
$NO_x=50$ ppt and minimum sum of $VOCs$ (blue solid lines), 
$NO_x=50$ ppt and maximum sum of $VOCs$ (red solid lines), 
$NO_x=5$ ppb and minimum sum of $VOCs$ (blue dashed lines),
$NO_x=5$ ppb and maximum sum of $VOCs$ (red dashed lines), Minimum sum of $VOCs$ corresponds to case A in Table 5, maximum sum of $VOCs$ - case K. Simulation time 24 hours.}
\label{fig:res0}
\end{figure}

### 2015-07-27 diss_titlepage.tex
Thesis title page

\thispagestyle{empty}
\vskip 3cm
\begin{center}
    {\Huge\bfseries\sf \mytitle \par}
    \vskip 1cm
    \sf by
    \vskip 1cm
    \large \sf \authorname \\
    \sf \authornumber
\end{center}

\begin{center}
	\large \sf Thesis presented in part-fulfilment of the degree of \degname \\ in accordance with the regulations of the \univname
\end{center}

\vskip 5cm

\begin{minipage}{0.4\textwidth}
\begin{flushleft} 
\large \sf 
\depname \\
\univname \\
\univaddr
\vskip 0.5cm
\copyright~\mydate~\authorname
\end{flushleft}
\end{minipage}

%\vskip 0.2cm

\begin{minipage}{\textwidth}
\begin{flushleft}
%\noindent
\sf This copy of the dissertation has been supplied on condition that anyone who consults it is understood to recognise that its copyright rests with the author and that no quotation from the dissertation, nor any information derived there from, may be published without the author's prior written consent. Moreover, it is supplied on the understanding that it represents an internal University document and that neither the University nor the author are responsible for the factual or interpretative correctness of the dissertation.
\end{flushleft}
\end{minipage}

\vfill

\newpage

### 2015-07-28 MSc_diss.tex
Minor changes
Writing text about model validation

--- 2015-07-27 MSc_diss.tex
@@
 
-\input{MSc_diss_title}
+\input{diss_titlepage}
 
@@
 \subsection{Model validation}
-Time evolution of species concentrations in the model agrees reasonably well with photochemical theory as it can be seen, for example, from Figure ???.
-At low $NO_x$ levels (50 ppt) $O_3$ is being removed from the system because there is not enough $NO$ that could react with the peroxy radicals ($HO_2$ and $RO_2$) to produce $NO_2$ (which photolysis in conjunction with the reaction $O(^3P)+O_2\xrightarrow{M} O_3$ is the main source of $O_3$ in the troposphere).
+Due to imperfection of the chemical mechanism time evolution of species concentrations given by the model does not completely agree with the photochemical theory, but reproduces it to some extent.
+Figure \ref{fig:model_validation} shows an example of output from the model, namely the changes in mixing ratios of the main species within 24-hours of simulation time under contrasting $NO_x$ and $VOC$ conditions.
+The values shown in Figure \ref{fig:model_validation} have been justified from the theoretical point of view and also compared with those derived from the Atmospheric Chemistry Learner's Box Model (ACLBM, http://guille.zicer.uea.ac.uk/ACLBM).
+ACLBM is an on-line box model for atmospheric chemistry developed by Prof.
+Roland von Glasow (University of East Anglia) that allows the selection of several chemical reaction mechanisms, choice of initial concentrations and emission fluxes.
+For comparison with our modelling data the ACLBM was set up and initialized using values from Tables \ref{tab:setupO3CO}, \ref{tab:setupNOx} and \ref{tab:setupVOCs}, and the '$O_3$-$NO_x$-$CO$-$CH_4$' chemical mechanism was chosen since it is the closest version of the chemical mechanism available in the ACLBM to our model.
+
+In theory at low $NO_x$ levels a net $O_3$ removal takes place because there is not enough $NO$ around that could react with the peroxy radicals ($HO_2$ and $RO_2$) to produce $NO_2$, which photolysis in conjunction with the reaction (\ref{reac:O3P+O2}) is the main source of $O_3$ in the troposphere.
 Some additional $O_3$ removal occurs because $HO_2$ radicals can react with $O_3$ leading to the regeneration of $OH$ radicals as part of an $O_3$ depleting $OH$-$HO_2$ ($HO_x$) cycle.
-In other words, the following reaction sequence dominates:
-\begin{itemize}
-\item[] $O_3 + hv \rightarrow O(^1D) + O_2$
-\item[] $O(^1D) + H_2O \rightarrow OH + OH$
-\item[] $O(^1D) \xrightarrow{M} O(^3P)$
-\item[] $O(^3P) + O_2 \xrightarrow{M} O_3$
-\item[] $OH + CO \rightarrow HO_2 + CO_2$
-\item[] $OH + RH \rightarrow RO_2 + H_2O$
-\item[] $HO_2 + HO_2 \rightarrow H_2O_2 + O_2$
-\item[] $RO_2 + HO_2 \rightarrow ROOH + O_2$
-\item[] $O_3 + HO_2 \rightarrow OH + 2O_2$.
-\end{itemize}
-A more pronounced $O_3$ loss takes place under lower $VOC$ concentrations because in this case more $OH$ radicals are left in the system after oxidation of $CO$ and alkanes leading to a greater concentration of $HO_2$ radicals, and therefore to a greater chance of the last reaction in the sequence.
+In other words, the following reaction sequence prevails:
+\begin{equation}\label{reac:O3+hv}
+O_3 + hv \rightarrow O(^1D) + O_2
+\end{equation}
+\begin{equation}\label{reac:O1D+H2O}
+O(^1D) + H_2O \rightarrow OH + OH
+\end{equation}
+\begin{equation}\label{reac:O1D=O3P}
+O(^1D) \xrightarrow{M} O(^3P
+\end{equation}
+\begin{equation}\label{reac:O3P+O2}
+O(^3P) + O_2 \xrightarrow{M} O_3
+\end{equation}
+\begin{equation}\label{reac:OH+CO}
+OH + CO \rightarrow HO_2 + CO_2
+\end{equation}
+\begin{equation}\label{reac:OH+RH}
+OH + RH \rightarrow RO_2 + H_2O
+\end{equation}
+\begin{equation}\label{reac:HO2+HO2}
+HO_2 + HO_2 \rightarrow H_2O_2 + O_2
+\end{equation}
+\begin{equation}\label{reac:RO2+HO2}
+RO_2 + HO_2 \rightarrow ROOH + O_2
+\end{equation}
+\begin{equation}\label{reac:O3+HO2}
+O_3 + HO_2 \rightarrow OH + 2O_2
+\end{equation}
+In this case our modelling data agrees well with the theory (Figure \ref{fig:O3valid}), because $O_3$ mixing ratio decreases with time, and this decrease is comparable to one derived from ACLBM (by 6 ppb and 4.5 ppb at low and high $VOCs$ concentrations respectively).
+
+The concentration of $NO_x$ is the main factor which determines whether $O_3$ is produced or removed in the atmosphere \citep{Fowler2008}.
+Theoretically there are three regimes of $O_3$ production that depend on $NO_x$ levels: low $NO_x$ is characterized by net $O_3$ removal (as seen previously), intermediate $NO_x$ - by net $O_3$ formation, and high $NO_x$ - by net $O_3$ formation or removal depending on $VOCs$ concentration.
+Since the chemical mechanism of our model does not include reactions of nitric acid ($HNO_3$) we won't expect to see a good representation of high $NO_x$ regime in our model.
+Therefore, further we consider just a contrasting case.
+
+At higher $NO_x$ levels, reactions of the peroxy radicals with $NO$ provide a link required for catalytic generation of $O_3$ via
+\begin{equation}\label{reac:HO2+NO}
+HO_2 + NO \rightarrow OH + NO_2
+\end{equation}
+\begin{equation}\label{reac:RO2+NO}
+RO_2 + NO \rightarrow RO + NO_2
+\end{equation}
+\begin{equation}\label{eq:NO2+hv}
+NO_2 + hv \rightarrow NO + O(^3P)
+\end{equation}
+\begin{equation} %same
+!!!!! same O(^3P) + O_2 \xrightarrow{M} O_3
+\end{equation}
+The competition between reactions (\ref{reac:HO2+NO}) and (\ref{reac:HO2+HO2}) for $HO_2$, and reactions (\ref{reac:RO2+NO}) and (\ref{reac:RO2+HO2}) for $RO_2$, allows a greater number of free-radical propagated $O_3$ forming cycles to happen prior to termination, therefore resulting in a net $O_3$ formation (as it seen in Figure \ref{fig:O3valid}).
+It appears that in this case $O_3$ formation is somewhat sensitive to changes in the concentrations of $VOCs$ as well.
+
+Under the prevailing conditions, NOx levels remain sufficiently low that peroxide formation via...
+continues to be the major radical sink.
+Ozone formation in this regime is often referred to as 'NOx limited' or 'NOx sensitive', because the O3 formation rate increases with increasing NOx concentrations.
+
+In contrast, the O3 formation is insensitive to changes in the concentration of CH4 or CO and to inputs of other VOC.
+This is because OH reacts with the organics exclusively as part of the free-radical propagated O3 forming cycles, with no competing radical termination reaction being available under the prevailing conditions.
 

### 2015-07-28 MSc_diss.tex
Continued writing text about model validation

@@
 \end{table}
-\subsection{Model validation}
+\subsection{Model performance}
 Due to imperfection of the chemical mechanism time evolution of species concentrations given by the model does not completely agree with the photochemical theory, but reproduces it to some extent.
-Figure \ref{fig:model_validation} shows an example of output from the model, namely the changes in mixing ratios of the main species within 24-hours of simulation time under contrasting $NO_x$ and $VOC$ conditions.
+Figure \ref{fig:model_validation} shows an example of output from the model, namely the changes in mixing ratios of the main species within 24-hours of simulation time under contrasting $NO_x$ and $VOCs$ conditions.
 The values shown in Figure \ref{fig:model_validation} have been justified from the theoretical point of view and also compared with those derived from the Atmospheric Chemistry Learner's Box Model (ACLBM, http://guille.zicer.uea.ac.uk/ACLBM).
@@
 Roland von Glasow (University of East Anglia) that allows the selection of several chemical reaction mechanisms, choice of initial concentrations and emission fluxes.
-For comparison with our modelling data the ACLBM was set up and initialized using values from Tables \ref{tab:setupO3CO}, \ref{tab:setupNOx} and \ref{tab:setupVOCs}, and the '$O_3$-$NO_x$-$CO$-$CH_4$' chemical mechanism was chosen since it is the closest version of the chemical mechanism available in the ACLBM to our model.
+For comparison with our modelling data the ACLBM was set up and initialized with values from Tables \ref{tab:setupO3CO}, \ref{tab:setupNOx} and \ref{tab:setupVOCs}, and the '$O_3$-$NO_x$-$CO$-$CH_4$' chemical mechanism was chosen as it is the closest version of the chemical mechanism available in the ACLBM to our model.
 
-In theory at low $NO_x$ levels a net $O_3$ removal takes place because there is not enough $NO$ around that could react with the peroxy radicals ($HO_2$ and $RO_2$) to produce $NO_2$, which photolysis in conjunction with the reaction (\ref{reac:O3P+O2}) is the main source of $O_3$ in the troposphere.
-Some additional $O_3$ removal occurs because $HO_2$ radicals can react with $O_3$ leading to the regeneration of $OH$ radicals as part of an $O_3$ depleting $OH$-$HO_2$ ($HO_x$) cycle.
-In other words, the following reaction sequence prevails:
+The concentration of $NO_x$ is the main factor which determines whether $O_3$ is produced or removed in the atmosphere, and theoretically there are three regimes of $O_3$ production that depend on $NO_x$ levels: at low $NO_x$ levels the regime is characterized by net $O_3$ removal, at intermediate $NO_x$ (also called $NO_x$-sensitive) - by net $O_3$ formation, and at high $NO_x$ - by net $O_3$ formation or removal depending more on $VOCs$ concentration (therefore called $VOC$-sensitive) (e.g., \citep{Fowler2008}).
+When the prevailing $NO_x$ levels are low, a net $O_3$ removal takes place because there is not enough $NO$ around to react with the peroxy radicals ($HO_2$ and $RO_2$) to produce $NO_2$, which photolysis in conjunction with the reaction (\ref{reac:O3P+O2}) is the main source of $O_3$ in the troposphere.
+Some additional $O_3$ removal in this regime occurs due to reaction of $HO_2$ radicals with $O_3$ leading to the regeneration of $OH$ radicals as part of an $O_3$ depleting $OH$-$HO_2$ ($HO_x$) cycle.
+In other words, the following reaction sequence predominates:
 \begin{equation}\label{reac:O3+hv}
@@
 \end{equation}
-In this case our modelling data agrees well with the theory (Figure \ref{fig:O3valid}), because $O_3$ mixing ratio decreases with time, and this decrease is comparable to one derived from ACLBM (by 6 ppb and 4.5 ppb at low and high $VOCs$ concentrations respectively).
+At low $NO_x$ regime our modelling data agrees well with the theory (Figure \ref{fig:O3valid}).
+$O_3$ mixing ratio decreases with time, and this decrease is comparable to one derived from ACLBM (by 6 ppb and 4.5 ppb at low and high $VOCs$ concentrations respectively).
+$NO$ is rapidly converted to $NO_2$ representing a state equilibrium between the two molecules in the troposphere (Figures \ref{fig:NOvalid}, \ref{fig:NO2valid}).
+$CO$ is lost due to its reaction with $OH$ (Figure \ref{fig:COvalid}).
+Mixing ratios of other species in our model are of the order of magnitude as in ACLBM.
 
-The concentration of $NO_x$ is the main factor which determines whether $O_3$ is produced or removed in the atmosphere \citep{Fowler2008}.
-Theoretically there are three regimes of $O_3$ production that depend on $NO_x$ levels: low $NO_x$ is characterized by net $O_3$ removal (as seen previously), intermediate $NO_x$ - by net $O_3$ formation, and high $NO_x$ - by net $O_3$ formation or removal depending on $VOCs$ concentration.
-Since the chemical mechanism of our model does not include reactions of nitric acid ($HNO_3$) we won't expect to see a good representation of high $NO_x$ regime in our model.
-Therefore, further we consider just a contrasting case.
+\input{diss_valid_vs_time}
 
-At higher $NO_x$ levels, reactions of the peroxy radicals with $NO$ provide a link required for catalytic generation of $O_3$ via
+When the prevailing $NO_x$ levels are more than 0.5 ppb \citep{Sillman1999}, but less than certain critical value, reactions (\ref{reac:HO2+NO}) and (\ref{reac:RO2+NO}) represent the dominant reaction pathways for $HO_2$ and $RO_2$ providing a link for catalytic generation of $O_3$:
 \begin{equation}\label{reac:HO2+NO}
@@
 \end{equation}
-The competition between reactions (\ref{reac:HO2+NO}) and (\ref{reac:HO2+HO2}) for $HO_2$, and reactions (\ref{reac:RO2+NO}) and (\ref{reac:RO2+HO2}) for $RO_2$, allows a greater number of free-radical propagated $O_3$ forming cycles to happen prior to termination, therefore resulting in a net $O_3$ formation (as it seen in Figure \ref{fig:O3valid}).
+The competition between reactions (\ref{reac:HO2+NO}) and (\ref{reac:HO2+HO2}) for $HO_2$, and reactions (\ref{reac:RO2+NO}) and (\ref{reac:RO2+HO2}) for $RO_2$, allows a greater number of free-radical propagated $O_3$ forming cycles to happen prior to termination (reactions (\ref{reac:HO2+HO2}) and (\ref{reac:RO2+HO2})), therefore resulting in a net $O_3$ formation.
+
+When the prevailing $NO_x$ levels are high, reactions (\ref{reac:HO2+NO}) and (\ref{reac:RO2+NO}) dominate for $HO_2$ and $RO_2$, making the rate of $O_3$ production being controlled by the availability of these radicals.
+The availability of $OH$ radical is also important at high $NO_x$ levels because $OH$ connects the $O_3$ production rate with rates of its reactions with $CO$ and $RH$ (reactions (\ref{reac:OH+CO}) and (\ref{reac:OH+RH})), therefore making this regime $VOC$-sensitive.
+Plus at this point the reaction of $OH$ with $NO_2$ to form nitric acid ($HNO_3$) becomes the major termination reaction for free radicals:
+\begin{equation}\label{reac:OH+NO2}
+OH + NO_2 \xrightarrow{M} HNO_3
+\end{equation}
+Under these circumstances, the number of free-radical propagated $O_3$ forming cycles decreases with increasing $NO_x$, and the production rate of $O_3$ decreases.
+However, if the concentrations of $VOCs$ are high, reactions of $OH$ with $CO$ and $RH$ ((\ref{reac:OH+CO}), (\ref{reac:OH+RH})) can compete with reaction (\ref{reac:OH+NO2}), leading to an increase in the $O_3$ production rate.
+
+Since the chemical mechanism of our model does not include the reaction of $HNO_3$ formation, a good representation of $VOC$-sensitive regime by our model is not expected.
+It also makes it difficult to split $NO_x$-sensitive and $VOC$-sensitive regimes in our model, because this split is determined by the recurrence (size) of the peroxide- and nitric-acid forming reactions \citep{Sillman1999}, while our mechanism lacks one of these reactions.
+However, needing to evaluate performance of our model under higher $NO_x$ conditions anyway below we consider just a contrasting case of $NO_x$ levels.
+
+In our model a net $O_3$ formation first occurs at 250 ppt of $NO_x$ and remains so at 10 ppb of $NO_x$ too (Figure \ref{fig:noAN_netO3_rate}).
+
+\begin{figure}[h] % noAN_netO3rate_withINORG_ppbC
+\includegraphics[width=0.5\linewidth]{D:/FACSIMILE/ANsCBmodel/matlab/ANsCB_pics/ANsNOxVOC/noAN/chem_noAN_netO3rate_withINORG_ppbC.png}
+\caption{noAN, netO3 rate}\label{fig:noAN_netO3_rate}
+\end{figure}
+
+\begin{figure}[h] % Sillman1999
+\includegraphics[width=0.5\linewidth]{{./pics/Sillman1999_Fig1_adapted}.png}
+\caption{Isopleths giving net rate of ozone production (ppb/h, solid lines) as a function of $VOC$ (ppbC) and $NO_x$ (ppb) for mean summer daytime meteorology and clear skies.
+The solid lines represent production rates of 1, 2.5, 5, 10, 15, 20 and 30 ppb/h.
+Adapted from \citep{Sillman1999}.}\label{fig:Sillman1999}
+\end{figure}
+
 It appears that in this case $O_3$ formation is somewhat sensitive to changes in the concentrations of $VOCs$ as well.
-
-Under the prevailing conditions, NOx levels remain sufficiently low that peroxide formation via...
-continues to be the major radical sink.
-Ozone formation in this regime is often referred to as 'NOx limited' or 'NOx sensitive', because the O3 formation rate increases with increasing NOx concentrations.
 
@@
 This is because OH reacts with the organics exclusively as part of the free-radical propagated O3 forming cycles, with no competing radical termination reaction being available under the prevailing conditions.
-
-The concentration of NOx is the main factor which determines whether O3 is produced or removed in the atmosphere.
-changing balance of the free-radical reactions, and their impact on O3, as the level of NOx increases.
-
-NO and NO2 exist in a state equilibrium with rapid interconversion between the two molecules in the troposphere.
 
@@
 
-\newpage
-\clearpage
-\thispagestyle{empty}% empty page style (?)
-\input{diss_valid_vs_time}
-
 \section{Results} \label{sec:res}
-\begin{figure}[h] % Sillman1999
-\includegraphics[width=0.5\linewidth]{{./pics/Sillman1999_Fig1_adapted}.png}
-\caption{Isopleths giving net rate of ozone production (ppb/h, solid lines) as a function of $VOC$ (ppbC) and $NO_x$ (ppb) for mean summer daytime meteorology and clear skies.
-The solid lines represent production rates of 1, 2.5, 5, 10, 15, 20 and 30 ppb/h.
-Adapted from \citep{Sillman1999}.}\label{fig:Sillman1999}
-\end{figure}
 \begin{figure}[h] % allAN_netO3_rate

### 2015-07-28 diss_valid_vs_time.tex
Minor changes
Changes in caption

@@
 
-\caption{Changes in mixing ratios of $O_3$, $O(^1D)$, $OH$, $NO$, $NO_2$, $HO_2$, $H_2O_2$, total $RO_2$ and $CO$ with time (in ppb) under different $NO_x$ and $VOC$ conditions: 
+\caption{Changes in mixing ratios of $O_3$, $O(^1D)$, $OH$, $NO$, $NO_2$, $HO_2$, $H_2O_2$, total $RO_2$ and $CO$ with time (in ppb) under different $NO_x$ and $VOC$ conditions with alkyl nitrate chemistry switched off:
 $NO_x=50$ ppt and minimum sum of $VOCs$ (blue solid lines), 
@@
 $NO_x=5$ ppb and maximum sum of $VOCs$ (red dashed lines), Minimum sum of $VOCs$ corresponds to case A in Table 5, maximum sum of $VOCs$ - case K. Simulation time 24 hours.}
-\label{fig:res0}
+\label{fig:model_validation}
 \end{figure}

### 2015-07-28 O3CycleScheme.tex
Minor changes

@@
-\documentclass[]{standalone}
-\usepackage[utf8]{inputenc}
-\usepackage[english]{babel}
-\usepackage{amsmath}
-\usepackage{amsfonts}
-\usepackage{amssymb}
+%\documentclass[]{standalone}
+%\usepackage[utf8]{inputenc}
+%\usepackage[english]{babel}
+%\usepackage{amsmath}
+%\usepackage{amsfonts}
+%\usepackage{amssymb}
+%
+%\usepackage{graphicx}
+%%\usepackage[left=2cm,right=2cm,top=2cm,bottom=2cm]{geometry}
+%
+%% Figures
+%\usepackage{tikz}
+%\usetikzlibrary{shapes,arrows}
+%\usetikzlibrary{positioning}
+%\usetikzlibrary{calc}
+%%\usepackage{chemfig}
+%
+%% Tikz styles
+%\tikzstyle{reg} = [rounded rectangle, text width=2cm, fill=black!5, align=center, anchor=west,font=\sffamily\Large\bfseries, minimum height=1cm, inner sep=0]
+%\tikzstyle{rad} = [rounded rectangle, draw=none, text width=2cm, fill=red!10, align=center, anchor=west,font=\sffamily\Large\bfseries, minimum height=1cm, inner sep=0]
+%\tikzstyle{long} = [rounded rectangle, draw=none, text width=3.5cm, fill=black!5, align=center, anchor=west,font=\sffamily\Large\bfseries, minimum height=1cm, inner sep=0]   
+%\tikzstyle{add} = [scale=1.5, draw=none,fill=none,align=center]
+%\tikzstyle{emp} = [draw=none,fill=none]
+%\tikzstyle{ar} = [->,draw=black!70,line width=2]
+%\tikzstyle{ln} = [-,draw=black!70,line width=2]
 
-\usepackage{graphicx}
-%\usepackage[left=2cm,right=2cm,top=2cm,bottom=2cm]{geometry}
-
-% Figures
-\usepackage{tikz}
-\usetikzlibrary{shapes,arrows}
-\usetikzlibrary{positioning}
-\usetikzlibrary{calc}
-%\usepackage{chemfig}
-
-% Tikz styles
-\tikzstyle{reg} = [rounded rectangle, text width=2cm, fill=black!5, align=center, anchor=west,font=\sffamily\Large\bfseries, minimum height=1cm, inner sep=0]
-\tikzstyle{rad} = [rounded rectangle, draw=none, text width=2cm, fill=red!10, align=center, anchor=west,font=\sffamily\Large\bfseries, minimum height=1cm, inner sep=0]
-\tikzstyle{long} = [rounded rectangle, draw=none, text width=3.5cm, fill=black!5, align=center, anchor=west,font=\sffamily\Large\bfseries, minimum height=1cm, inner sep=0]   
-\tikzstyle{add} = [scale=1.5, draw=none,fill=none,align=center]
-\tikzstyle{emp} = [draw=none,fill=none]
-\tikzstyle{ar} = [->,draw=black!70,line width=2]
-\tikzstyle{ln} = [-,draw=black!70,line width=2]
-
-\begin{document}
+%\begin{document}
 \begin{tikzpicture}[node distance = 4cm, auto]
@@
 
-\end{document}
+%\end{document}

### 2015-07-29 MSc_diss.tex
Finished section about model performance

@@
 \end{equation}
-At low $NO_x$ regime our modelling data agrees well with the theory (Figure \ref{fig:O3valid}).
+At low $NO_x$ regime (for example, 50 ppt) our modelling data agrees well with the theory (Figure \ref{fig:O3valid}).
 $O_3$ mixing ratio decreases with time, and this decrease is comparable to one derived from ACLBM (by 6 ppb and 4.5 ppb at low and high $VOCs$ concentrations respectively).
-$NO$ is rapidly converted to $NO_2$ representing a state equilibrium between the two molecules in the troposphere (Figures \ref{fig:NOvalid}, \ref{fig:NO2valid}).
+$NO$ is rapidly (almost instantly) converted to $NO_2$ representing a state equilibrium between the two molecules in the troposphere (Figures \ref{fig:NOvalid}, \ref{fig:NO2valid}).
 $CO$ is lost due to its reaction with $OH$ (Figure \ref{fig:COvalid}).
@@
 
-When the prevailing $NO_x$ levels are more than 0.5 ppb \citep{Sillman1999}, but less than certain critical value, reactions (\ref{reac:HO2+NO}) and (\ref{reac:RO2+NO}) represent the dominant reaction pathways for $HO_2$ and $RO_2$ providing a link for catalytic generation of $O_3$:
+When the prevailing $NO_x$ levels are more than 0.5 ppb \citep{Sillman1999}, but less than a certain critical value, reactions (\ref{reac:HO2+NO}) and (\ref{reac:RO2+NO}) represent the dominant reaction pathways for $HO_2$ and $RO_2$ providing a link for catalytic generation of $O_3$:
 \begin{equation}\label{reac:HO2+NO}
@@
 It also makes it difficult to split $NO_x$-sensitive and $VOC$-sensitive regimes in our model, because this split is determined by the recurrence (size) of the peroxide- and nitric-acid forming reactions \citep{Sillman1999}, while our mechanism lacks one of these reactions.
-However, needing to evaluate performance of our model under higher $NO_x$ conditions anyway below we consider just a contrasting case of $NO_x$ levels.
+However, needing to evaluate performance of our model under higher $NO_x$ conditions anyway below we consider just a contrasting case of $NO_x$ levels, for example, 5 ppb (Figure \ref{fig:model_validation}).
 
-In our model a net $O_3$ formation first occurs at 250 ppt of $NO_x$ and remains so at 10 ppb of $NO_x$ too (Figure \ref{fig:noAN_netO3_rate}).
+At higher $NO_x$ regime (5 ppb) $O_3$ mixing ratio in the model increases with time (Figure \ref{fig:O3valid}) which agrees with the photochemical theory.
+The net increase is by 186 ppb and 171 ppb at low and high $VOCs$ concentrations respectively.
+$NO$ is rapidly converted to $NO_2$, but slower than at low $NO_x$ levels because of their higher initial concentrations (Figure \ref{fig:NOvalid}, \ref{fig:NO2valid}).
+Due to a higher efficiency of reaction (\ref{reac:HO2+NO}) $OH$ mixing ratio is higher at higher $NO_x$ levels, resulting in a more pronounced loss of $CO$ (Figure \ref{fig:COvalid}).
 
-\begin{figure}[h] % noAN_netO3rate_withINORG_ppbC
-\includegraphics[width=0.5\linewidth]{D:/FACSIMILE/ANsCBmodel/matlab/ANsCB_pics/ANsNOxVOC/noAN/chem_noAN_netO3rate_withINORG_ppbC.png}
-\caption{noAN, netO3 rate}\label{fig:noAN_netO3_rate}
+The best way to illustrate the central features of the relationship between $O_3$, $NO_x$ and $VOC$ mentioned above is to construct a ozone isopleth plot in a form shown in Figure \ref{fig:Sillman1999}.
+The isopleth plot shows that the rate of $O_3$ production is a nonlinear function of the $NO_x$ and $VOC$ concentrations.
+When $VOC$/$NO_x$ ratios are high, the $O_3$ production rate increases with increasing $NO_x$ eventually reaching a local maximum, and decreasing afterwards as $NO_x$ concentrations continue to increase.
+The local maximum (the 'ridge line') defines the split into $NO_x$-sensitive and $VOC$-sensitive regimes.
+Under the 'ridge line' the $VOC$/$NO_x$ ratios are high, and the $O_3$ production rate increases with increasing $NO_x$ and almost insensitive to $VOC$.
+Above the 'ridge line' the $VOC$/$NO_x$ ratios are low, and the $O_3$ production rate increases with increasing $VOC$ and decreases with increasing $NO_x$ \citep{Sillman2013}.
+
+\begin{figure} % Sillman1999withdotsline and netO3rate
+\centering
+\begin{minipage}{.5\textwidth}
+  \centering
+  \includegraphics[width=.99\linewidth]{{./pics/Sillman1999_Fig1_adapted_withdotssquare}.png}
+  \captionof{figure}{Isopleths giving net rate of $O_3$ production (ppb/h, solid lines) as a function of $VOC$ (ppbC) and $NO_x$ (ppb) for mean summer daytime meteorology and clear skies.
+The solid lines represent production rates of 1, 2.5, 5, 10, 15, 20 and 30 ppb/h.
+Adapted from \citep{Sillman1999}.}\label{fig:Sillman1999}
+\end{minipage}%
+\begin{minipage}{.5\textwidth}
+  \centering
+  \includegraphics[width=.99\linewidth]{D:/FACSIMILE/ANsCBmodel/matlab/ANsCB_pics/ANsNOxVOC/noAN/chem_noAN_netO3rate_withINORG_ppbC.png}
+  \captionof{figure}{Isopleths giving net rate of $O_3$ production (ppb/h, color) as a function of $VOC$ (ppbC) and $NO_x$ (ppb) in the first modelling experiment with alkyl nitrate chemistry switched off combined with an additional run with inorganic chemistry only (i.e.
+the only $VOC$ is $CO$).}\label{fig:noAN_netO3_rate}
+\end{minipage}
 \end{figure}
 
-\begin{figure}[h] % Sillman1999
-\includegraphics[width=0.5\linewidth]{{./pics/Sillman1999_Fig1_adapted}.png}
-\caption{Isopleths giving net rate of ozone production (ppb/h, solid lines) as a function of $VOC$ (ppbC) and $NO_x$ (ppb) for mean summer daytime meteorology and clear skies.
-The solid lines represent production rates of 1, 2.5, 5, 10, 15, 20 and 30 ppb/h.
-Adapted from \citep{Sillman1999}.}\label{fig:Sillman1999}
-\end{figure}
+The same kind of plot can be found in Figure \ref{fig:noAN_netO3_rate}, which is produced using data from our model from the first experiment with alkyl nitrate chemistry switched off.
+Taking into account the fact that the maximum $NO_x$ mixing ratio considered in the experiment is 10 ppb, our results are directly comparable only with a part of the data from Figure \ref{fig:Sillman1999}, namely with those inside the green dashed square.
+Moreover, since the minimum $NO_x$ mixing ratio considered in the experiment is 5 ppt, while the corresponding value in Figure \ref{fig:Sillman1999} is 1 ppb, a negative $O_3$ production rate is seen in our results, whereas it stays out of range of Figure \ref{fig:Sillman1999}.
 
-It appears that in this case $O_3$ formation is somewhat sensitive to changes in the concentrations of $VOCs$ as well.
-
-In contrast, the O3 formation is insensitive to changes in the concentration of CH4 or CO and to inputs of other VOC.
-This is because OH reacts with the organics exclusively as part of the free-radical propagated O3 forming cycles, with no competing radical termination reaction being available under the prevailing conditions.
-
-!! from \citep{Finlayson-Pitts2000}
-Urban:
-Certain reproducible features of time-concentration profiles for  pollutants are observed in "smoggy" ambient air.
-Figure 1.3, a classic example of historical interest, shows such profiles for NO, NO2, and total oxidant (mainly O3) in Pasadena, California, during a  severe photochemical air pollution episode in July 1973.
-Reproducible  features  include  the  following:
-\begin{itemize}
-\item In the early morning, the concentration of NO rises and  reaches a maximum at a time that approximately coincides with the  maximum emissions of NO, in this case, peak  automobile  traffic; 
-\item Subsequently, NO2 rises to a maximum;
-\item Oxidant (e.g., O3) levels, which are relatively low in the  early morning, increase significantly about noon when the NO concentration drops to a low value.
-\end{itemize}
+One can see the same pattern in these two figures.
+At high $VOC$/$NO_x$ ratios, the rate of $O_3$ production increases with increasing $NO_x$ and reaches its maximum at the highest $NO_x$ and $VOC$ levels (which corresponds to the top right corner of the green dashed square).
+The change in $O_3$ production rate at high $VOC$/$NO_x$ ratios in our model is largely insensitive to $VOC$, which fits into the theory.
+At low $VOC$/$NO_x$ ratios the $O_3$ production rate starts to depend on $VOC$ concentration also resembling the behaviour of ozone isopleths at the top left corner of the green dashed square.
+The overall tendency of $O_3$ production rate to increase with increasing $VOC$ and $NO_x$ is present in both figures showing that the model reproduces this feature according to the theory.
+However, the modelled values of $O_3$ production rates are much lower than those in Figure \ref{fig:Sillman1999}: in the model the maximum production rate is 9 ppb/h, while in theory it is 30 ppb/h, meaning that our model underestimates the $O_3$ production rate, but reproduces some features of the relationship between $O_3$, $NO_x$ and $VOC$.
 

### 2015-07-30 MSc_diss.tex
New figure in section about model validation

@@
 Figure \ref{fig:model_validation} shows an example of output from the model, namely the changes in mixing ratios of the main species within 24-hours of simulation time under contrasting $NO_x$ and $VOCs$ conditions.
-The values shown in Figure \ref{fig:model_validation} have been justified from the theoretical point of view and also compared with those derived from the Atmospheric Chemistry Learner's Box Model (ACLBM, http://guille.zicer.uea.ac.uk/ACLBM).
+The values shown in Figure \ref{fig:model_validation} have been justified from the theoretical point of view and also compared with those derived from the Atmospheric Chemistry Learner's Box Model (ACLBM, http://guille.uea.ac.uk/ACLBM/).
 ACLBM is an on-line box model for atmospheric chemistry developed by Prof.
@@
 
-\begin{figure} % Sillman1999withdotsline and netO3rate
+\begin{figure} % Sillman1999withdotsline and netO3rate_withINORG_NOx_1ppb_100ppb
 \centering
@@
   \centering
-  \includegraphics[width=.99\linewidth]{D:/FACSIMILE/ANsCBmodel/matlab/ANsCB_pics/ANsNOxVOC/noAN/chem_noAN_netO3rate_withINORG_ppbC.png}
+  \includegraphics[width=.99\linewidth]{D:/FACSIMILE/ANsCBmodel/matlab/ANsCB_pics/ANsNOxVOC/noAN/chem_noAN_netO3rate_withINORG_NOx_1ppb_100ppb.png}
   \captionof{figure}{Isopleths giving net rate of $O_3$ production (ppb/h, color) as a function of $VOC$ (ppbC) and $NO_x$ (ppb) in the first modelling experiment with alkyl nitrate chemistry switched off combined with an additional run with inorganic chemistry only (i.e.
@@
 \end{figure}
+
+Rewrite!
 
@@
 \section{Results} \label{sec:res}
-\begin{figure}[h] % allAN_netO3_rate
-\includegraphics[width=\linewidth]{D:/FACSIMILE/ANsCBmodel/matlab/ANsCB_pics/ANsNOxVOC/allAN/chem_allAN_netO3_rate.png}
-\caption{allAN, netO3 rate}\label{fig:allAN_netO3_rate}
-\end{figure}
+
+%\begin{figure}[h] % allAN_netO3_rate
+%\includegraphics[width=\linewidth]{D:/FACSIMILE/ANsCBmodel/matlab/ANsCB_pics/ANsNOxVOC/allAN/chem_allAN_netO3_rate.png}
+%\caption{allAN, netO3 rate}\label{fig:allAN_netO3_rate}
+%\end{figure}
 

### 2015-07-30 MSc_diss.tex
Minor changes

@@
   \centering
-  \includegraphics[width=.99\linewidth]{{./pics/Sillman1999_Fig1_adapted_withdotssquare}.png}
+  \includegraphics[width=.99\linewidth]{{./pics/Sillman1999_Fig1_adapted_withdotsrectangle}.png}
   \captionof{figure}{Isopleths giving net rate of $O_3$ production (ppb/h, solid lines) as a function of $VOC$ (ppbC) and $NO_x$ (ppb) for mean summer daytime meteorology and clear skies.
@@
   \captionof{figure}{Isopleths giving net rate of $O_3$ production (ppb/h, color) as a function of $VOC$ (ppbC) and $NO_x$ (ppb) in the first modelling experiment with alkyl nitrate chemistry switched off combined with an additional run with inorganic chemistry only (i.e.
-the only $VOC$ is $CO$).}\label{fig:noAN_netO3_rate}
+the only $VOC$ is $CO$).}\label{fig:noAN_netO3rate}
 \end{minipage}
@@
 
-The same kind of plot can be found in Figure \ref{fig:noAN_netO3_rate}, which is produced using data from our model from the first experiment with alkyl nitrate chemistry switched off.
+The same kind of plot can be found in Figure \ref{fig:noAN_netO3rate}, which is produced using data from our model from the first experiment with alkyl nitrate chemistry switched off.
 Taking into account the fact that the maximum $NO_x$ mixing ratio considered in the experiment is 10 ppb, our results are directly comparable only with a part of the data from Figure \ref{fig:Sillman1999}, namely with those inside the green dashed square.
@@
 \section{Results} \label{sec:res}
+\subsection{Sensitivity of $O_3$ production to alkyl nitrate chemistry}
+
+A number of simulations for evaluating the sensitivity of the $O_3$ production to alkyl nitrate chemistry were carried out.
+This was accomplished by switching on and off all reactions present in the chemical mechanism that describe formation or destruction of all alkyl nitrates.
+
+\begin{figure}
+\centering
+\begin{minipage}{.3\textwidth}
+  \centering
+  \includegraphics[width=.99\linewidth]{D:/FACSIMILE/ANsCBmodel/matlab/ANsCB_pics/ANsNOxVOC/noAN/chem_noAN_netO3rate_noINORG_NOx_5ppt_100ppb.png}
+\end{minipage}
+\begin{minipage}{.3\textwidth}
+  \centering
+  \includegraphics[width=.99\linewidth]{D:/FACSIMILE/ANsCBmodel/matlab/ANsCB_pics/ANsNOxVOC/allAN/chem_allAN_netO3rate_noINORG_NOx_5ppt_100ppb.png}
+\end{minipage}
+\begin{minipage}{.3\textwidth}
+  \centering
+  \includegraphics[width=.99\linewidth]{D:/FACSIMILE/ANsCBmodel/matlab/ANsCB_pics/ANsNOxVOC/allAN/chem_allAN_netO3ratediff_noINORG_NOx_5ppt_100ppb.png}
+\end{minipage}
+\caption{Isopleths giving net rate of $O_3$ production (ppb/h, color) as a function of $VOC$ (ppbC) and $NO_x$ (ppb) with alkyl nitrate chemistry switched off (left) and switched on (middle).
+Difference between them is shown on the right.}
+\end{figure}
+
+The simulations that compared ... provided insight into which reactions... under changing $NO_x$ mixing ratios (Table ...).
 

### 2015-07-31 MSc_diss.tex
Updates in results, rewriting literature review

@@
 \subsection{Alkyl nitrates}
+\subsubsection{Alkyl nitrates chemistry}
 Alkyl nitrates are important tropospheric trace gases which are formed following the same chemistry that leads to the production of ozone \citep{Reeves2007}.
@@
 
+\subsubsection{Observational studies}
+
+mean ANs concentrations: \citep{Reeves2007}, \citep{Roberts1998}
+
+\subsubsection{Modelling studies}
+
 To sum up, alkyl nitrates are produced as a minor branch of the reaction of peroxy radicals with NO \citep{Day2003}.
@@
 Despite their relative importance alkyl nitrates are usually not included in the climate-chemistry models, for example in UKCA (Reeves, personal communication), which potentially could lead to discrepancies between models and observations described before.
+
+[Source characterization and photochemical processes in
+transported air masses have been analyzed by examining
+RONO2 formation from anthropogenic emissions of alkane
+precursors [Simpson et al., 2003; Madronich, 2006; Simpson
+et al., 2006; Reeves et al., 2007; Worton et al., 2010]. (Neuman2012)]
 
@@
 
-A number of simulations for evaluating the sensitivity of the $O_3$ production to alkyl nitrate chemistry were carried out.
-This was accomplished by switching on and off all reactions present in the chemical mechanism that describe formation or destruction of all alkyl nitrates.
+To evaluate the sensitivity of $O_3$ production (efficiency) to alkyl nitrate ($RONO_2$) chemistry under different $NO_x$ and $VOC$ conditions a number of simulations were carried out.
+In these simulations the chemical mechanism of the model was adjusted to either include or exclude the formation and loss of alkyl nitrates, which was accomplished by switching on and off all reactions that involve $RONO_2$ while keeping the rest of the mechanism unchanged.
 
-\begin{figure}
+As it has been noted in Section \ref{sec:intro}, the relationship between $O_3$ and $RONO_2$ production rates is controlled by the branching ratio ($\alpha_3$) between reactions:
+\begin{equation}\label{reac:RO2+NO=RO+NO2}
+RO_2 + NO \xrightarrow{1-\alpha_3} RO + NO_2
+\end{equation}
+\begin{equation}\label{reac:RO2+NO=RONO2}
+RO_2 + NO \xrightarrow{\alpha_3} RONO_2
+\end{equation}
+Reaction (\ref{reac:RO2+NO=RO+NO2}) leads to $O_3$ production and $NO$ regeneration, resulting in continuing $O_3$ production, while reaction (\ref{reac:RO2+NO=RONO2}) terminates this process.
+Branching ratios used in the model are taken from the MCM v3.3, and their values (Table \ref{tab:ANbranching}) were kept constant in all model runs.
+
+\begin{table} %  AN branching ratios
+\caption{Branching ratios for the formation of alkyl nitrates from their precursor peroxy radicals and $NO$ (taken from the MCM v3.3 (2015)).}\label{tab:ANbranching}
+\centering
+\begin{tabular}{cc}
+\hline
+Alkyl nitrate    & Branching ratio \\
+\hline
+methyl           & 0.001 \\
+ethyl            & 0.009 \\
+iso-propyl       & 0.042 \\
+n-propyl         & 0.020 \\
+n-butyl          & 0.033 \\
+sec-butyl        & 0.090 \\
+iso-butyl        & 0.033 \\ 
+tert-butyl       & 0.025 \\
+n-pentyl         & 0.052 \\
+2-pentyl         & 0.129 \\
+3-pentyl         & 0.131 \\
+2-methylbutyl    & 0.052 \\
+3-methyl-2-butyl & 0.141 \\
+2-methyl-2-butyl & 0.047 \\
+\hline
+\end{tabular}
+\end{table}
+
+Figure \ref{fig:netO3rate_noAN_withAN_diff} shows isopleths of net $O_3$ production rate as a function of $NO_x$ and sum of $VOCs$ initial concentrations from two series of model runs (with and without alkyl chemistry) and the difference between them.
+Since the actual formation of $O_3$ in our model starts at 250 ppt of $NO_x$, Figure \ref{fig:netO3rate_noAN_withAN_diff} summarises modelling data from experiments with $NO_x$ initial mixing ratios from 250 ppt till 100 ppb.
+
+\begin{figure} % netO3rate noAN withAN diff
 \centering
@@
   \centering
-  \includegraphics[width=.99\linewidth]{D:/FACSIMILE/ANsCBmodel/matlab/ANsCB_pics/ANsNOxVOC/noAN/chem_noAN_netO3rate_noINORG_NOx_5ppt_100ppb.png}
+  \includegraphics[width=.99\linewidth]{D:/FACSIMILE/ANsCBmodel/matlab/ANsCB_pics/ANsNOxVOC/noAN/chem_noAN_netO3rate_noINORG_NOx_250ppt_100ppb.png}
 \end{minipage}
@@
   \centering
-  \includegraphics[width=.99\linewidth]{D:/FACSIMILE/ANsCBmodel/matlab/ANsCB_pics/ANsNOxVOC/allAN/chem_allAN_netO3rate_noINORG_NOx_5ppt_100ppb.png}
+  \includegraphics[width=.99\linewidth]{D:/FACSIMILE/ANsCBmodel/matlab/ANsCB_pics/ANsNOxVOC/allAN/chem_allAN_netO3rate_noINORG_NOx_250ppt_100ppb.png}
 \end{minipage}
@@
   \centering
-  \includegraphics[width=.99\linewidth]{D:/FACSIMILE/ANsCBmodel/matlab/ANsCB_pics/ANsNOxVOC/allAN/chem_allAN_netO3ratediff_noINORG_NOx_5ppt_100ppb.png}
+  \includegraphics[width=.99\linewidth]{D:/FACSIMILE/ANsCBmodel/matlab/ANsCB_pics/ANsNOxVOC/allAN/chem_allAN_netO3ratediff_noINORG_NOx_250ppt_100ppb.png}
 \end{minipage}
 \caption{Isopleths giving net rate of $O_3$ production (ppb/h, color) as a function of $VOC$ (ppbC) and $NO_x$ (ppb) with alkyl nitrate chemistry switched off (left) and switched on (middle).
-Difference between them is shown on the right.}
+Difference between them is shown on the right.}\label{fig:netO3rate_noAN_withAN_diff}
+\end{figure}
+
+As expected, the inclusion of alkyl nitrate chemistry slows down the net $O_3$ production rate.
+At low $NO_x$ and $VOCs$ initial levels, this slowdown is -0.2...-0.6 ppb of $O_3$ per hour.
+As $NO_x$ and $VOCs$ levels increase the $O_3$ production rate decreases, reaching the maximum decrease of -1.94 ppb of $O_3$ per hour at 100 ppb of $NO_x$ and $VOCs$ level from case I (not the highest $VOCs$ level).
+At the range of $NO_x$ levels from 1 ppb to 10 ppb the $O_3$ production rate is slightly more sensitive to $VOCs$ levels, while it is almost insensitive to $VOCs$ at $NO_x$ levels higher than 10 ppb and 2050 ppbC of $VOCs$.
+
+\begin{figure} % netO3mixrat noAN withAN diff
+\centering
+\begin{minipage}{.3\textwidth}
+  \centering
+  \includegraphics[width=.99\linewidth]{D:/FACSIMILE/ANsCBmodel/matlab/ANsCB_pics/ANsNOxVOC/noAN/chem_noAN_netO3mixrat_noINORG_NOx_250ppt_100ppb.png}
+\end{minipage}
+\begin{minipage}{.3\textwidth}
+  \centering
+  \includegraphics[width=.99\linewidth]{D:/FACSIMILE/ANsCBmodel/matlab/ANsCB_pics/ANsNOxVOC/allAN/chem_allAN_netO3mixrat_noINORG_NOx_250ppt_100ppb.png}
+\end{minipage}
+\begin{minipage}{.3\textwidth}
+  \centering
+  \includegraphics[width=.99\linewidth]{D:/FACSIMILE/ANsCBmodel/matlab/ANsCB_pics/ANsNOxVOC/allAN/chem_allAN_netO3mixratdiff_noINORG_NOx_250ppt_100ppb.png}
+\end{minipage}
+\caption{Isopleths giving net $O_3$ production (ppb, color) as a function of $VOC$ (ppbC) and $NO_x$ (ppb) with alkyl nitrate chemistry switched off (left) and switched on (middle).
+Difference between them is shown on the right.
+Net production = final - initial value of $O_3$ mixing ratio.}\label{fig:netO3mixrat_noAN_withAN_diff}
+\end{figure}
+
+The change in net rate of $O_3$ production due to inclusion of alkyl nitrate chemistry into the chemical mechanism has a direct impact on absolute $O_3$ levels calculated by the model.
+Figure \ref{fig:netO3mixrat_noAN_withAN_diff} shows that the net $O_3$ production (difference between final and initial $O_3$ mixing ratio) is smaller when alkyl nitrate chemistry is switched on.
+At low $NO_x$ and $VOCs$ levels the (negative) change in the net $O_3$ production is 5-10 ppb, at high $NO_x$ and $VOCs$ levels it is 35-45 ppb (per 24 hours) with maximum value of 46.5 ppb.
+However, these changes in the net $O_3$ production should be treated with care, because the model due to reasons described in Section \ref{sec:method} gives unrealistically high $O_3$ mixing ratios (maximum 650 ppb while typical maximum urban $O_3$ mixing ratio (Los Angeles, Mexico City) is 490 ppb (http://www-personal.umich.edu/~sillman/ozone.htm).
+Therefore, in this case a more useful estimate of effect of these changes is to express them in percentage form (i.e., compare like with like), and it appears that these changes are very small: ???
+respectively.
+
+To see what caused these changes it is useful to look at the net production of the sum of the alkyl nitrates themselves as a function of $NO_x$ and $VOCs$ levels (Figure \ref{fig:endtotalANs}).
+Since alkyl nitrates are the only sink for $NO_x$ in our model the high production of these species in total was expected.
+The range is ppb.
+However, real concentrations reported from observational studies are of different order, tens of ppt.
+\citep{Reeves2007} 10 ppt (ITOP) \citep{Roberts1998} 14 ppt
+
+\citep{Reeves2007}
+[In the 'marine' air masses methyl nitrate shows a slightly higher mean concentration to 'other' air masses (2.40 compared to 2.29 ppt), but the concentrations of the other alkyl
+nitrates are lower by around 30–50% with the exception of
+ethyl nitrate which is lower by about 15% (1.33 compared to
+1.57 ppt), such that it becomes the second most abundant alkyl
+nitrate.
+The highest alkyl
+nitrate concentration was observed during flight B031 (air
+from New York) for 2-methyl-3-butyl nitrate (13.7 ppt),
+when the maximum concentration of 2 + 3-pentyl nitrate
+(9.0 ppt) was also measured. It is quite possible that some of
+the ‘‘marine’’ maximum concentrations of the alkyl nitrates
+were impacted by terrestrial emissions (e.g., 5.4 pptv of
+2-methyl-3-butyl nitrate and 1.7 pptv for 2 + 3-pentyl nitrate
+observed during flight B030).
+
+The mean NO concentration observed during ITOP (20 plus minus 31 pptv (Lewis et al., 2007)) is below the limit of 100–200 pptv at which the efficiency of the alkyl nitrate formation is reduced because of competition between the reaction of ROO with NO and ROO self/cross reactions (Roberts et al., 1998)]
+
+\begin{figure} % endtotalANs
+\includegraphics[width=.5\linewidth]{D:/FACSIMILE/ANsCBmodel/matlab/ANsCB_pics/ANsNOxVOC/allAN/chem_allAN_endtotalANs_noINORG_NOx_250ppt_100ppb.png}
+\caption{Isopleths of net production of alkyl nitrates ($\Sigma RONO_2$) as a function of $NO_x$ and $VOCs$ levels.}\label{fig:endtotalANs}
 \end{figure}
@@
 
-%\begin{figure}[h] % allAN_netO3_rate
-%\includegraphics[width=\linewidth]{D:/FACSIMILE/ANsCBmodel/matlab/ANsCB_pics/ANsNOxVOC/allAN/chem_allAN_netO3_rate.png}
-%\caption{allAN, netO3 rate}\label{fig:allAN_netO3_rate}
-%\end{figure}
-
 !
@@
 
+\subsection{$O_3$ production and carbon bonds}
+
 \section{Discussion} \label{sec:discuss}
+Future plans and possible improvements in the model.
+
 The experimental results presented in the previous section provide an opportunity to speculate about the behaviour of our simple model of alkyl nitrate production and loss with time in comparison with the observational data.

### 2015-07-31 MSc_diss.tex
Correcting the structure

@@
 Table \ref{tab:O3budget} demonstrates that the range of the global tropospheric ozone budget estimates is approximately 60 Tg which is 18\% of the mean.
-In terms of ozone production the variation is even higher and account for 28\% of the corresponding value.
+In terms of ozone production the variation is even higher and accounts for 28\% of the corresponding value.
 In that regard, two questions arise: why does the tropospheric ozone budget differ between models and observations?
@@
 
+! add something from report!
+
 The second way to understanding the discrepancies between models and observations is to conduct an idealised theoretical study using a conceptual model.
 The advantage of this approach is that such kind of models can be quite easily constructed, and more importantly can be focused on representation of a certain chemical reactions which presumably could explain the discrepancy between models and observations.
-An example of such set of reactions is alkyl nitrate chemistry.
 
-\subsection{Alkyl nitrates}
-\subsubsection{Alkyl nitrates chemistry}
-Alkyl nitrates are important tropospheric trace gases which are formed following the same chemistry that leads to the production of ozone \citep{Reeves2007}.
-The simplest example, in case of $CH_4$ oxidation, can be described by the set of reactions below (also shown in Figure 2).
-
+\subsubsection*{Ozone tropospheric chemistry}
 %\input{O3CycleScheme}
-
+Having clarified the role of tropospheric ozone, uncertainties in its budget and having identified the ways of revealing the cause of these uncertainties, it is time to remind ourselves the chemistry that controls the abundance of ozone in the troposphere.
+The key chemical processes that lead to $O_3$ production and destruction are driven by reaction cycles involving free-radical intermediates, which are mainly formed from the photolysis of $O_3$ itself (\ref{reac:O3+hv}).
 In the presence of UV light shorter than 320 nm and water vapour $O_3$ photolysis is the main source of OH radicals in the troposphere:
-\begin{equation} \label{eq:O3hv}
+\begin{equation} \label{reac:O3+hv}
 O_3 + hv \rightarrow O(^1D) + O_2
@@
 \end{equation}
-Once formed the OH radicals react mainly with $CO$ and $CH_4$ to initiate the $O_3$ formation or removal cycles by producing peroxy radicals ($HO_2$, $CH_3O_2$):
+Once formed the OH radicals react mainly with $CO$ and $CH_4$ to initiate the $O_3$ formation or removal cycles by producing the peroxy radicals ($HO_2$, $CH_3O_2$):
 \begin{equation} \label{eq:OH+CO}
@@
 
-The fate of the peroxy radicals depends on the presence of NOx.
-At low NOx levels (remote regions), the $HO_2$ radical reacts with another HO2 to form hydrogen peroxide, H2O2, or with the methyl peroxy radical, $CH_3O_2$, to form methyl hydrogen peroxide ($CH_3OOH$):
+The fate of the peroxy radicals depends on the presence of $NO_x$ ($=NO + NO_2$).
+At low $NO_x$ levels (remote regions), the $HO_2$ radical reacts with another $HO_2$ to form hydrogen peroxide ($H_2O_2$) or reacts with the methyl peroxy radical ($CH_3O_2$) to form methyl hydrogen peroxide ($CH_3OOH$):
 \begin{equation} \label{eq:HO2+HO2}
@@
 \end{equation}
-On a whole, at low NOx levels a net $O_3$ loss takes place, because the reaction sequence is initiated by $O_3$ photolysis.
-Plus under the same conditions, some additional $O_3$ removal occurs due to the re-action of $HO_2$ radicals with $O_3$,
+On a whole, at low $NO_x$ levels a net $O_3$ loss takes place, because the reaction sequence is initiated by $O_3$ photolysis.
+Plus under the same conditions, some additional $O_3$ removal occurs due to the reaction of $HO_2$ radicals with $O_3$,
 \begin{equation} \label{eq:HO2+O3}
@@
 \end{equation}
-leading to the regeneration of OH radicals as part of an $O_3$ depleting HOx catalytic cycle.
+leading to the regeneration of OH radicals as part of an $O_3$ depleting $OH-HO_2$ ($HO_x$) catalytic cycle.
 
@@
 In other words, at high NOx regime the $O_3$ formation rate becomes sensitive to the concentration of $CH_4$ or $CO$ and to inputs of other VOCs \citep{Fowler2008}.
+
+\subsection{Alkyl nitrates}
+\subsubsection{Alkyl nitrates chemistry}
+(An example of such set of reactions is alkyl nitrate chemistry)
+
+Alkyl nitrates ($RONO_2$) are organic tropospheric trace gases, which   due to one of their formation mechanisms are closely linked with the production of tropospheric ozone \citep{Reeves2007}.
 

### 2015-08-01 MSc_diss.tex
Finished section about ozone tropospheric chemistry

@@
 The oxidation of these compounds is the main process taking place in the atmosphere that keeps their concentrations at levels that do not distort the chemical balance of the atmosphere \citep{Prinn2003}.
-In other words, oxidation is a removal “cleansing” mechanism for these compounds (various pollutants and greenhouse gases), and its rate is often referred to as the oxidizing capacity of the atmosphere.
-This capacity depends mainly on two factors: the global burden of the principal oxidants such as ozone ($O_3$), the hydroxyl radical ($OH$) and hydrogen peroxide ($H_2O_2$), and climate \citep{Prinn2003,Thompson1992}.
+In other words, oxidation is a removal 'cleansing' mechanism for these compounds (various pollutants and greenhouse gases), and its rate is often referred to as the oxidizing capacity of the atmosphere.
+This capacity depends mainly on two factors: climate and the global burden of the principal oxidants such as ozone ($O_3$), the hydroxyl radical ($OH$) and hydrogen peroxide ($H_2O_2$) \citep{Prinn2003,Thompson1992}.
 
@@
 %\input{O3CycleScheme}
-Having clarified the role of tropospheric ozone, uncertainties in its budget and having identified the ways of revealing the cause of these uncertainties, it is time to remind ourselves the chemistry that controls the abundance of ozone in the troposphere.
-The key chemical processes that lead to $O_3$ production and destruction are driven by reaction cycles involving free-radical intermediates, which are mainly formed from the photolysis of $O_3$ itself (\ref{reac:O3+hv}).
-In the presence of UV light shorter than 320 nm and water vapour $O_3$ photolysis is the main source of OH radicals in the troposphere:
-\begin{equation} \label{reac:O3+hv}
+Having identified the role of tropospheric ozone, it is time to remind ourselves the chemistry that controls abundance of it.
+
+The key chemical processes that lead to $O_3$ formation and destruction in the troposphere are driven by reaction cycles involving free-radical intermediates, which are mainly formed from the photolysis of $O_3$ itself (\ref{reac:O3+hv=O1D+O2}) \citep{Fowler2008}.
+At UV wavelengths shorter than 320 nm, $O_3$ photolysis generates electronically excited oxygen atoms ($O(^1D)$), which than can either react with water vapour to form $OH$ radicals (\ref{reac:O1D+H2O=OH+OH}), or collide with an inert molecule (denoted '$M$') to reform $O_3$ (\ref{reac:O1D=O3P} and \ref{reac:O3P+O2=O3}):
+\begin{equation}\label{reac:O3+hv=O1D+O2}
 O_3 + hv \rightarrow O(^1D) + O_2
 \end{equation}
-\begin{equation} \label{eq:O1D+H2O}
-O(^1D) + H_2O \rightarrow 2OH
+\begin{equation}\label{reac:O1D+H2O=OH+OH}
+O(^1D) + H_2O \rightarrow OH + OH
 \end{equation}
-Once formed the OH radicals react mainly with $CO$ and $CH_4$ to initiate the $O_3$ formation or removal cycles by producing the peroxy radicals ($HO_2$, $CH_3O_2$):
-\begin{equation} \label{eq:OH+CO}
-OH + CO \rightarrow HO_2 + CO_2
+\begin{equation}\label{reac:O1D=O3P}
+O(^1D) \xrightarrow{M} O(^3P)
 \end{equation}
-\begin{equation} \label{eq:OH+CH4}
-OH + CH_4 \rightarrow CH_3O_2 + H_2O
+\begin{equation}\label{reac:O3P+O2=O3}
+O(^3P) + O_2 \xrightarrow{M} O_3
+\end{equation}
+Once formed the OH radicals mainly react with $CO$ and $CH_4$ and initiate the $O_3$ formation or removal cycles by producing the peroxy radicals ($HO_2$, $CH_3O_2$):
+\begin{equation}\label{reac:OH+CO=H+CO2}
+OH + CO \rightarrow H + CO_2
+\end{equation}
+\begin{equation}\label{reac:H+O2=HO2}
+H + O_2 \xrightarrow{M} HO_2
+\end{equation}
+\begin{equation}\label{reac:OH+CH4=CH3+H2O}
+OH + CH_4 \rightarrow CH_3 + H_2O
+\end{equation}
+\begin{equation}\label{reac:CH3O+O2=CH3O2}
+CH_3 + O_2 \xrightarrow{M} CH_3O_2
+\end{equation}
+The fate of the peroxy radicals depends on the presence of $NO_x$ ($=NO + NO_2$).
+In low-$NO_x$ conditions (remote regions), the $HO_2$ radical reacts with another $HO_2$ to form hydrogen peroxide ($H_2O_2$) or reacts with the methyl peroxy radical ($CH_3O_2$) to form methyl hydrogen peroxide ($CH_3OOH$):
+\begin{equation} \label{reac:HO2+HO2=H2O2+O2}
+HO_2 + HO_2 \rightarrow H_2O_2 + O_2
+\end{equation}
+\begin{equation} \label{reac:HO2+CH3O2=CH3OOH+O2}
+HO_2 + CH_3O_2 \rightarrow CH_3OOH + O_2
+\end{equation}
+On a whole, in low-$NO_x$ conditions, a net $O_3$ destruction takes place, because the reaction sequence is initiated by $O_3$ photolysis.
+Plus in the same conditions some additional $O_3$ removal can occur because of the reaction of the $HO_2$ radical with $O_3$ leading to further destruction of ozone in a chain sequence involving formation of $OH$ radicals:
+\begin{equation}\label{reac:HO2+O3=OH+2O2}
+HO_2 + O_3 \rightarrow OH + 2O_2
+\end{equation}
+\begin{equation}\label{reac:OH+O3=HO2+O2}
+OH + O_3 \rightarrow HO_2 + O_2
 \end{equation}
 
-The fate of the peroxy radicals depends on the presence of $NO_x$ ($=NO + NO_2$).
-At low $NO_x$ levels (remote regions), the $HO_2$ radical reacts with another $HO_2$ to form hydrogen peroxide ($H_2O_2$) or reacts with the methyl peroxy radical ($CH_3O_2$) to form methyl hydrogen peroxide ($CH_3OOH$):
-\begin{equation} \label{eq:HO2+HO2}
-HO_2 + HO_2 \rightarrow H_2O_2 + O_2
-\end{equation}
-\begin{equation} \label{eq:HO2+CH3O2}
-HO_2 + CH_3O_2 \rightarrow CH_3OOH + O_2
-\end{equation}
-On a whole, at low $NO_x$ levels a net $O_3$ loss takes place, because the reaction sequence is initiated by $O_3$ photolysis.
-Plus under the same conditions, some additional $O_3$ removal occurs due to the reaction of $HO_2$ radicals with $O_3$,
-\begin{equation} \label{eq:HO2+O3}
-HO_2 + O_3 \rightarrow OH + 2O_2
-\end{equation}
-leading to the regeneration of OH radicals as part of an $O_3$ depleting $OH-HO_2$ ($HO_x$) catalytic cycle.
-
-At intermediate NOx levels (rural areas of most industrialised countries), the peroxy radicals react with nitrogen monoxide ($NO$) to form nitrogen dioxide ($NO_2$) with the subsequent photolysis of $NO_2$ generating $O_3$:
-\begin{equation} \label{eq:HO2+NO}
+In intermediate $NO_x$ conditions (rural areas of most industrialised countries), the peroxy radicals react with nitrogen monoxide ($NO$) to form nitrogen dioxide ($NO_2$), which subsequent photolysis leads to $O_3$ generation:
+\begin{equation}\label{reac:HO2+NO=NO2+OH}
 HO_2 + NO \rightarrow NO_2 + OH
 \end{equation}
-\begin{equation} \label{eq:CH_3O_2+NO}
+\begin{equation}\label{reac:CH3O2+NO=NO2+CH3O}
 CH_3O_2 + NO \rightarrow NO_2 + CH_3O
 \end{equation}
-\begin{equation} \label{eq:NO2hv}
+\begin{equation}\label{reac:NO2+hv=NO+O3P}
 NO_2 + hv \rightarrow NO + O(^3P)
 \end{equation}
-\begin{equation} \label{eq:O3P+O2}
-O(^3P) + O_2 + M \rightarrow O_3 + M
+\begin{equation}\label{reac:O3P+O2=O3}
+!!!(R4)sameO(^3P) + O_2 \xrightarrow{M} O_3
 \end{equation}
+As shown in Figure ???
+reactions (\ref{reac:HO2+NO=NO2+OH}) and (\ref{reac:CH3O2+NO=NO2+CH3O}) form parts of the free-radical propagated $O_3$ forming cycles, which may occur a number of times before being halted by radical termination reactions (\ref{reac:HO2+HO2=H2O2+O2} and \ref{reac:HO2+CH3O2=CH3OOH+O2}), and the higher $NO_x$ level is the greater number of these cycles can occur.
+Therefore, in intermediate $NO_x$ conditions a net $O_3$ formation takes place and it is sensitive to $NO_x$.
+
+In high-$NO_x$ conditions (urban environment), the $OH$ radical may have another fate rather than reacting with $CO$, $CH_4$ and $O_3$.
+Depending on the concentration of $NO_2$ it can react with it to form nitric acid ($HNO_3$):
+\begin{equation}\label{reac:OH+NO2=HNO3}
+OH + NO_2 \xrightarrow{M} HNO_3
+\end{equation}
+The formation of $HNO_3$ terminates free-radical propagated $O_3$ forming cycles leading to a decrease in $O_3$ production rate.
+However, elevated emissions of $CO$ and $CH_4$ (and other nonmethane volatile organic compounds (nmVOCs)) may make reactions (\ref{reac:OH+CO=H+CO2})-(\ref{reac:OH+CH4=CH3+H2O}) more competitive with the reaction (\ref{reac:OH+NO2=HNO3}) and lead to an increase in the $O_3$ production rate.
+Therefore, in high-$NO_x$ conditions the $O_3$ production rate can either decrease or increase depending on $NO_x$ and $VOCs$ levels.
+
+The free-radical propagated oxidation of hydrocarbons leads to the generation of carbonyl compounds, among which formaldehyde ($HCHO$) is the most common.
+Its photolysis and further oxidation contributes to radical generation:
+\begin{equation}\label{reac:HCHO+hv=H+HCO}
+HCHO + hv \rightarrow H + HCO
+\end{equation}
+\begin{equation}\label{reac:HCO+O2=HO2+CO}
+HCO + O_2 \rightarrow HO_2 + CO
+\end{equation}
+\begin{equation}\label{reac:H+O2=HO2}
+!!!sameH + O_2 \xrightarrow{M} HO_2
+\end{equation}
+therefore also having an impact on the rates of $O_3$ production \citep{Fowler2008} as do some other species, which are not mentioned in this section.
+Their role is described in other sections as fas as it is necessary.
+
+\subsection{Alkyl nitrates}
+\subsubsection{Alkyl nitrates chemistry}
+(An example of such set of reactions is alkyl nitrate chemistry)
+
+Alkyl nitrates ($RONO_2$) are organic tropospheric trace gases, which   due to one of their formation mechanisms are closely linked with the production of tropospheric ozone \citep{Reeves2007}.
+
 As it is shown in Figure 2 reactions (8) and (9) form part of the free-radical propagated $O_3$ forming cycles, which may occur a number of times before being halted by a radical termination reactions (not shown).
@@
 Being a reservoir for the short lived NOx and having atmospheric lifetimes ranging from several days to about a month \citep{Reeves2007} alkyl nitrates can be transported over long distances, and then lost through photolysis and reaction with $OH$:
-\begin{equation} \label{eq:CH3NO3hv}
+\begin{equation} \label{eq:CH3NO3+hv}
 CH_3ONO_2 + hv \rightarrow CH_3O + NO_2
@@
 It means that alkyl nitrates can ‘release’ $NO_2$ back to the atmosphere after a while, and in association with reactions (10) and (11) start a new $O_3$ formation cycle being quite away from the source.
-
-The free-radical propagated oxidation of hydrocarbons leads to the generation of carbonyl compounds, in this case formaldehyde ($HCHO$).
-Its further oxidation and photolysis contributes to radical generation:
-\begin{equation} \label{eq:HCHOhv1}
-HCHO + hv \rightarrow 2HO_2 + CO
-\end{equation}
-\begin{equation} \label{eq:HCHOhv2}
-HCHO + hv \rightarrow CO + H_2
-\end{equation}
-\begin{equation} \label{eq:HCHO+OH}
-HCHO + OH \rightarrow HO_2 + CO
-\end{equation}
-As a result, the formation of $HCHO$ has an impact on the rates of the oxidation cycles described above, through secondary radical generation \citep{Fowler2008}.
-At high NOx levels (urban environment) reactions (8) and (9) dominate for $CH_3O_2$ and $HO_2$, but $O_3$ formation becomes inhibited by further increases in NOx.
-In other words, at high NOx regime the $O_3$ formation rate becomes sensitive to the concentration of $CH_4$ or $CO$ and to inputs of other VOCs \citep{Fowler2008}.
-
-\subsection{Alkyl nitrates}
-\subsubsection{Alkyl nitrates chemistry}
-(An example of such set of reactions is alkyl nitrate chemistry)
-
-Alkyl nitrates ($RONO_2$) are organic tropospheric trace gases, which   due to one of their formation mechanisms are closely linked with the production of tropospheric ozone \citep{Reeves2007}.
 

### 2015-08-01 MSc_diss.tex
Changes in structure

@@
 \subsection{Alkyl nitrates}
-\subsubsection{Alkyl nitrates chemistry}
-(An example of such set of reactions is alkyl nitrate chemistry)
+Alkyl nitrates ($RONO_2$) are organic tropospheric trace gases, which   due to one of their formation mechanisms are closely linked with the production of tropospheric ozone \citep{Reeves2007}.
+This mechanism is present in the following section.
 
-Alkyl nitrates ($RONO_2$) are organic tropospheric trace gases, which   due to one of their formation mechanisms are closely linked with the production of tropospheric ozone \citep{Reeves2007}.
+\subsubsection*{Alkyl nitrates chemistry}
 
@@
 
-\subsubsection{Observational studies}
-
-mean ANs concentrations: \citep{Reeves2007}, \citep{Roberts1998}
-
-\subsubsection{Modelling studies}
-
-To sum up, alkyl nitrates are produced as a minor branch of the reaction of peroxy radicals with NO \citep{Day2003}.
-Since these compounds are removed from the troposphere by photolysis and OH oxidation they can provide information on photochemical evolution of an air mass through relationships with their parent hydrocarbons \citep{Bertman1995}.
-Acting as reservoir species for reactive nitrogen (NOy) and having a relatively long atmospheric lifetimes alkyl nitrates play an important role in long range transport of NOx.
-Furthermore, alkyl nitrates link to ozone and carbonyl production making them useful tracers for both species \citep{Worton2010}.
-Despite their relative importance alkyl nitrates are usually not included in the climate-chemistry models, for example in UKCA (Reeves, personal communication), which potentially could lead to discrepancies between models and observations described before.
-
-[Source characterization and photochemical processes in
-transported air masses have been analyzed by examining
-RONO2 formation from anthropogenic emissions of alkane
-precursors [Simpson et al., 2003; Madronich, 2006; Simpson
-et al., 2006; Reeves et al., 2007; Worton et al., 2010]. (Neuman2012)]
-
-!! from \citep{Finlayson-Pitts2000}
-The relationship between the peroxy radical concentration and the ozone photolysis rate constant for these higher NO conditions can be again approximated using steady-state analysis [..].
-While OH is recycled in its reactions with CO and CH4 via HO2, it is pernanetly removed at higher NOx concentrations by the reaction of OH with NO2, forming nitric acid:
-OH + NO2 = M = HONO2 (113).
-This reaction is primarily a daytime reaction because most OH sources are photolytic in nature. As a result, the NO2 reaction with OH competes!!!! with NO2 photolysis (page 290/993)!!
-         Rate coeff (OH+NO2 = HNO3)  J(NO2 = NO+O)
-Book     2.47e-30          7.00e-3
-My model 1.15e-11          4.39e-03
-Thus, the reaction with OH is not usually a dominant loss process for NO2, but it is still sufficiently fact to from significant amounts of HNO3 during the day, particularly in polluted regions with relatively large NO2 concentrations.
-Nitric acid undergoes both wet and dry deposition rapidly and can be neutralized by ammonia, the major gaseous base found in the atmosphere.
-
-\subsection{Motivation}
-lalalaep
-
-\section{Overall objective and specific aims}
-\subsection{Overall objective}
-The overall purpose of this research is to investigate the impact of alkyl nitrate chemistry on ozone production efficiency.
-\subsection{Specific aims}
-Specific aims are:
-\begin{itemize}
-\item to investigate of the relationship between ozone production and carbon bond emissions (see Data and Methods);
-\item to analyse the impact of inclusion of alkyl nitrates chemistry on this relationship;
-\item to analyse everything mentioned above in various NOx regimes.
-\end{itemize}
-
-\section{Literature review} \label{sec:data}
-Alkyl nitrates ($RONO_2$) are organic tropospheric trace gases.
-They are produced photochemically through the oxidation of their parent hydrocarbons \citep{Roberts1990} or emitted directly from equatorial oceans \citep{Blake2003} and biomass burning \citep{Simpson2002}.
-Being involved in the ozone production chemistry, these species can act as reservoirs of nitrogen oxides which availability largely affects the production of ozone \citep{Reeves2007}.
-Since the alkyl nitrates have atmospheric lifetimes ranging from several days to about a month, they can redistribute nitrogen oxides presented over polluted continental regions to clean remote marine environments, and therefore are considered as one of the tracers for determining the anthropogenic influence \citep{Atherton1989, Reeves2007, Worton2005}.
-
-Alkyl nitrates are removed from the atmosphere by photolysis \citep{Turberg1990} or through oxidation by hydroxyl radical ($OH$) \citep{Talukdar1997}.
-Photolysis is the dominant loss mechanism for the compounds with short carbon chains and becomes more important with decreasing carbon number.
-Reaction with $OH$ follows the opposite trend and is most important for the pentyl nitrates and higher.
-
-Owing to the existence of the relationship between alkyl nitrates and their parent hydrocarbons, measurements of their concentrations can provide information on the extent of photochemical processing in the air mass.
-\cite{Bertman1995} developed a kinetic approach to describe this relationship through the evolution of the ratio of alkyl nitrate to its parent hydrocarbon as a function of time.
-This approach has been extensively used to study air mass photochemical ageing at several sites, in North America \citep{Bertman1995, Roberts1998}, China \citep{Simpson2006}, the British Isles \citep{Worton2010} and over the Pacific \citep{Simpson2003} and North Atlantic Ocean \citep{Reeves2007}.
-
-\cite{Bertman1995} found that the theoretical ratios of  2-butyl nitrate to n-butane and 2- and 3-pentyl nitrates to pentane strongly agree with those measured from the ambient air showing no evidence for primary sources of these nitrates.
-On the contrary, measurements of the other ratios, ethyl/ethane, n-propyl nitrate/propane, and 2-propyl nitrate/propane, appeared to be significantly higher than predicted, relative to the 2-butyl nitrate/n-butane ratio, meaning that there should be an additional source of these alkyl nitrates.
-\cite{Roberts1998} generally approved the theory by comparing Bertman's conclusions with the results from a new observational dataset.
-
-Using methyl nitrate as a tracer of marine $RONO_2$ \cite{Simpson2006} revealed that the South China Sea in not a region of strong $RONO_2$ emissions suggesting that photochemical rather marine production of alkyl nitrates is dominant at the site.
-
-\cite{Worton2010} found that the photochemical source of the short-chain alkyl nitrates originates mostly from the photochemical oxidation and decomposition of longer-chain compounds rather than from the oxidation of the parent hydrocarbons as it is thought in the previous studies.
-
-The brief description of the measurement sites and data is presented in section \ref{sec:data}.
-In section \ref{sec:method} the methodology used in this work is defined.
-Section \ref{sec:res} contains the results.
-Section \ref{sec:discuss} discusses the results, particularly focusing attention on the analysis of the discrepancy between the observational data and the results of theoretical modelling.
-Conclusions constitute section \ref{sec:conclusion}.
-
-!
-Since the most alkoxy radicals ($RO$) react rapidly with $O_2$ under atmospheric conditions and most large (>=C4) alkoxy radicals also rapidly decompose or isomerize, alkyl nitrate formation from reaction RO+NO2 = RONO2 is relatively unimportant for most organics under atmospheric conditions while observations showed that the reaction RO2+NO = RONO2 gives a higher yield (from \citep{Atkinson2000}, actually from 1982).
-
-! Since one of the main mechanisms of the alkyl nitrate production (>C3!!) is the gas gas phase oxidation of alkanes by OH in the presence of NOx (\citep{Newland2013}).
-
-\subsection{Bertman}
+\subsubsection*{Bertman}
 Below is the simplified scheme of atmospheric hydrocarbon oxidation leading to the formation of alkyl nitrates (see also Fig.
@@
 \eqref{eq:integral} and \eqref{eq:model} the same \citep{Reeves2007}.
+
+\subsection{Ozone, alkyl nitrates and carbon bonds}
+
+\subsection{Motivation}
+lalalaep
+
+\section{Overall objective and specific aims}
+\subsection{Overall objective}
+The overall purpose of this research is to investigate the impact of alkyl nitrate chemistry on ozone production efficiency.
+\subsection{Specific aims}
+Specific aims are:
+\begin{itemize}
+\item to investigate of the relationship between ozone production and carbon bond emissions (see Data and Methods);
+\item to analyse the impact of inclusion of alkyl nitrates chemistry on this relationship;
+\item to analyse everything mentioned above in various NOx regimes.
+\end{itemize}
+
+\section{Literature review} \label{sec:data}
+\subsubsection{Observational studies}
+
+mean ANs concentrations: \citep{Reeves2007}, \citep{Roberts1998}
+
+\subsubsection{Modelling studies}
+
+To sum up, alkyl nitrates are produced as a minor branch of the reaction of peroxy radicals with NO \citep{Day2003}.
+Since these compounds are removed from the troposphere by photolysis and OH oxidation they can provide information on photochemical evolution of an air mass through relationships with their parent hydrocarbons \citep{Bertman1995}.
+Acting as reservoir species for reactive nitrogen (NOy) and having a relatively long atmospheric lifetimes alkyl nitrates play an important role in long range transport of NOx.
+Furthermore, alkyl nitrates link to ozone and carbonyl production making them useful tracers for both species \citep{Worton2010}.
+Despite their relative importance alkyl nitrates are usually not included in the climate-chemistry models, for example in UKCA (Reeves, personal communication), which potentially could lead to discrepancies between models and observations described before.
+
+[Source characterization and photochemical processes in
+transported air masses have been analyzed by examining
+RONO2 formation from anthropogenic emissions of alkane
+precursors [Simpson et al., 2003; Madronich, 2006; Simpson
+et al., 2006; Reeves et al., 2007; Worton et al., 2010]. (Neuman2012)]
+
+!! from \citep{Finlayson-Pitts2000}
+The relationship between the peroxy radical concentration and the ozone photolysis rate constant for these higher NO conditions can be again approximated using steady-state analysis [..].
+While OH is recycled in its reactions with CO and CH4 via HO2, it is pernanetly removed at higher NOx concentrations by the reaction of OH with NO2, forming nitric acid:
+OH + NO2 = M = HONO2 (113).
+This reaction is primarily a daytime reaction because most OH sources are photolytic in nature. As a result, the NO2 reaction with OH competes!!!! with NO2 photolysis (page 290/993)!!
+         Rate coeff (OH+NO2 = HNO3)  J(NO2 = NO+O)
+Book     2.47e-30          7.00e-3
+My model 1.15e-11          4.39e-03
+Thus, the reaction with OH is not usually a dominant loss process for NO2, but it is still sufficiently fact to from significant amounts of HNO3 during the day, particularly in polluted regions with relatively large NO2 concentrations.
+Nitric acid undergoes both wet and dry deposition rapidly and can be neutralized by ammonia, the major gaseous base found in the atmosphere.
+
+Alkyl nitrates ($RONO_2$) are organic tropospheric trace gases.
+They are produced photochemically through the oxidation of their parent hydrocarbons \citep{Roberts1990} or emitted directly from equatorial oceans \citep{Blake2003} and biomass burning \citep{Simpson2002}.
+Being involved in the ozone production chemistry, these species can act as reservoirs of nitrogen oxides which availability largely affects the production of ozone \citep{Reeves2007}.
+Since the alkyl nitrates have atmospheric lifetimes ranging from several days to about a month, they can redistribute nitrogen oxides presented over polluted continental regions to clean remote marine environments, and therefore are considered as one of the tracers for determining the anthropogenic influence \citep{Atherton1989, Reeves2007, Worton2005}.
+
+Alkyl nitrates are removed from the atmosphere by photolysis \citep{Turberg1990} or through oxidation by hydroxyl radical ($OH$) \citep{Talukdar1997}.
+Photolysis is the dominant loss mechanism for the compounds with short carbon chains and becomes more important with decreasing carbon number.
+Reaction with $OH$ follows the opposite trend and is most important for the pentyl nitrates and higher.
+
+Owing to the existence of the relationship between alkyl nitrates and their parent hydrocarbons, measurements of their concentrations can provide information on the extent of photochemical processing in the air mass.
+\cite{Bertman1995} developed a kinetic approach to describe this relationship through the evolution of the ratio of alkyl nitrate to its parent hydrocarbon as a function of time.
+This approach has been extensively used to study air mass photochemical ageing at several sites, in North America \citep{Bertman1995, Roberts1998}, China \citep{Simpson2006}, the British Isles \citep{Worton2010} and over the Pacific \citep{Simpson2003} and North Atlantic Ocean \citep{Reeves2007}.
+
+\cite{Bertman1995} found that the theoretical ratios of  2-butyl nitrate to n-butane and 2- and 3-pentyl nitrates to pentane strongly agree with those measured from the ambient air showing no evidence for primary sources of these nitrates.
+On the contrary, measurements of the other ratios, ethyl/ethane, n-propyl nitrate/propane, and 2-propyl nitrate/propane, appeared to be significantly higher than predicted, relative to the 2-butyl nitrate/n-butane ratio, meaning that there should be an additional source of these alkyl nitrates.
+\cite{Roberts1998} generally approved the theory by comparing Bertman's conclusions with the results from a new observational dataset.
+
+Using methyl nitrate as a tracer of marine $RONO_2$ \cite{Simpson2006} revealed that the South China Sea in not a region of strong $RONO_2$ emissions suggesting that photochemical rather marine production of alkyl nitrates is dominant at the site.
+
+\cite{Worton2010} found that the photochemical source of the short-chain alkyl nitrates originates mostly from the photochemical oxidation and decomposition of longer-chain compounds rather than from the oxidation of the parent hydrocarbons as it is thought in the previous studies.
+
+The brief description of the measurement sites and data is presented in section \ref{sec:data}.
+In section \ref{sec:method} the methodology used in this work is defined.
+Section \ref{sec:res} contains the results.
+Section \ref{sec:discuss} discusses the results, particularly focusing attention on the analysis of the discrepancy between the observational data and the results of theoretical modelling.
+Conclusions constitute section \ref{sec:conclusion}.
+
+!
+Since the most alkoxy radicals ($RO$) react rapidly with $O_2$ under atmospheric conditions and most large (>=C4) alkoxy radicals also rapidly decompose or isomerize, alkyl nitrate formation from reaction RO+NO2 = RONO2 is relatively unimportant for most organics under atmospheric conditions while observations showed that the reaction RO2+NO = RONO2 gives a higher yield (from \citep{Atkinson2000}, actually from 1982).
+
+! Since one of the main mechanisms of the alkyl nitrate production (>C3!!) is the gas gas phase oxidation of alkanes by OH in the presence of NOx (\citep{Newland2013}).
 

### 2015-08-01 MSc_diss.tex
Started subsection about ANs formation

@@
 therefore also having an impact on the rates of $O_3$ production \citep{Fowler2008} as do some other species, which are not mentioned in this section.
-Their role is described in other sections as fas as it is necessary.
+Their role is described in other sections as far as it is necessary.
 
 \subsection{Alkyl nitrates}
-Alkyl nitrates ($RONO_2$) are organic tropospheric trace gases, which   due to one of their formation mechanisms are closely linked with the production of tropospheric ozone \citep{Reeves2007}.
+Alkyl nitrates ($RONO_2$) are organic tropospheric trace gases, which through one of their formation mechanisms are closely linked with the production of tropospheric ozone \citep{Reeves2007}.
 This mechanism is present in the following section.
@@
 \subsubsection*{Alkyl nitrates chemistry}
+
+The photochemical formation of the alkyl nitrates starts when the chemistry of the troposphere reaches the regime of $O_3$ production, i.e.
+when reactions of the peroxy radicals with $NO$ start to compete with their self- and cross-reactions more effectively due to an increase in $NO_x$ concentration.
+In this case the reaction between of the alkyl peroxy radicals ($RO_2$ where $R$ denotes methyl group, $CH_3$) and $NO$ can reveal its two-channel nature \citep{Day2003}, which is controlled by the branching ratio $\alpha_3$ in the following way:
+\begin{subequations} \label{eq:peroxy_no0}
+\begin{align}
+RO_2 + NO &\xrightarrow{1-\alpha_3} RO + NO_2 \label{eq:peroxy_no1}\\
+RO_2 + NO &\xrightarrow{\alpha_3} RONO_2 \label{eq:peroxy_no2}
+\end{align}
+\end{subequations}
+The former reaction is a crucial since it can serve as a termination step
+
+in the net production of tropospheric ozone, since photolysis of $NO_2$ is the main photochemical source of $O_3$ in the troposphere.
+
+In other words, a branching ratio $\alpha_3$ of the reactions between alkyl peroxy radicals and $NO$ leads to formation of alkyl nitrates via rearrangement of a high-energy intermediate \citep{Bertman1995}.
+The branching ratio $\alpha_3$ is larger when the parent alkane is larger and is a function of temperature and pressure.
 
@@
 \subsubsection*{Bertman}
-Below is the simplified scheme of atmospheric hydrocarbon oxidation leading to the formation of alkyl nitrates (see also Fig.
-\ref{fig:scheme_bertman}).
-It begins from the process of $H$ abstraction by the $OH$ radical where the fraction of the primary or secondary radical is denoted by $\alpha_1$.
-\begin{equation} \label{eq:alkane_oh}
-RH + \bullet OH \xrightarrow{k_1, \alpha_1} R\bullet + H_2O
-\end{equation}
-The resulting radical rapidly reacts with $O_2$ and forms an alkyl peroxy radical ($R\bullet$):
-\begin{equation} \label{eq:rad_o2}
-R\bullet + O_2 \xrightarrow{k_2} ROO\bullet
-\end{equation}
 Then the alkyl peroxy radical reacts with $NO$, when it can either lose an oxygen atom to form alkoxy radical ($RO\bullet$) or bond with $NO$ to form an alkyl nitrate.
@@
 \end{subequations}
-In other words, a branching ratio $\alpha_3$ of the reactions between alkyl peroxy radicals and $NO$ leads to formation of nitrate products via rearrangement of a high-energy intermediate \citep{Bertman1995}.
+In other words, a branching ratio $\alpha_3$ of the reactions between alkyl peroxy radicals and $NO$ leads to formation of alkyl nitrates via rearrangement of a high-energy intermediate \citep{Bertman1995}.
 The branching ratio $\alpha_3$ is larger when the parent alkane is larger and is a function of temperature and pressure.
@@
 
-Assuming that the only source of alkyl nitrates is the $OH$ reaction with alkane, and that reactions take place in $NO$-rich atmosphere (so no $ROO\bullet$ self-reactions occur), the alkyl nitrate part in this generalised scheme can be reduced to the two equations of formation and loss:
+[Assuming that the only source of alkyl nitrates is the $OH$ reaction with alkane, and that reactions take place in $NO$-rich atmosphere (so no $ROO\bullet$ self-reactions occur), the alkyl nitrate part in this generalised scheme can be reduced to the two equations of formation and loss:
 \begin{subequations} \label{eq:an_form_loss0}
@@
 This simplification is reasonable only when mixing affects both the parent alkane and the corresponding alkyl nitrate similarly, keeping the ratios in eqs.
-\eqref{eq:integral} and \eqref{eq:model} the same \citep{Reeves2007}.
+\eqref{eq:integral} and \eqref{eq:model} the same \citep{Reeves2007}].
 

### 2015-08-02 MSc_diss.tex
Rewriting literature review

@@
 when reactions of the peroxy radicals with $NO$ start to compete with their self- and cross-reactions more effectively due to an increase in $NO_x$ concentration.
-In this case the reaction between of the alkyl peroxy radicals ($RO_2$ where $R$ denotes methyl group, $CH_3$) and $NO$ can reveal its two-channel nature \citep{Day2003}, which is controlled by the branching ratio $\alpha_3$ in the following way:
-\begin{subequations} \label{eq:peroxy_no0}
+In this case the reaction between the alkyl peroxy radical ($RO_2$, where $R$ denotes methyl group, $CH_3$) and $NO$ can reveal its two-channel nature \citep{Day2003}, which is controlled by the branching ratio $\alpha_3$ in the following way \citep{Roberts1990}
+\begin{subequations} \label{reac:RONO2ab}
 \begin{align}
@@
 \end{subequations}
-The former reaction is a crucial since it can serve as a termination step
-
-in the net production of tropospheric ozone, since photolysis of $NO_2$ is the main photochemical source of $O_3$ in the troposphere.
-
-In other words, a branching ratio $\alpha_3$ of the reactions between alkyl peroxy radicals and $NO$ leads to formation of alkyl nitrates via rearrangement of a high-energy intermediate \citep{Bertman1995}.
+The latter reaction can serve as a termination step of the propagated $O_3$ forming cycle, and as \citep{Roberts1998} estimated this likely begin to occur when the $NO$ concentration increases up to 100-200 ppt.
+
+\subsubsection*{Branching ratio}
 The branching ratio $\alpha_3$ is larger when the parent alkane is larger and is a function of temperature and pressure.
-
-As it is shown in Figure 2 reactions (8) and (9) form part of the free-radical propagated $O_3$ forming cycles, which may occur a number of times before being halted by a radical termination reactions (not shown).
-However, the reaction of $CH_3O_2$ with NO is a two-channel reaction \citep{Day2003}, and its second branch can terminate the $O_3$ forming cycle earlier by producing an alkyl nitrate, in this case methyl nitrate ($CH_3ONO_2$):
-\begin{equation} \label{eq:CH3NO3form}
-CH_3O_2 + NO \rightarrow CH_3ONO_2
-\end{equation}
-
+\citep{Newland2013}
+[The yield of this secondary pathway increases with increasing carbon chain length (Table 1.3)]
+
+\subsubsection*{Yield}
+\citep{Newland2013}
+[Since the yield of alkyl nitrates from Reaction 1.9 increases with chain length but the mixing ratios of the alkanes tends to decrease with chain length, propyl and butyl nitrates are generally present at the highest concentrations in polluted air (Beyersdorf et al., 2010)]
+\citep{Sommariva2008}
+[The yield of alkyl nitrate formation is temperature and pressure dependent and increases with the number of carbons in the peroxy radical \citep{Roberts1990}!]
+
+\subsubsection*{Alternative photochemical chemical sources}
+\citep{Newland2013}
+[Another photochemical pathway has been shown to become important for the production of the lower (methyl) alkyl nitrates in polluted air.
+Reaction 1.10 shows the reaction of the alkoxy radical, formed in in Reaction 1.9a, with NO2.
+\begin{equation}\label{reac:RO+NO2=RONO2}
+RO + NO_2 \rightarrow RONO_2
+\end{equation}
+This photochemical pathway has been shown to become dominant over Reaction 1.9b for the production of methyl nitrate in polluted air masses where NOx mixing ratios exceed 40 ppb (Archibald et al., 2007).
+This reaction is only likely to be important for methyl nitrate (!) rather than the higher nitrates because of the very small branching ratio ($\alpha_3$) for the production of methyl nitrate via the reaction of the peroxy radical with NO]
+\citep{Newland2013}
+[in highly polluted environments, the reaction of the alkoxy radical with the NO2 radical (Reaction 1.10) may be important for formation of the C1 – C2 alkyl nitrates (Flocke et al., 1998; Simpson et al., 2006; Archibald et al., 2007).
+Archibald et al.
+(2007) reported a model run which suggested that the formation of methyl nitrate via the two production routes would be equal at NOx values of 37 ppb.
+It is unclear from the paper whether they mean the NOx mixing ratio or NO2 mixing ratio but either would suggest a value of a similar magnitude to that shown in Table 6.6.]
+
+\subsubsection*{Loss reactions}
 Being a reservoir for the short lived NOx and having atmospheric lifetimes ranging from several days to about a month \citep{Reeves2007} alkyl nitrates can be transported over long distances, and then lost through photolysis and reaction with $OH$:
-\begin{equation} \label{eq:CH3NO3+hv}
-CH_3ONO_2 + hv \rightarrow CH_3O + NO_2
-\end{equation}
-\begin{equation} \label{eq:CH3NO3+OH}
-CH_3ONO_2 + OH \rightarrow HCHO + NO_2
+\begin{equation} \label{reac:RONO2+hv=RO+NO2}
+RONO_2 + hv \rightarrow RO + NO_2
+\end{equation}
+\begin{equation} \label{reac:RONO2+OH=RCHO+NO2}
+RONO_2 + OH \rightarrow RCHO + NO_2
 \end{equation}
 It means that alkyl nitrates can ‘release’ $NO_2$ back to the atmosphere after a while, and in association with reactions (10) and (11) start a new $O_3$ formation cycle being quite away from the source.
-
-\subsubsection*{Bertman}
-Then the alkyl peroxy radical reacts with $NO$, when it can either lose an oxygen atom to form alkoxy radical ($RO\bullet$) or bond with $NO$ to form an alkyl nitrate.
-The former reaction is a crucial in the net production of tropospheric ozone, since photolysis of $NO_2$ is the main photochemical source of $O_3$ in the troposphere.
-\begin{subequations} \label{eq:peroxy_no0}
-\begin{align}
-ROO\bullet + NO &\xrightarrow{k_3, (1-\alpha_3)} RO\bullet + NO_2 \label{eq:peroxy_no1}\\
-ROO\bullet + NO &\xrightarrow{k_3, \alpha_3} RONO_2 \label{eq:peroxy_no2}
-\end{align}
-\end{subequations}
-In other words, a branching ratio $\alpha_3$ of the reactions between alkyl peroxy radicals and $NO$ leads to formation of alkyl nitrates via rearrangement of a high-energy intermediate \citep{Bertman1995}.
-The branching ratio $\alpha_3$ is larger when the parent alkane is larger and is a function of temperature and pressure.
-
-\begin{figure}[h]
-\centering
-\begin{tikzpicture}[node distance = 4cm, auto]
-\node[reg] (rh) {$RH$};
-\node[rad] (roo) at (4,0){$ROO\bullet$};
-\node[add] (oh) at (3,0.3) {$\bullet OH$};
-\node[rad] (roor) at (4,-2){$ROOR'$};
-\node[add] (r_oo) at (3.7,-1) {$R'OO\bullet$};
-\node[reg] (rono2) at (2,3){$RONO_2$};
-\node[rad] (ro) at (6,3){$RO\bullet$};
-\node[add] (alpha) at ($(roo)!.5!(rono2) + (-1,-0.2)$) {$\alpha_3$};
-\node[add] (alpha2) at ($(roo)!.5!(ro) + (1,-0.2)$) {$1-\alpha_3$};
-\node[add] (no) at ($(roo)+(0,1.5)$) {$NO$};
-\node[reg] (rcho) at (9,3){$RCHO$};
-\node[add] (prod) at (0,3){\textit{products}};
-
-\draw[ar] (rh.east) to [out=0,in=180](roo.west);
-\draw[ar] (roo.south) to [out=270,in=90](roor.north);
-\draw[ar] (roo.north) to [out=90,in=270](rono2.south);
-\draw[ar] (roo.north) to [out=90,in=270](ro.south);
-\draw[ar] (rono2.west) to [out=180,in=0](prod.east);
-\draw[ar] (ro.east) to [out=0,in=180](rcho.west);
-\end{tikzpicture}
-\caption{Schematic representation of the alkyl nitrate chemistry and their formation through the photooxydation of alkanes. Adapted from \citep{Bertman1995}.}
-\label{fig:scheme_bertman}
-\end{figure}
-
-Apart from $NO$, alkyl peroxy radicals can react with each other and give peroxides, without formation of nitrates.
-As the eqs.
-\eqref{eq:peroxy_no0} outline, whether $ROO\bullet$ will bond with nitrates or with other peroxy radicals depends partially on $NO$ concentration.
-The reaction between $ROO\bullet$ and $NO$ is more likely when nitrate concentration is high.
+\citep{Newland2013}
+[The main loss processes for the alkyl nitrates are oxidation by the OH radical and photodissociation (e.g.
+Talukdar et al., 1997a, 1997b).
+The importance of the OH sink relative to photolysis increases with increasing carbon chain length to become the dominant loss process in C5 alkyl nitrates and above (Roberts and Fajer, 1989).]
 
 The alkyl nitrates are removed from the atmosphere through photolysis and reaction with hydroxyl radical.
-\begin{equation} \label{eq:an_photo}
-RONO_2 + h\nu \xrightarrow{j_4} RO\bullet + NO_2
-\end{equation}
-\begin{equation} \label{eq:an_oh}
-RONO_2 + OH \xrightarrow{k_5} \mathit{products}
-\end{equation}
 The relative importance of these two loss reactions depends generally on the net UV intensity and the size of the carbon chain in alkyl nitrates.
@@
 
-[Assuming that the only source of alkyl nitrates is the $OH$ reaction with alkane, and that reactions take place in $NO$-rich atmosphere (so no $ROO\bullet$ self-reactions occur), the alkyl nitrate part in this generalised scheme can be reduced to the two equations of formation and loss:
-\begin{subequations} \label{eq:an_form_loss0}
-\begin{align}
-RH &\xrightarrow{\beta k_A} RONO2 \label{eq:an_form_loss1}\\
-RONO_2 &\xrightarrow{k_B} \mathit{products} \label{eq:an_form_loss2}
-\end{align}
-\end{subequations}
-where $k_A = k_1[OH]$ and $k_B=j_4 + k_5[OH]$.
-The factor $\beta=\alpha_1\alpha_3$ takes into account the fraction of $H$ atom abstraction from the particular hydrocarbon (with the reaction rate $k_1$) and the branching ratios leading to the nitrate.
-The photolytic loss rate of $RONO_2$ is denoted by $j_4$, and $k_5$ is the rate of alkyl nitrate destruction by $OH$.
-
-These kinetic equations lead to the ordinary differential equations which can be integrated in time ($t$) giving a relationship between the alkyl nitrates and their parent alkanes concentrations:
-\begin{equation} \label{eq:integral}
-\frac{[RONO_2]}{[RH]} = \frac{\beta k_A}{k_B - k_A}\left(1-e^{(k_A - k_B)t}\right)+\frac{[RONO_2]_0}{[RH]_0}e^{(k_A - k_B)t}
-\end{equation}
-where subscript $0$ denotes the initial concentrations.
-
-A further good approximation is that there is no alkyl nitrate initially (\textit{i.e.} no direct emissions), hence, the second term in eq.
-(\eqref{eq:integral}) becomes zero.
-Finally, the evolution of alkyl nitrates in an air mass can be modelled using the following formula.
-\begin{equation} \label{eq:model}
-\frac{[RONO_2]}{[RH]} = \frac{\beta k_A}{k_B - k_A}\left(1-e^{(k_A - k_B)t}\right)
-\end{equation}
-It has to be noted that eq.
-\eqref{eq:model} assumes no mixing with surrounding air.
-This simplification is reasonable only when mixing affects both the parent alkane and the corresponding alkyl nitrate similarly, keeping the ratios in eqs.
-\eqref{eq:integral} and \eqref{eq:model} the same \citep{Reeves2007}].
+Alkyl nitrates are removed from the atmosphere by photolysis \citep{Turberg1990} or through oxidation by hydroxyl radical ($OH$) \citep{Talukdar1997}.
+Photolysis is the dominant loss mechanism for the compounds with short carbon chains and becomes more important with decreasing carbon number.
+Reaction with $OH$ follows the opposite trend and is most important for the pentyl nitrates and higher.
 
@@
 \subsection{Motivation}
-lalalaep
+\citep{Sommariva2008}
+[Typically, chemical mechanisms do not contain
+enough information about the various formation routes of secondary organic compounds, leading to the underestimation of their concentrations, with possible consequences on the estimates of ozone production in remote regions, or to ineffective ozone-control policies.
+In this work, a quasi-explicit chemical model based upon the Master Chemical Mechanism (MCM) has been used to study in detail the photochemistry of a number of alkyl nitrates and determine quantitatively their formation processes.
+This kind of study shows the potential of using mechanisms with a high level of chemical detail to study the formation of pollutants in the atmosphere (Madronich, 2006).]
 
@@
 
+\citep{Sommariva2008}
+[Alkyl nitrates typically constitute only a small
+fraction of total NOy <10\%, even though they can be a much larger component in the equatorial marine boundary layer (Blake et al., 2003; Simpson et al., 2006 and references therein).
+Due to their long lifetime (41 week) alkyl nitrates can transport NOx to remote regions, with potential impacts on regional O3 formation (e.g., Murphy et al., 2006; Simpson et al., 2006)]
+
+\subsubsection*{Oceanic sources}
+\citep{Newland2013}
+[An oceanic source for C1 – C3 alkyl nitrates was first postulated by Atlas et al.
+(1993) during the SAGA campaign in which it was observed that elevated mixing ratios of methyl and ethyl nitrates correlated well with equatorial regions of high productivity.
+This has since been confirmed by field measurements (e.g.
+Chuck et al., 2002; Blake et al., 2003; Dahl et al., 2005).
+Chuck et al.
+(2002) estimate an annual emission flux from equatorial Atlantic and Pacific Ocean regions of 230 and 68 Gg yr-1 for methyl and ethyl nitrate respectively.
+The oceanic source of more or equal C4 alkyl nitrates is insignificant compared to anthropogenic sources (Blake et al., 2003).
+The production mechanism of the alkyl nitrates is thought to be by the reaction ROO + NO in the aqueous phase with the peroxy radical coming from the photolysis of coloured dissolved organic matter (CDOM) and the NO coming from the photolysis of nitrite (Dahl et al., 2003).
+Direct emission from algae (Chuck et al., 2002) or from bacteria at depth (Dahl et al., 2007) has also been postulated.]
+
+\citep{Atlas1993}
+Alkyl Nitrates, Nonmethane Hydrocarbons, and Halocarbon Gases Over the Equatorial Pacific Ocean During Saga 3
+February March 1990
+The cruise covered a region of the equatorial Pacific Ocean from 15N to 10S latitude and 144 to 165W longitude.
+On this cruise we collected samples for the measurement of alkyl nitrates (RONO2) (C2-C5), nonmethane hydrocarbons (NMHC) and several halocarbon gases.
+Results:
+- Alkyl nitrates were found at levels higher than predicted from steady state calculations based on measured mixing ratios of hydrocarbons and NO.
+The measured levels of RONO2 suggest long-range transport as one mechanism contributing to elevated concentrations of alkyl nitrates in the remote troposphere.
+However, the distributions of C2 and C3 alkyl nitrates over the equator were  similar to the organobromine gases.
+This distribution suggests
+a possible oceanic source for alkyl nitrates to the atmosphere.
+Model: 1D 41 constituents incl. 3 ANs
+
+\citep{Chuck2002} September, October
+Measurements of methyl and ethyl nitrate in seawater and air samples along two Atlantic Ocean transects provide the FIRST direct evidence for an oceanic source of these compounds.
+A simple box model calculation suggests that the equatorial source could be a major component of the local atmospheric alkyl nitrate budget.
+In the remote atmosphere, the concentration of methyl nitrate (MeONO2) is expected to be low because of low concentrations of NOx and a branching ratio of less than 1\% for the formation of MeONO2
+from the reaction of alkylperoxy radicals with NO (8).
+However, unexpect- edly high levels of RONO2 in marine air and observations of a strong positive correlation between ethyl and propyl nitrate and the marine biogenically produced halocarbon, bromoform, in the marine boundary layer over the equatorial Pacific Ocean led Atlas et al.
+to propose an oceanic source for alkyl nitrates in the atmo- sphere (9).
+Subsequent atmospheric sampling campaigns, mostly conducted over the remote Pacific Ocean, have provided yet more evidence to support the hypothesis of an oceanic source (2, 10), with the contribution of alkyl nitrates to the NOy
+sum reaching 80\% in equatorial and high-latitude regions over the South Pacific (Blake, 1999).
+Although there have been no pub- lished measurements confirming the oceanic emission of these species, other studies have also invoked a seawater source for methyl, eth- yl, and propyl nitrates (11, 12).
+The supersaturation of surface waters over large areas of the Atlantic provides the first direct evidence for an oceanic source of MeONO2 and EtONO2.
+dominant source region is found between the equator and 40S.
+distinct mechanisms may be respon- sible for the production of methyl and ethyl nitrate, with different water types supporting different mechanisms.
+Methyl nitrate generally showed profiles more typically associated with algal production, with decreasing concentrations be- low the chlorophyll maximum.
+In contrast, in waters more strongly affected by continental influences, such as increased loading of dis- solved organic matter, the alkyl nitrates showed maximum concentrations in surface waters, which may suggest that photochemical production is important in shelf and coastal regions.
+
+\citep{Blake2003} oceanic+urban!!!
+Observations: concentration distributions of C1-C4 alkyl nitrates observed during the NASA airborne campaigns Pacific Exploratory Mission (PEM) -Tropics A (September– October 1996) and PEM-Tropics B (March–April 1999).
+The total geographic range for PEM-Tropics A was 45?N–72?S latitude and 153?E–75?W longitude, and for PEM- Tropics B was 40?N–36?S latitude and 149?E–75?W longitude.
+We observed high methyl nitrate (MeONO2,CH3ONO2) mixing ratios (approximately 50 pptv) at low altitudes in a latitude band between 8?Nto13?S stretching across the equatorial Pacific, illustrating the oceanic source of MeONO2.
+This source may be associated with the high-nutrient, low-chlorophyll character of equatorial Pacific waters.
+We discuss MeONO2 and ethyl nitrate (EtONO2,C2H5ONO2), whose abundance is dominated by equatorial oceanic sources, 2-Propyl nitrate (2-PrONO2,2-C3H7ONO2), which has significant oceanic and northern hemispheric (NH) sources associated with urban/ industrial hydrocarbon emissions, and 2-butyl nitrate (2-BuONO2 2-C4H8ONO2), which has mostly NH (northern hemisphere) sources.
+the ocean as the major source for MeONO2
+Ethyl nitrate has a dominant oceanic source in equatorial regions, but at mid NH latitudes a small additional source is associated with urban/industrial emissions.
+
+\subsubsection*{Biomass burning source}
+\citep{Newland2013}
+[Alkyl nitrates also have a biomass burning source (Friedli et al., 2001; Simpson et al., 2002).
+Simpson et al.
+(2002) report emissions of C1 – C4 alkyl nitrates from savanna fires in Australia, with methyl nitrate being primarily emitted during the ’flaming’ stage and ethyl – butyl nitrates mainly emitted during the ‘smouldering’ stage.
+They estimate global emissions from this source to be on the order of 8 Gg yr-1]
+
+\subsubsection*{Lifetime}
+\citep{Newland2013}
+[Species within this group provide a transport reservoir for the short lived NOx compounds due to their relatively longer lifetimes.
+This allows NOx to be transported from polluted areas to remote background sites.
+Alkyl nitrates have been reported as making up only a few percent of the NOy ('odd' nitrogen) budget in urban and rural areas (Shepson et al., 1993; Thornberry et al., 2001), however due to their relatively long lifetimes and their continued production during transport of polluted air masses, as alkanes are converted to alkyl nitrates (e.g.
+Bertman et al., 1995), they can comprise up to tens of percent of the NOy budget in polar regions and in the remote marine troposphere (Talbot et al., 2000).
+Therefore alkyl nitrates could potentially be a substantial source of NOx in remote background regions contributing to tropospheric ozone formation in these regions.]
+\citep{Newland2013}
+[The lifetimes of the alkyl nitrates vary with latitude and season due to the seasonality of the OH sink and photolysis.
+Clemitshaw et al.
+(1997) report a range of lifetimes with respect to OH and to photolysis.
+Photolysis lifetimes vary from about ten days for butyl and pentyl nitrate in the summer to 1 – 2 years in the winter at 65N with lifetimes also decreasing with altitude.
+The lifetimes due to reaction with OH (tauOH) at 65N range from tens of years in
+winter to about 10 days in the summer. tauOH increases with altitude as the OH concentration decreases. At the equator tauOH is similar all year round, on the order of a few days.]
+\citep{Chuck2002}
+Use of the mean sea-to-air fluxes for 0 to 40S (Table 1) and measured atmospheric concentrations results in a lifetime of be- tween 4.5 and 25 days for MeONO2 and between 5 and 10 days for EtONO2 (22).
+Literature estimates of the atmospheric life- time of MeONO2, based on known atmospheric destruction processes, are poorly con- strained, ranging from 6 to 29 days (23, 24).
+For EtONO2
+, the atmospheric lifetime with respect to photodissociation at the equator and at 40S has been estimated to be 7 and 12 days, respectively, for October (25).
+
+\subsubsection*{Global distribution}
+\citep{Newland2013}
+[There have been a number of studies of the latitudinal and seasonal variations of alkyl nitrates in marine air (Blake et al., 2003), in the Arctic (Swanson et al., 2003) and in the Antarctic (Beyersdorf et al., 2010).
+Blake et al.
+(2003) report extensive measurements of C1 – C4 alkyl nitrates from the Pacific between 45N and 72S during the PEM-Tropics A and B projects.
+This study observed an equatorial band between about 8N and 13S in which the oceans are a source of alkyl nitrates.
+The highest mean methyl and ethyl nitrate measurements were found in this region.
+The highest mixing ratios of 2-propyl nitrate were found in the most northerly regions of the study (i.e.
+40-45N) but it also showed significantly elevated mixing ratios in the same equatorial band as the C1 – C2 nitrates.
+The high mixing ratios in the northern mid-latitudes are explained by production from anthropogenic polluted outflow.
+2-butyl nitrate showed almost no enhancement in equatorial regions and again had the highest mixing ratios in the most northerly regions of the study.
+Swanson et al.
+(2003) and Beyersdorf et al.
+(2010) observe a strong seasonality of alkyl nitrate mixing ratios in the Arctic and Antarctic respectively.
+Polar alkyl nitrate mixing ratios show a similar seasonality to the alkanes, with a spring-time peak just before polar sunrise and lows in the summer time.
+Winter time mixing ratios of 2-propyl and 2-butyl nitrate are about 10 ppt, while methyl and ethyl nitrate are about 5 – 6 ppt.
+In the summer 2-propyl and 2-butyl nitrate mixing ratios fall close to zero, while methyl and ethyl nitrate mixing ratios fall to about 3 ppt.
+The amplitude of the seasonal cycle reflects the reaction rate with OH, i.e.
+it increases with increasing carbon number.]
+
+\subsection{Mean AN concentrations}
 mean ANs concentrations: \citep{Reeves2007}, \citep{Roberts1998}
+\citep{Atherton1989}
+Alkyl nitrate concentrations of 0.8-37 pptv (2-butyl and 2- and 3-pentyl nitrates) in remote marine environments [Atlas, 1988] and approximately 2 ppbv in urban environments [Hanst et al., 1982] have been measured.
+\citep{Chuck2002}
+The atmospheric mole fraction of MeONO2 ranged from 1.2 to 43 pptv (Fig.
+2,Aand B) and 0.8 to 10 pptv for EtONO2 (Fig.
+2, D and E).
+These values lie within the range of concentrations reported in the literature for MeONO2 (0.2 to 50 pptv) and EtONO2 (0.3 to 12.2 pptv) in the remote marine troposphere (2, 8, 9).
+\citep{Blake2003}
+Pacific Ocean methyl nitrate 50 pptv
+
+\subsubsection{Comparison methodology}
+\citep{Sommariva2008}
+[The ratios between the alkyl nitrates and their
+main alkane precursors can also be plotted against one another, for example against 2-butylnitrate/ n-butane as in Fig.
+4.
+In this way the modelled and the measured concentrations can be compared without taking into account their variation with time.
+A similar approach was used in previous studies to investigate the sources of alkyl nitrates and the photochemical processing of polluted air masses (e.g., Bertman et al., 1995; Flocke et al., 1998; Roberts et al., 1998; Stroud et al., 2001; Simpson et al., 2003; Reeves et al., 2007).
+These studies compared plots of the measured ratios against 2-butylnitrate/n-butane to the ratios calcu- lated with a simple sequential reaction model which assumed a single precursors for each alkyl nitrate.
+Usually, the underestimation of the ratios for short- chain compounds (propylnitrates/propane, ethylni- trate/ethane) was interpreted as an indication of photochemical sources other than the main alkane precursor.]
 
 \subsubsection{Modelling studies}
+A huge number of modelling studies has been done about alkyl nitrates.
+
+\citep{Atherton1989}
+The transport from urban areas can lead to alkyl nitrate concentrations of 4-83 pptv after 120- 240 hours.
+These levels are far higher than those possible from marine in-situ production.
+Additionally, alkyl nitrates may be more abundant than PAN in remote marine environments.
+Model: box
+Chemistry: explicit chemical reaction mechanism
+[Carter et al., 1986.
+A variety of alkyl nitrates, including methyl-, butyl-, pentyl-, hexyl-, heptyl-, octyl-, and nonyl-nitrates are allowed to form from fourteen different alkyl peroxy radicals.
++ ocean emission sources fluxes of c
+for ethene, ethane, propene, and propane, NO, NO2
+Initial conditions: clear sky conditions, a quantum yield of 1, solar zenith angle of 45 degrees, and total ozone column of 8*10 cm-2.
+Daytime+nigthtime, invertion height, dilution.
+298 K.
+No deposition.
+Scenarios: Two, MLOW and MHIGH, represent the marine in-situ production of alkyl nitrates.
+The total hydrocarbon concentration was 17.6 ppbC.
+MLOW had NO = 4 pptv and NO2 = 4.6 pptv based on measurements over the Pacific Ocean.
+MHIGH had NO = 46 pptv, NO2 = 54 pptv, and PAN = 50 pptv.
+The two other scenarios, UHIGH and ULOW, simulate the transport of species from urban to remote environments.
+UHIGH contained 2150 ppbC and 140 ppbv NOx, and ULOW contained 270 ppbC and 18 ppbv NOx, giving each a HC/NOx ratio of 15.
+Simulation time: 12 hours MLOW and MHIGH, 240 hours UHIGH and ULOW.
+
+Results:
+- The total alkyl nitrate concentration present after 12 hours is less than 0.4 pptv for MHIGH and less than 0.1 pptv for MLOW.
+These levels are much lower than observations, indicating that marine in-situ production can account for only a fraction of the total marine alkyl nitrate concentration.
+- IHIGH The total alkyl nitrate concentration peaks during the first 6 hours and then drops to 3.1 ppbv after 12 hours.
+This value is com- parable with the 2 ppbv alkyl nitrate concentration de- termined in Los Angeles [Hanst et al., 1982]
+- The model predicts alkyl nitrate concentrations of 21-83 pptv for UHIGH and 4-12 pptv for ULOW for the hours 120-240.
+- General: The model predictions of 12-83 pptv (af- ter 120 hours) and 4-21 pptv (after 240 hours) are of the same order as observations.
+
+\citep{Sommariva2008}
+Validation: typical urban plume in the North-East of the United States, measurements taken during three flights of the NOAA WP-3D aircraft, which sampled a plume from the New York City area during the New England Air Quality Study (NEAQS) 2004 campaign
+- Seven alkyl nitrates (ethyl- nitrate, i-propylnitrate, n-propylnitrate, 2-butylni- trate, 2-pentylnitrate, 3-pentylnitrate, 3-methyl-2- butylnitrate) were measured during these flights together with a full suite of VOCs and other chemical parameters, such as CO, O3,NOx, HNO3 and PANs (Neuman et al., 2006).
+- The model results have been qualitatively compared to measurements of alkyl nitrates taken on 20–22 July 2004, when the NOAA WP-3D aircraft made three flights aimed to inter- cept the New York City plume.
+Model: subset! of MCM v3.1 (The MCM is one of the most detailed chemical mechanism for tropospheric chemistry available)
+- to study the formation processes and photochemical evolution of alkyl nitrates
+Chemistry: oxidation mechanisms of 64 VOCs (50 hydrocar-
+bons and 14 oxygenates) plus CH4 and CO, The inorganic chemistry mechanism was taken from the IUPAC recommendation (Atkinson et al., 2003).
+A dimethyl sulphide (DMS) mechanism (Sommariva et al., 2004) was also included.
+The photolysis rates were calculated using the MCM parametrization (Saunders et al., 2003) for end of July 2004 at the latitude and longitude of Boston, MA.
+only gas-phase!!!
+Measurements made during NEAQS 2004 showed that there was very little removal of, for example,HNO3 by dry deposition in plumes decoupled from the marine boundary layer during transport from New York City (NYC) to the North Atlantic (Neuman et al., 2006).
+Therefore, removal by dry deposition was neglected in the model.
+Initialisation: measurements taken in the Boston metropolitan area during the 2002 and 2004 NEAQS campaigns.
+The initial concentrations of VOCs were determined using emission ratios; CH4 and O3 were set to typical values for urban conditions (1800 and 90 ppb) (Table 1).
+The model has been initialized using measurements and VOCs emission ratios from the 2002 and 2004 New England Air Quality Studies.
+Simulation time: 5 days.
+Scenarios: 2: A (higher OH, good with obs, reference), B
+Time: 20-22 July.
+
+Results:
+- While long-chain (C5) alkyl nitrates are produced for 90\% or more from the oxidation of a single parent alkane, short-chain <C4 alkyl nitrates can be formed from several precursors.
+The relative importance of each production route has been quantitatively determined thanks to the high level of chemical detail provided by the MCM.
+These secondary routes
+to the formation of alkyl nitrates include the oxidation of longer-chain alkanes and oxygenated intermediates, like carbonyls, peroxides and carboxylic acids.
+- The best agreement is in the case of 3-methyl-2-butylnitrate and the discrepancy between the modelled and the measured ratios clearly increases with decreasing carbon number .
+- RH/AN, Except in the case of ethylnitrate, the model can reproduce the measured rate of increase in the alkyl nitrate to alkane ratios fairly well, which suggests that the kinetic parameters used in the MCMmodel are overall correct.
+- By considering together Figs. 3 and 4 it can be reasonably concluded that the chemistry of alkyl nitrates is well represented in the MCM model.
+- Using the explicit description of the chemistry provided by the MCM, it has been possible to determine quantitatively the photochemical formation and the evolution of alkyl nitrates and the relative importance of their different formation pathways.
+The results are summarized in Table 2.
+Large alkyl nitrates are produced almost exclusively from the reaction of a single parent
+alkane.
+Smaller <=C4 alkyl nitrates also have other significant precursors, like longer-chain alkanes and alkenes.
+For n-propylnitrate and ethylnitrate, pro- panal and butanal, which are largely emitted, are important sources within the first 24 h.
+For all alkyl nitrates, the role of organic peroxides and car- boxylic acids formed by the reactions of peroxy radicals with HO2 at low [NOx] increases with time.
+These species contribute to sustain the production of alkyl nitrates far from the emission points of the parent VOCs.
+The relative importance of the different pathways
+and how they change with time are dependent on the initial conditions of the model. In a scenario with lower O3 and CO and higher NOx the
+importance of the direct routes (alkane ?OH) is typically lower during the first two–three days, and the contributions of peroxides and carboxylic acids become significant later and are on average less important.
+These results suggest that secondary chemistry in the formation of the alkyl nitrates, particularly oC4, cannot be neglected and should be included in the chemical models, due to its potential impact on ozone-controlling strategies.
+
+\citep{Newland2013}
+Model1: 2-D atmospheric model
+- to study the formation, loss and transport of the alkyl nitrates to the Arctic
+- to derive northern hemisphere emissions of the six alkanes.
+Chemistry:
+- (with fixed NOx and OH field taken from full chemistry run with model by Hough and Johnson (1990))
+Validation of ANs: comparison with measurements of alkyl nitrates in the Arctic.
+Performance: recreate both the measured seasonality of the alkyl nitrate, and the peak mixing ratio, reasonably well.
+Simulation time: 6 days?
+
+Model2: box
+Chemistry:
+- alkanesmethane, ethane, propane, butane and pentane and ANs explicitly described (RO+NO2 too)
+
+Model3: firn model
+- to derive atmospheric histories of six alkanes and six alkyl nitrates from measurements in firn air from the NEEM site, Greenland (methyl, ethyl, 2-propyl, 2-butyl, 2+3-pentyl and 3- methyl-2-butyl nitrate), 1950-2008.
+
+Results:
+- There are two ways in which NOx mixing ratios can limit alkyl nitrate production: either through air masses being NO limited with respect to the formation of a given alkyl nitrate or by the NOx in an air mass being 'used up' (track back to Reeves)
+-- Alkyl nitrate production becomes NOx saturated above a given NO mixing ratio. This limit increases with increasing carbon number.
+-- The time taken for an air mass to 'use up' all the initial NOx is dependent mainly on the initial NOx mixing ratio.
+- At high [NO2] (> 27 ppb), methyl nitrate production is dominated by the reaction of the alkoxy radical with NO2
+- The regions that contribute most to modeled alkane and alkyl nitrate mixing ratios in the Arctic are Europe and Russia
+- The importance of the RO + NO2 pathway compared to RO2 + NO is entirely dependent on the branching ratio (alpha3) of the RO2 + NO reaction pathway.
+The smaller alpha3 is, the more important the RO + NO2 pathway is.
+-- At 10 ppb of NO2 the RO + NO2 pathway accounts for 30 – 60\% of methyl nitrate production.
+At 100 ppb it accounts for 85 – 95\% of methyl nitrate production and almost 20\% of ethyl nitrate production.
+- The rate of alkyl nitrate production in a polluted air mass is proportional to the OH concentration in that air mass.
+
+\subsubsection*{Other}
 
@@
 
-!! from \citep{Finlayson-Pitts2000}
-The relationship between the peroxy radical concentration and the ozone photolysis rate constant for these higher NO conditions can be again approximated using steady-state analysis [..].
-While OH is recycled in its reactions with CO and CH4 via HO2, it is pernanetly removed at higher NOx concentrations by the reaction of OH with NO2, forming nitric acid:
-OH + NO2 = M = HONO2 (113).
-This reaction is primarily a daytime reaction because most OH sources are photolytic in nature. As a result, the NO2 reaction with OH competes!!!! with NO2 photolysis (page 290/993)!!
-         Rate coeff (OH+NO2 = HNO3)  J(NO2 = NO+O)
-Book     2.47e-30          7.00e-3
-My model 1.15e-11          4.39e-03
-Thus, the reaction with OH is not usually a dominant loss process for NO2, but it is still sufficiently fact to from significant amounts of HNO3 during the day, particularly in polluted regions with relatively large NO2 concentrations.
-Nitric acid undergoes both wet and dry deposition rapidly and can be neutralized by ammonia, the major gaseous base found in the atmosphere.
-
-Alkyl nitrates ($RONO_2$) are organic tropospheric trace gases.
 They are produced photochemically through the oxidation of their parent hydrocarbons \citep{Roberts1990} or emitted directly from equatorial oceans \citep{Blake2003} and biomass burning \citep{Simpson2002}.
@@
 Since the alkyl nitrates have atmospheric lifetimes ranging from several days to about a month, they can redistribute nitrogen oxides presented over polluted continental regions to clean remote marine environments, and therefore are considered as one of the tracers for determining the anthropogenic influence \citep{Atherton1989, Reeves2007, Worton2005}.
-
-Alkyl nitrates are removed from the atmosphere by photolysis \citep{Turberg1990} or through oxidation by hydroxyl radical ($OH$) \citep{Talukdar1997}.
-Photolysis is the dominant loss mechanism for the compounds with short carbon chains and becomes more important with decreasing carbon number.
-Reaction with $OH$ follows the opposite trend and is most important for the pentyl nitrates and higher.
 
@@
 Conclusions constitute section \ref{sec:conclusion}.
-
-!
-Since the most alkoxy radicals ($RO$) react rapidly with $O_2$ under atmospheric conditions and most large (>=C4) alkoxy radicals also rapidly decompose or isomerize, alkyl nitrate formation from reaction RO+NO2 = RONO2 is relatively unimportant for most organics under atmospheric conditions while observations showed that the reaction RO2+NO = RONO2 gives a higher yield (from \citep{Atkinson2000}, actually from 1982).
-
-! Since one of the main mechanisms of the alkyl nitrate production (>C3!!) is the gas gas phase oxidation of alkanes by OH in the presence of NOx (\citep{Newland2013}).
 
@@
 
+!! C2H5NO3 ethylnitrate, IC3H7NO3 i-propylnitrate, NC3H7NO3  n-propylnitrate, SC4H9NO3 2-butylnitrate, PEBNO3 2-pentylnitrate;PECNO3 3-pentylnitrate, IPEBNO3 3-methyl-2-butylnitrate
+
 \begin{table} % RH and ANs
@@
 
-The simulations that compared ... provided insight into which reactions... under changing $NO_x$ mixing ratios (Table ...).
-
 !

### 2015-08-02 MSc_diss.tex
Rewriting literature review

@@
 \subsection{Ozone, alkyl nitrates and carbon bonds}
+The question that is always important for ozone-policies is the ozone production efficiency.
+Historically people noticed that with decreasing NOx emission the ozone production?
+concentration?
+drops, so the way of calculating the efficiency of ozone production due to drop in NOx emissions was proposed in the following way:
+
+In addition to a traditional way of calculating ozone production efficiency, $\epsilon$, usually defined as the 
+number of $O_3$ molecules produced per molecule of $NO_x$ consumed (Jacob, 1999): 
+\begin{equation}
+\epsilon = \dfrac{P_{O_3}}{L_{NO_x}}
+\end{equation}
+where $P_{O_3}$ – the $O_3$ production rate, and $L_{NO_x}$ – the $NO_x$ loss rate, a novel method of calculation of $\epsilon$ is applied.
+
+This formula is not very useful if we want to estimate the impact of AN chemistry on O3, but there is another method, proposed by Professor Matthew Evans (University of York) suggests that the ozone production efficiency can be estimated via calculation of the number of carbon bonds emitted \citep{Evans2014} (personal communication).
+For example, if a given set of reactions contains only methane, $CH_4$, as 'a carbon bond reservoir', it means that this system has 4 carbon bonds (4 C-H).
+If  ethene, $C_2H_4$, is considered the system has 5 bonds (4 C-H bonds and 1 C=C) \citep{Edwards2013}.
+Assuming that every carbon bond is broken chemically and every peroxy radical ($RO_2$) reacts with an NO the overall efficiency of carbon bonds conversion to ozone can be calculated as:
+\begin{equation}
+\epsilon = \dfrac{P_{O_3}}{E_{C_{bons}}} = \chi*nu
+\end{equation}
+where $P_{O_3}$ – the O3 production rate (mol hr-1), $E_{C_{bons}}$  – the carbon bond initial abundance (mol hr-1), $\chi$ – ratio between rate of carbon bonds loss to make $RO_2$ and rate of all carbon bonds loss, $nu$ – ratio between rate of $RO_2$ loss in reaction with NO and rate of all $RO_2$ loss.
+In other words, chemical loss of 1 carbon bond produces 1 peroxy radical, and photolysis loss of 1 carbon bond produces 0 or 2 peroxy.
+In other words,  it means that essentially we need to calculate the overall number of peroxy radicals remained in the chemical system by the end of simulation.
+
+The same kind of approach is used in Common Representative Intermediates (CRI) mechanism for VOC degradation \citep{Jenkin2002, Jenkin2008}.
+A reduced mechanism describing ozone formation from the tropospheric degradation of methane and 115 emitted non-methane hydrocarbons and oxygenated volatile organic compounds has been developed, using the Master Chemical Mechanism version 3.1 (MCM v3.1) as a reference benchmark.
+The Common Representative Intermediates mechanism version 2 (CRI v2) has been built up on a compound-by-compound basis, with the performance of its chemistry optimised for each compound in turn by comparison with that of MCM v3.1, using a series of five-day box model simulations.
+The resultant mech- anism contains 1183 reactions of 434 chemical species, i.e., ca.
+10\% of the number of reactions and species to degrade the same set of VOCs in MCM v3.1.
+Similarly to CRI v1 (Jenkin et al., 2002a), a key assumption in the CRI v2 construction methodology is that the potential for ozone formation from a given volatile organic compound (VOC) is related to the number of reactive (i.e., C-C and C-H) bonds it contains.
+This index allows a series of generic intermediates to be defined, with each being used as a 'common representative' for a large set of species possessing the same index, as formed in detailed mechanisms such as the MCM.
+On this basis, the number of reactive (C-H and C-C)
+bonds in a closed-shell intermediate can be used as an index to represent the number of NO-to-NO2 conversions which can potentially result from its subsequent OH-initiated and NOx-catalysed degradation.
 
@@
 Due to their long lifetime (41 week) alkyl nitrates can transport NOx to remote regions, with potential impacts on regional O3 formation (e.g., Murphy et al., 2006; Simpson et al., 2006)]
+
+\subsubsection*{Continent}
+\citep{Day2003}
+We describe measurements of NO2, total peroxy nitrates (?PNs), total alkyl nitrates (?ANs), and HNO3 using thermal dissociation followed by laser-induced fluorescence detection of NO2 at three continental locations.
+At a rural site in California, measurements over a full annual cycle show that ?ANs are routinely 10–20\% of NOy.
+rural site, at a suburban site in California and an urban site in Houston, Texas [rural (UC-BFRS), suburban (Granite Bay) and urban (La Porte)]
+SumANs a more broad term than usual.
+?AN mixing ratios at the UC-BFRS exhibit a strong seasonal cycle with a minimum in winter and a maximum in summer.
+The seasonal transition is sharp.
+Concentrations of ?ANs rise rapidly from 150 ppt at the end of April to 500 ppt by early June.
+We also show that ?ANs are a large part, if not all, of the ‘‘missing NOy’’.
 
@@
 Bertman et al., 1995), they can comprise up to tens of percent of the NOy budget in polar regions and in the remote marine troposphere (Talbot et al., 2000).
-Therefore alkyl nitrates could potentially be a substantial source of NOx in remote background regions contributing to tropospheric ozone formation in these regions.]
-\citep{Newland2013}
-[The lifetimes of the alkyl nitrates vary with latitude and season due to the seasonality of the OH sink and photolysis.
+Therefore alkyl nitrates could potentially be a substantial source of NOx in remote background regions contributing to tropospheric ozone formation in these regions.
+The lifetimes of the alkyl nitrates vary with latitude and season due to the seasonality of the OH sink and photolysis.
 Clemitshaw et al.
@@
 winter to about 10 days in the summer. tauOH increases with altitude as the OH concentration decreases. At the equator tauOH is similar all year round, on the order of a few days.]
+
 \citep{Chuck2002}
@@
 A similar approach was used in previous studies to investigate the sources of alkyl nitrates and the photochemical processing of polluted air masses (e.g., Bertman et al., 1995; Flocke et al., 1998; Roberts et al., 1998; Stroud et al., 2001; Simpson et al., 2003; Reeves et al., 2007).
-These studies compared plots of the measured ratios against 2-butylnitrate/n-butane to the ratios calcu- lated with a simple sequential reaction model which assumed a single precursors for each alkyl nitrate.
+These studies compared plots of the measured ratios against 2-butylnitrate/n-butane to the ratios calculated with a simple sequential reaction model which assumed a single precursors for each alkyl nitrate.
 Usually, the underestimation of the ratios for short- chain compounds (propylnitrates/propane, ethylni- trate/ethane) was interpreted as an indication of photochemical sources other than the main alkane precursor.]
+
+\citep{Day2003}
+We also show that sumANs are a large part, if not all, of the 'missing NOy' reported in many prior experiments and emphasize that the ratios of sumANs/NOz and of O3/sumANs are especially useful for evaluating chemical models and comparing observations at different sites.
 
@@
 A huge number of modelling studies has been done about alkyl nitrates.
+
+\citep{Neu2008}
+Observations have revealed widespread emissions of alkyl nitrates from the tropical and Southern Oceans.
+We present the first chemical transport model simulations to examine the global impact of these emissions.
+Matching observed atmospheric abundances, we derive a total oceanic flux of methyl nitrate (MeONO2) and ethyl nitrate (EtONO2) equivalent to 0.35 Tg of N per year, which contributes as much as 1 DU to the tropospheric ozone column in the Western Pacific and is responsible for about 3\% of the global oxidative capacity of the troposphere.
+Atmospheric observations indicate that low molecular weight (C1-C4) alkyl nitrates are ubiquitous over the tropical and Southern Oceans [Atlas et al., 1993; Blake et al., 1999, 2003] and can be a significant component (20–80\%) of the total reactive nitrogen in remote marine air [Jones et al., 1999; Talbot et al., 2000].
+Recent ship-borne measurements show supersaturation of alkyl nitrates in tropical surface waters, confirming an oceanic source [Chuck et al., 2002; Moore and Blough, 2002; Dahl et al., 2005].
+This work presents the first 3-D chemical transport model (CTM) simulations to examine the role of oceanic alkyl nitrates in global chemistry, using observational constraints.
+In this study, we incorporate MeONO2 and EtONO2 in the chemical mechanism of the UCI CTM [Carver et al., 1997] and compare their simulated abundances to observations to derive new estimates of their oceanic fluxes.
+We assess the relative importance of oceanic alkyl nitrates as a natural source of NOx in remote regions by evaluating their impact on tropospheric ozone (O3) and the oxidative ca- pacity of the atmosphere.
+The UCI CTM [Wild and Prather, 2000; Hsu et al.,2005] includes 30 chemical species in the troposphere.
+We define a control run with no alkyl nitrate emis-
+sions and compare it to two simulations in which we impose a constant, spatially uniform flux of MeONO2 and EtONO2 from: 1) the tropical oceans (10?S–10?N) and 2) the Southern Ocean (south of 45?S).
+Results: impact on NOx and O3 chemistry
+Oceanic alkyl nitrates cause measurable increases in
+both NOx and O3 in marine air.
+Their largest impact is seen over the equatorial Western Pacific (Figure 3).
+There is a persistent minimum in both NOx and O3 (Figure 3a) in this region, where the circulation isolates the marine boundary layer, resulting in sustained photochemical loss of O3 and a deep minimum throughout the tropospheric O3 column.
+At the same time, the circulation acts to accumulate alkyl nitrates in theWestern Pacific (see Figure 2), and they drive up boundary layer NOx abundances by as much as 250\%.
+This leads to O3 increases of up to 20\% (Figures 3b and 3f), with the largest increases occurring from May to October.
+From December to March (not shown), the relative increases in NOx and O3 are smaller, but are still concen- trated in the Western Pacific even though the alkyl nitrates are spread more evenly across the Pacific during these months.
+The geographic pattern of the alkyl nitrate influence persists into the middle and upper troposphere (Figures 3c-3d and 3g-3h), with reduced relative impact because of the larger lightning NOx source.
+However, the largest zonal mean increases in mixing ratio, 3–8 ppt for NOx and 1 ppb for O3, occur in the tropical upper and middle troposphere, respectively (Figure 4).
+Pacific for example, ignore Southern ocean.
+The O3 produced by NOx from alkyl nitrates fills in
+the minimum in the tropospheric O3 column over the Western Pacific, adding more than 1 DU, with a maximum in October (Figure 3e).
+The global average increase in tropospheric O3 column varies with season from 0.35 to 0.42 DU.!!!!
+WOW!!!
+Oceanic alkyl nitrates are necessary to accurately
+model NOx in the tropical remote marine environment, since they are responsible for 15–20\% of the O3 in the Western Pacific.
+The coincident location of the maximum alkyl nitrate impact and the deep minimum in O3 over the Western Pacific is a fascinating consequence of the convergence gence of the trade winds, but will make it difficult to detect an alkyl nitrate signature in O3.
+[22] At current levels, alkyl nitrates have a modest climate impact.
+From a simple scaling [Wild et al., 2001], the global increase in tropospheric column O3 will have a radiative forcing of about 16 mW/m2.
+In steady state, this will be almost entirely offset by the reduced CH4 lifetime, which produces a roughly equal negative radiative forcing.
+Surprisingly, a comparison to Wild et al.
+[2001] shows that on a per-molecule basis, alkyl nitrates have the largest impact on O3 and CH4 of any NOx source analyzed to date.
+This seems to be a consequence of their long lifetime combined with efficient transport to the tropical upper troposphere.
+With a 2.4\% effect on CH4 lifetime, a 5-fold increase in alkyl nitrate emissions during glacial periods would account for roughly 1/3 of the 300 ppb drop in CH4 seen between interglacials and glacials.
 
@@
 Simulation time: 12 hours MLOW and MHIGH, 240 hours UHIGH and ULOW.
-
 Results:
@@
 Time: 20-22 July.
-
 Results:
@@
 Simulation time: 6 days?
-
 Model2: box
@@
 - alkanesmethane, ethane, propane, butane and pentane and ANs explicitly described (RO+NO2 too)
-
 Model3: firn model
 - to derive atmospheric histories of six alkanes and six alkyl nitrates from measurements in firn air from the NEEM site, Greenland (methyl, ethyl, 2-propyl, 2-butyl, 2+3-pentyl and 3- methyl-2-butyl nitrate), 1950-2008.
-
 Results:

### 2015-08-02 MSc_diss.tex
Rewriting literature review mess

@@
 
+\citep{Neuman2012}
+Ozone production efficiency (OPE), which can be determined observationally from the ratio of O3 to NOx oxidation products, is used regularly as a metric to assess O3 photochemistry [e.g., Trainer et al., 1993; Neuman et al.,2009]
+
 In addition to a traditional way of calculating ozone production efficiency, $\epsilon$, usually defined as the 
@@
 We also show that ?ANs are a large part, if not all, of the ‘‘missing NOy’’.
+
+\citep{Simpson2006} Pearl River Delta
+Mixing ratios of seven C1–C5 alkyl nitrates (RONO2) were measured during a 16-month study (August 2001–December
+2002) at Tai O, a coastal site 30km west of central Hong Kong in the Pearl River Delta, the fastest-growing industrial region in the world.
+The C3–C4 (rather than C1–C2) RONO2 were most abundant throughout the study, showing the importance of photochemical (rather than marine) RONO2 production in the sampled air.
+A lack of methyl nitrate (MeONO2) enhancement during summer, when the prevailing winds are from the ocean, indicates that the South China Sea is not a region of strong RONO2 emissions.
+By contrast, MeONO2 levels during pollution episodes (up to 25 parts per trillion by volume (pptv)) were the highest that our group has recorded during urban photochemical RONO2 production, as opposed to marine emissions or biomass burning.
+The highest summed RONO2 level of the study (204 pptv) was measured in the afternoon of 7 November 2002, during an intense pollution episode that captured the highest ozone (O3) level ever recorded in Hong Kong (203 ppbv).
+During pollution episodes, the average ratio of O3 to summed RONO2 was
+roughly 1000:1 in freshly polluted air (ethyne/CO?3–5 pptv/ppbv) and 500:1 in very freshly polluted air (ethyne/ CO?6–8 pptv/ppbv).
+Ozone and RONO2 share a common photochemical source, and their good correlation in pollution plumes shows that RONO2 can be used as a tracer of photochemical O3 production.
+Even MeONO2 showed similar
+diurnal variations as the C2–C5 RONO2, indicating a strong photochemical source despite its very slow photochemical production from methane oxidation.
+The decomposition of longer-chain alkoxy radicals also does not explain the high MeONO2 levels, and rough calculations show that methoxy radical reaction with NO2 appears to be a viable alternate pathway for MeONO2 production in polluted atmospheres, though further measurements and modeling are required to confirm this mechanism.
+\citep{Comment}
+It has generally been assumed that reaction (3) dominates over reaction (2) and that production of CH3ONO2 would be small.
+Simpson et al.
+(2006) suggest that model studies would be useful to test their theory.
+In this short comment on their paper, some model simulations are presented that confirm their hypothesis.The model simulations have been carried out using a box model running the CRI v1.0 mechanism (Jenkin et al., 2002), which is a reduced scheme of the Master Chemical Mechanism (Jenkin et al., 1997).
+The box model was integrated using emissions of NOx and VOC taken from the UK National emissions inventory and would be repre- sentative therefore of European rather than Asian emission profiles but are still instructive.
+The model is designed to simulate regional photochemical episodes and runs during European summer time (mid June) under moderately high temperature (approximately 25 1C) and relative humidity (70\%).
+The concentration of CH3ONO2 from the two routes, reaction (1b) and reaction (2), have been compared as a function of NOx and are summarised in Fig.
+1.
+It is clear that above about 10 ppb of NO2 channel (2) becomes important as a source of CH3ONO2 and at around 35 ppb starts to become the dominant channel in the simulations performed.
+
+\citep{Stroud2001} plane+colorado, plane+North Atlantic
+During both campaigns alkyl nitrates were measured aboard the NOAA Orion WP-3 aircraft with a vertical range from the boundary layer to the midtroposphere (7 km) [Hiibler et al.,1998].
+During STERAO 1996 the aircraft was based at Buckley Air Force Base in Denver, Colorado.
+Flights were conducted from June 13, 1996, to July 16, 1996, over the northeastern plains of Colorado to study how thunderstorms affect the chemical composition of the atmosphere.
+Alkyl nitrates were measured over the northeastern plains of Colorado during the Stratospheric Tropospheric Exchange: Radiation, Aerosols and Ozone (STERAO) 1996 campaign and over the North Atlantic during the North Atlantic Regional Experiment (NARE) 1997.
+A comparison of alkyl nitrate to hydrocarbon ratios with a sequential reaction scheme model suggests our understanding of C4-Cs alkyl nitrate photochemistry is adequate to reproduce the observations.
+Air masses with larger ratios of alkyl nitrate to hydrocarbons were observed during NARE 1997 than STERAO 1996, consistent with the NARE 1997 measurements being in the more remote North Atlantic.
+These results suggest that information related to the photochemical age of an air  mass can be derived from alkyl nitrate to hydrocarbon ratios.
+
+\citep{Worton2010} TORCH
+Alkyl nitrates (C1eC5) were measured at two sites (near urban and rural) in southeast England during the Tropospheric Organic Chemistry Experiment (TORCH).
+Methyl nitrate was the dominant species during both campaigns accounting for on average about one third of the total measured alkyl nitrates.
+High mixing ratios (>50 pptv) and variability of methyl nitrate were observed at the near urban site (TORCH1) that were not seen at the rural site (TORCH2) and which could not be explained by local photochemical production or direct emissions.
+The diurnal variation of methyl nitrate during TORCH1 showed a morning maximum that would be consistent with nighttime chemistry followed by transport to the surface by boundary layer dynamics.
+Similarly, elevated morning mixing ratios were also observed during TORCH2 although the magnitudes were much smaller.
+As a result, methyl nitrate could represent a tracer for nighttime chemistry seen at the ground the following day.
+At both campaigns, the dominant source of short chain alkyl nitrates and carbonyl precursor radicals (?C4) were from decomposition of larger compounds.
+The magnitude of the source increased with decreasing carbon number consistent with increasing total precursor abundance.
+Non-photochemical emissions of acetaldehyde and acetone could not be accounted for by automobile exhaust emissions alone and indicated that other direct sources are likely important in this environment.
+Alkyl nitrate measurements were made at two sites (near urban
+and rural) as part of the TORCH experiment. Observed mixing ratios
+were higher (factor of ?2) during TORCH1, especially during the identified heatwave period.
+Methyl nitrate accounted for w30\% of the total measured C1eC5 species and during TORCH1 was often present at mixing ratios > 50 pptv.
+The diurnal profile of methyl nitrate showed an early morning maximum that we attributed to nighttime production from NO3 chemistry in the residual layer followed by transportation to the surface by boundary layer dynamics.
+Further work is needed to evaluate our hypothesis of a nighttime source of methyl nitrate but if proven, methyl nitrate could represent an interesting tracer of nighttime chemistry aloft.
+The predominant photochemical source of the short chain alkyl nitrates and carbonyl compounds was from the photochemical oxidation and decomposition of longer chain compounds as opposed to the oxidation of the parent hydrocarbons similar to previous studies.
+The magnitude of this source shows a preference for the shortest chain species consistent with the increasing number of precursor compounds identified in the MCM.
+Primary sources of acetaldehyde and acetone were also significant during TORCH1 and could not be accounted for by automobile exhaust emissions alone.
+Further work to identify and quantify emissions from additional direct sources is necessary if appropriate control strategies are to be implemented.
+
+\citep{deKock1994}
+Air samples were collected at Cape Recife (34 ° 1' S 25 ° 40'E) situated on the south-east coast of Southern Afi'ica (Figure 1).
+The suburban, urban and industrial areas of Port Elizabeth lie approximately 10-40 kin north to north-west of the sampling site.
+Mean sumANs 12.7.
+Variations in absolute and relative concentrations of the alkyl nitrates could be related to changes in meteorological conditions.
+It is considered that the differences observed in chemical composition may represent different sources of trace gas emissions or different degrees of photochemical processing of similar sources.
+The meas~ed levels of the alkyl nitrates suggest long-range transport to play a major role in regulating the distribution of alkyl nitrates in the troposphere.
+
+\citep{Beyersdorf2010}
+South Pole, other Antarctic, ANs only
 
@@
 
+\citep{Reeves2007} ocean+urban+fire
+This paper is based on alkyl nitrate measurements made over the North Atlantic as part of the International Consortium for Research on Atmospheric Transport and Transformation (ICARTT).
+The focus is on the analysis of air samples collected on the UK BAe-146 aircraft during the Intercontinental Transport of Ozone and Precursors (ITOP) project, but air samples collected on board the NASA DC-8 and NOAA WP-3D aircraft as part of a Lagrangian experiment are also used.
+The ratios between the alkyl nitrates and their parent hydrocarbons are compared with those expected from chemical theory.
+Further, a box model is run to investigate the temporal evolution of the alkyl nitrates in three Lagrangian case studies and compared to observations.
+The general agreement between the alkyl nitrate data and photochemical theory suggests that during the first few days of transport from the source region, photochemical production of alkyl nitrates, and thus ozone, had taken place.
+Concentrations of methyl, ethyl, 1-propyl, 2-propyl, 2-butyl, 2-methyl-3-butyl and the sum of 2- and 3-pentyl nitrates (hereafter referred to as 2 + 3-pentyl nitrate) are presented.
+The alkyl nitrate concentrations observed during
+ITOP are typical of those found a few days downwind of precursor source regions.
+Although there is some evidence for an oceanic source of methyl nitrate, for the most part the air samples collected during ITOP do not appear to be strongly influenced by oceanic sources of the alkyl nitrates.
+Elevated concentrations were observed in air masses which have been influenced by emissions from the N.E.
+United States and from Alaskan fires.
+The large variations in ratios between the C5 and C2-C4 nitrates suggest that emission ratios for the parent hydrocarbons vary between different regions of the N.E.
+United States.
+[49] In general the alkyl nitrate data agrees well with photochemical theory.
+This suggests that at least during the first few days of transport from the source region photo- chemical production of alkyl nitrates had taken place.
+This implies that ozone production had also occurred.
+There also appears to be a widespread common source of ethyl nitrate and 1-propyl nitrate other than from their parent hydro- carbons.
+This might be from longer-chain hydrocarbons, possibly pentanes.
+[50] The observations in the more photochemically pro- cessed air masses are consistent with the concentration of NO dropping to a point where the rates of the alkyl nitrate production reactions no longer dominate the peroxy radical self/cross reactions.
+However, in these low levels of NO, where the balance between ozone production and destruc- tion is highly sensitive to NOX, the alkyl nitrates might contribute to maintaining the ozone levels in air as it undergoes long-range transport.
+[51] The observed patterns of the alkyl nitrates also suggest that the rates of photochemical processing in the Alaskan smoke plumes were small and that the rates of ozone formation in these air masses were low.
+
+\citep{Simpson2003} airborne too
+TRACE-P samples that did not show influence from marine alkyl nitrate sources were used to investigate photochemical alkyl nitrate evolution.
+Relative to 2-butyl nitrate/n-butane, the measured ratios of ethyl nitrate/ethane and 2-propyl nitrate/propane showed notable deviations from modeled values based on laboratory kinetic data, suggesting additional Asian sources of their alkyl peroxy radical precursors.
+By contrast, the measured ratios of 1-propyl-, 2-pentyl-, and 3-pentyl nitrate to their respective parent hydrocarbons were fairly close to modeled values.
+
+\citep{Roberts1998} marine MBL+urban ANs
+The measurements described herein were made as part of the
+Organic nitrates are a class of product species that contain
+North Atlantic Regional Experiment (NARE) during August and September 1993.
+These measurements consisted of NO, NO 2, particulate nitrate, HNO3, PAN, PPN, RONO 2, NOy, 0 3, CO, SO 2, and C2-C6 hydrocarbons at Chebogue Point.
+This site was ideally located to study the evolution of photochemically processed air masses in the marine boundary layer (MBL).
+Chebogue Point is approximately 500 km downwind from the eastern sea coast of the United States, and is often impacted by transport of polluted air from the continental source regions.
+Northwest Territories, Canada, have been made in the winter to
+The correlations between organic nitrates and CO or O3 have been presented by Roberts et al.
+[1996].
+It has been established that CO is a good tracer of continentally derived pollution at this site [Parrish et al., 1993].
+This continentally derived pollution contains hydrocarbons (HC) and NO x, the precursors of organic nitrates and ozone, and serves to link the sources of CO with the sources of organic nitrates and ozone.
+There are distinct differences in the lifetimes of CO, O3 , and organic nitrates, and the factors that they are dependent on.
+For these reasons, organic nitrates correlate to varying degrees with CO and O3.
+Measurements of the organic nitrates; PAN, PPN, methyl-,
+ethyl-, 1-propyl-, 2-propyl-, 2-butyl, and 3-pentyl nitrates were made at Chebogue Point, Nova Scotia, during the 1993 NARE campaign.
+The relationships of these compounds with other measures of continental pollution, for example, CO and 03, are consistent with our view of these compounds as products of HC-NO x photochemistry.
+Comparisons of the ratios with that of [2-butyl nitrate]/[butane] were used to test whether the actual data fit the line predicted from rate constants, branching ratios, photolysis rates, and assumed [OH].
+Abstract. Measurements of six alkyl nitrates, methyl-, ethyl-, 1-propyl-, 2-propyl-, 2-butyl, and 3-pentyl nitrates, and two peroxycarboxylic nitric anhydrides, peroxyacetic nitric anhydride
+(PAN) and peroxypropionic nitric anhydride (PPN) were made
+at Chebogue Point, Nova Scotia, during the 1993 North Atlantic
+Regional Experiment campaign. The correlation of alkyl nitrates
+(RONO2) with carbon monoxide makes it clear that the chemistry at this site is being driven by continental emissions, since RONO2 compounds are known to be by-products of HC-NOx photochemistry.
+Comparison of the ratios of alkyl nitrates to their parent hydrocarbons to that of 2-butyl nitrate/butane showed significant
+deviations from trends predicted from rate constants, branching
+ratios, and loss rates. For ethyl nitrate this deviation is likely due to a combination of many additional pathways  o the intermediate
+peroxy radical from the decomposition of larger alkoxy radicals.
+The propyl nitrates showed trends that differed from one another; 2-propyl nitrate was a factor of 2 to 3 higher than predicted, while 1-propyl nitrate was a factor of 8 to 10 higher than predicted.
+The trend of 3-pentyl nitrate/pentane was relatively close (about
+50\% lower) to the predicted trend.
+
 \subsubsection*{Biomass burning source}
@@
 They estimate global emissions from this source to be on the order of 8 Gg yr-1]
+\citep{Simpson2002}
+We report the first observations of the emission of five C1–C4 alkyl nitrates (methyl-, ethyl-, n-propyl-, i-propyl-, and 2-butyl nitrate) from savanna burning.
+Average alkyl nitrate mixing ratios in the immediate vicinity of three bushfires in Northern Australia were 47–122 times higher than local background mixing ratios.
+These are the highest alkyl nitrate mixing ratios we have ever detected, with maximum mixing ratios exceeding 3 ppbv for methyl nitrate.
+Methyl nitrate dominated the alkyl nitrate emissions during the flaming stage of savanna burning, whereas C2–C4 alkyl nitrates were mostly emitted during the smoldering stage.
+To explain the formation of alkyl nitrates from biomass burning, we propose a reaction mechanism involving the combination of reactive radicals at high temperature.
+Bearing in mind the uncertainties associated with extrapolating small data sets to much larger scales, alkyl nitrate emissions from global savanna burning are estimated to be on the order of 8 Gg/yr.
 
@@
 \subsection{Mean AN concentrations}
-mean ANs concentrations: \citep{Reeves2007}, \citep{Roberts1998}
+mean ANs concentrations: \citep{Reeves2007}, 10 ppt \citep{Roberts1998} 14 ppt
 \citep{Atherton1989}
@@
 Pacific Ocean methyl nitrate 50 pptv
+\citep{Neuman2012}
+approx 1 ppb as difference between NOy and (NOx+PAN+HNO3) is sample closest to the plume center
 
@@
 A huge number of modelling studies has been done about alkyl nitrates.
+In particular, O3 and the RONO2 oxida- tion products formed from reactions between these NOx and VOC emissions are examined in detail...
+
+\citep{Neuman2012}
+Source characterization and photochemical processes in transported air masses have been analyzed by examining RONO2 formation from anthropogenic emissions of alkane precursors [Simpson et al., 2003; Madronich, 2006; Simpson et al., 2006; Reeves et al., 2007; Worton et al., 2010].
+Reactive nitrogen species and volatile organic compounds (VOCs) were measured from an instrumented aircraft during daytime flights in the marine boundary layer downwind from the area of surfacing oil.
+A unique VOC mixture, where alkanes dominated the hydroxyl radical (OH) loss rate, was emitted into a clean marine environment, enabling a focused examination of O3 and RONO2 formation processes.
+In the atmospheric plume from DWH, the OH loss rate, an indicator of potential O3 formation, was large and dominated by alkanes with between 5 and 10 carbons per molecule (C5–C10).
+Observations showed that NOx was oxidized very rapidly with a 0.8 h lifetime, producing primarily C6–C10 RONO2 that accounted for 78\% of the reactive nitrogen enhancements in the atmospheric plume 2.5 h downwind from DWH.
+Both observations and calculations of RONO2 and O3 production rates show that alkane oxidation dominated O3 formation chemistry in the plume.
+Rapid and nearly complete oxidation of NOx to RONO2 effectively terminated O3 production, with O3 formation
+yields of 6.0 plus minus 0.5 ppbv O3 per ppbv of NOx oxidized.
+Eight RONO2 compounds with between one and five carbon atoms
+per molecule were measured, and in a few samples the sum of C6–C8 RONO2.
+NOx loss by OH + NO2 was negligible here, as demonstrated by the absence of HNO3 production in the plumes (Figure 4). Instead, the rapid NOx loss was caused by the reaction of peroxy radicals with NO.
+Figure 4 illustrates temporal changes in reactive nitrogen partitioning in the DWH plume.
+In fresh plumes (e.g., Figure 4a), NOx accounted for over 80\% of plume NOy.
+In aged plumes (e.g., Figure 4b), NOx oxidation was nearly complete, but NOx, PAN, HNO3, and the sum of eight C1–C5 RONO2 accounted for only 22\% of plume NOy.
+This result differs dramatically from previous studies, where the sum of NOx + PAN + HNO3 accounted for over 96\% of NOy in urban and petrochemical plumes [Neuman et al., 2002; Ryerson et al., 2003] and for approximately 90\% of NOy in rural areas [Parrish et al., 1993].
+In urban air masses, RONO2 often accounted for a few percent of NOy [O’Brien et al., 1997; Flocke et al.,1998; Schneider et al., 1998; Talbot et al., 2003; Ryerson et al., 2003], and up to 10–20\% of NOy in some studies [Day et al., 2003; Perring et al., 2010].
+The largest contribution to the total calculated RONO2 production rate is from C6–C10 alkanes, which were responsible for 80\% of the RONO2 production.
+OPE.
+In the DWH plume, OPE was 6.0 plus minus 0.5 ppbv/ppbv after 1 h of transport (Figure 3b), and similar to values measured in U.S.
+urban areas [Trainer et al., 1995; Neuman et al., 2009].
+In contrast, petrochemical plumes with OH loss rates dominated by alkenes had OPEs ranging from 10 to 18 [Ryerson et al., 2003] (Figure 3b).
+On 10 June, O3 enhancements in the DWH plumes were less than 15 ppbv (Figure 5b), even after NOx was completely oxidized.
+Although the ratio of VOC to NOx was very large, O3 production was limited by OH oxidation of C5–C10 alkanes that preferentially formed RONO2 and by NOx levels less than 10 ppbv.
 

### 2015-08-02 MSc_diss.tex
Finished section about ANs chemistry

@@
 \subsection{Alkyl nitrates}
-Alkyl nitrates ($RONO_2$) are organic tropospheric trace gases, which through one of their formation mechanisms are closely linked with the production of tropospheric ozone \citep{Reeves2007}.
+Alkyl nitrates ($RONO_2$) are organic trace gases, which through one of their formation mechanisms are closely linked with the production of tropospheric ozone \citep{Reeves2007}.
 This mechanism is present in the following section.
@@
 \subsubsection*{Alkyl nitrates chemistry}
-
-The photochemical formation of the alkyl nitrates starts when the chemistry of the troposphere reaches the regime of $O_3$ production, i.e.
-when reactions of the peroxy radicals with $NO$ start to compete with their self- and cross-reactions more effectively due to an increase in $NO_x$ concentration.
-In this case the reaction between the alkyl peroxy radical ($RO_2$, where $R$ denotes methyl group, $CH_3$) and $NO$ can reveal its two-channel nature \citep{Day2003}, which is controlled by the branching ratio $\alpha_3$ in the following way \citep{Roberts1990}
-\begin{subequations} \label{reac:RONO2ab}
+The photochemical formation of the alkyl nitrates starts when chemistry of the troposphere reaches the regime of $O_3$ production, i.e.
+when reactions of the peroxy radicals with $NO$ start to compete with their self- and cross-reactions more effectively due to an increase in $NO_x$ mixing ratio.
+In this case the reaction between the alkyl peroxy radical ($RO_2$, where $R$ denotes methyl group, $CH_3$) and $NO$ can reveal its two-channel nature \citep{Day2003}:
+\begin{subequations} \label{reac:RONO2alpha}
 \begin{align}
-RO_2 + NO &\xrightarrow{1-\alpha_3} RO + NO_2 \label{eq:peroxy_no1}\\
-RO_2 + NO &\xrightarrow{\alpha_3} RONO_2 \label{eq:peroxy_no2}
+RO_2 + NO &\xrightarrow{1-\alpha_3} RO + NO_2 \label{reac:RO+NO=aphaRO+NO2}\\
+RO_2 + NO &\xrightarrow{\alpha_3} RONO_2 \label{reac:RO+NO=aphaRONO2}
 \end{align}
 \end{subequations}
-The latter reaction can serve as a termination step of the propagated $O_3$ forming cycle, and as \citep{Roberts1998} estimated this likely begin to occur when the $NO$ concentration increases up to 100-200 ppt.
-
-\subsubsection*{Branching ratio}
-The branching ratio $\alpha_3$ is larger when the parent alkane is larger and is a function of temperature and pressure.
-\citep{Newland2013}
-[The yield of this secondary pathway increases with increasing carbon chain length (Table 1.3)]
-
-\subsubsection*{Yield}
-\citep{Newland2013}
-[Since the yield of alkyl nitrates from Reaction 1.9 increases with chain length but the mixing ratios of the alkanes tends to decrease with chain length, propyl and butyl nitrates are generally present at the highest concentrations in polluted air (Beyersdorf et al., 2010)]
-\citep{Sommariva2008}
-[The yield of alkyl nitrate formation is temperature and pressure dependent and increases with the number of carbons in the peroxy radical \citep{Roberts1990}!]
-
-\subsubsection*{Alternative photochemical chemical sources}
-\citep{Newland2013}
-[Another photochemical pathway has been shown to become important for the production of the lower (methyl) alkyl nitrates in polluted air.
-Reaction 1.10 shows the reaction of the alkoxy radical, formed in in Reaction 1.9a, with NO2.
+The latter reaction can serve as a termination step of the propagated $O_3$ forming cycle (Figure ???), and as \citep{Roberts1998} estimated it is likely to begin to occur when the $NO$ mixing ratio increases up to 100-200 ppt.
+The yield of alkyl nitrate formation is controlled by the branching ratio ($\alpha_3$), which is temperature and pressure dependent and increases with the number of carbons in the alkyl peroxy radical \citep{Roberts1990}.
+
+An alternative photochemical pathway of the formation of alkyl nitrates is the reaction of the alkoxy radical ($RO$) with $NO_2$:
 \begin{equation}\label{reac:RO+NO2=RONO2}
@@
 \end{equation}
-This photochemical pathway has been shown to become dominant over Reaction 1.9b for the production of methyl nitrate in polluted air masses where NOx mixing ratios exceed 40 ppb (Archibald et al., 2007).
-This reaction is only likely to be important for methyl nitrate (!) rather than the higher nitrates because of the very small branching ratio ($\alpha_3$) for the production of methyl nitrate via the reaction of the peroxy radical with NO]
-\citep{Newland2013}
-[in highly polluted environments, the reaction of the alkoxy radical with the NO2 radical (Reaction 1.10) may be important for formation of the C1 – C2 alkyl nitrates (Flocke et al., 1998; Simpson et al., 2006; Archibald et al., 2007).
-Archibald et al.
-(2007) reported a model run which suggested that the formation of methyl nitrate via the two production routes would be equal at NOx values of 37 ppb.
-It is unclear from the paper whether they mean the NOx mixing ratio or NO2 mixing ratio but either would suggest a value of a similar magnitude to that shown in Table 6.6.]
-
-\subsubsection*{Loss reactions}
-Being a reservoir for the short lived NOx and having atmospheric lifetimes ranging from several days to about a month \citep{Reeves2007} alkyl nitrates can be transported over long distances, and then lost through photolysis and reaction with $OH$:
+This pathway has been shown to become dominant over reaction (\ref{reac:RO+NO=aphaRONO2}) only for the production of lower (methyl ($C_1$), ethyl ($C_2$)) nitrates when $NO_x$ mixing ratio exceeds 35 ppb \citep{Archibald2007}, i.e.
+in highly polluted environments.
+
+Alkyl nitrates are removed from the atmosphere by photolysis \citep{Turberg1990} and $OH$ radical oxidation \citep{Talukdar1997}:
 \begin{equation} \label{reac:RONO2+hv=RO+NO2}
@@
 \end{equation}
-It means that alkyl nitrates can ‘release’ $NO_2$ back to the atmosphere after a while, and in association with reactions (10) and (11) start a new $O_3$ formation cycle being quite away from the source.
-\citep{Newland2013}
-[The main loss processes for the alkyl nitrates are oxidation by the OH radical and photodissociation (e.g.
-Talukdar et al., 1997a, 1997b).
-The importance of the OH sink relative to photolysis increases with increasing carbon chain length to become the dominant loss process in C5 alkyl nitrates and above (Roberts and Fajer, 1989).]
-
-The alkyl nitrates are removed from the atmosphere through photolysis and reaction with hydroxyl radical.
-The relative importance of these two loss reactions depends generally on the net UV intensity and the size of the carbon chain in alkyl nitrates.
-Photolysis is relatively more important for nitrates with $< C_4$, while nitrates with longer chains ($\geq C_5$) tend to react faster with $OH$.
-In case of butyl nitrates both mechanisms are of similar significance \citep{Worton2010}.
-
-Alkyl nitrates are removed from the atmosphere by photolysis \citep{Turberg1990} or through oxidation by hydroxyl radical ($OH$) \citep{Talukdar1997}.
-Photolysis is the dominant loss mechanism for the compounds with short carbon chains and becomes more important with decreasing carbon number.
-Reaction with $OH$ follows the opposite trend and is most important for the pentyl nitrates and higher.
-
+The relative importance of these two loss reactions depends generally on the net UV intensity and the size of the carbon chain in the alkyl nitrate.
+Photolysis is relatively more important for nitrates with $< C_4$, while nitrates with longer chains ($\geq C_5$) tend to react faster with $OH$ \citep{Worton2010}.
+
+The lifetimes of alkyl nitrates depend on the seasonality of the photolysis and $OH$ sink, which in turn vary with latitude and altitude.
+According to \citep{Clemitshaw1997} and \citep{Turberg1990} alkyl nitrates have atmospheric lifetimes of several days to several years, which makes them temporary reservoirs for the short lived $NO_x$ compounds.
+Being transported from polluted areas to remote background regions and eventually destroyed alkyl nitrates can 'release' $NO_2$ back to the atmosphere, thereby contributing to tropospheric ozone formation.
 \subsection{Ozone, alkyl nitrates and carbon bonds}
@@
 \citep{deKock1994}
-Air samples were collected at Cape Recife (34 ° 1' S 25 ° 40'E) situated on the south-east coast of Southern Afi'ica (Figure 1).
+Air samples were collected at Cape Recife (34 1' S 25 40'E) situated on the south-east coast of Southern Afi'ica (Figure 1).
 The suburban, urban and industrial areas of Port Elizabeth lie approximately 10-40 kin north to north-west of the sampling site.
@@
 Bearing in mind the uncertainties associated with extrapolating small data sets to much larger scales, alkyl nitrate emissions from global savanna burning are estimated to be on the order of 8 Gg/yr.
-
-\subsubsection*{Lifetime}
-\citep{Newland2013}
-[Species within this group provide a transport reservoir for the short lived NOx compounds due to their relatively longer lifetimes.
-This allows NOx to be transported from polluted areas to remote background sites.
-Alkyl nitrates have been reported as making up only a few percent of the NOy ('odd' nitrogen) budget in urban and rural areas (Shepson et al., 1993; Thornberry et al., 2001), however due to their relatively long lifetimes and their continued production during transport of polluted air masses, as alkanes are converted to alkyl nitrates (e.g.
-Bertman et al., 1995), they can comprise up to tens of percent of the NOy budget in polar regions and in the remote marine troposphere (Talbot et al., 2000).
-Therefore alkyl nitrates could potentially be a substantial source of NOx in remote background regions contributing to tropospheric ozone formation in these regions.
-The lifetimes of the alkyl nitrates vary with latitude and season due to the seasonality of the OH sink and photolysis.
-Clemitshaw et al.
-(1997) report a range of lifetimes with respect to OH and to photolysis.
-Photolysis lifetimes vary from about ten days for butyl and pentyl nitrate in the summer to 1 – 2 years in the winter at 65N with lifetimes also decreasing with altitude.
-The lifetimes due to reaction with OH (tauOH) at 65N range from tens of years in
-winter to about 10 days in the summer. tauOH increases with altitude as the OH concentration decreases. At the equator tauOH is similar all year round, on the order of a few days.]
 

### 2015-08-02 MSc_diss.tex
Finished section about carbon bonds theory

@@
 Tropospheric ozone ($O_3$) plays an important role in the chemical and thermal balance of the atmosphere.
-Although it accounts only for about 10\% of the total ozone, it is a primary precursor to the formation of hydroxyl radicals ($OH$) which in association with hydrogen peroxide ($H_2O_2$) determines the ‘cleansing’ capacity of the atmosphere \citep{Prinn2003,Tarasick2008,Thompson1992}.
+Although it accounts only for about 10\% of the total ozone, it is a primary precursor to the formation of hydroxyl radicals ($OH$) which in association with hydrogen peroxide ($H_2O_2$) determines the 'cleansing' capacity of the atmosphere \citep{Prinn2003,Tarasick2008,Thompson1992}.
 On the other hand, being a source of main pollutant removal agents, ozone is an air pollutant itself.
@@
 This mechanism is present in the following section.
-
 \subsubsection*{Alkyl nitrates chemistry}
@@
 \subsection{Ozone, alkyl nitrates and carbon bonds}
-The question that is always important for ozone-policies is the ozone production efficiency.
-Historically people noticed that with decreasing NOx emission the ozone production?
-concentration?
-drops, so the way of calculating the efficiency of ozone production due to drop in NOx emissions was proposed in the following way:
-
-\citep{Neuman2012}
-Ozone production efficiency (OPE), which can be determined observationally from the ratio of O3 to NOx oxidation products, is used regularly as a metric to assess O3 photochemistry [e.g., Trainer et al., 1993; Neuman et al.,2009]
-
-In addition to a traditional way of calculating ozone production efficiency, $\epsilon$, usually defined as the 
-number of $O_3$ molecules produced per molecule of $NO_x$ consumed (Jacob, 1999): 
+In the previous sections we covered the chemistry that controls the production of $O_3$ in the troposphere. A traditional way to assess the efficiency of this production is to calculate the ratio:
+\begin{equation}\label{eq:OPE}
+OPE = \dfrac{k_{HO_2+NO}[HO_2][NO]}{k_{OH+NO_2}[OH][NO_2]} = \dfrac{P_{O_3}}{L_{NO_x}}
+\end{equation}
+where $OPE$ is the ozone production efficiency, $k_{HO_2+NO}$ - the rate coefficient of reaction (\ref{reac:HO2+NO=NO2+OH}), $k_{OH+NO_2}$ - the rate coefficient of reaction (\ref{reac:OH+NO2=HNO3}), $[X]$ - mixing ratios of corresponding species, $P_{O_3}$ – the $O_3$ production rate, $L_{NO_x}$ – the $NO_x$ loss rate.
+$OPE$ is determined by the amount of $HO_2$ reacting with $NO$ to form $NO_2$ and $OH$ (i.e., $k_{HO_2+NO}[HO_2][NO]$, (\ref{reac:HO2+NO=NO2+OH})) and the amount of $NO_x$ removed from the system by reaction of $OH$ with $NO_2$ to form nitric acid ($HNO_3$, reaction (\ref{reac:OH+NO2=HNO3})).
+In other words, $OPE$ shows the amount of ozone created ($P_{O_3}$) per molecule of $NO_x$ removed ($L_{NO_x}$) from the system \citep{Jacob1999}.
+Due to the reasons described in section \ref{sec:chem_mech} (comment at the bottom of the page!"No representation of the nitric acid ($HNO_3$) in the model"), we use another expression for calculations of the $O_3$ production efficiency.
+This expression has been recently proposed by Professor Matthew Evans (University of York) amid the mounting concerns about the need of thorough intercomparison of chemical mechanisms of global atmospheric models (including those used in the IPCC reports), why it is more universal and applicable to any kind of model.
+
+According to \citep{Evans2014} (personal communication), the $O_3$ production efficiency can be estimated via calculation of the number of carbon bonds emitted into the system.
+For example, if the system contains one molecule of methane, $CH_4$, as 'a carbon bond reservoir', it means that this system has 4 carbon bonds (4 C-H), if  instead of methane it contains one molecule of ethane, $C_2H_6$, the system has 7 carbon bonds (6 C-H bonds and 1 C-C), and so on \citep{Edwards2013}.
+Assuming that every carbon bond is broken chemically and every organic peroxy radical ($RO_2$) reacts with $NO$, the overall efficiency of carbon bonds conversion to $O_3$, $\epsilon$, can be calculated as:
 \begin{equation}
-\epsilon = \dfrac{P_{O_3}}{L_{NO_x}}
-\end{equation}
-where $P_{O_3}$ – the $O_3$ production rate, and $L_{NO_x}$ – the $NO_x$ loss rate, a novel method of calculation of $\epsilon$ is applied.
-
-This formula is not very useful if we want to estimate the impact of AN chemistry on O3, but there is another method, proposed by Professor Matthew Evans (University of York) suggests that the ozone production efficiency can be estimated via calculation of the number of carbon bonds emitted \citep{Evans2014} (personal communication).
-For example, if a given set of reactions contains only methane, $CH_4$, as 'a carbon bond reservoir', it means that this system has 4 carbon bonds (4 C-H).
-If  ethene, $C_2H_4$, is considered the system has 5 bonds (4 C-H bonds and 1 C=C) \citep{Edwards2013}.
-Assuming that every carbon bond is broken chemically and every peroxy radical ($RO_2$) reacts with an NO the overall efficiency of carbon bonds conversion to ozone can be calculated as:
-\begin{equation}
-\epsilon = \dfrac{P_{O_3}}{E_{C_{bons}}} = \chi*nu
-\end{equation}
-where $P_{O_3}$ – the O3 production rate (mol hr-1), $E_{C_{bons}}$  – the carbon bond initial abundance (mol hr-1), $\chi$ – ratio between rate of carbon bonds loss to make $RO_2$ and rate of all carbon bonds loss, $nu$ – ratio between rate of $RO_2$ loss in reaction with NO and rate of all $RO_2$ loss.
-In other words, chemical loss of 1 carbon bond produces 1 peroxy radical, and photolysis loss of 1 carbon bond produces 0 or 2 peroxy.
-In other words,  it means that essentially we need to calculate the overall number of peroxy radicals remained in the chemical system by the end of simulation.
-
-The same kind of approach is used in Common Representative Intermediates (CRI) mechanism for VOC degradation \citep{Jenkin2002, Jenkin2008}.
-A reduced mechanism describing ozone formation from the tropospheric degradation of methane and 115 emitted non-methane hydrocarbons and oxygenated volatile organic compounds has been developed, using the Master Chemical Mechanism version 3.1 (MCM v3.1) as a reference benchmark.
-The Common Representative Intermediates mechanism version 2 (CRI v2) has been built up on a compound-by-compound basis, with the performance of its chemistry optimised for each compound in turn by comparison with that of MCM v3.1, using a series of five-day box model simulations.
-The resultant mech- anism contains 1183 reactions of 434 chemical species, i.e., ca.
-10\% of the number of reactions and species to degrade the same set of VOCs in MCM v3.1.
-Similarly to CRI v1 (Jenkin et al., 2002a), a key assumption in the CRI v2 construction methodology is that the potential for ozone formation from a given volatile organic compound (VOC) is related to the number of reactive (i.e., C-C and C-H) bonds it contains.
-This index allows a series of generic intermediates to be defined, with each being used as a 'common representative' for a large set of species possessing the same index, as formed in detailed mechanisms such as the MCM.
-On this basis, the number of reactive (C-H and C-C)
-bonds in a closed-shell intermediate can be used as an index to represent the number of NO-to-NO2 conversions which can potentially result from its subsequent OH-initiated and NOx-catalysed degradation.
-
+\epsilon = \dfrac{P_{O_3}}{E_{C_{bons}}} = \chi\cdot\upsilon
+\end{equation}
+where $P_{O_3}$ – the rate of $O_3$ production (mol yr-1), $E_{C_{bons}}$  – the rate of carbon bonds emission (mol yr-1), $\chi$ – ratio between the rate of carbon bonds loss to form $RO_2$ and the loss rate of all carbon bonds, $\upsilon$ – ratio between the rate of $RO_2$ loss in reaction with $NO$ resulting in $NO$-to-$NO_2$ conversion and the loss rate of all $RO_2$.
+In other words, chemical loss of 1 carbon bond produces 1 peroxy radical:
+\begin{itemize}
+\centering
+\item[] $OH + {CH_4}|_{4 bonds} \rightarrow {CH_3O_2}|_{3 bonds} + H_2O$
+\item[] 4 carbon bonds = 3 carbon bonds, 1 peroxy radical
+\end{itemize}
+Loss of 1 carbon bond via photolysis produces 2 peroxy radicals:
+\begin{itemize}
+\centering
+\item[] ${HCHO}|_{2bonds} + hv \rightarrow H + {HCO}|_{1bond}$
+\item[] ${HCO}|_{1bond} + O_2 \rightarrow HO_2 + {CO}|_{1bond}$
+\item[] $H + O_2 \xrightarrow{M} HO_2$
+\item[] $horizontal line$
+\item[] 2 carbon bonds = 1 carbon bond, 2 peroxy radicals
+\end{itemize}
+or no peroxy radicals because of dry or wet deposition:
+\begin{itemize}
+\centering
+\item[] $HCHO \xrightarrow{deposition}$
+\item[] 2 carbon bonds = 2 carbon bonds, 0 peroxy radicals
+\end{itemize}
+
+The fact that $\epsilon$ does not account for reactions where no $NO$-to-$NO_2$ conversion happens, as, for example, in case of the formation of the alkyl nitrates, $\epsilon$ provides a way to relate the rate of $O_3$ production to the alkyl nitrate chemistry.
+Knowing the rate of carbon bond emissions ($E_{C_{bons}}$) from emissions inventory data and knowing the rate of $O_3$ production from a model run ($P_{O_3}$), one can estimate the efficiency of $O_3$ production per carbon bond emitted in their model, and than compare it with other such values from different models potentially revealing the true cause of discrepancies between models and observations of tropospheric $O_3$ budgets.
+
+For link at the bottom?
+The same kind of approach is used in the Common Representative Intermediates (CRI) mechanism for VOC degradation \citep{Jenkin2002, Jenkin2008}.
+It is a reduced chemical mechanism that describes $O_3$ formation from the tropospheric degradation of methane and 115 emitted non-methane hydrocarbons and oxygenated volatile organic compounds developed using the Master Chemical Mechanism version 3.1 (MCM v3.1) as a reference benchmark.
 \subsection{Motivation}
@@
 
-\subsection{Chemical mechanism}
+\subsection{Chemical mechanism}\label{sec:chem_mech}
 The chemical mechanism of the model is based on recommendations from \citep{Atkinson2004} and information from the Master Chemical Mechanism, MCM v3.3 \citep{Jenkin1997,Saunders2003}, retrieved via website: http://mcm.leeds.ac.uk/MCM.

### 2015-08-02 MSc_diss.tex
Changes in objectives and aims

@@
 This capacity depends mainly on two factors: climate and the global burden of the principal oxidants such as ozone ($O_3$), the hydroxyl radical ($OH$) and hydrogen peroxide ($H_2O_2$) \citep{Prinn2003,Thompson1992}.
-
 \subsection{Tropospheric ozone}
@@
 The only way to answer these questions is to perform a rigorous investigation of the factors that drive tropospheric ozone and attribute those differences to respective factors \citep{Young2012}.
+
+MAKE GRAPHIC FROM TABLE ABOUT O3 BUDGETS!
 
@@
 Assuming that every carbon bond is broken chemically and every organic peroxy radical ($RO_2$) reacts with $NO$, the overall efficiency of carbon bonds conversion to $O_3$, $\epsilon$, can be calculated as:
-\begin{equation}
+\begin{equation}\label{eq:epsilon}
 \epsilon = \dfrac{P_{O_3}}{E_{C_{bons}}} = \chi\cdot\upsilon
@@
 The same kind of approach is used in the Common Representative Intermediates (CRI) mechanism for VOC degradation \citep{Jenkin2002, Jenkin2008}.
-It is a reduced chemical mechanism that describes $O_3$ formation from the tropospheric degradation of methane and 115 emitted non-methane hydrocarbons and oxygenated volatile organic compounds developed using the Master Chemical Mechanism version 3.1 (MCM v3.1) as a reference benchmark.
-\subsection{Motivation}
+It is a reduced chemical mechanism that describes $O_3$ formation from the tropospheric degradation of methane and 115 emitted non-methane hydrocarbons and oxygenated volatile organic compounds developed using the Master Chemical Mechanism version 3.1 (MCM v3.1) as a reference benchmark
+\section{Overall objective and specific aims}
+As the reader might anticipateit is mentioned in section \ref{sec:lit_review}
 \citep{Sommariva2008}
@@
 This kind of study shows the potential of using mechanisms with a high level of chemical detail to study the formation of pollutants in the atmosphere (Madronich, 2006).]
-
-\section{Overall objective and specific aims}
 \subsection{Overall objective}
@@
 \subsection{Specific aims}
-Specific aims are:
+Specific aims are to investigate:
 \begin{itemize}
-\item to investigate of the relationship between ozone production and carbon bond emissions (see Data and Methods);
-\item to analyse the impact of inclusion of alkyl nitrates chemistry on this relationship;
-\item to analyse everything mentioned above in various NOx regimes.
+\item[(1)] the relationship between ozone production and carbon bond emissions;
+\item[(2)] the impact of inclusion of alkyl nitrates chemistry on this relationship;
+\item[(3)] how (1) and (2) vary in different $NO_x$ and $VOC$ regimes;
+\item[(4)] build different version of the box model;
+\item[(5)] to evaluate the results.
 \end{itemize}
 
-\section{Literature review} \label{sec:data}
+\section{Literature review} \label{sec:lit_review}
 \subsubsection{Observational studies}

### 2015-08-03 MSc_diss.tex
Minor changes

@@
 \section{Literature review} \label{sec:lit_review}
+A lot of work is has been dedicated to measurements of alkyl nitrates in different environments and even more to measurements of ozone.
+The majority of these studies have been done in a observational-modelling framework.
+
+During/After one of these investigations, namely the SAGA campaign, \citep{Atlas1993} proposed by that the short-chained alkyl nitrates ($C_1$-$C_3$) can be produced not only photochemically through the oxidation of their parent hydrocarbons \citep{Roberts1990}, but also emitted directly from equatorial oceans.
+This has since been
+confirmed by field measurements \citep{Chuck2002,Blake2003,Dalh2005}.
+
+There have been a number of studies of the latitudial and seasonal variations of alkyl nitrates in marine air (Blake et al., 2003), in the Arctic (Swanson et al., 2003) and in the
+Antarctic (Beyersdorf et al., 2010)
 \subsubsection{Observational studies}
@@
 
-\subsection{Mean AN concentrations}
-mean ANs concentrations: \citep{Reeves2007}, 10 ppt \citep{Roberts1998} 14 ppt
-\citep{Atherton1989}
-Alkyl nitrate concentrations of 0.8-37 pptv (2-butyl and 2- and 3-pentyl nitrates) in remote marine environments [Atlas, 1988] and approximately 2 ppbv in urban environments [Hanst et al., 1982] have been measured.
-\citep{Chuck2002}
-The atmospheric mole fraction of MeONO2 ranged from 1.2 to 43 pptv (Fig.
-2,Aand B) and 0.8 to 10 pptv for EtONO2 (Fig.
-2, D and E).
-These values lie within the range of concentrations reported in the literature for MeONO2 (0.2 to 50 pptv) and EtONO2 (0.3 to 12.2 pptv) in the remote marine troposphere (2, 8, 9).
-\citep{Blake2003}
-Pacific Ocean methyl nitrate 50 pptv
-\citep{Neuman2012}
-approx 1 ppb as difference between NOy and (NOx+PAN+HNO3) is sample closest to the plume center
+\subsubsection{Mean AN concentrations}
+\begin{table}
+\begin{tabular}{cccc}
+\hline
+Observation & $RONO_2$                                  & Comments & Reference \\
+\hline
+ITOP?       & mean $\Sigma RONO_2$ 10 ppt               &          & \citep{Reeves2007} \\
+???         & mean $\Sigma RONO_2$ 14 ppt               &          & \citep{Roberts1998} \\
+remote marine environments & 2-butyl and 2- and 3-pentyl: 0.8-37 pptv & &\citep{Atherton1989} [Atlas, 1988]\\
+urban environments         & 2-butyl and 2- and 3-pentyl?: approx. 2 ppb & & \citep{Atherton1989} [Hanst et al., 1982] \\
+??                         & MeONO2: 1.2 to 43 pptv     &          & \citep{Chuck2002} \\
+??                         & EtONO2: 0.8 to 10 pptv     &          & \citep{Chuck2002} \\
+remote marine troposphere  & MeONO2: 0.2 to 50 pptv     & in lit. (2, 8, 9) & \citep{Chuck2002} \\
+remote marine troposphere  & EtONO2: 0.3 to 12.2 pptv   & in lit. (2, 8, 9) & \citep{Chuck2002} \\
+Pacific Ocean              & MeONO2: 50 pptv            &                   & \citep{Blake2003} \\
+airborne                   & $\Sigma RONO_2$: approx 1 ppb         & difference between NOy and (NOx+PAN+HNO3) 
+                                                                     is sample closest to the plume center & \citep{Neuman2012} \\
+\hline
+\end{tabular}
+\end{table}
 
@@
 
-\citep{Neu2008}
+The first 3-D chemical transport  model simulations examining the role of methyl and ethyl nitrates in tropospheric ozone chemistry have been conducted by \citep{Neu2008}.
+In their study, they incorporated methyl and ethyl nitrates in the chemical mechanism of the UCI CTM [Carver et al., 1997] and compared their simulated abundances to observations.
+To assess  the relative importance of
+oceanic alkyl nitrates as a natural source of $NO_x$ in remote regions they evaluated their impact on tropospheric $O_3$ and the oxidative capacity of the atmosphere.
+
 Observations have revealed widespread emissions of alkyl nitrates from the tropical and Southern Oceans.

### 2015-08-03 MSc_diss.tex
Added figures from Neu2008

@@
 
-The first 3-D chemical transport  model simulations examining the role of methyl and ethyl nitrates in tropospheric ozone chemistry have been conducted by \citep{Neu2008}.
-In their study, they incorporated methyl and ethyl nitrates in the chemical mechanism of the UCI CTM [Carver et al., 1997] and compared their simulated abundances to observations.
-To assess  the relative importance of
-oceanic alkyl nitrates as a natural source of $NO_x$ in remote regions they evaluated their impact on tropospheric $O_3$ and the oxidative capacity of the atmosphere.
-
-Observations have revealed widespread emissions of alkyl nitrates from the tropical and Southern Oceans.
-We present the first chemical transport model simulations to examine the global impact of these emissions.
-Matching observed atmospheric abundances, we derive a total oceanic flux of methyl nitrate (MeONO2) and ethyl nitrate (EtONO2) equivalent to 0.35 Tg of N per year, which contributes as much as 1 DU to the tropospheric ozone column in the Western Pacific and is responsible for about 3\% of the global oxidative capacity of the troposphere.
-Atmospheric observations indicate that low molecular weight (C1-C4) alkyl nitrates are ubiquitous over the tropical and Southern Oceans [Atlas et al., 1993; Blake et al., 1999, 2003] and can be a significant component (20–80\%) of the total reactive nitrogen in remote marine air [Jones et al., 1999; Talbot et al., 2000].
-Recent ship-borne measurements show supersaturation of alkyl nitrates in tropical surface waters, confirming an oceanic source [Chuck et al., 2002; Moore and Blough, 2002; Dahl et al., 2005].
-This work presents the first 3-D chemical transport model (CTM) simulations to examine the role of oceanic alkyl nitrates in global chemistry, using observational constraints.
-In this study, we incorporate MeONO2 and EtONO2 in the chemical mechanism of the UCI CTM [Carver et al., 1997] and compare their simulated abundances to observations to derive new estimates of their oceanic fluxes.
-We assess the relative importance of oceanic alkyl nitrates as a natural source of NOx in remote regions by evaluating their impact on tropospheric ozone (O3) and the oxidative ca- pacity of the atmosphere.
-The UCI CTM [Wild and Prather, 2000; Hsu et al.,2005] includes 30 chemical species in the troposphere.
-We define a control run with no alkyl nitrate emis-
-sions and compare it to two simulations in which we impose a constant, spatially uniform flux of MeONO2 and EtONO2 from: 1) the tropical oceans (10?S–10?N) and 2) the Southern Ocean (south of 45?S).
-Results: impact on NOx and O3 chemistry
-Oceanic alkyl nitrates cause measurable increases in
-both NOx and O3 in marine air.
-Their largest impact is seen over the equatorial Western Pacific (Figure 3).
-There is a persistent minimum in both NOx and O3 (Figure 3a) in this region, where the circulation isolates the marine boundary layer, resulting in sustained photochemical loss of O3 and a deep minimum throughout the tropospheric O3 column.
-At the same time, the circulation acts to accumulate alkyl nitrates in theWestern Pacific (see Figure 2), and they drive up boundary layer NOx abundances by as much as 250\%.
-This leads to O3 increases of up to 20\% (Figures 3b and 3f), with the largest increases occurring from May to October.
-From December to March (not shown), the relative increases in NOx and O3 are smaller, but are still concen- trated in the Western Pacific even though the alkyl nitrates are spread more evenly across the Pacific during these months.
-The geographic pattern of the alkyl nitrate influence persists into the middle and upper troposphere (Figures 3c-3d and 3g-3h), with reduced relative impact because of the larger lightning NOx source.
-However, the largest zonal mean increases in mixing ratio, 3–8 ppt for NOx and 1 ppb for O3, occur in the tropical upper and middle troposphere, respectively (Figure 4).
-Pacific for example, ignore Southern ocean.
-The O3 produced by NOx from alkyl nitrates fills in
-the minimum in the tropospheric O3 column over the Western Pacific, adding more than 1 DU, with a maximum in October (Figure 3e).
-The global average increase in tropospheric O3 column varies with season from 0.35 to 0.42 DU.!!!!
-WOW!!!
-Oceanic alkyl nitrates are necessary to accurately
-model NOx in the tropical remote marine environment, since they are responsible for 15–20\% of the O3 in the Western Pacific.
-The coincident location of the maximum alkyl nitrate impact and the deep minimum in O3 over the Western Pacific is a fascinating consequence of the convergence gence of the trade winds, but will make it difficult to detect an alkyl nitrate signature in O3.
-[22] At current levels, alkyl nitrates have a modest climate impact.
-From a simple scaling [Wild et al., 2001], the global increase in tropospheric column O3 will have a radiative forcing of about 16 mW/m2.
-In steady state, this will be almost entirely offset by the reduced CH4 lifetime, which produces a roughly equal negative radiative forcing.
-Surprisingly, a comparison to Wild et al.
-[2001] shows that on a per-molecule basis, alkyl nitrates have the largest impact on O3 and CH4 of any NOx source analyzed to date.
-This seems to be a consequence of their long lifetime combined with efficient transport to the tropical upper troposphere.
-With a 2.4\% effect on CH4 lifetime, a 5-fold increase in alkyl nitrate emissions during glacial periods would account for roughly 1/3 of the 300 ppb drop in CH4 seen between interglacials and glacials.
+The first 3-D chemical transport  model simulations examining the impact of methyl and ethyl nitrates on tropospheric ozone chemistry in remote regions have been conducted by \citep{Neu2008}.
+In their study, they incorporated methyl and ethyl nitrates in the chemical mechanism of the UCI CTM [Carver et al., 1997] and by matching their observed atmospheric abundances derived a total oceanic emission flux of these two nitrates (0.35 Tg of N per year).
+Comparing two simulations with and without spatially uniform emissions of methyl and ethyl nitrates they found that the largest impact of oceanic alkyl nitrates is seen over the equatorial Western Pacific (Figure ???).
+Atmospheric circulation (convergence of the trade winds) in this region isolates the marine boundary layer, resulting in sustained photochemical loss of $O_3$ (Figure ???
+c) and at the same time accumulation of alkyl nitrates (Figure ???
+a), which drive up boundary layer $NO_x$ abundances by as much as 250\% (Figure ???
+b).
+Such combination of factors leads to increases in $O_3$ of up to 20\% (Figures ???
+c and d), with the largest increases occurring from May to October.
+
+\begin{figure}\label{fig:Neu2008}
+\caption{1}
+\includegraphics[width=.5\linewidth]{{D:/UEA/ENV-MB4X_Dissertation/ENV-MB4X_Dissertation_thesis/ENV-MB4X_Dissertation_thesis_latex/pics/Neu2008_Fig.2_MeONO2_OCT_colorbar}.png}
+\end{figure}
+\begin{figure}
+\caption{1}
+\includegraphics[width=.5\linewidth]{{D:/UEA/ENV-MB4X_Dissertation/ENV-MB4X_Dissertation_thesis/ENV-MB4X_Dissertation_thesis_latex/pics/Neu2008_Fig.2_MeONO2_OCT}.png}
+\end{figure}
+\begin{figure}
+\caption{1}
+\includegraphics[width=.5\linewidth]{{D:/UEA/ENV-MB4X_Dissertation/ENV-MB4X_Dissertation_thesis/ENV-MB4X_Dissertation_thesis_latex/pics/Neu2008_Fig.3_OCT_NOxdiff}.png}
+\end{figure}
+\begin{figure}
+\caption{October monthly mean mixing ratios of (a) methyl nitrate ($MeONO_2$, $CH_3ONO_2$) and (c) $O_3$, and percent difference (perturbation - control) in (b) $NO_x$, (d) $O_3$.
+All averaged over the boundary layer (below 850 mb).
+Adapted from \citep{Neu2008}}
+\includegraphics[width=.5\linewidth]{{D:/UEA/ENV-MB4X_Dissertation/ENV-MB4X_Dissertation_thesis/ENV-MB4X_Dissertation_thesis_latex/pics/Neu2008_Fig.3_OCT_O3contol_O3diff}.png}
+\end{figure}
 

### 2015-08-03 MSc_diss.tex
Updates in Neuman2012 review

@@
 \subsubsection{Observational studies}
+Anthropogenic!!! Neuman
+Source characterization and photochemical processes in transported air masses have been analyzed by examining RONO2 formation from anthropogenic emissions of alkane precursors [Simpson et al., 2003; Madronich, 2006; Simpson et al., 2006; Reeves et al., 2007; Worton
+et al., 2010]
 
@@
 A huge number of modelling studies has been done about alkyl nitrates.
-In particular, O3 and the RONO2 oxida- tion products formed from reactions between these NOx and VOC emissions are examined in detail...
-
-\citep{Neuman2012}
-Source characterization and photochemical processes in transported air masses have been analyzed by examining RONO2 formation from anthropogenic emissions of alkane precursors [Simpson et al., 2003; Madronich, 2006; Simpson et al., 2006; Reeves et al., 2007; Worton et al., 2010].
-Reactive nitrogen species and volatile organic compounds (VOCs) were measured from an instrumented aircraft during daytime flights in the marine boundary layer downwind from the area of surfacing oil.
-A unique VOC mixture, where alkanes dominated the hydroxyl radical (OH) loss rate, was emitted into a clean marine environment, enabling a focused examination of O3 and RONO2 formation processes.
-In the atmospheric plume from DWH, the OH loss rate, an indicator of potential O3 formation, was large and dominated by alkanes with between 5 and 10 carbons per molecule (C5–C10).
-Observations showed that NOx was oxidized very rapidly with a 0.8 h lifetime, producing primarily C6–C10 RONO2 that accounted for 78\% of the reactive nitrogen enhancements in the atmospheric plume 2.5 h downwind from DWH.
-Both observations and calculations of RONO2 and O3 production rates show that alkane oxidation dominated O3 formation chemistry in the plume.
-Rapid and nearly complete oxidation of NOx to RONO2 effectively terminated O3 production, with O3 formation
-yields of 6.0 plus minus 0.5 ppbv O3 per ppbv of NOx oxidized.
-Eight RONO2 compounds with between one and five carbon atoms
-per molecule were measured, and in a few samples the sum of C6–C8 RONO2.
-NOx loss by OH + NO2 was negligible here, as demonstrated by the absence of HNO3 production in the plumes (Figure 4). Instead, the rapid NOx loss was caused by the reaction of peroxy radicals with NO.
-Figure 4 illustrates temporal changes in reactive nitrogen partitioning in the DWH plume.
-In fresh plumes (e.g., Figure 4a), NOx accounted for over 80\% of plume NOy.
-In aged plumes (e.g., Figure 4b), NOx oxidation was nearly complete, but NOx, PAN, HNO3, and the sum of eight C1–C5 RONO2 accounted for only 22\% of plume NOy.
-This result differs dramatically from previous studies, where the sum of NOx + PAN + HNO3 accounted for over 96\% of NOy in urban and petrochemical plumes [Neuman et al., 2002; Ryerson et al., 2003] and for approximately 90\% of NOy in rural areas [Parrish et al., 1993].
-In urban air masses, RONO2 often accounted for a few percent of NOy [O’Brien et al., 1997; Flocke et al.,1998; Schneider et al., 1998; Talbot et al., 2003; Ryerson et al., 2003], and up to 10–20\% of NOy in some studies [Day et al., 2003; Perring et al., 2010].
-The largest contribution to the total calculated RONO2 production rate is from C6–C10 alkanes, which were responsible for 80\% of the RONO2 production.
-OPE.
-In the DWH plume, OPE was 6.0 plus minus 0.5 ppbv/ppbv after 1 h of transport (Figure 3b), and similar to values measured in U.S.
-urban areas [Trainer et al., 1995; Neuman et al., 2009].
-In contrast, petrochemical plumes with OH loss rates dominated by alkenes had OPEs ranging from 10 to 18 [Ryerson et al., 2003] (Figure 3b).
-On 10 June, O3 enhancements in the DWH plumes were less than 15 ppbv (Figure 5b), even after NOx was completely oxidized.
-Although the ratio of VOC to NOx was very large, O3 production was limited by OH oxidation of C5–C10 alkanes that preferentially formed RONO2 and by NOx levels less than 10 ppbv.
+In particular, O3 and the RONO2 oxidation products formed from reactions between these NOx and VOC emissions are examined in detail...
+
+A study that enabled us to focus on examination of $O_3$ and $RONO_2$ formation processes has been conducted by \citep{Neuman2012}.
+After an explosion at the Deepwater Horizon (DWH) offshore oil platform on 20 April 2010 large amounts of oil from the DWH spill reached the ocean surface.
+Subsequent oil evaporation caused a release of little methane and other light alkanes to the atmosphere forming a $VOC$ mixture atypical for the remote marine environment.
+At the same the $NO_x$ levels were enhanced in the vicinity of the oil spill due to spill response efforts (ship exhaust, surface oil burning, flaring of recovered gas).
+Airborne measurements of reactive nitrogen species ($NO_y$ = $NO_x$+$HNO_3$+$PAN$(peroxyacetul nitrate at the bottom)+$RONO_2$) and $VOCs$ in the marine boundary layer during daytime downwind from the DWH area showed that oxidation of alkanes with between 5 and 10 carbons per molecule ($C_5$–$C_{10}$) dominated $O_3$ formation chemistry in the plume.
+This circumstance led to a rapid and nearly complete oxidation of $NO_x$ to $RONO_2$ with mostly 6-10 carbons per molecule ($C_6$–$C_{10}$) effectively terminating $O_3$ production.
+On 10 June, $O_3$ enhancements in the DWH plumes were less than 15 ppbv, even after $NO_x$ was completely oxidized.
+
+$O_3$ production efficiencies in the DWH plume were 6 plus minus 0.5 ppbv $O_3$ per ppbv of $NO_x$ oxidized after 1 h of transport.
+These values are comparable with those measured in U.S.
+urban areas [Trainer et al., 1995; Neuman et al., 2009], but smaller than those measured in the petrochemical plumes with $OH$ loss rates dominated by alkenes (10-18 ppbv $O_3$ per ppbv of $NO_x$) [Ryerson et al., 2003].
+
+Temporal changes in reactive nitrogen partitioning in the DWH plume (Figure \ref{fig:Neuman2012}) nicely illustrate that in fresh part of the plume (Figure ???
+a) $NO_x$ accounted for over 80\% of plume $NO_y$, in aged plumes (Figure ???
+b), $NO_x$ oxidation was nearly complete, but NOx, PAN, HNO3, and the sum of eight $C_5$–$C_{10}$ $RONO_2$ accounted for only 22\% of plume $NO_y$.
+It means that $C_6$–$C_{10}$ $RONO_2$ account for 78\% of the aged plume $NO_y$\citep{Neuman2012}.
+This result differs dramatically from previous studies, where the estimated contribution of $RONO_2$ to $NO_y$ was much lower: approximately 4\% in urban and petrochemical plumes [Neuman et al., 2002; Ryerson et al., 2003], approx.
+10\% in rural areas [Parrish et al., 1993], a few percent in urban air masses [O’Brien et al., 1997; Flocke et al.,1998; Schneider et al., 1998; Talbot et al., 2003; Ryerson et al., 2003], and up to 10–20\% of NOy in some studies [Day et al., 2003; Perring et al., 2010].
+
+\begin{figure}[h!]\label{fig:Neuman2012}
+\includegraphics[width=.5\linewidth]{{D:/UEA/ENV-MB4X_Dissertation/ENV-MB4X_Dissertation_thesis/ENV-MB4X_Dissertation_thesis_latex/pics/Neuman2012_Fig.4}.png}
+\caption{Time series of reactive nitrogen species measured on 10 June 2010 in the MBL in plumes (a) 10 km and (b) 50 km downwind from DWH.
+$NO_x$, $NO_y$, and $HNO_3$ were measured once per s and $PAN$ was measured once per 2 s.
+$C_1$–$C_5$ alkyl nitrates were measured in whole air samples at locations indicated by the circles.
+The sum of alkyl nitrates (black lines) is inferred from the difference between $NO_y$ and ($NO_x$+$HNO_3$+$PAN$) \citep{Neuman2012}.}
+\end{figure}
 
@@
 \begin{figure}
+\includegraphics[width=.5\linewidth]{{D:/UEA/ENV-MB4X_Dissertation/ENV-MB4X_Dissertation_thesis/ENV-MB4X_Dissertation_thesis_latex/pics/Neu2008_Fig.3_OCT_O3contol_O3diff}.png}
 \caption{October monthly mean mixing ratios of (a) methyl nitrate ($MeONO_2$, $CH_3ONO_2$) and (c) $O_3$, and percent difference (perturbation - control) in (b) $NO_x$, (d) $O_3$.
@@
 Adapted from \citep{Neu2008}}
-\includegraphics[width=.5\linewidth]{{D:/UEA/ENV-MB4X_Dissertation/ENV-MB4X_Dissertation_thesis/ENV-MB4X_Dissertation_thesis_latex/pics/Neu2008_Fig.3_OCT_O3contol_O3diff}.png}
 \end{figure}

### 2015-08-03 MSc_diss.tex
Updates in Sommariva2008 review

@@
 
-\citep{Sommariva2008}
-Validation: typical urban plume in the North-East of the United States, measurements taken during three flights of the NOAA WP-3D aircraft, which sampled a plume from the New York City area during the New England Air Quality Study (NEAQS) 2004 campaign
-- Seven alkyl nitrates (ethyl- nitrate, i-propylnitrate, n-propylnitrate, 2-butylni- trate, 2-pentylnitrate, 3-pentylnitrate, 3-methyl-2- butylnitrate) were measured during these flights together with a full suite of VOCs and other chemical parameters, such as CO, O3,NOx, HNO3 and PANs (Neuman et al., 2006).
-- The model results have been qualitatively compared to measurements of alkyl nitrates taken on 20–22 July 2004, when the NOAA WP-3D aircraft made three flights aimed to inter- cept the New York City plume.
-Model: subset! of MCM v3.1 (The MCM is one of the most detailed chemical mechanism for tropospheric chemistry available)
-- to study the formation processes and photochemical evolution of alkyl nitrates
-Chemistry: oxidation mechanisms of 64 VOCs (50 hydrocar-
-bons and 14 oxygenates) plus CH4 and CO, The inorganic chemistry mechanism was taken from the IUPAC recommendation (Atkinson et al., 2003).
-A dimethyl sulphide (DMS) mechanism (Sommariva et al., 2004) was also included.
-The photolysis rates were calculated using the MCM parametrization (Saunders et al., 2003) for end of July 2004 at the latitude and longitude of Boston, MA.
-only gas-phase!!!
-Measurements made during NEAQS 2004 showed that there was very little removal of, for example,HNO3 by dry deposition in plumes decoupled from the marine boundary layer during transport from New York City (NYC) to the North Atlantic (Neuman et al., 2006).
-Therefore, removal by dry deposition was neglected in the model.
-Initialisation: measurements taken in the Boston metropolitan area during the 2002 and 2004 NEAQS campaigns.
-The initial concentrations of VOCs were determined using emission ratios; CH4 and O3 were set to typical values for urban conditions (1800 and 90 ppb) (Table 1).
-The model has been initialized using measurements and VOCs emission ratios from the 2002 and 2004 New England Air Quality Studies.
-Simulation time: 5 days.
-Scenarios: 2: A (higher OH, good with obs, reference), B
-Time: 20-22 July.
-Results:
-- While long-chain (C5) alkyl nitrates are produced for 90\% or more from the oxidation of a single parent alkane, short-chain <C4 alkyl nitrates can be formed from several precursors.
-The relative importance of each production route has been quantitatively determined thanks to the high level of chemical detail provided by the MCM.
-These secondary routes
-to the formation of alkyl nitrates include the oxidation of longer-chain alkanes and oxygenated intermediates, like carbonyls, peroxides and carboxylic acids.
-- The best agreement is in the case of 3-methyl-2-butylnitrate and the discrepancy between the modelled and the measured ratios clearly increases with decreasing carbon number .
-- RH/AN, Except in the case of ethylnitrate, the model can reproduce the measured rate of increase in the alkyl nitrate to alkane ratios fairly well, which suggests that the kinetic parameters used in the MCMmodel are overall correct.
-- By considering together Figs. 3 and 4 it can be reasonably concluded that the chemistry of alkyl nitrates is well represented in the MCM model.
-- Using the explicit description of the chemistry provided by the MCM, it has been possible to determine quantitatively the photochemical formation and the evolution of alkyl nitrates and the relative importance of their different formation pathways.
-The results are summarized in Table 2.
-Large alkyl nitrates are produced almost exclusively from the reaction of a single parent
-alkane.
-Smaller <=C4 alkyl nitrates also have other significant precursors, like longer-chain alkanes and alkenes.
-For n-propylnitrate and ethylnitrate, pro- panal and butanal, which are largely emitted, are important sources within the first 24 h.
-For all alkyl nitrates, the role of organic peroxides and car- boxylic acids formed by the reactions of peroxy radicals with HO2 at low [NOx] increases with time.
-These species contribute to sustain the production of alkyl nitrates far from the emission points of the parent VOCs.
-The relative importance of the different pathways
-and how they change with time are dependent on the initial conditions of the model. In a scenario with lower O3 and CO and higher NOx the
-importance of the direct routes (alkane ?OH) is typically lower during the first two–three days, and the contributions of peroxides and carboxylic acids become significant later and are on average less important.
-These results suggest that secondary chemistry in the formation of the alkyl nitrates, particularly oC4, cannot be neglected and should be included in the chemical models, due to its potential impact on ozone-controlling strategies.
+Probably the most detailed investigation of the formation processes and photochemical evolution of alkyl nitrates (but not their impact on $O_3$) has been done by \citep{Sommariva2008}.
+Using a subset of the Master Chemical Mechanism v3.1 (MCM) they constructed a highly detailed chemical model and run it for 5 days under two scenarios (Table \ref{tab:Sommariva2008init}) to explore the sensitivity of the evolution of the chemical composition of an air mass emitted from an urban environment to the initial conditions.
+As it seen in Figure ???
+according to the modelling results $O_3$ is produced and $NO_x$ is removed from the system with the consequent formation of $HNO_3$ via the reaction of $NO_2$ with $OH$, which broadly agrees with the observations [(Neuman et al., 2006).].
+In scenario A the concentration of $NO_x$ after the first day is about 0.5–1 ppb for $NO_2$ and approximately 100 ppt for $NO$, which is high enough to sustain the production of $O_3$ to up to 135 ppb and the formation of alkyl nitrates during the simulation.
+In scenario B (Figure ???
+right) the increase in $O_3$ is slower (note that initial concentration is 0) and the maximum concentration after five days is about 100 ppb.
+The concentration of $NO_x$ remains relatively high during the second (10–15 ppb) and third day (3–5 ppb) and then decreases to $<$200 ppt of $NO$ and $<$1ppb of $NO_2$ on the fourth and fifth days, therefore allowing the formation of alkyl nitrates too.
+
+\begin{table}\label{tab:Sommariva2008init}
+\caption{Initial values of the some model parameters (in ppb) for two scenarios used in \citep{Sommariva2008}}
+\begin{tabular}{ccc}
+\hline
+Parameter & Initial value Scenario A & Scenario B \\
+\hline
+$CO$      & 332.5 & 232.5 \\
+$NO_x$    & 34.9  & 46.5  \\
+$O_3$     & 90    & 0   \\
+\hline
+\end{tabular}
+\end{table}
+\begin{figure}
+\includegraphics[width=\linewidth]{{D:/UEA/ENV-MB4X_Dissertation/ENV-MB4X_Dissertation_thesis/ENV-MB4X_Dissertation_thesis_latex/pics/Sommariva2008_Fig.1}.png}
+\caption{Modelled concentrations of $O_3$, $NO_x$ and $HNO_3$ for scenario A (left) and B (right). Source: \citep{Sommariva2008}.}
+\end{figure}
+
+Comparison of modelling results with the values observed inside a typical urban plume sampled using the NOAA WP-3D aircraft from the New York City area during the New England Air Quality Study (NEAQS) 2004 campaign enabled \citep{Sommariva2008} to conclude that the chemistry of alkyl nitrates is well represented in the MCM model.
+They found that short-chain $<C_4$ alkyl nitrates can be formed from several precursors, while long-chain ($C_5$) alkyl nitrates are produced for 90\% or more from the oxidation of a single parent alkane.
+They also found that at lower $O_3$ and $CO$ initial concentrations and higher $NOx$ (scenario A) the importance of the reactions between alkane and $OH$ is typically lower during the first two-three days, and the contributions of peroxides and carboxylic acids become significant later and are on average less important.
 

### 2015-08-04 MSc_diss.tex
Updates in mean RONO2 concentrations, clean up literature review

@@
 \newcommand{\mydate}{\the\year}
+
+\newcommand{\picspath}{D:/UEA/ENV-MB4X_Dissertation/ENV-MB4X_Dissertation_thesis/ENV-MB4X_Dissertation_thesis_latex/pics/}
 %% Remove author and date from the maketitle command
@@
 
-During/After one of these investigations, namely the SAGA campaign, \citep{Atlas1993} proposed by that the short-chained alkyl nitrates ($C_1$-$C_3$) can be produced not only photochemically through the oxidation of their parent hydrocarbons \citep{Roberts1990}, but also emitted directly from equatorial oceans.
-This has since been
-confirmed by field measurements \citep{Chuck2002,Blake2003,Dalh2005}.
-
-There have been a number of studies of the latitudial and seasonal variations of alkyl nitrates in marine air (Blake et al., 2003), in the Arctic (Swanson et al., 2003) and in the
-Antarctic (Beyersdorf et al., 2010)
-\subsubsection{Observational studies}
-Anthropogenic!!! Neuman
-Source characterization and photochemical processes in transported air masses have been analyzed by examining RONO2 formation from anthropogenic emissions of alkane precursors [Simpson et al., 2003; Madronich, 2006; Simpson et al., 2006; Reeves et al., 2007; Worton
-et al., 2010]
-
-\citep{Sommariva2008}
-[Alkyl nitrates typically constitute only a small
-fraction of total NOy <10\%, even though they can be a much larger component in the equatorial marine boundary layer (Blake et al., 2003; Simpson et al., 2006 and references therein).
-Due to their long lifetime (41 week) alkyl nitrates can transport NOx to remote regions, with potential impacts on regional O3 formation (e.g., Murphy et al., 2006; Simpson et al., 2006)]
+During/After one of these investigations, namely the SAGA campaign, \citep{Atlas1993} first postulated that the short-chained alkyl nitrates ($C_1$-$C_3$) can be produced not only photochemically through the oxidation of their parent hydrocarbons \citep{Roberts1990}, but also emitted directly from equatorial oceans.
+This has since been confirmed by field measurements \citep{Chuck2002,Blake2003,Dahl2005}.
+\citep{Chuck2002} provided the first direct evidence of that based on detection of supersaturation of surface waters by methyl and ethyl nitrates and their high concentrations in the marine air over equatorial areas (0-40$^{\circ}$) of the Atlantic.
+Later during two extensive NASA airborne campaigns Pacific Exploratory Mission (PEM)-Tropics A (September-October 1996) and PEM-Tropics B (March-April 1999) \citep{Blake2003} observed high levels of methyl and ethyl nitrates (approximately 50 ppt) in the air at the band between 8$^{\circ}$N-13$^{\circ}$S stretching across the equatorial Pacific, and pointed out that oceanic source of $\geq C_4$ alkyl nitrates is insignificant compared to anthropogenic sources.
+
+Alkyl nitrates also have a biomass burning source which have been first reported by \citep{Simpson2002}.
+They found that in the immediate vicinity of savanna fires in Northern Australia the average alkyl nitrate mixing ratios were 47–122 times higher than local background mixing ratios, exceeding 3 ppbv for methyl nitrate.
+Methyl nitrate was primarily emitted during the 'flaming' stage, while $C_2$-$C_4$ ethyl-butyl nitrates dominated during the 'smoldering' stage.
+
+There have been a number of studies of the seasonal variations of alkyl nitrates in the Arctic \citep{Swanson2003} and in the Antarctic \citep{Beyersdorf2010}.
+Polar alkyl nitrate follow the seasonality of the alkanes displaying a clear late winter maximum and a broad summer minimum.
+
+Owing to the existence of the relationship between alkyl nitrates and their parent hydrocarbons, measurements of their concentrations can provide information on the extent of photochemical processing in the air mass.
+\cite{Bertman1995} developed a kinetic approach to describe this relationship through the evolution of the ratio of alkyl nitrate to its parent hydrocarbon as a function of time.
+This approach has been extensively used to study air mass photochemical ageing at several sites, in North America \citep{Bertman1995,Roberts1998}, China \citep{Simpson2006}, the British Isles \citep{Worton2010} and over the Pacific \citep{Simpson2003} and North Atlantic Ocean \citep{Reeves2007,Stroud2001}.
 
@@
 
-\citep{Stroud2001} plane+colorado, plane+North Atlantic
-During both campaigns alkyl nitrates were measured aboard the NOAA Orion WP-3 aircraft with a vertical range from the boundary layer to the midtroposphere (7 km) [Hiibler et al.,1998].
-During STERAO 1996 the aircraft was based at Buckley Air Force Base in Denver, Colorado.
-Flights were conducted from June 13, 1996, to July 16, 1996, over the northeastern plains of Colorado to study how thunderstorms affect the chemical composition of the atmosphere.
-Alkyl nitrates were measured over the northeastern plains of Colorado during the Stratospheric Tropospheric Exchange: Radiation, Aerosols and Ozone (STERAO) 1996 campaign and over the North Atlantic during the North Atlantic Regional Experiment (NARE) 1997.
-A comparison of alkyl nitrate to hydrocarbon ratios with a sequential reaction scheme model suggests our understanding of C4-Cs alkyl nitrate photochemistry is adequate to reproduce the observations.
-Air masses with larger ratios of alkyl nitrate to hydrocarbons were observed during NARE 1997 than STERAO 1996, consistent with the NARE 1997 measurements being in the more remote North Atlantic.
-These results suggest that information related to the photochemical age of an air  mass can be derived from alkyl nitrate to hydrocarbon ratios.
-
-\citep{Worton2010} TORCH
-Alkyl nitrates (C1eC5) were measured at two sites (near urban and rural) in southeast England during the Tropospheric Organic Chemistry Experiment (TORCH).
-Methyl nitrate was the dominant species during both campaigns accounting for on average about one third of the total measured alkyl nitrates.
-High mixing ratios (>50 pptv) and variability of methyl nitrate were observed at the near urban site (TORCH1) that were not seen at the rural site (TORCH2) and which could not be explained by local photochemical production or direct emissions.
-The diurnal variation of methyl nitrate during TORCH1 showed a morning maximum that would be consistent with nighttime chemistry followed by transport to the surface by boundary layer dynamics.
-Similarly, elevated morning mixing ratios were also observed during TORCH2 although the magnitudes were much smaller.
-As a result, methyl nitrate could represent a tracer for nighttime chemistry seen at the ground the following day.
-At both campaigns, the dominant source of short chain alkyl nitrates and carbonyl precursor radicals (?C4) were from decomposition of larger compounds.
-The magnitude of the source increased with decreasing carbon number consistent with increasing total precursor abundance.
-Non-photochemical emissions of acetaldehyde and acetone could not be accounted for by automobile exhaust emissions alone and indicated that other direct sources are likely important in this environment.
-Alkyl nitrate measurements were made at two sites (near urban
-and rural) as part of the TORCH experiment. Observed mixing ratios
-were higher (factor of ?2) during TORCH1, especially during the identified heatwave period.
-Methyl nitrate accounted for w30\% of the total measured C1eC5 species and during TORCH1 was often present at mixing ratios > 50 pptv.
-The diurnal profile of methyl nitrate showed an early morning maximum that we attributed to nighttime production from NO3 chemistry in the residual layer followed by transportation to the surface by boundary layer dynamics.
-Further work is needed to evaluate our hypothesis of a nighttime source of methyl nitrate but if proven, methyl nitrate could represent an interesting tracer of nighttime chemistry aloft.
-The predominant photochemical source of the short chain alkyl nitrates and carbonyl compounds was from the photochemical oxidation and decomposition of longer chain compounds as opposed to the oxidation of the parent hydrocarbons similar to previous studies.
-The magnitude of this source shows a preference for the shortest chain species consistent with the increasing number of precursor compounds identified in the MCM.
-Primary sources of acetaldehyde and acetone were also significant during TORCH1 and could not be accounted for by automobile exhaust emissions alone.
-Further work to identify and quantify emissions from additional direct sources is necessary if appropriate control strategies are to be implemented.
-
 \citep{deKock1994}
-Air samples were collected at Cape Recife (34 1' S 25 40'E) situated on the south-east coast of Southern Afi'ica (Figure 1).
+Air samples were collected at Cape Recife (34 1' S 25 40'E) situated on the south-east coast of Southern Afirica (Figure 1).
 The suburban, urban and industrial areas of Port Elizabeth lie approximately 10-40 kin north to north-west of the sampling site.
@@
 
-\citep{Beyersdorf2010}
-South Pole, other Antarctic, ANs only
-
 \subsubsection*{Oceanic sources}
-\citep{Newland2013}
-[An oceanic source for C1 – C3 alkyl nitrates was first postulated by Atlas et al.
-(1993) during the SAGA campaign in which it was observed that elevated mixing ratios of methyl and ethyl nitrates correlated well with equatorial regions of high productivity.
-This has since been confirmed by field measurements (e.g.
-Chuck et al., 2002; Blake et al., 2003; Dahl et al., 2005).
-Chuck et al.
-(2002) estimate an annual emission flux from equatorial Atlantic and Pacific Ocean regions of 230 and 68 Gg yr-1 for methyl and ethyl nitrate respectively.
-The oceanic source of more or equal C4 alkyl nitrates is insignificant compared to anthropogenic sources (Blake et al., 2003).
-The production mechanism of the alkyl nitrates is thought to be by the reaction ROO + NO in the aqueous phase with the peroxy radical coming from the photolysis of coloured dissolved organic matter (CDOM) and the NO coming from the photolysis of nitrite (Dahl et al., 2003).
-Direct emission from algae (Chuck et al., 2002) or from bacteria at depth (Dahl et al., 2007) has also been postulated.]
-
-\citep{Atlas1993}
-Alkyl Nitrates, Nonmethane Hydrocarbons, and Halocarbon Gases Over the Equatorial Pacific Ocean During Saga 3
-February March 1990
-The cruise covered a region of the equatorial Pacific Ocean from 15N to 10S latitude and 144 to 165W longitude.
-On this cruise we collected samples for the measurement of alkyl nitrates (RONO2) (C2-C5), nonmethane hydrocarbons (NMHC) and several halocarbon gases.
-Results:
-- Alkyl nitrates were found at levels higher than predicted from steady state calculations based on measured mixing ratios of hydrocarbons and NO.
-The measured levels of RONO2 suggest long-range transport as one mechanism contributing to elevated concentrations of alkyl nitrates in the remote troposphere.
-However, the distributions of C2 and C3 alkyl nitrates over the equator were  similar to the organobromine gases.
-This distribution suggests
-a possible oceanic source for alkyl nitrates to the atmosphere.
-Model: 1D 41 constituents incl. 3 ANs
-
-\citep{Chuck2002} September, October
-Measurements of methyl and ethyl nitrate in seawater and air samples along two Atlantic Ocean transects provide the FIRST direct evidence for an oceanic source of these compounds.
-A simple box model calculation suggests that the equatorial source could be a major component of the local atmospheric alkyl nitrate budget.
-In the remote atmosphere, the concentration of methyl nitrate (MeONO2) is expected to be low because of low concentrations of NOx and a branching ratio of less than 1\% for the formation of MeONO2
-from the reaction of alkylperoxy radicals with NO (8).
-However, unexpect- edly high levels of RONO2 in marine air and observations of a strong positive correlation between ethyl and propyl nitrate and the marine biogenically produced halocarbon, bromoform, in the marine boundary layer over the equatorial Pacific Ocean led Atlas et al.
-to propose an oceanic source for alkyl nitrates in the atmo- sphere (9).
-Subsequent atmospheric sampling campaigns, mostly conducted over the remote Pacific Ocean, have provided yet more evidence to support the hypothesis of an oceanic source (2, 10), with the contribution of alkyl nitrates to the NOy
-sum reaching 80\% in equatorial and high-latitude regions over the South Pacific (Blake, 1999).
-Although there have been no pub- lished measurements confirming the oceanic emission of these species, other studies have also invoked a seawater source for methyl, eth- yl, and propyl nitrates (11, 12).
-The supersaturation of surface waters over large areas of the Atlantic provides the first direct evidence for an oceanic source of MeONO2 and EtONO2.
-dominant source region is found between the equator and 40S.
-distinct mechanisms may be respon- sible for the production of methyl and ethyl nitrate, with different water types supporting different mechanisms.
-Methyl nitrate generally showed profiles more typically associated with algal production, with decreasing concentrations be- low the chlorophyll maximum.
-In contrast, in waters more strongly affected by continental influences, such as increased loading of dis- solved organic matter, the alkyl nitrates showed maximum concentrations in surface waters, which may suggest that photochemical production is important in shelf and coastal regions.
-
-\citep{Blake2003} oceanic+urban!!!
-Observations: concentration distributions of C1-C4 alkyl nitrates observed during the NASA airborne campaigns Pacific Exploratory Mission (PEM) -Tropics A (September– October 1996) and PEM-Tropics B (March–April 1999).
-The total geographic range for PEM-Tropics A was 45?N–72?S latitude and 153?E–75?W longitude, and for PEM- Tropics B was 40?N–36?S latitude and 149?E–75?W longitude.
-We observed high methyl nitrate (MeONO2,CH3ONO2) mixing ratios (approximately 50 pptv) at low altitudes in a latitude band between 8?Nto13?S stretching across the equatorial Pacific, illustrating the oceanic source of MeONO2.
-This source may be associated with the high-nutrient, low-chlorophyll character of equatorial Pacific waters.
-We discuss MeONO2 and ethyl nitrate (EtONO2,C2H5ONO2), whose abundance is dominated by equatorial oceanic sources, 2-Propyl nitrate (2-PrONO2,2-C3H7ONO2), which has significant oceanic and northern hemispheric (NH) sources associated with urban/ industrial hydrocarbon emissions, and 2-butyl nitrate (2-BuONO2 2-C4H8ONO2), which has mostly NH (northern hemisphere) sources.
-the ocean as the major source for MeONO2
-Ethyl nitrate has a dominant oceanic source in equatorial regions, but at mid NH latitudes a small additional source is associated with urban/industrial emissions.
-
 \citep{Reeves2007} ocean+urban+fire
@@
 
-\citep{Simpson2003} airborne too
-TRACE-P samples that did not show influence from marine alkyl nitrate sources were used to investigate photochemical alkyl nitrate evolution.
-Relative to 2-butyl nitrate/n-butane, the measured ratios of ethyl nitrate/ethane and 2-propyl nitrate/propane showed notable deviations from modeled values based on laboratory kinetic data, suggesting additional Asian sources of their alkyl peroxy radical precursors.
-By contrast, the measured ratios of 1-propyl-, 2-pentyl-, and 3-pentyl nitrate to their respective parent hydrocarbons were fairly close to modeled values.
-
-\citep{Roberts1998} marine MBL+urban ANs
-The measurements described herein were made as part of the
-Organic nitrates are a class of product species that contain
-North Atlantic Regional Experiment (NARE) during August and September 1993.
-These measurements consisted of NO, NO 2, particulate nitrate, HNO3, PAN, PPN, RONO 2, NOy, 0 3, CO, SO 2, and C2-C6 hydrocarbons at Chebogue Point.
-This site was ideally located to study the evolution of photochemically processed air masses in the marine boundary layer (MBL).
-Chebogue Point is approximately 500 km downwind from the eastern sea coast of the United States, and is often impacted by transport of polluted air from the continental source regions.
-Northwest Territories, Canada, have been made in the winter to
-The correlations between organic nitrates and CO or O3 have been presented by Roberts et al.
-[1996].
-It has been established that CO is a good tracer of continentally derived pollution at this site [Parrish et al., 1993].
-This continentally derived pollution contains hydrocarbons (HC) and NO x, the precursors of organic nitrates and ozone, and serves to link the sources of CO with the sources of organic nitrates and ozone.
-There are distinct differences in the lifetimes of CO, O3 , and organic nitrates, and the factors that they are dependent on.
-For these reasons, organic nitrates correlate to varying degrees with CO and O3.
-Measurements of the organic nitrates; PAN, PPN, methyl-,
-ethyl-, 1-propyl-, 2-propyl-, 2-butyl, and 3-pentyl nitrates were made at Chebogue Point, Nova Scotia, during the 1993 NARE campaign.
-The relationships of these compounds with other measures of continental pollution, for example, CO and 03, are consistent with our view of these compounds as products of HC-NO x photochemistry.
-Comparisons of the ratios with that of [2-butyl nitrate]/[butane] were used to test whether the actual data fit the line predicted from rate constants, branching ratios, photolysis rates, and assumed [OH].
-Abstract. Measurements of six alkyl nitrates, methyl-, ethyl-, 1-propyl-, 2-propyl-, 2-butyl, and 3-pentyl nitrates, and two peroxycarboxylic nitric anhydrides, peroxyacetic nitric anhydride
-(PAN) and peroxypropionic nitric anhydride (PPN) were made
-at Chebogue Point, Nova Scotia, during the 1993 North Atlantic
-Regional Experiment campaign. The correlation of alkyl nitrates
-(RONO2) with carbon monoxide makes it clear that the chemistry at this site is being driven by continental emissions, since RONO2 compounds are known to be by-products of HC-NOx photochemistry.
-Comparison of the ratios of alkyl nitrates to their parent hydrocarbons to that of 2-butyl nitrate/butane showed significant
-deviations from trends predicted from rate constants, branching
-ratios, and loss rates. For ethyl nitrate this deviation is likely due to a combination of many additional pathways  o the intermediate
-peroxy radical from the decomposition of larger alkoxy radicals.
-The propyl nitrates showed trends that differed from one another; 2-propyl nitrate was a factor of 2 to 3 higher than predicted, while 1-propyl nitrate was a factor of 8 to 10 higher than predicted.
-The trend of 3-pentyl nitrate/pentane was relatively close (about
-50\% lower) to the predicted trend.
-
-\subsubsection*{Biomass burning source}
-\citep{Newland2013}
-[Alkyl nitrates also have a biomass burning source (Friedli et al., 2001; Simpson et al., 2002).
-Simpson et al.
-(2002) report emissions of C1 – C4 alkyl nitrates from savanna fires in Australia, with methyl nitrate being primarily emitted during the ’flaming’ stage and ethyl – butyl nitrates mainly emitted during the ‘smouldering’ stage.
-They estimate global emissions from this source to be on the order of 8 Gg yr-1]
-\citep{Simpson2002}
-We report the first observations of the emission of five C1–C4 alkyl nitrates (methyl-, ethyl-, n-propyl-, i-propyl-, and 2-butyl nitrate) from savanna burning.
-Average alkyl nitrate mixing ratios in the immediate vicinity of three bushfires in Northern Australia were 47–122 times higher than local background mixing ratios.
-These are the highest alkyl nitrate mixing ratios we have ever detected, with maximum mixing ratios exceeding 3 ppbv for methyl nitrate.
-Methyl nitrate dominated the alkyl nitrate emissions during the flaming stage of savanna burning, whereas C2–C4 alkyl nitrates were mostly emitted during the smoldering stage.
-To explain the formation of alkyl nitrates from biomass burning, we propose a reaction mechanism involving the combination of reactive radicals at high temperature.
-Bearing in mind the uncertainties associated with extrapolating small data sets to much larger scales, alkyl nitrate emissions from global savanna burning are estimated to be on the order of 8 Gg/yr.
-
-\citep{Chuck2002}
-Use of the mean sea-to-air fluxes for 0 to 40S (Table 1) and measured atmospheric concentrations results in a lifetime of be- tween 4.5 and 25 days for MeONO2 and between 5 and 10 days for EtONO2 (22).
-Literature estimates of the atmospheric life- time of MeONO2, based on known atmospheric destruction processes, are poorly con- strained, ranging from 6 to 29 days (23, 24).
-For EtONO2
-, the atmospheric lifetime with respect to photodissociation at the equator and at 40S has been estimated to be 7 and 12 days, respectively, for October (25).
-
-\subsubsection*{Global distribution}
-\citep{Newland2013}
-[There have been a number of studies of the latitudinal and seasonal variations of alkyl nitrates in marine air (Blake et al., 2003), in the Arctic (Swanson et al., 2003) and in the Antarctic (Beyersdorf et al., 2010).
-Blake et al.
-(2003) report extensive measurements of C1 – C4 alkyl nitrates from the Pacific between 45N and 72S during the PEM-Tropics A and B projects.
-This study observed an equatorial band between about 8N and 13S in which the oceans are a source of alkyl nitrates.
-The highest mean methyl and ethyl nitrate measurements were found in this region.
-The highest mixing ratios of 2-propyl nitrate were found in the most northerly regions of the study (i.e.
-40-45N) but it also showed significantly elevated mixing ratios in the same equatorial band as the C1 – C2 nitrates.
-The high mixing ratios in the northern mid-latitudes are explained by production from anthropogenic polluted outflow.
-2-butyl nitrate showed almost no enhancement in equatorial regions and again had the highest mixing ratios in the most northerly regions of the study.
-Swanson et al.
-(2003) and Beyersdorf et al.
-(2010) observe a strong seasonality of alkyl nitrate mixing ratios in the Arctic and Antarctic respectively.
-Polar alkyl nitrate mixing ratios show a similar seasonality to the alkanes, with a spring-time peak just before polar sunrise and lows in the summer time.
-Winter time mixing ratios of 2-propyl and 2-butyl nitrate are about 10 ppt, while methyl and ethyl nitrate are about 5 – 6 ppt.
-In the summer 2-propyl and 2-butyl nitrate mixing ratios fall close to zero, while methyl and ethyl nitrate mixing ratios fall to about 3 ppt.
-The amplitude of the seasonal cycle reflects the reaction rate with OH, i.e.
-it increases with increasing carbon number.]
-
-\subsubsection{Mean AN concentrations}
+\subsubsection*{Mean AN concentrations}
 \begin{table}
@@
 \hline
-Observation & $RONO_2$                                  & Comments & Reference \\
+Environment   & Mean $RONO_2$ mixing ratio                & $RONO_2$ breakdown & Reference \\
 \hline
@@
 ???           & mean $\Sigma RONO_2$ 14 ppt               &          & \citep{Roberts1998} \\
-remote marine environments & 2-butyl and 2- and 3-pentyl: 0.8-37 pptv & &\citep{Atherton1989} [Atlas, 1988]\\
-urban environments         & 2-butyl and 2- and 3-pentyl?: approx. 2 ppb & & \citep{Atherton1989} [Hanst et al., 1982] \\
+remote marine & 2-butyl and 2- and 3-pentyl: 0.8-37 pptv & &\citep{Atherton1989} [Atlas, 1988]\\
+urban         & 2-butyl and 2- and 3-pentyl?: approx. 2 ppb & & \citep{Atherton1989} [Hanst et al., 1982] \\
 ??                         & MeONO2: 1.2 to 43 pptv     &          & \citep{Chuck2002} \\
 ??                         & EtONO2: 0.8 to 10 pptv     &          & \citep{Chuck2002} \\
-remote marine troposphere  & MeONO2: 0.2 to 50 pptv     & in lit. (2, 8, 9) & \citep{Chuck2002} \\
-remote marine troposphere  & EtONO2: 0.3 to 12.2 pptv   & in lit. (2, 8, 9) & \citep{Chuck2002} \\
+remote marine & MeONO2: 0.2 to 50 pptv     & in lit. (2, 8, 9) & \citep{Chuck2002} \\
+remote marine & EtONO2: 0.3 to 12.2 pptv   & in lit. (2, 8, 9) & \citep{Chuck2002} \\
 Pacific Ocean & MeONO2: 50 pptv            &                   & \citep{Blake2003} \\
@@
                                                         is sample closest to the plume center & \citep{Neuman2012} \\
+rural         & mean $\Sigma RONO_2$ (C1-C8): range: 30-630 ppt, mean over the whole period 120 ppt & & \citep{Flocke1998} \\
 \hline
@@
 
-\subsubsection{Comparison methodology}
+\subsubsection*{Comparison methodology}
 \citep{Sommariva2008}
@@
 
-\subsubsection{Modelling studies}
+\subsubsection*{Modelling studies}
 A huge number of modelling studies has been done about alkyl nitrates.
@@
 
-\begin{figure}\label{fig:Neu2008}
-\caption{1}
-\includegraphics[width=.5\linewidth]{{D:/UEA/ENV-MB4X_Dissertation/ENV-MB4X_Dissertation_thesis/ENV-MB4X_Dissertation_thesis_latex/pics/Neu2008_Fig.2_MeONO2_OCT_colorbar}.png}
-\end{figure}
-\begin{figure}
-\caption{1}
-\includegraphics[width=.5\linewidth]{{D:/UEA/ENV-MB4X_Dissertation/ENV-MB4X_Dissertation_thesis/ENV-MB4X_Dissertation_thesis_latex/pics/Neu2008_Fig.2_MeONO2_OCT}.png}
-\end{figure}
-\begin{figure}
-\caption{1}
-\includegraphics[width=.5\linewidth]{{D:/UEA/ENV-MB4X_Dissertation/ENV-MB4X_Dissertation_thesis/ENV-MB4X_Dissertation_thesis_latex/pics/Neu2008_Fig.3_OCT_NOxdiff}.png}
-\end{figure}
-\begin{figure}
-\includegraphics[width=.5\linewidth]{{D:/UEA/ENV-MB4X_Dissertation/ENV-MB4X_Dissertation_thesis/ENV-MB4X_Dissertation_thesis_latex/pics/Neu2008_Fig.3_OCT_O3contol_O3diff}.png}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{figure}[t!]\label{fig:Neu2008}
+	\centering
+	\begin{subfigure}[t]{0.5\textwidth}
+        \centering
+        \includegraphics[width=\linewidth]{{\picspath Neu2008_Fig.2_MeONO2_OCT}.png}
+        \caption{}
+    \end{subfigure}%
+    ~ % glues subplots on the same line somehow...
+    \begin{subfigure}[t]{0.5\textwidth}
+        \centering
+        \includegraphics[width=\linewidth]{{\picspath Neu2008_Fig.3_OCT_O3contol}.png}
+        \caption{}
+    \end{subfigure}
+    \begin{subfigure}[t]{0.5\textwidth}
+        \centering
+        \includegraphics[width=\linewidth]{{\picspath Neu2008_Fig.3_OCT_NOxdiff}.png}
+        \caption{}
+    \end{subfigure}%
+    ~
+    \begin{subfigure}[t]{0.5\textwidth}
+        \centering
+        \includegraphics[width=\linewidth]{{\picspath Neu2008_Fig.3_OCT_O3diff}.png}
+        \caption{}
+    \end{subfigure}
 \caption{October monthly mean mixing ratios of (a) methyl nitrate ($MeONO_2$, $CH_3ONO_2$) and (c) $O_3$, and percent difference (perturbation - control) in (b) $NO_x$, (d) $O_3$.
@@
 \end{figure}
-
-\citep{Atherton1989}
-The transport from urban areas can lead to alkyl nitrate concentrations of 4-83 pptv after 120- 240 hours.
-These levels are far higher than those possible from marine in-situ production.
-Additionally, alkyl nitrates may be more abundant than PAN in remote marine environments.
-Model: box
-Chemistry: explicit chemical reaction mechanism
-[Carter et al., 1986.
-A variety of alkyl nitrates, including methyl-, butyl-, pentyl-, hexyl-, heptyl-, octyl-, and nonyl-nitrates are allowed to form from fourteen different alkyl peroxy radicals.
-+ ocean emission sources fluxes of c
-for ethene, ethane, propene, and propane, NO, NO2
-Initial conditions: clear sky conditions, a quantum yield of 1, solar zenith angle of 45 degrees, and total ozone column of 8*10 cm-2.
-Daytime+nigthtime, invertion height, dilution.
-298 K.
-No deposition.
-Scenarios: Two, MLOW and MHIGH, represent the marine in-situ production of alkyl nitrates.
-The total hydrocarbon concentration was 17.6 ppbC.
-MLOW had NO = 4 pptv and NO2 = 4.6 pptv based on measurements over the Pacific Ocean.
-MHIGH had NO = 46 pptv, NO2 = 54 pptv, and PAN = 50 pptv.
-The two other scenarios, UHIGH and ULOW, simulate the transport of species from urban to remote environments.
-UHIGH contained 2150 ppbC and 140 ppbv NOx, and ULOW contained 270 ppbC and 18 ppbv NOx, giving each a HC/NOx ratio of 15.
-Simulation time: 12 hours MLOW and MHIGH, 240 hours UHIGH and ULOW.
-Results:
-- The total alkyl nitrate concentration present after 12 hours is less than 0.4 pptv for MHIGH and less than 0.1 pptv for MLOW.
-These levels are much lower than observations, indicating that marine in-situ production can account for only a fraction of the total marine alkyl nitrate concentration.
-- IHIGH The total alkyl nitrate concentration peaks during the first 6 hours and then drops to 3.1 ppbv after 12 hours.
-This value is com- parable with the 2 ppbv alkyl nitrate concentration de- termined in Los Angeles [Hanst et al., 1982]
-- The model predicts alkyl nitrate concentrations of 21-83 pptv for UHIGH and 4-12 pptv for ULOW for the hours 120-240.
-- General: The model predictions of 12-83 pptv (af- ter 120 hours) and 4-21 pptv (after 240 hours) are of the same order as observations.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 Using a subset of the Master Chemical Mechanism v3.1 (MCM) they constructed a highly detailed chemical model and run it for 5 days under two scenarios (Table \ref{tab:Sommariva2008init}) to explore the sensitivity of the evolution of the chemical composition of an air mass emitted from an urban environment to the initial conditions.
-As it seen in Figure ???
-according to the modelling results $O_3$ is produced and $NO_x$ is removed from the system with the consequent formation of $HNO_3$ via the reaction of $NO_2$ with $OH$, which broadly agrees with the observations [(Neuman et al., 2006).].
+As it seen in Figure \ref{fig:Sommariva2008fig} according to the modelling results $O_3$ is produced and $NO_x$ is removed from the system with the consequent formation of $HNO_3$ via the reaction of $NO_2$ with $OH$, which broadly agrees with the observations [(Neuman et al., 2006).].
 In scenario A the concentration of $NO_x$ after the first day is about 0.5–1 ppb for $NO_2$ and approximately 100 ppt for $NO$, which is high enough to sustain the production of $O_3$ to up to 135 ppb and the formation of alkyl nitrates during the simulation.
-In scenario B (Figure ???
-right) the increase in $O_3$ is slower (note that initial concentration is 0) and the maximum concentration after five days is about 100 ppb.
+In scenario B (Figure \ref{fig:Sommariva2008fig} right) the increase in $O_3$ is slower (note that initial concentration is 0) and the maximum concentration after five days is about 100 ppb.
 The concentration of $NO_x$ remains relatively high during the second (10–15 ppb) and third day (3–5 ppb) and then decreases to $<$200 ppt of $NO$ and $<$1ppb of $NO_2$ on the fourth and fifth days, therefore allowing the formation of alkyl nitrates too.
@@
 \end{table}
-\begin{figure}
+\begin{figure}\label{fig:Sommariva2008fig}
 \includegraphics[width=\linewidth]{{D:/UEA/ENV-MB4X_Dissertation/ENV-MB4X_Dissertation_thesis/ENV-MB4X_Dissertation_thesis_latex/pics/Sommariva2008_Fig.1}.png}
@@
 They also found that at lower $O_3$ and $CO$ initial concentrations and higher $NOx$ (scenario A) the importance of the reactions between alkane and $OH$ is typically lower during the first two-three days, and the contributions of peroxides and carboxylic acids become significant later and are on average less important.
-
-\citep{Newland2013}
-Model1: 2-D atmospheric model
-- to study the formation, loss and transport of the alkyl nitrates to the Arctic
-- to derive northern hemisphere emissions of the six alkanes.
-Chemistry:
-- (with fixed NOx and OH field taken from full chemistry run with model by Hough and Johnson (1990))
-Validation of ANs: comparison with measurements of alkyl nitrates in the Arctic.
-Performance: recreate both the measured seasonality of the alkyl nitrate, and the peak mixing ratio, reasonably well.
-Simulation time: 6 days?
-Model2: box
-Chemistry:
-- alkanesmethane, ethane, propane, butane and pentane and ANs explicitly described (RO+NO2 too)
-Model3: firn model
-- to derive atmospheric histories of six alkanes and six alkyl nitrates from measurements in firn air from the NEEM site, Greenland (methyl, ethyl, 2-propyl, 2-butyl, 2+3-pentyl and 3- methyl-2-butyl nitrate), 1950-2008.
-Results:
-- There are two ways in which NOx mixing ratios can limit alkyl nitrate production: either through air masses being NO limited with respect to the formation of a given alkyl nitrate or by the NOx in an air mass being 'used up' (track back to Reeves)
--- Alkyl nitrate production becomes NOx saturated above a given NO mixing ratio. This limit increases with increasing carbon number.
--- The time taken for an air mass to 'use up' all the initial NOx is dependent mainly on the initial NOx mixing ratio.
-- At high [NO2] (> 27 ppb), methyl nitrate production is dominated by the reaction of the alkoxy radical with NO2
-- The regions that contribute most to modeled alkane and alkyl nitrate mixing ratios in the Arctic are Europe and Russia
-- The importance of the RO + NO2 pathway compared to RO2 + NO is entirely dependent on the branching ratio (alpha3) of the RO2 + NO reaction pathway.
-The smaller alpha3 is, the more important the RO + NO2 pathway is.
--- At 10 ppb of NO2 the RO + NO2 pathway accounts for 30 – 60\% of methyl nitrate production.
-At 100 ppb it accounts for 85 – 95\% of methyl nitrate production and almost 20\% of ethyl nitrate production.
-- The rate of alkyl nitrate production in a polluted air mass is proportional to the OH concentration in that air mass.
 
@@
 
-[Source characterization and photochemical processes in
-transported air masses have been analyzed by examining
-RONO2 formation from anthropogenic emissions of alkane
-precursors [Simpson et al., 2003; Madronich, 2006; Simpson
-et al., 2006; Reeves et al., 2007; Worton et al., 2010]. (Neuman2012)]
-
-They are produced photochemically through the oxidation of their parent hydrocarbons \citep{Roberts1990} or emitted directly from equatorial oceans \citep{Blake2003} and biomass burning \citep{Simpson2002}.
-Being involved in the ozone production chemistry, these species can act as reservoirs of nitrogen oxides which availability largely affects the production of ozone \citep{Reeves2007}.
-Since the alkyl nitrates have atmospheric lifetimes ranging from several days to about a month, they can redistribute nitrogen oxides presented over polluted continental regions to clean remote marine environments, and therefore are considered as one of the tracers for determining the anthropogenic influence \citep{Atherton1989, Reeves2007, Worton2005}.
-
-Owing to the existence of the relationship between alkyl nitrates and their parent hydrocarbons, measurements of their concentrations can provide information on the extent of photochemical processing in the air mass.
-\cite{Bertman1995} developed a kinetic approach to describe this relationship through the evolution of the ratio of alkyl nitrate to its parent hydrocarbon as a function of time.
-This approach has been extensively used to study air mass photochemical ageing at several sites, in North America \citep{Bertman1995, Roberts1998}, China \citep{Simpson2006}, the British Isles \citep{Worton2010} and over the Pacific \citep{Simpson2003} and North Atlantic Ocean \citep{Reeves2007}.
-
-\cite{Bertman1995} found that the theoretical ratios of  2-butyl nitrate to n-butane and 2- and 3-pentyl nitrates to pentane strongly agree with those measured from the ambient air showing no evidence for primary sources of these nitrates.
-On the contrary, measurements of the other ratios, ethyl/ethane, n-propyl nitrate/propane, and 2-propyl nitrate/propane, appeared to be significantly higher than predicted, relative to the 2-butyl nitrate/n-butane ratio, meaning that there should be an additional source of these alkyl nitrates.
-\cite{Roberts1998} generally approved the theory by comparing Bertman's conclusions with the results from a new observational dataset.
-
 Using methyl nitrate as a tracer of marine $RONO_2$ \cite{Simpson2006} revealed that the South China Sea in not a region of strong $RONO_2$ emissions suggesting that photochemical rather marine production of alkyl nitrates is dominant at the site.
-
-\cite{Worton2010} found that the photochemical source of the short-chain alkyl nitrates originates mostly from the photochemical oxidation and decomposition of longer-chain compounds rather than from the oxidation of the parent hydrocarbons as it is thought in the previous studies.
 

### 2015-08-04 MSc_diss.tex
Updates in mean ANs mixing ratios (before renaming ANs)

@@
 \usepackage{epstopdf}
+% Tables
+\usepackage{multirow}
 % Tikz styles
@@
 The only way to answer these questions is to perform a rigorous investigation of the factors that drive tropospheric ozone and attribute those differences to respective factors \citep{Young2012}.
-
-MAKE GRAPHIC FROM TABLE ABOUT O3 BUDGETS!
 
@@
 
-\subsubsection*{Mean AN concentrations}
-\begin{table}
+\begin{table}%[ht] % mean AN mixing ratios
 \begin{tabular}{cccc}
 \hline
-Environment   & Mean $RONO_2$ mixing ratio                & $RONO_2$ breakdown & Reference \\
-\hline
-ITOP?         & mean $\Sigma RONO_2$ 10 ppt               &          & \citep{Reeves2007} \\
-???           & mean $\Sigma RONO_2$ 14 ppt               &          & \citep{Roberts1998} \\
-remote marine & 2-butyl and 2- and 3-pentyl: 0.8-37 pptv & &\citep{Atherton1989} [Atlas, 1988]\\
-urban         & 2-butyl and 2- and 3-pentyl?: approx. 2 ppb & & \citep{Atherton1989} [Hanst et al., 1982] \\
-??                         & MeONO2: 1.2 to 43 pptv     &          & \citep{Chuck2002} \\
-??                         & EtONO2: 0.8 to 10 pptv     &          & \citep{Chuck2002} \\
-remote marine & MeONO2: 0.2 to 50 pptv     & in lit. (2, 8, 9) & \citep{Chuck2002} \\
-remote marine & EtONO2: 0.3 to 12.2 pptv   & in lit. (2, 8, 9) & \citep{Chuck2002} \\
-Pacific Ocean & MeONO2: 50 pptv            &                   & \citep{Blake2003} \\
-airborne      & $\Sigma RONO_2$: approx 1 ppb         & difference between NOy and (NOx+PAN+HNO3) 
-                                                        is sample closest to the plume center & \citep{Neuman2012} \\
-rural         & mean $\Sigma RONO_2$ (C1-C8): range: 30-630 ppt, mean over the whole period 120 ppt & & \citep{Flocke1998} \\
+Environment     & Mean $\Sigma RONO_2$ & $\Sigma RONO_2$ breakdown & Reference \\
+\hline
+remote marine   & 10 pptv    & methyl                   & \citep{Reeves2007} \\
+(North Atlantic)&            & ethyl                    & \\
+airborne        &            & 1-propyl                 & \\   
+                             && 2-propyl                & \\   
+                             && 2-butyl                 & \\                                                                                               
+                             && 2-methyl-3-butyl        & \\
+                             && sum of 2- and 3-pentyl  & \\
+\hline
+rural           & 14 pptv    & methyl                   & \citep{Roberts1998} \\
+(Nova Scotia, Canada)        &&  ethyl   & \\
+stationary                   && n-propyl & \\
+Chebogue Point               && 2-propyl & \\
+                             && 2-butyl  & \\ 
+\hline
+rural         & 120 ppt (30-630 ppt) & $C_1$-$C_8$       & \citep{Flocke1998} \\
+(Black Forest, Germany)      &&& \\ 
+stationary                   &&& \\
+\hline
+polluted marine & approx 1 ppb        & $NO_y$-($NO_x$+$HNO_3$+$PAN$) & \citep{Neuman2012} \\
+(Gulf of Mexico)&(plume center) && \\
+airborne &&& \\
 \hline
@@
 \end{table}
+%??                         & MeONO2: 1.2 to 43 pptv     &          & \citep{Chuck2002} \\
+%??                         & EtONO2: 0.8 to 10 pptv     &          & \citep{Chuck2002} \\
+%remote marine & MeONO2: 0.2 to 50 pptv     & in lit. (2, 8, 9) & \citep{Chuck2002} \\
+%remote marine & EtONO2: 0.3 to 12.2 pptv   & in lit. (2, 8, 9) & \citep{Chuck2002} \\
+%Pacific Ocean & MeONO2: 50 pptv            &                   & \citep{Blake2003} \\
 

### 2015-08-04 MSc_diss.tex
Finished draft of literature review and introduction

@@
 
+The brief description of the measurement sites and data is presented in section \ref{sec:data}.
+In section \ref{sec:method} the methodology used in this work is defined.
+Section \ref{sec:res} contains the results.
+Section \ref{sec:discuss} discusses the results, particularly focusing attention on the analysis of the discrepancy between the observational data and the results of theoretical modelling.
+Conclusions constitute section \ref{sec:conclusion}.
+
 \section{Literature review} \label{sec:lit_review}
 A lot of work is has been dedicated to measurements of alkyl nitrates in different environments and even more to measurements of ozone.
-The majority of these studies have been done in a observational-modelling framework.
-
-During/After one of these investigations, namely the SAGA campaign, \citep{Atlas1993} first postulated that the short-chained alkyl nitrates ($C_1$-$C_3$) can be produced not only photochemically through the oxidation of their parent hydrocarbons \citep{Roberts1990}, but also emitted directly from equatorial oceans.
+Here we focus our attention on some major investigations that lead us to our current understanding of the relationship between tropospheric ozone and alkyl nitrates.
+
+During one of these investigations, namely the SAGA campaign, \citep{Atlas1993} first postulated that the short-chained alkyl nitrates ($C_1$-$C_3$) can be produced not only photochemically through the oxidation of their parent hydrocarbons \citep{Roberts1990}, but also emitted directly from equatorial oceans.
 This has since been confirmed by field measurements \citep{Chuck2002,Blake2003,Dahl2005}.
@@
 There have been a number of studies of the seasonal variations of alkyl nitrates in the Arctic \citep{Swanson2003} and in the Antarctic \citep{Beyersdorf2010}.
-Polar alkyl nitrate follow the seasonality of the alkanes displaying a clear late winter maximum and a broad summer minimum.
-
-Owing to the existence of the relationship between alkyl nitrates and their parent hydrocarbons, measurements of their concentrations can provide information on the extent of photochemical processing in the air mass.
-\cite{Bertman1995} developed a kinetic approach to describe this relationship through the evolution of the ratio of alkyl nitrate to its parent hydrocarbon as a function of time.
-This approach has been extensively used to study air mass photochemical ageing at several sites, in North America \citep{Bertman1995,Roberts1998}, China \citep{Simpson2006}, the British Isles \citep{Worton2010} and over the Pacific \citep{Simpson2003} and North Atlantic Ocean \citep{Reeves2007,Stroud2001}.
-
-\subsubsection*{Continent}
-\citep{Day2003}
-We describe measurements of NO2, total peroxy nitrates (?PNs), total alkyl nitrates (?ANs), and HNO3 using thermal dissociation followed by laser-induced fluorescence detection of NO2 at three continental locations.
-At a rural site in California, measurements over a full annual cycle show that ?ANs are routinely 10–20\% of NOy.
-rural site, at a suburban site in California and an urban site in Houston, Texas [rural (UC-BFRS), suburban (Granite Bay) and urban (La Porte)]
-SumANs a more broad term than usual.
-?AN mixing ratios at the UC-BFRS exhibit a strong seasonal cycle with a minimum in winter and a maximum in summer.
-The seasonal transition is sharp.
-Concentrations of ?ANs rise rapidly from 150 ppt at the end of April to 500 ppt by early June.
-We also show that ?ANs are a large part, if not all, of the ‘‘missing NOy’’.
-
-\citep{Simpson2006} Pearl River Delta
-Mixing ratios of seven C1–C5 alkyl nitrates (RONO2) were measured during a 16-month study (August 2001–December
-2002) at Tai O, a coastal site 30km west of central Hong Kong in the Pearl River Delta, the fastest-growing industrial region in the world.
-The C3–C4 (rather than C1–C2) RONO2 were most abundant throughout the study, showing the importance of photochemical (rather than marine) RONO2 production in the sampled air.
-A lack of methyl nitrate (MeONO2) enhancement during summer, when the prevailing winds are from the ocean, indicates that the South China Sea is not a region of strong RONO2 emissions.
-By contrast, MeONO2 levels during pollution episodes (up to 25 parts per trillion by volume (pptv)) were the highest that our group has recorded during urban photochemical RONO2 production, as opposed to marine emissions or biomass burning.
-The highest summed RONO2 level of the study (204 pptv) was measured in the afternoon of 7 November 2002, during an intense pollution episode that captured the highest ozone (O3) level ever recorded in Hong Kong (203 ppbv).
-During pollution episodes, the average ratio of O3 to summed RONO2 was
-roughly 1000:1 in freshly polluted air (ethyne/CO?3–5 pptv/ppbv) and 500:1 in very freshly polluted air (ethyne/ CO?6–8 pptv/ppbv).
-Ozone and RONO2 share a common photochemical source, and their good correlation in pollution plumes shows that RONO2 can be used as a tracer of photochemical O3 production.
-Even MeONO2 showed similar
-diurnal variations as the C2–C5 RONO2, indicating a strong photochemical source despite its very slow photochemical production from methane oxidation.
-The decomposition of longer-chain alkoxy radicals also does not explain the high MeONO2 levels, and rough calculations show that methoxy radical reaction with NO2 appears to be a viable alternate pathway for MeONO2 production in polluted atmospheres, though further measurements and modeling are required to confirm this mechanism.
-\citep{Comment}
-It has generally been assumed that reaction (3) dominates over reaction (2) and that production of CH3ONO2 would be small.
-Simpson et al.
-(2006) suggest that model studies would be useful to test their theory.
-In this short comment on their paper, some model simulations are presented that confirm their hypothesis.The model simulations have been carried out using a box model running the CRI v1.0 mechanism (Jenkin et al., 2002), which is a reduced scheme of the Master Chemical Mechanism (Jenkin et al., 1997).
-The box model was integrated using emissions of NOx and VOC taken from the UK National emissions inventory and would be repre- sentative therefore of European rather than Asian emission profiles but are still instructive.
-The model is designed to simulate regional photochemical episodes and runs during European summer time (mid June) under moderately high temperature (approximately 25 1C) and relative humidity (70\%).
-The concentration of CH3ONO2 from the two routes, reaction (1b) and reaction (2), have been compared as a function of NOx and are summarised in Fig.
-1.
-It is clear that above about 10 ppb of NO2 channel (2) becomes important as a source of CH3ONO2 and at around 35 ppb starts to become the dominant channel in the simulations performed.
-
-\citep{deKock1994}
-Air samples were collected at Cape Recife (34 1' S 25 40'E) situated on the south-east coast of Southern Afirica (Figure 1).
-The suburban, urban and industrial areas of Port Elizabeth lie approximately 10-40 kin north to north-west of the sampling site.
-Mean sumANs 12.7.
-Variations in absolute and relative concentrations of the alkyl nitrates could be related to changes in meteorological conditions.
-It is considered that the differences observed in chemical composition may represent different sources of trace gas emissions or different degrees of photochemical processing of similar sources.
-The meas~ed levels of the alkyl nitrates suggest long-range transport to play a major role in regulating the distribution of alkyl nitrates in the troposphere.
-
-\subsubsection*{Oceanic sources}
-\citep{Reeves2007} ocean+urban+fire
-This paper is based on alkyl nitrate measurements made over the North Atlantic as part of the International Consortium for Research on Atmospheric Transport and Transformation (ICARTT).
-The focus is on the analysis of air samples collected on the UK BAe-146 aircraft during the Intercontinental Transport of Ozone and Precursors (ITOP) project, but air samples collected on board the NASA DC-8 and NOAA WP-3D aircraft as part of a Lagrangian experiment are also used.
-The ratios between the alkyl nitrates and their parent hydrocarbons are compared with those expected from chemical theory.
-Further, a box model is run to investigate the temporal evolution of the alkyl nitrates in three Lagrangian case studies and compared to observations.
-The general agreement between the alkyl nitrate data and photochemical theory suggests that during the first few days of transport from the source region, photochemical production of alkyl nitrates, and thus ozone, had taken place.
-Concentrations of methyl, ethyl, 1-propyl, 2-propyl, 2-butyl, 2-methyl-3-butyl and the sum of 2- and 3-pentyl nitrates (hereafter referred to as 2 + 3-pentyl nitrate) are presented.
-The alkyl nitrate concentrations observed during
-ITOP are typical of those found a few days downwind of precursor source regions.
-Although there is some evidence for an oceanic source of methyl nitrate, for the most part the air samples collected during ITOP do not appear to be strongly influenced by oceanic sources of the alkyl nitrates.
-Elevated concentrations were observed in air masses which have been influenced by emissions from the N.E.
-United States and from Alaskan fires.
-The large variations in ratios between the C5 and C2-C4 nitrates suggest that emission ratios for the parent hydrocarbons vary between different regions of the N.E.
-United States.
-[49] In general the alkyl nitrate data agrees well with photochemical theory.
-This suggests that at least during the first few days of transport from the source region photo- chemical production of alkyl nitrates had taken place.
-This implies that ozone production had also occurred.
-There also appears to be a widespread common source of ethyl nitrate and 1-propyl nitrate other than from their parent hydro- carbons.
-This might be from longer-chain hydrocarbons, possibly pentanes.
-[50] The observations in the more photochemically pro- cessed air masses are consistent with the concentration of NO dropping to a point where the rates of the alkyl nitrate production reactions no longer dominate the peroxy radical self/cross reactions.
-However, in these low levels of NO, where the balance between ozone production and destruc- tion is highly sensitive to NOX, the alkyl nitrates might contribute to maintaining the ozone levels in air as it undergoes long-range transport.
-[51] The observed patterns of the alkyl nitrates also suggest that the rates of photochemical processing in the Alaskan smoke plumes were small and that the rates of ozone formation in these air masses were low.
-
-\begin{table}%[ht] % mean AN mixing ratios
+They show that polar alkyl nitrate follow the seasonality of the alkanes displaying a clear late winter maximum and a broad summer minimum.
+In temperate latitudes, seasonal variations of alkyl nitrate depend more on the origin of an air mass.
+In polluted air masses, the mixing ratios of alkyl nitrates exhibited summer maximum and strong daily variations, with maximum in the early evening.
+In clean air masses, alkyl nitrate mixing ratios showed a summer minimum and no pronounced diurnal variation \citep{Flocke1998}.
+Mean mixing ratios of the sum of alkyl nitrates measured in different environments can be found in Table \ref{tab:ANmean}.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{table}\label{tab:ANmean} % mean AN mixing ratios
 \begin{tabular}{cccc}
@@
 (Nova Scotia, Canada)        && ethyl   & \\
-stationary                   && n-propyl & \\
-Chebogue Point               && 2-propyl & \\
+ground                       && n-propyl & \\
+                             && 2-propyl & \\
                              && 2-butyl  & \\ 
 \hline
-rural         & 120 ppt (30-630 ppt) & $C_1$-$C_8$       & \citep{Flocke1998} \\
-(Black Forest, Germany)      &&& \\ 
-stationary                   &&& \\
+suburban                     & 17.48 pptv  & 2-propyl    & \citep{DeKock1994} \\
+(Cape Retire, South Africa)  &(3.62-122.23)& 1-propyl & \\
+ground                                    && 2-butyl  & \\
+                                          && 1-butyl  & \\
+                                          && 3-pentyl & \\
+                                          && 2-pentyl & \\
+                                          && 2-methyl-1-butyl & \\
+                                          && i-pentyl & \\
+\hline
+suburban                     & 119.8 ppt  & $C_1$-$C_8$       & \citep{Flocke1998} \\
+(Black Forest, Germany)      & (30-630 ppt) && \\ 
+ground                                     &&& \\
 \hline
@@
 \end{table}
-%??                         & MeONO2: 1.2 to 43 pptv     &          & \citep{Chuck2002} \\
-%??                         & EtONO2: 0.8 to 10 pptv     &          & \citep{Chuck2002} \\
-%remote marine & MeONO2: 0.2 to 50 pptv     & in lit. (2, 8, 9) & \citep{Chuck2002} \\
-%remote marine & EtONO2: 0.3 to 12.2 pptv   & in lit. (2, 8, 9) & \citep{Chuck2002} \\
-%Pacific Ocean & MeONO2: 50 pptv            &                   & \citep{Blake2003} \\
-
-\subsubsection*{Comparison methodology}
-\citep{Sommariva2008}
-[The ratios between the alkyl nitrates and their
-main alkane precursors can also be plotted against one another, for example against 2-butylnitrate/ n-butane as in Fig.
-4.
-In this way the modelled and the measured concentrations can be compared without taking into account their variation with time.
-A similar approach was used in previous studies to investigate the sources of alkyl nitrates and the photochemical processing of polluted air masses (e.g., Bertman et al., 1995; Flocke et al., 1998; Roberts et al., 1998; Stroud et al., 2001; Simpson et al., 2003; Reeves et al., 2007).
-These studies compared plots of the measured ratios against 2-butylnitrate/n-butane to the ratios calculated with a simple sequential reaction model which assumed a single precursors for each alkyl nitrate.
-Usually, the underestimation of the ratios for short- chain compounds (propylnitrates/propane, ethylni- trate/ethane) was interpreted as an indication of photochemical sources other than the main alkane precursor.]
-
-\citep{Day2003}
-We also show that sumANs are a large part, if not all, of the 'missing NOy' reported in many prior experiments and emphasize that the ratios of sumANs/NOz and of O3/sumANs are especially useful for evaluating chemical models and comparing observations at different sites.
-
-\subsubsection*{Modelling studies}
-A huge number of modelling studies has been done about alkyl nitrates.
-In particular, O3 and the RONO2 oxidation products formed from reactions between these NOx and VOC emissions are examined in detail...
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+Owing to the existence of the relationship between alkyl nitrates and their parent hydrocarbons, measurements of their concentrations can provide information on the extent of photochemical processing in the air mass.
+\cite{Bertman1995} developed a kinetic approach to describe this relationship through the evolution of the ratio of alkyl nitrate to its parent hydrocarbon as a function of time.
+This approach has been extensively used to study air mass photochemical ageing at several sites, in North America \citep{Bertman1995,Roberts1998}, China \citep{Simpson2006}, the British Isles \citep{Worton2010} and over the Pacific \citep{Simpson2003} and North Atlantic Ocean \citep{Reeves2007,Stroud2001}.
 
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \begin{figure}[h!]\label{fig:Neuman2012}
@@
 \end{figure}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+Probably the most detailed investigation of the formation processes and photochemical evolution of alkyl nitrates (but not their impact on $O_3$) has been done by \citep{Sommariva2008}.
+Using a subset of the Master Chemical Mechanism v3.1 (MCM) they constructed a highly detailed chemical model and run it for 5 days under two scenarios (Table \ref{tab:Sommariva2008init}) to explore the sensitivity of the evolution of the chemical composition of an air mass emitted from an urban environment to the initial conditions.
+As it seen in Figure \ref{fig:Sommariva2008fig} according to the modelling results $O_3$ is produced and $NO_x$ is removed from the system with the consequent formation of $HNO_3$ via the reaction of $NO_2$ with $OH$, which broadly agrees with the observations [(Neuman et al., 2006).].
+In scenario A the concentration of $NO_x$ after the first day is about 0.5–1 ppb for $NO_2$ and approximately 100 ppt for $NO$, which is high enough to sustain the production of $O_3$ to up to 135 ppb and the formation of alkyl nitrates during the simulation.
+In scenario B (Figure \ref{fig:Sommariva2008fig} right) the increase in $O_3$ is slower (note that initial concentration is 0) and the maximum concentration after five days is about 100 ppb.
+The concentration of $NO_x$ remains relatively high during the second (10–15 ppb) and third day (3–5 ppb) and then decreases to $<$200 ppt of $NO$ and $<$1ppb of $NO_2$ on the fourth and fifth days, therefore allowing the formation of alkyl nitrates too.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{table}\label{tab:Sommariva2008init}
+\caption{Initial values of the some model parameters (in ppb) for two scenarios used in \citep{Sommariva2008}}
+\begin{tabular}{ccc}
+\hline
+Parameter & Initial value Scenario A & Scenario B \\
+\hline
+$CO$      & 332.5 & 232.5 \\
+$NO_x$    & 34.9  & 46.5  \\
+$O_3$     & 90    & 0   \\
+\hline
+\end{tabular}
+\end{table}
+\begin{figure}\label{fig:Sommariva2008fig}
+\includegraphics[width=\linewidth]{{D:/UEA/ENV-MB4X_Dissertation/ENV-MB4X_Dissertation_thesis/ENV-MB4X_Dissertation_thesis_latex/pics/Sommariva2008_Fig.1}.png}
+\caption{Modelled concentrations of $O_3$, $NO_x$ and $HNO_3$ for scenario A (left) and B (right). Source: \citep{Sommariva2008}.}
+\end{figure}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+Comparison of modelling results with the values observed inside a typical urban plume sampled using the NOAA WP-3D aircraft from the New York City area during the New England Air Quality Study (NEAQS) 2004 campaign enabled \citep{Sommariva2008} to conclude that the chemistry of alkyl nitrates is well represented in the MCM model.
+They found that short-chain $<C_4$ alkyl nitrates can be formed from several precursors, while long-chain ($C_5$) alkyl nitrates are produced for 90\% or more from the oxidation of a single parent alkane.
+They also found that at lower $O_3$ and $CO$ initial concentrations and higher $NOx$ (scenario A) the importance of the reactions between alkane and $OH$ is typically lower during the first two-three days, and the contributions of peroxides and carboxylic acids become significant later and are on average less important.
 
@@
 
-Probably the most detailed investigation of the formation processes and photochemical evolution of alkyl nitrates (but not their impact on $O_3$) has been done by \citep{Sommariva2008}.
-Using a subset of the Master Chemical Mechanism v3.1 (MCM) they constructed a highly detailed chemical model and run it for 5 days under two scenarios (Table \ref{tab:Sommariva2008init}) to explore the sensitivity of the evolution of the chemical composition of an air mass emitted from an urban environment to the initial conditions.
-As it seen in Figure \ref{fig:Sommariva2008fig} according to the modelling results $O_3$ is produced and $NO_x$ is removed from the system with the consequent formation of $HNO_3$ via the reaction of $NO_2$ with $OH$, which broadly agrees with the observations [(Neuman et al., 2006).].
-In scenario A the concentration of $NO_x$ after the first day is about 0.5–1 ppb for $NO_2$ and approximately 100 ppt for $NO$, which is high enough to sustain the production of $O_3$ to up to 135 ppb and the formation of alkyl nitrates during the simulation.
-In scenario B (Figure \ref{fig:Sommariva2008fig} right) the increase in $O_3$ is slower (note that initial concentration is 0) and the maximum concentration after five days is about 100 ppb.
-The concentration of $NO_x$ remains relatively high during the second (10–15 ppb) and third day (3–5 ppb) and then decreases to $<$200 ppt of $NO$ and $<$1ppb of $NO_2$ on the fourth and fifth days, therefore allowing the formation of alkyl nitrates too.
-
-\begin{table}\label{tab:Sommariva2008init}
-\caption{Initial values of the some model parameters (in ppb) for two scenarios used in \citep{Sommariva2008}}
-\begin{tabular}{ccc}
-\hline
-Parameter & Initial value Scenario A & Scenario B \\
-\hline
-$CO$      & 332.5 & 232.5 \\
-$NO_x$    & 34.9  & 46.5  \\
-$O_3$     & 90    & 0   \\
-\hline
-\end{tabular}
-\end{table}
-\begin{figure}\label{fig:Sommariva2008fig}
-\includegraphics[width=\linewidth]{{D:/UEA/ENV-MB4X_Dissertation/ENV-MB4X_Dissertation_thesis/ENV-MB4X_Dissertation_thesis_latex/pics/Sommariva2008_Fig.1}.png}
-\caption{Modelled concentrations of $O_3$, $NO_x$ and $HNO_3$ for scenario A (left) and B (right). Source: \citep{Sommariva2008}.}
-\end{figure}
-
-Comparison of modelling results with the values observed inside a typical urban plume sampled using the NOAA WP-3D aircraft from the New York City area during the New England Air Quality Study (NEAQS) 2004 campaign enabled \citep{Sommariva2008} to conclude that the chemistry of alkyl nitrates is well represented in the MCM model.
-They found that short-chain $<C_4$ alkyl nitrates can be formed from several precursors, while long-chain ($C_5$) alkyl nitrates are produced for 90\% or more from the oxidation of a single parent alkane.
-They also found that at lower $O_3$ and $CO$ initial concentrations and higher $NOx$ (scenario A) the importance of the reactions between alkane and $OH$ is typically lower during the first two-three days, and the contributions of peroxides and carboxylic acids become significant later and are on average less important.
-
-\subsubsection*{Other}
-
 To sum up, alkyl nitrates are produced as a minor branch of the reaction of peroxy radicals with NO \citep{Day2003}.
@@
 Despite their relative importance alkyl nitrates are usually not included in the climate-chemistry models, for example in UKCA (Reeves, personal communication), which potentially could lead to discrepancies between models and observations described before.
-
-Using methyl nitrate as a tracer of marine $RONO_2$ \cite{Simpson2006} revealed that the South China Sea in not a region of strong $RONO_2$ emissions suggesting that photochemical rather marine production of alkyl nitrates is dominant at the site.
-
-The brief description of the measurement sites and data is presented in section \ref{sec:data}.
-In section \ref{sec:method} the methodology used in this work is defined.
-Section \ref{sec:res} contains the results.
-Section \ref{sec:discuss} discusses the results, particularly focusing attention on the analysis of the discrepancy between the observational data and the results of theoretical modelling.
-Conclusions constitute section \ref{sec:conclusion}.
 
@@
 Above the 'ridge line' the $VOC$/$NO_x$ ratios are low, and the $O_3$ production rate increases with increasing $VOC$ and decreases with increasing $NO_x$ \citep{Sillman2013}.
-
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \begin{figure} % Sillman1999withdotsline and netO3rate_withINORG_NOx_1ppb_100ppb
@@
 \end{figure}
-
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 Rewrite!
@@
 Branching ratios used in the model are taken from the MCM v3.3, and their values (Table \ref{tab:ANbranching}) were kept constant in all model runs.
-
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \begin{table} %  AN branching ratios
@@
 \end{table}
-
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 Figure \ref{fig:netO3rate_noAN_withAN_diff} shows isopleths of net $O_3$ production rate as a function of $NO_x$ and sum of $VOCs$ initial concentrations from two series of model runs (with and without alkyl chemistry) and the difference between them.
 Since the actual formation of $O_3$ in our model starts at 250 ppt of $NO_x$, Figure \ref{fig:netO3rate_noAN_withAN_diff} summarises modelling data from experiments with $NO_x$ initial mixing ratios from 250 ppt till 100 ppb.
-
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \begin{figure} % netO3rate noAN withAN diff
@@
 \end{figure}
-
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 As expected, the inclusion of alkyl nitrate chemistry slows down the net $O_3$ production rate.
@@
 
-\citep{Reeves2007}
-[In the 'marine' air masses methyl nitrate shows a slightly higher mean concentration to 'other' air masses (2.40 compared to 2.29 ppt), but the concentrations of the other alkyl
-nitrates are lower by around 30–50% with the exception of
-ethyl nitrate which is lower by about 15% (1.33 compared to
-1.57 ppt), such that it becomes the second most abundant alkyl
-nitrate.
-The highest alkyl
-nitrate concentration was observed during flight B031 (air
-from New York) for 2-methyl-3-butyl nitrate (13.7 ppt),
-when the maximum concentration of 2 + 3-pentyl nitrate
-(9.0 ppt) was also measured. It is quite possible that some of
-the ‘‘marine’’ maximum concentrations of the alkyl nitrates
-were impacted by terrestrial emissions (e.g., 5.4 pptv of
-2-methyl-3-butyl nitrate and 1.7 pptv for 2 + 3-pentyl nitrate
-observed during flight B030).
-
-The mean NO concentration observed during ITOP (20 plus minus 31 pptv (Lewis et al., 2007)) is below the limit of 100–200 pptv at which the efficiency of the alkyl nitrate formation is reduced because of competition between the reaction of ROO with NO and ROO self/cross reactions (Roberts et al., 1998)]
-
 \begin{figure} % endtotalANs

### 2015-08-04 MSc_diss.tex
Finished section about O3 and CBs

@@
 This capacity depends mainly on two factors: climate and the global burden of the principal oxidants such as ozone ($O_3$), the hydroxyl radical ($OH$) and hydrogen peroxide ($H_2O_2$) \citep{Prinn2003,Thompson1992}.
-\subsection{Tropospheric ozone}
+\subsection{Tropospheric ozone}\label{sec:into_O3}
 Tropospheric ozone ($O_3$) plays an important role in the chemical and thermal balance of the atmosphere.
@@
 
-\subsubsection*{Ozone tropospheric chemistry}
+\subsubsection*{Ozone tropospheric chemistry}\label{intro_O3chem}
 %\input{O3CycleScheme}
@@
 
-\subsection{Alkyl nitrates}
+\subsection{Alkyl nitrates}\label{sec:intro_AN}
 Alkyl nitrates ($RONO_2$) are organic trace gases, which through one of their formation mechanisms are closely linked with the production of tropospheric ozone \citep{Reeves2007}.
 This mechanism is present in the following section.
-\subsubsection*{Alkyl nitrates chemistry}
+\subsubsection*{Alkyl nitrates chemistry}\label{into_ANchem}
 The photochemical formation of the alkyl nitrates starts when chemistry of the troposphere reaches the regime of $O_3$ production, i.e.
@@
 It is a reduced chemical mechanism that describes $O_3$ formation from the tropospheric degradation of methane and 115 emitted non-methane hydrocarbons and oxygenated volatile organic compounds developed using the Master Chemical Mechanism version 3.1 (MCM v3.1) as a reference benchmark
-\section{Overall objective and specific aims}
+\section{Overall objective and specific aims}\label{sec:objams}
 As the reader might anticipateit is mentioned in section \ref{sec:lit_review}
@@
 This kind of study shows the potential of using mechanisms with a high level of chemical detail to study the formation of pollutants in the atmosphere (Madronich, 2006).]
-\subsection{Overall objective}
+\subsection{Overall objective}\label{sec:objams_obj}
 The overall purpose of this research is to investigate the impact of alkyl nitrate chemistry on ozone production efficiency.
-\subsection{Specific aims}
+\subsubsection*{Specific aims}\label{sec:objams_ams}
 Specific aims are to investigate:
@@
 \section{Methodology} \label{sec:method}
-\subsection{Model development and description}
+\subsection{Model development and description}\label{sec:method_develop}
 A zero-dimensional (box) model is developed to investigate the impact of alkyl nitrate chemistry on ozone production efficiency.
@@
 
-\subsection{Chemical mechanism}\label{sec:chem_mech}
+\subsection{Chemical mechanism}\label{sec:method_chem}
 The chemical mechanism of the model is based on recommendations from \citep{Atkinson2004} and information from the Master Chemical Mechanism, MCM v3.3 \citep{Jenkin1997,Saunders2003}, retrieved via website: http://mcm.leeds.ac.uk/MCM.
@@
 Despite that halogens (chlorine ($Cl$), bromine ($Br$) and iodine ($I$)) are very reactive chemical compounds mostly known by their role in anthropogenic stratospheric ozone-depletion chemistry, and that they are directly (for example, reaction of the $Cl$ radical with hydrocarbons (e.g., $CH_4$)) and indirectly (coupling with the cycles of sulfur and nitrogen that eventually influence $O_3$, $OH$ and $NO_x$ concentrations) involved in the tropospheric chemistry \citep{VonGlasow2014,Platt2003}, they are excluded from the chemical mechanism to keep it more transparent.
-\subsection{Photolysis}
+\subsection{Photolysis}\label{sec:method_photolysis}
 To estimate photolysis rates for a given geographical location, one must take into account the latitude and season, as well as the time of the day.
@@
 
-\subsection{Design of experiments}
+\subsection{Design of experiments}\label{sec:method_expdesign}
 To investigate the influence of alkyl nitrate chemistry on (potential changes in) tropospheric $O_3$ production, box model runs were performed at a series of different $NO_x$ and $VOCs$ initial concentrations, so that an $O_3$ isopleth plot could be constructed, similar to those found in \citep{Dodge1977} and \citep{Sillman1999}.
@@
 \end{table}
-
-!! For long-lived species, averaging may be reasonable.
 
@@
 \section{Results} \label{sec:res}
-\subsection{Sensitivity of $O_3$ production to alkyl nitrate chemistry}
-
-To evaluate the sensitivity of $O_3$ production (efficiency) to alkyl nitrate ($RONO_2$) chemistry under different $NO_x$ and $VOC$ conditions a number of simulations were carried out.
+\subsection{Sensitivity of $O_3$ production to alkyl nitrate chemistry}\label{sec:res_O3ANsensetivity}
+
+To evaluate the sensitivity of $O_3$ production to alkyl nitrate chemistry under different $NO_x$ and $VOC$ conditions a number of simulations were carried out.
 In these simulations the chemical mechanism of the model was adjusted to either include or exclude the formation and loss of alkyl nitrates, which was accomplished by switching on and off all reactions that involve $RONO_2$ while keeping the rest of the mechanism unchanged.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 Figure \ref{fig:netO3rate_noAN_withAN_diff} shows isopleths of net $O_3$ production rate as a function of $NO_x$ and sum of $VOCs$ initial concentrations from two series of model runs (with and without alkyl chemistry) and the difference between them.
@@
 At the range of $NO_x$ levels from 1 ppb to 10 ppb the $O_3$ production rate is slightly more sensitive to $VOCs$ levels, while it is almost insensitive to $VOCs$ at $NO_x$ levels higher than 10 ppb and 2050 ppbC of $VOCs$.
-
-\begin{figure} % netO3mixrat noAN withAN diff
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{figure} % netO3mixrat noAN allAN diff
 \centering
@@
 \end{figure}
-
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 The change in net rate of $O_3$ production due to inclusion of alkyl nitrate chemistry into the chemical mechanism has a direct impact on absolute $O_3$ levels calculated by the model.
@@
 To see what caused these changes it is useful to look at the net production of the sum of the alkyl nitrates themselves as a function of $NO_x$ and $VOCs$ levels (Figure \ref{fig:endtotalANs}).
-Since alkyl nitrates are the only sink for $NO_x$ in our model the high production of these species in total was expected.
-The range is ppb.
+Since alkyl nitrates are the only sink for $NO_x$ in our model high production of these species was expected.
+The range is from 0.2 ppb to 1.4 ppb for the range of $NO_x$ levels inspected from 250 ppt to 100 ppb.
 However, real concentrations reported from observational studies are of different order, tens of ppt.
-\citep{Reeves2007} 10 ppt (ITOP) \citep{Roberts1998} 14 ppt
-
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \begin{figure} % endtotalANs
@@
 \end{figure}
-
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 !
@@
 
-\subsection{$O_3$ production and carbon bonds}
-
+\subsection{$O_3$ production and carbon bonds}\label{sec:res_O3ANCB}
+
+As it has been noted in section \ref{sec:intro_ANs}, there is a way to estimate the $O_3$ production per carbon bond emitted into the system and relate it to the alkyl nitrate chemistry.
+Such calculations have been first conducted by \citep{Evans2014} for the global 3D chemistry-climate model GEOS-Chem\citep{Bey2001}.
+They estimated that $\epsilon$ value for the GEOS-Chem model in simulation with full chemistry is equal to 0.2 meaning that to produce 1 molecule of $O_3$ 0.2 carbon bonds should have been emitted.
+They also estimated that in simulation with $CH_4$ only $\epsilon$ is equal to 0.3, meaning that more carbon bonds should have been emitted to produce the same amount of $O_3$.
+
+We tried to calculate the same value for our model, however since it does not include any emissions, the $\epsilon$'s denominator has been treated a little bit differently.
+To be exact the term $E_{C_{bons}}$ is calculated as a rate of change of the difference between final and initial sum of carbon bonds ($\Sigma Cbonds$) in the system at constant sum of carbon bonds contained in all of the alkanes and $CO$ ($\Sigma Cbonds_{alkanes, CO}$):
+\begin{equation}\label{eq:ECbonds}
+E_{C_{bons}} = {\dfrac{\Sigma Cbonds|_{final} - \Sigma Cbonds|_{initial}}{24 hours}}\bigg|_{\Sigma Cbonds_{alkanes, CO} =const}
+\end{equation}
+In other words, we calculate an increment in the amount of carbon bonds contained in all organic species except for alkanes and $CO$ (e.g., in $ROOH$, $RONO_2$) in the system by the end of simulation and divide it by simulation time (24 hours).
+Despite that we somewhat excluded the impact of alkanes and $CO$ on $O_3$ production efficiency, we have managed to save an opportunity to take into account the impact of alkyl nitrates, because the carbon bonds contained in them are still in this formula.
+
+Figure \ref{fig:epsilon250ppt100ppb} shows $\epsilon$ values calculated from our model as a function of $NO_x$ (250 ppt - 100 ppb) and $VOCs$ mixing ratios.
+It appears that in both cases with and without alkyl nitrates $\epsilon$ is insensitive to $NO_x$ in the range of 1-100 ppb of $NO_x$, but quickly drops within this $NO_x$ range with increasing $VOCs$.
+At the rest of $\epsilon$ isopleth field it is equally sensitive to $NO_x$ and $VOCs$.
+The difference in $\epsilon$ between two series of model runs is small.
+The average $\epsilon$ value in model runs without alkyl nitrates and $NO_x$ of 250 ppt - 100 ppb is equal to 0.44, in case with them it is equal to 0.41, which means that 2.3 carbon bonds should have been emitted to produce 1 molecule of $O_3$ in case without alkyl nitrates, while in case with alkyl nitrates 2.4 carbon bonds should have been emitted to produce the same amount of $O_3$.
+
+If we construct the same kind of plot for all $NO_x$ conditions considered in the model, from 5 ppt to 100 ppb, we find that $\epsilon$ could be negative (Figure \ref{fig:epsilon5ppt100ppb}), which could be explained by a net $O_3$ removal at $NO_x$ levels lower 250 ppt in our model.
+The lower $VOCs$ level is more negative $epsilon$ becomes.
+The average $\epsilon$ value in all model runs without alkyl nitrates is equal to 0.25, in case with them it is equal to 0.22, meaning that 4 carbon bonds should have been emitted to produce 1 molecule of $O_3$ in case without alkyl nitrates, while in case with alkyl nitrates 4.5 carbon bonds should have been emitted to produce the same amount of $O_3$.
+The difference between $\epsilon$ values between two series of model runs is small.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{figure} % epsilon noANs allANs diff 250 ppt 100 ppb
+\centering
+\begin{minipage}{.3\textwidth}
+  \centering
+  \includegraphics[width=.99\linewidth]{D:/FACSIMILE/ANsCBmodel/matlab/ANsCB_pics/ANsNOxVOC/noAN/chem_noAN_epsilon_noINORG_NOx_250ppt_100ppb.png}
+\end{minipage}
+\begin{minipage}{.3\textwidth}
+  \centering
+  \includegraphics[width=.99\linewidth]{D:/FACSIMILE/ANsCBmodel/matlab/ANsCB_pics/ANsNOxVOC/allAN/chem_allAN_epsilon_noINORG_NOx_250ppt_100ppb.png}
+\end{minipage}
+\begin{minipage}{.3\textwidth}
+  \centering
+  \includegraphics[width=.99\linewidth]{D:/FACSIMILE/ANsCBmodel/matlab/ANsCB_pics/ANsNOxVOC/allAN/chem_allAN_epsilondiff_noINORG_NOx_250ppt_100ppb.png}
+\end{minipage}
+\caption{Isopleths giving net $O_3$ production efficiency (or $\epsilon$, molecules of $O_3$ per carbon bond, color) as a function of $VOC$ (ppbC) and $NO_x$ (250ppt - 100 ppb) with alkyl nitrate chemistry switched off (left) and switched on (middle).
+Difference between them is shown on the right.}\label{fig:epsilon250ppt100ppb}
+\end{figure}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{figure} % epsilon noANs allANs diff 5 ppt 100 ppb
+\centering
+\begin{minipage}{.3\textwidth}
+  \centering
+  \includegraphics[width=.99\linewidth]{D:/FACSIMILE/ANsCBmodel/matlab/ANsCB_pics/ANsNOxVOC/noAN/chem_noAN_epsilon_noINORG_NOx_5ppt_100ppb.png}
+\end{minipage}
+\begin{minipage}{.3\textwidth}
+  \centering
+  \includegraphics[width=.99\linewidth]{D:/FACSIMILE/ANsCBmodel/matlab/ANsCB_pics/ANsNOxVOC/allAN/chem_allAN_epsilon_noINORG_NOx_5ppt_100ppb.png}
+\end{minipage}
+\begin{minipage}{.3\textwidth}
+  \centering
+  \includegraphics[width=.99\linewidth]{D:/FACSIMILE/ANsCBmodel/matlab/ANsCB_pics/ANsNOxVOC/allAN/chem_allAN_epsilondiff_noINORG_NOx_5ppt_100ppb.png}
+\end{minipage}
+\caption{Isopleths giving net $O_3$ production efficiency (or $\epsilon$, molecules of $O_3$ per carbon bond, color) as a function of $VOC$ (ppbC) and $NO_x$ (5 ppt - 100 ppb) with alkyl nitrate chemistry switched off (left) and switched on (middle).
+Difference between them is shown on the right.}\label{fig:epsilon5ppt100ppb}
+\end{figure}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Discussion} \label{sec:discuss}
+Use CRI.
 Future plans and possible improvements in the model.

### 2015-08-05 MSc_diss.tex
Almost finished section about O3 sensitivity to total ANs

@@
 Being transported from polluted areas to remote background regions and eventually destroyed alkyl nitrates can 'release' $NO_2$ back to the atmosphere, thereby contributing to tropospheric ozone formation.
-\subsection{Ozone, alkyl nitrates and carbon bonds}
+\subsection{Ozone, alkyl nitrates and carbon bonds}\label{sec:intro_CB}
 In the previous sections we covered the chemistry that controls the production of $O_3$ in the troposphere. A traditional way to assess the efficiency of this production is to calculate the ratio:
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 Rewrite!
@@
 Since the actual formation of $O_3$ in our model starts at 250 ppt of $NO_x$, Figure \ref{fig:netO3rate_noAN_withAN_diff} summarises modelling data from experiments with $NO_x$ initial mixing ratios from 250 ppt till 100 ppb.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 As expected, the inclusion of alkyl nitrate chemistry slows down the net $O_3$ production rate.
@@
 As $NO_x$ and $VOCs$ levels increase the $O_3$ production rate decreases, reaching the maximum decrease of -1.94 ppb of $O_3$ per hour at 100 ppb of $NO_x$ and $VOCs$ level from case I (not the highest $VOCs$ level).
-At the range of $NO_x$ levels from 1 ppb to 10 ppb the $O_3$ production rate is slightly more sensitive to $VOCs$ levels, while it is almost insensitive to $VOCs$ at $NO_x$ levels higher than 10 ppb and 2050 ppbC of $VOCs$.
+At the range of $NO_x$ levels from 1 ppb to 10 ppb the difference in $O_3$ production rate is slightly more sensitive to $VOCs$ levels, while it is almost insensitive to $VOCs$ at $NO_x$ levels higher than 10 ppb and 2050 ppbC of $VOCs$.
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{figure} % netO3mixrat noAN allAN diff
-\centering
-\begin{minipage}{.3\textwidth}
+\begin{figure} % netO3mixrat noAN allAN diff diffpercent
+\centering
+\begin{minipage}{.4\textwidth} % noAN
   \centering
@@
 \end{minipage}
-\begin{minipage}{.3\textwidth}
+\begin{minipage}{.4\textwidth} % diff
+  \centering
+  \includegraphics[width=.99\linewidth]{D:/FACSIMILE/ANsCBmodel/matlab/ANsCB_pics/ANsNOxVOC/allAN/chem_allAN_netO3mixratdiff_noINORG_NOx_250ppt_100ppb.png}
+\end{minipage}
+\begin{minipage}{.4\textwidth} % allAN
   \centering
@@
 \end{minipage}
-\begin{minipage}{.3\textwidth}
-  \centering
-  \includegraphics[width=.99\linewidth]{D:/FACSIMILE/ANsCBmodel/matlab/ANsCB_pics/ANsNOxVOC/allAN/chem_allAN_netO3mixratdiff_noINORG_NOx_250ppt_100ppb.png}
-\end{minipage}
-\caption{Isopleths giving net $O_3$ production (ppb, color) as a function of $VOC$ (ppbC) and $NO_x$ (ppb) with alkyl nitrate chemistry switched off (left) and switched on (middle).
-Difference between them is shown on the right.
+\begin{minipage}{.4\textwidth} % diffpercent
+  \centering
+  \includegraphics[width=.99\linewidth]{D:/FACSIMILE/ANsCBmodel/matlab/ANsCB_pics/ANsNOxVOC/allAN/chem_allAN_netO3mixratdiffpercent_noINORG_NOx_250ppt_100ppb.png}
+\end{minipage}
+\caption{Isopleths giving net $O_3$ production (ppb, color) as a function of $VOC$ (ppbC) and $NO_x$ (ppb, range 250 ppt - 100 ppb) with alkyl nitrate chemistry switched off (a) and switched on (b).
+Absolute difference (with alkyl nitrates minus without them) between two series of model runs in ppb (c), same in \% (d).
 Net production = final - initial value of $O_3$ mixing ratio.}\label{fig:netO3mixrat_noAN_withAN_diff}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 The change in net rate of $O_3$ production due to inclusion of alkyl nitrate chemistry into the chemical mechanism has a direct impact on absolute $O_3$ levels calculated by the model.
 Figure \ref{fig:netO3mixrat_noAN_withAN_diff} shows that the net $O_3$ production (difference between final and initial $O_3$ mixing ratio) is smaller when alkyl nitrate chemistry is switched on.
-At low $NO_x$ and $VOCs$ levels the (negative) change in the net $O_3$ production is 5-10 ppb, at high $NO_x$ and $VOCs$ levels it is 35-45 ppb (per 24 hours) with maximum value of 46.5 ppb.
-However, these changes in the net $O_3$ production should be treated with care, because the model due to reasons described in Section \ref{sec:method} gives unrealistically high $O_3$ mixing ratios (maximum 650 ppb while typical maximum urban $O_3$ mixing ratio (Los Angeles, Mexico City) is 490 ppb (http://www-personal.umich.edu/~sillman/ozone.htm).
-Therefore, in this case a more useful estimate of effect of these changes is to express them in percentage form (i.e., compare like with like), and it appears that these changes are very small: ???
-respectively.
-
-To see what caused these changes it is useful to look at the net production of the sum of the alkyl nitrates themselves as a function of $NO_x$ and $VOCs$ levels (Figure \ref{fig:endtotalANs}).
+At low $NO_x$ and $VOCs$ levels the negative (!) change in the net $O_3$ production is 5-10 ppb, at high $NO_x$ and $VOCs$ levels it is 35-45 ppb (per 24 hours) with maximum value of 46.5 ppb.
+
+These changes in the net $O_3$ production should be treated with care, because the model due to reasons described in Section \ref{sec:method} gives unrealistically high $O_3$ mixing ratios.
+The maximum $O_3$ mixing ratio in the model is equal to 650 ppb (at the highest $NO_x$ level (100 ppb) and the lowest $VOCs$ level (case A)), while maximum typical urban (Los Angeles, Mexico City) $O_3$ mixing ratio is 490 ppb (http://www-personal.umich.edu/~sillman/ozone.htm).
+Therefore, a more useful estimate is a percentage difference in $O_3$ mixing ratios between two series of model runs shown in Figure \ref{fig:netO3rate_noAN_withAN_diff} DDDDD!.
+It appears that the largest absolute changes in $O_3$ mixing ratios (40-45 ppb) are changes by only 20-30\%, while changes up to 90\% take place the model is in low $NO_x$ and high $VOCs$ conditions.
+
+To see what caused these changes it is useful to look at the net production of the sum of the alkyl nitrates themselves as a function of $NO_x$ (250 ppt - 100 ppb) and $VOCs$ (Figure \ref{fig:endtotalANs}).
 Since alkyl nitrates are the only sink for $NO_x$ in our model high production of these species was expected.
-The range is from 0.2 ppb to 1.4 ppb for the range of $NO_x$ levels inspected from 250 ppt to 100 ppb.
-However, real concentrations reported from observational studies are of different order, tens of ppt.
+The $NO_x$-$VOC$ field of total $RONO_2$ isopleths is mostly characterized by mixing ratios of ppb level (0.2-1.4 ppb), while the observations show that their mean mixing ratios lay at a level of tens-thousands of ppt (Table \ref{tab:ANmean}).
+Nevertheless, Figure \ref{fig:endtotalANs} will be of interest to other researches, because it is probably the first ever attempt to derive the distribution of $\Sigma RONO_2$ mixing ratios in $NO_x$-$VOC$ field.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-As it has been noted in section \ref{sec:intro_ANs}, there is a way to estimate the $O_3$ production per carbon bond emitted into the system and relate it to the alkyl nitrate chemistry.
-Such calculations have been first conducted by \citep{Evans2014} for the global 3D chemistry-climate model GEOS-Chem\citep{Bey2001}.
+As it has been noted in section \ref{sec:intro_CB}, there is a way to estimate the $O_3$ production per carbon bond emitted into the system and relate it to the alkyl nitrate chemistry.
+Such calculations have been first conducted by \citep{Evans2014} for the global 3D chemistry-climate model GEOS-Chem \citep{Bey2001}.
 They estimated that $\epsilon$ value for the GEOS-Chem model in simulation with full chemistry is equal to 0.2 meaning that to produce 1 molecule of $O_3$ 0.2 carbon bonds should have been emitted.

### 2015-08-05 MSc_diss.tex
Updates in discussion and conclusion

@@
 \section{Discussion} \label{sec:discuss}
-Use CRI.
+The experimental results presented in the previous section provide an opportunity to speculate about the behaviour of our simple chemistry box conceptual model and its sensitivity to alkyl nitrate production and loss.
+
+This research has limitations link with the disadvantages of the model, not enough experiments.
+
+Due to obvious limitations
+
+Future plans include illumination of all weaknesses from our model, particularly adding other sinks for $NO_x$ into the chemical mechanism.
+
+Probably it is reasonable to conduct a similar kind of research using the Common Representative Mechanism (CRI) which includes representation of alkyl nitrates and is traceable to the MCM.
+
 Future plans and possible improvements in the model.
 
-The experimental results presented in the previous section provide an opportunity to speculate about the behaviour of our simple model of alkyl nitrate production and loss with time in comparison with the observational data.
+\subsubsection*{Future research}
+A greater depth of information may have been obtained by conducting more experiments.
+Another possible improvement to the study could have been
 
 \section{Conclusion} \label{sec:conclusion}
+The present study provides an assessment of of
+
+The present study is purely theoretical, investigates the impact of alkyl nitrates on tropospheric ozone production.
+We examine the sensitivity of the modelled ozone production to the alkyl nitrate chemistry.
+
+The research hopes to contribute to the atmospheric chemistry research b
+
+We hope that our results concerning the sensitivity of ozone calculations to alkyl nitrate chemistry will be interesting to other researches, as well as the carbon bond efficiency calculations.
+
 \bibliography{diss_refs}

### 2015-08-05 MSc_diss.tex
Updates in section about O3 and groups of ANs

@@
 \subsection{Sensitivity of $O_3$ production to alkyl nitrate chemistry}\label{sec:res_O3ANsensetivity}
+\subsubsection{$O_3$ and total alkyl nitrates}
 
@@
 At the range of $NO_x$ levels from 1 ppb to 10 ppb the difference in $O_3$ production rate is slightly more sensitive to $VOCs$ levels, while it is almost insensitive to $VOCs$ at $NO_x$ levels higher than 10 ppb and 2050 ppbC of $VOCs$.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
+\subsubsection{$O_3$ and individual alkyl nitrates}
+
+To estimate the impact of individual alkyl nitrates or their groups on $O_3$ production an additional series of experiments have been conducted (Figure \ref{fig:timeO3diff10ppbNOx6VOC}).
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{figure}
+\centering
+\includegraphics[width=0.7\linewidth]{D:/FACSIMILE/ANsCBmodel/matlab/ANsCB_pics/ANsNOxVOC/allAN/chem_allAN_timeO3diff_10ppbNOx6VOC.eps}
+\caption{Time evolution of the difference in $O_3$ mixing ratio between base experiment (with all alkyl nitrate chemistry switched on) and experiments where chemistry of individual alkyl nitrates or their groups is switch on.
+Black line shows the difference between base experiment and an experiment with all alkyl nitrate chemistry switched off.
+$NO_x$ = 10 ppb, $VOCs$ = case F.}\label{fig:timeO3diff10ppbNOx6VOC}
+\end{figure}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{table}
+\caption{Difference in $O_3$ mixing ratio by the end of simulation depending alkyl nitrate chemistry.}\label{tab:timeO3diff10ppbNOx6VOC}
+\begin{tabular}{cc}
+\hline
+Alkyl nitrate            & $O_3$ difference (ppb) \\
+\hline
+$CH_3ONO_2$              & -11.5 \\
+$C_2H_5ONO_2$            & -11.6 \\
+$\Sigma C_3H7ONO_2$      & -10.5 \\
+$\Sigma nC_4H_9ONO_2$    & -8.7  \\
+$\Sigma iC_4H_9ONO_2$    & -11.0 \\
+$\Sigma nC_5H_{11}ONO_2$ & -10.1 \\
+$\Sigma iC_5H_{11}ONO_2$ & -8.0  \\
+$\Sigma RONO_2$          & -11.9 \\
+\hline
+\end{tabular}
+\end{table}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \subsection{$O_3$ production and carbon bonds}\label{sec:res_O3ANCB}

### 2015-08-05 MSc_diss.tex
Updates in section about O3 and individual ANs

@@
 \begin{table} % RH and ANs
-\caption{Alkyl nitrates and their parent alkanes}\label{tab:RHandANS}
+\caption{Alkyl nitrates and their parent alkanes}\label{tab:RHandANs}
 \centering
@@
 
-To estimate the impact of individual alkyl nitrates or their groups on $O_3$ production an additional series of experiments have been conducted (Figure \ref{fig:timeO3diff10ppbNOx6VOC}).
+To estimate the impact of individual alkyl nitrates on $O_3$ production an additional series of experiments have been conducted.
+Prior to calculations some of the alkyl nitrates have been gathered into groups according to their parent alkanes as it shown in Table \ref{tab:RHandANs}.
+Setting initial $NO_x$ mixing ratio to 10 ppb and taking the middlemost $VOCs$ initial values (i.e., case F), the model has been run with certain alkyl nitrate chemistry switched on.
+For example, to estimate the impact of $CH_3ONO_2$ the reactions involving this nitrate were switched on, while the other reactions with nitrates were switched off.
+Experiment with all alkyl nitrate chemistry is considered as a base run.
+
+To estimate the impact of certain alkyl nitrates on $O_3$ the difference between base run and other runs including one with all alkyl nitrate chemistry switched off has been calculated.
+Figure \ref{fig:timeO3diff10ppbNOx6VOC} shows the time evolution of this difference.
+This difference becomes more and more negative with time as the mixing ratio of alkyl nitrates themselves rises (Figure \ref{fig:timeANs10ppbNOx6VOC}).
+By the end of simulation the drop in $O_3$ mixing ratio due to the rise in mixing ratio of total alkyl nitrates is equal to -11.9 ppb (Table \ref{tab:timeO3diffANs10ppbNOx6VOC}.
+
+The contribution of different alkyl nitrates to drop in $O_3$ is not the same.
+By the end of simulation, the drop in $O_3$ mixing ratio due to ethyl nitrate only ($C_2H_5ONO_2$) is 11.6 ppb, and is almost the same drop is caused by methyl nitrate.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{figure}
+\centering
+\includegraphics[width=0.7\linewidth]{D:/FACSIMILE/ANsCBmodel/matlab/ANsCB_pics/ANsNOxVOC/allAN/chem_allAN_timeANs_10ppbNOx6VOC.eps}
+\caption{Time evolution of alkyl nitrates mixing ratios. $NO_x$ = 10 ppb, $VOCs$ = case F.}\label{fig:timeANs10ppbNOx6VOC}
+\end{figure}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \begin{table}
-\caption{Difference in $O_3$ mixing ratio by the end of simulation depending alkyl nitrate chemistry.}\label{tab:timeO3diff10ppbNOx6VOC}
-\begin{tabular}{cc}
-\hline
-Alkyl nitrate            & $O_3$ difference (ppb) \\
-\hline
-$CH_3ONO_2$              & -11.5 \\
-$C_2H_5ONO_2$            & -11.6 \\
-$\Sigma C_3H7ONO_2$      & -10.5 \\
-$\Sigma nC_4H_9ONO_2$    & -8.7  \\
-$\Sigma iC_4H_9ONO_2$    & -11.0 \\
-$\Sigma nC_5H_{11}ONO_2$ & -10.1 \\
-$\Sigma iC_5H_{11}ONO_2$ & -8.0  \\
-$\Sigma RONO_2$          & -11.9 \\
+\caption{Difference in $O_3$ mixing ratio by the end of simulation depending alkyl nitrate chemistry.}\label{tab:timeO3diffANs10ppbNOx6VOC}
+\begin{tabular}{ccc}
+\hline
+Alkyl nitrate            & $RONO_2$ mixing ratio (ppt) & $O_3$ difference (ppb) \\
+\hline
+$CH_3ONO_2$              & 43  & -11.5 \\
+$C_2H_5ONO_2$            & 38  & -11.6 \\
+$\Sigma C_3H7ONO_2$      & 115 & -10.5 \\
+$\Sigma nC_4H_9ONO_2$    & 192 & -8.7  \\
+$\Sigma iC_4H_9ONO_2$    & 81  & -11.0 \\
+$\Sigma nC_5H_{11}ONO_2$ & 108 & -10.1 \\
+$\Sigma iC_5H_{11}ONO_2$ & 290 & -8.0  \\
+$\Sigma RONO_2$          & 867 & -11.9 \\
 \hline

### 2015-08-05 MSc_diss.tex
New layout

@@
 \documentclass[11pt,a4paper]{article}
+% Font
+\usepackage{helvet}
+\renewcommand{\familydefault}{\sfdefault}
 % Encoding and language
@@
 \usepackage{afterpage}
-% Indenting
+% Indenting and line spacing
+\linespread{1.5}
+%\usepackage{setspace}
 \usepackage{indentfirst}
@@
 \usepackage{mathtools} % for arrows
+
+\newcounter{matriz}
+\newenvironment{matriz}{\refstepcounter{matriz}\equation}{\tag{\thematriz}\endequation}
 \renewcommand{\theequation}{R\arabic{equation}}
+
 % References
@@
 \usepackage{epstopdf}
+\usepackage{wrapfig}
 % Tables
@@
 \input{diss_titlepage}
-
-\tableofcontents
-\newpage
+%\onehalfspacing
 
@@
 \end{abstract}
+
+\tableofcontents
+\newpage
 
@@
 \begin{figure}[h]
+\centering
 \includegraphics[width=\linewidth]{{./pics/Isaksen2009_O3}.png}
@@
 However, despite general agreement on how the drivers impact ozone concentrations at global and regional scales, its budget varies considerably between different modelling and observational studies \citep{Stevenson2006,Wild2007,Young2012}.
-Table \ref{tab:O3budget} demonstrates that the range of the global tropospheric ozone budget estimates is approximately 60 Tg which is 18\% of the mean.
+Figure \ref{fig:O3budget} demonstrates that the range of the global tropospheric ozone budget estimates is approximately 60 Tg which is 18\% of the mean.
 In terms of ozone production the variation is even higher and accounts for 28\% of the corresponding value.
@@
 
-\begin{table}[h] % O3 budget
-\centering
+\begin{figure}[t] % O3 budget
 \caption{Summary of tropospheric ozone global budget model and observation estimates for present (about 2000) conditions.
@@
 Adapted from \citep{Myhre2013}}
+\centering
 \label{tab:O3budget}
-\begin{tabular}{ccl}
-\hline
-Burden, Tg & Production, Tg yr–1 & Reference \\
-\hline
-\multicolumn{3}{c}{Modelling studies} \\
-$337\pm23$  & $4877\pm853$ & Young et al. (2013); ACCMIP \\
-$323$	    & -	           & Archibald et al. (2011) \\
-$330$	    & $4876$       & Kawase et al. (2011) \\
-$312$       & $4289$       & Huijnen et al. (2010) \\
-$334$       & $3826$       & Zeng et al. (2010) \\
-$324$       & $4870$       & Wild and Palmer (2008) \\
-$314$       & -	           & Zeng et al. (2008) \\
-$319$	    & $4487$	   & Wu et al. (2007) \\
-$372$	    & $5042$       & Horowitz (2006) \\
-$349$	    & $4384$	   & Liao et al. (2006) \\
-$344\pm39$	& $5110\pm606$ & Stevenson et al. (2006); ACCENT \\
-$314\pm33$	& $4465\pm514$ & Wild (2007) (post-2000 studies) \\
-\hline
-\multicolumn{3}{c}{Observational studies} \\
-$333$   	& -	           & Fortuin and Kelder (1998) \\
-$327$	    & -	           & Logan (1999) \\
-$325$	    & -	           & Ziemke et al. (2011); 60S–60N \\
-$319–351$	& -	           & Osterman et al. (2008); 60S–60N \\
-\hline
-\end{tabular}
-\end{table}
+\includegraphics[width=0.7\linewidth]{o3_studies.eps}
+\end{figure}
 
@@
 \end{equation}
-\begin{equation}\label{reac:O3P+O2=O3}
-!!!(R4)sameO(^3P) + O_2 \xrightarrow{M} O_3
-\end{equation}
-As shown in Figure ???
-reactions (\ref{reac:HO2+NO=NO2+OH}) and (\ref{reac:CH3O2+NO=NO2+CH3O}) form parts of the free-radical propagated $O_3$ forming cycles, which may occur a number of times before being halted by radical termination reactions (\ref{reac:HO2+HO2=H2O2+O2} and \ref{reac:HO2+CH3O2=CH3OOH+O2}), and the higher $NO_x$ level is the greater number of these cycles can occur.
+\begin{equation}
+O(^3P) + O_2 \xrightarrow{M} O_3\tag{\ref{reac:O3P+O2=O3}}
+\end{equation}
+As shown in Figure \ref{fig:O3scheme} reactions (\ref{reac:HO2+NO=NO2+OH}) and (\ref{reac:CH3O2+NO=NO2+CH3O}) form parts of the free-radical propagated $O_3$ forming cycles, which may occur a number of times before being halted by radical termination reactions (\ref{reac:HO2+HO2=H2O2+O2} and \ref{reac:HO2+CH3O2=CH3OOH+O2}), and the higher $NO_x$ level is the greater number of these cycles can occur.
 Therefore, in intermediate $NO_x$ conditions a net $O_3$ formation takes place and it is sensitive to $NO_x$.
+
+\begin{figure}[h!]
+\centering
+\input{diss_NewO3scheme}
+\caption{$O_3$ chemistry scheme}\label{fig:O3scheme}
+\end{figure}
 
@@
 \end{equation}
-\begin{equation}\label{reac:H+O2=HO2}
-!!!sameH + O_2 \xrightarrow{M} HO_2
-\end{equation}
+\begin{equation}
+H + O_2 \xrightarrow{M} HO_2\tag{\ref{reac:H+O2=HO2}}
+\end{equation}
+
 therefore also having an impact on the rates of $O_3$ production \citep{Fowler2008} as do some other species, which are not mentioned in this section.
@@
 In the previous sections we covered the chemistry that controls the production of $O_3$ in the troposphere. A traditional way to assess the efficiency of this production is to calculate the ratio:
-\begin{equation}\label{eq:OPE}
+\begin{matriz}\label{eq:OPE}
 OPE = \dfrac{k_{HO_2+NO}[HO_2][NO]}{k_{OH+NO_2}[OH][NO_2]} = \dfrac{P_{O_3}}{L_{NO_x}}
-\end{equation}
+\end{matriz}
 where $OPE$ is the ozone production efficiency, $k_{HO_2+NO}$ - the rate coefficient of reaction (\ref{reac:HO2+NO=NO2+OH}), $k_{OH+NO_2}$ - the rate coefficient of reaction (\ref{reac:OH+NO2=HNO3}), $[X]$ - mixing ratios of corresponding species, $P_{O_3}$ – the $O_3$ production rate, $L_{NO_x}$ – the $NO_x$ loss rate.
@@
 Assuming that every carbon bond is broken chemically and every organic peroxy radical ($RO_2$) reacts with $NO$, the overall efficiency of carbon bonds conversion to $O_3$, $\epsilon$, can be calculated as:
-\begin{equation}\label{eq:epsilon}
+\begin{matriz}\label{eq:epsilon}
 \epsilon = \dfrac{P_{O_3}}{E_{C_{bons}}} = \chi\cdot\upsilon
-\end{equation}
+\end{matriz}
 where $P_{O_3}$ – the rate of $O_3$ production (mol yr-1), $E_{C_{bons}}$  – the rate of carbon bonds emission (mol yr-1), $\chi$ – ratio between the rate of carbon bonds loss to form $RO_2$ and the loss rate of all carbon bonds, $\upsilon$ – ratio between the rate of $RO_2$ loss in reaction with $NO$ resulting in $NO$-to-$NO_2$ conversion and the loss rate of all $RO_2$.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{table}\label{tab:ANmean} % mean AN mixing ratios
+\renewcommand{\arraystretch}{1}
+\begin{table}[h!]
+\caption{!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!}\label{tab:ANmean} % mean AN mixing ratios
 \begin{tabular}{cccc}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{figure}[h!]
+\centering
+\includegraphics[width=0.5\linewidth]{{\picspath Neuman2012_Fig.4}.png}
+\caption{Time series of reactive nitrogen species measured on 10 June 2010 in the MBL in plumes (a) 10 km and (b) 50 km downwind from DWH.
+$NO_x$, $NO_y$, and $HNO_3$ were measured once per s and $PAN$ was measured once per 2 s.
+$C_1$–$C_5$ alkyl nitrates were measured in whole air samples at locations indicated by the circles.
+The sum of alkyl nitrates (black lines) is inferred from the difference between $NO_y$ and ($NO_x$+$HNO_3$+$PAN$) \citep{Neuman2012}.}\label{fig:Neuman2012}
+\end{figure}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%}
+
 $O_3$ production efficiencies in the DWH plume were 6 plus minus 0.5 ppbv $O_3$ per ppbv of $NO_x$ oxidized after 1 h of transport.
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{figure}[h!]\label{fig:Neuman2012}
-\includegraphics[width=.5\linewidth]{{D:/UEA/ENV-MB4X_Dissertation/ENV-MB4X_Dissertation_thesis/ENV-MB4X_Dissertation_thesis_latex/pics/Neuman2012_Fig.4}.png}
-\caption{Time series of reactive nitrogen species measured on 10 June 2010 in the MBL in plumes (a) 10 km and (b) 50 km downwind from DWH.
-$NO_x$, $NO_y$, and $HNO_3$ were measured once per s and $PAN$ was measured once per 2 s.
-$C_1$–$C_5$ alkyl nitrates were measured in whole air samples at locations indicated by the circles.
-The sum of alkyl nitrates (black lines) is inferred from the difference between $NO_y$ and ($NO_x$+$HNO_3$+$PAN$) \citep{Neuman2012}.}
-\end{figure}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 Probably the most detailed investigation of the formation processes and photochemical evolution of alkyl nitrates (but not their impact on $O_3$) has been done by \citep{Sommariva2008}.
@@
 \caption{Initial values of the some model parameters (in ppb) for two scenarios used in \citep{Sommariva2008}}
+\centering
 \begin{tabular}{ccc}
@@
 	\centering
-	\begin{subfigure}[t]{0.5\textwidth}
-        \centering
-        \includegraphics[width=\linewidth]{{\picspath Neu2008_Fig.2_MeONO2_OCT}.png}
+	\begin{subfigure}[t]{0.45\textwidth}
+        \centering
+        \includegraphics[width=\linewidth]{{\picspath Neu2008_Fig.2_MeONO2_OCT_merge}.png}
         \caption{}
@@
 \subsection{Chemical mechanism}\label{sec:method_chem}
-The chemical mechanism of the model is based on recommendations from \citep{Atkinson2004} and information from the Master Chemical Mechanism, MCM v3.3 \citep{Jenkin1997,Saunders2003}, retrieved via website: http://mcm.leeds.ac.uk/MCM.
+The chemical mechanism of the model is based on recommendations from \citep{Atkinson2004} and information from the Master Chemical Mechanism, MCM v3.3 \citep{Jenkin1997,Saunders2003}, retrieved via website: \href{http://mcm.leeds.ac.uk/MCM}{http://mcm.leeds.ac.uk/MCM}.
 On a whole it includes 90 species and 165 reactions.
@@
 Their calculations were performed for clear sky conditions at an altitude of 0.5 km for 1 July at a latitude of 45$^{\circ}$N, resulting in the following formula:
-\begin{equation} \label{eq:MCMphotolysis}
-\renewcommand{\theequation}{\arabic{equation}}
+\begin{matriz} \label{eq:MCMphotolysis}
 J = l(cos\chi)^mexp(-nsec\chi)
-\end{equation}
+\end{matriz}
 where $J$ - photolysis rate coefficient, $\chi$ - solar zenith angle, $l$, $m$ and $n$ - optimising parameters \citep{Jenkin1997,Saunders2003}.
 
-The values of solar zenith angle are calculated using NOAA Solar Calculator (http://www.esrl.noaa.gov/gmd/grad/solcalc/).
+The values of solar zenith angle are calculated using \href{http://www.esrl.noaa.gov/gmd/grad/solcalc/}{NOAA Solar Calculator}.
 
@@
 Figure \ref{fig:model_validation} shows an example of output from the model, namely the changes in mixing ratios of the main species within 24-hours of simulation time under contrasting $NO_x$ and $VOCs$ conditions.
-The values shown in Figure \ref{fig:model_validation} have been justified from the theoretical point of view and also compared with those derived from the Atmospheric Chemistry Learner's Box Model (ACLBM, http://guille.uea.ac.uk/ACLBM/).
+The values shown in Figure \ref{fig:model_validation} have been justified from the theoretical point of view and also compared with those derived from the Atmospheric Chemistry Learner's Box Model (ACLBM, \href{http://guille.uea.ac.uk/ACLBM/}{http://guille.uea.ac.uk/ACLBM/}).
 ACLBM is an on-line box model for atmospheric chemistry developed by Prof.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{figure} % Sillman1999withdotsline and netO3rate_withINORG_NOx_1ppb_100ppb
-\centering
-\begin{minipage}{.5\textwidth}
-  \centering
-  \includegraphics[width=.99\linewidth]{{./pics/Sillman1999_Fig1_adapted_withdotsrectangle}.png}
-  \captionof{figure}{Isopleths giving net rate of $O_3$ production (ppb/h, solid lines) as a function of $VOC$ (ppbC) and $NO_x$ (ppb) for mean summer daytime meteorology and clear skies.
+
+\begin{figure}[t!]
+	\centering
+    \begin{subfigure}[t]{0.45\linewidth}
+        \centering
+        \includegraphics[width=\linewidth]{{./pics/Sillman1999_Fig1_adapted_withdotsrectangle}.png}
+        \caption{(ppb/h, solid lines) as a function of $VOC$ (ppbC) and $NO_x$ (ppb) for mean summer daytime meteorology and clear skies.
 The solid lines represent production rates of 1, 2.5, 5, 10, 15, 20 and 30 ppb/h.
-Adapted from \citep{Sillman1999}.}\label{fig:Sillman1999}
-\end{minipage}%
-\begin{minipage}{.5\textwidth}
-  \centering
-  \includegraphics[width=.99\linewidth]{D:/FACSIMILE/ANsCBmodel/matlab/ANsCB_pics/ANsNOxVOC/noAN/chem_noAN_netO3rate_withINORG_NOx_1ppb_100ppb.png}
-  \captionof{figure}{Isopleths giving net rate of $O_3$ production (ppb/h, color) as a function of $VOC$ (ppbC) and $NO_x$ (ppb) in the first modelling experiment with alkyl nitrate chemistry switched off combined with an additional run with inorganic chemistry only (i.e.
-the only $VOC$ is $CO$).}\label{fig:noAN_netO3rate}
-\end{minipage}
+Adapted from \citep{Sillman1999}.} \label{fig:Sillman1999}
+    \end{subfigure}%
+    \hfill
+    \begin{subfigure}[t]{0.45\linewidth}
+        \centering
+        \includegraphics[width=\linewidth]{D:/FACSIMILE/ANsCBmodel/matlab/ANsCB_pics/ANsNOxVOC/noAN/chem_noAN_netO3rate_withINORG_NOx_1ppb_100ppb.png}
+        \caption{as a function of $VOC$ (ppbC) and $NO_x$ (ppb) in the first modelling experiment with alkyl nitrate chemistry switched off combined with an additional run with inorganic chemistry only (i.e.
+the only $VOC$ is $CO$).} \label{fig:noAN_netO3rate}
+    \end{subfigure}
+    \caption{Isopleths giving net rate of $O_3$ production (ppb/h, color)}\label{fig:netO3rate}
 \end{figure}
@@
 These changes in the net $O_3$ production should be treated with care, because the model due to reasons described in Section \ref{sec:method} gives unrealistically high $O_3$ mixing ratios.
-The maximum $O_3$ mixing ratio in the model is equal to 650 ppb (at the highest $NO_x$ level (100 ppb) and the lowest $VOCs$ level (case A)), while maximum typical urban (Los Angeles, Mexico City) $O_3$ mixing ratio is 490 ppb (http://www-personal.umich.edu/~sillman/ozone.htm).
+The maximum $O_3$ mixing ratio in the model is equal to 650 ppb (at the highest $NO_x$ level (100 ppb) and the lowest $VOCs$ level (case A)), while maximum typical urban (Los Angeles, Mexico City) $O_3$ mixing ratio is 490 ppb (\href{http://www-personal.umich.edu/~sillman/ozone.htm}{http://www-personal.umich.edu/~sillman/ozone.htm}).
 Therefore, a more useful estimate is a percentage difference in $O_3$ mixing ratios between two series of model runs shown in Figure \ref{fig:netO3rate_noAN_withAN_diff} DDDDD!.
@@
 \end{figure}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \begin{figure}
@@
 \end{figure}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \begin{table}
 \caption{Difference in $O_3$ mixing ratio by the end of simulation depending alkyl nitrate chemistry.}\label{tab:timeO3diffANs10ppbNOx6VOC}
+\centering
 \begin{tabular}{ccc}
@@
 To be exact the term $E_{C_{bons}}$ is calculated as a rate of change of the difference between final and initial sum of carbon bonds ($\Sigma Cbonds$) in the system at constant sum of carbon bonds contained in all of the alkanes and $CO$ ($\Sigma Cbonds_{alkanes, CO}$):
-\begin{equation}\label{eq:ECbonds}
+\begin{matriz}\label{eq:ECbonds}
 E_{C_{bons}} = {\dfrac{\Sigma Cbonds|_{final} - \Sigma Cbonds|_{initial}}{24 hours}}\bigg|_{\Sigma Cbonds_{alkanes, CO} =const}
-\end{equation}
+\end{matriz}
 In other words, we calculate an increment in the amount of carbon bonds contained in all organic species except for alkanes and $CO$ (e.g., in $ROOH$, $RONO_2$) in the system by the end of simulation and divide it by simulation time (24 hours).

### 2015-08-05 MSc_diss.tex
Checked numbers of reactions and figure for the first time

@@
 \begin{figure}[t] % O3 budget
+\centering
+\includegraphics[width=0.7\linewidth]{o3_studies.eps}
 \caption{Summary of tropospheric ozone global budget model and observation estimates for present (about 2000) conditions.
 All uncertainties quoted as 1 standard deviation (68\% confidence interval).
-Adapted from \citep{Myhre2013}}
-\centering
-\label{tab:O3budget}
-\includegraphics[width=0.7\linewidth]{o3_studies.eps}
+Adapted from \citep{Myhre2013}}\label{fig:O3budget}
 \end{figure}
@@
 \begin{align}
-RO_2 + NO &\xrightarrow{1-\alpha_3} RO + NO_2 \label{reac:RO+NO=aphaRO+NO2}\\
-RO_2 + NO &\xrightarrow{\alpha_3} RONO_2 \label{reac:RO+NO=aphaRONO2}
+RO_2 + NO &\xrightarrow{1-\alpha_3} RO + NO_2 \label{reac:RO2+NO=aphaRO+NO2}\\
+RO_2 + NO &\xrightarrow{\alpha_3} RONO_2 \label{reac:RO2+NO=aphaRONO2}
 \end{align}
 \end{subequations}
-The latter reaction can serve as a termination step of the propagated $O_3$ forming cycle (Figure ???), and as \citep{Roberts1998} estimated it is likely to begin to occur when the $NO$ mixing ratio increases up to 100-200 ppt.
+The latter reaction can serve as a termination step of the propagated $O_3$ forming cycle (Figure \ref{fig:O3scheme}), and as \citep{Roberts1998} estimated it is likely to begin to occur when the $NO$ mixing ratio increases up to 100-200 ppt.
 The yield of alkyl nitrate formation is controlled by the branching ratio ($\alpha_3$), which is temperature and pressure dependent and increases with the number of carbons in the alkyl peroxy radical \citep{Roberts1990}.
@@
 \end{equation}
-This pathway has been shown to become dominant over reaction (\ref{reac:RO+NO=aphaRONO2}) only for the production of lower (methyl ($C_1$), ethyl ($C_2$)) nitrates when $NO_x$ mixing ratio exceeds 35 ppb \citep{Archibald2007}, i.e.
+This pathway has been shown to become dominant over reaction (\ref{reac:RO2+NO=aphaRONO2}) only for the production of lower (methyl ($C_1$), ethyl ($C_2$)) nitrates when $NO_x$ mixing ratio exceeds 35 ppb \citep{Archibald2007}, i.e.
 in highly polluted environments.
@@
 In other words, $OPE$ shows the amount of ozone created ($P_{O_3}$) per molecule of $NO_x$ removed ($L_{NO_x}$) from the system \citep{Jacob1999}.
-Due to the reasons described in section \ref{sec:chem_mech} (comment at the bottom of the page!"No representation of the nitric acid ($HNO_3$) in the model"), we use another expression for calculations of the $O_3$ production efficiency.
+Due to the reasons described in section \ref{sec:method_chem} \footnote{No representation of the nitric acid ($HNO_3$) in the model.}, we use another expression for calculations of the $O_3$ production efficiency.
 This expression has been recently proposed by Professor Matthew Evans (University of York) amid the mounting concerns about the need of thorough intercomparison of chemical mechanisms of global atmospheric models (including those used in the IPCC reports), why it is more universal and applicable to any kind of model.
@@
 
-Temporal changes in reactive nitrogen partitioning in the DWH plume (Figure \ref{fig:Neuman2012}) nicely illustrate that in fresh part of the plume (Figure ???
-a) $NO_x$ accounted for over 80\% of plume $NO_y$, in aged plumes (Figure ???
-b), $NO_x$ oxidation was nearly complete, but NOx, PAN, HNO3, and the sum of eight $C_5$–$C_{10}$ $RONO_2$ accounted for only 22\% of plume $NO_y$.
-It means that $C_6$–$C_{10}$ $RONO_2$ account for 78\% of the aged plume $NO_y$\citep{Neuman2012}.
+Temporal changes in reactive nitrogen partitioning in the DWH plume (Figure \ref{fig:Neuman2012}) nicely illustrate that in fresh part of the plume (Figure \ref{fig:Neuman2012}a) $NO_x$ accounted for over 80\% of plume $NO_y$, in aged plumes (Figure \ref{fig:Neuman2012}b), $NO_x$ oxidation was nearly complete, but NOx, PAN, HNO3, and the sum of eight $C_5$–$C_{10}$ $RONO_2$ accounted for only 22\% of plume $NO_y$.
+It means that $C_6$–$C_{10}$ $RONO_2$ account for 78\% of the aged plume $NO_y$ \citep{Neuman2012}.
 This result differs dramatically from previous studies, where the estimated contribution of $RONO_2$ to $NO_y$ was much lower: approximately 4\% in urban and petrochemical plumes [Neuman et al., 2002; Ryerson et al., 2003], approx.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{table}\label{tab:Sommariva2008init}
-\caption{Initial values of the some model parameters (in ppb) for two scenarios used in \citep{Sommariva2008}}
+\begin{table}
+\caption{Initial values of the some model parameters (in ppb) for two scenarios used in \citep{Sommariva2008}}\label{tab:Sommariva2008init}
 \centering
@@
 \end{table}
-\begin{figure}\label{fig:Sommariva2008fig}
+\begin{figure}
 \includegraphics[width=\linewidth]{{D:/UEA/ENV-MB4X_Dissertation/ENV-MB4X_Dissertation_thesis/ENV-MB4X_Dissertation_thesis_latex/pics/Sommariva2008_Fig.1}.png}
-\caption{Modelled concentrations of $O_3$, $NO_x$ and $HNO_3$ for scenario A (left) and B (right). Source: \citep{Sommariva2008}.}
+\caption{Modelled concentrations of $O_3$, $NO_x$ and $HNO_3$ for scenario A (left) and B (right). Source: \citep{Sommariva2008}.}\label{fig:Sommariva2008fig}
 \end{figure}
@@
 In their study, they incorporated methyl and ethyl nitrates in the chemical mechanism of the UCI CTM [Carver et al., 1997] and by matching their observed atmospheric abundances derived a total oceanic emission flux of these two nitrates (0.35 Tg of N per year).
-Comparing two simulations with and without spatially uniform emissions of methyl and ethyl nitrates they found that the largest impact of oceanic alkyl nitrates is seen over the equatorial Western Pacific (Figure ???).
-Atmospheric circulation (convergence of the trade winds) in this region isolates the marine boundary layer, resulting in sustained photochemical loss of $O_3$ (Figure ???
-c) and at the same time accumulation of alkyl nitrates (Figure ???
-a), which drive up boundary layer $NO_x$ abundances by as much as 250\% (Figure ???
-b).
-Such combination of factors leads to increases in $O_3$ of up to 20\% (Figures ???
-c and d), with the largest increases occurring from May to October.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{figure}[t!]\label{fig:Neu2008}
+Comparing two simulations with and without spatially uniform emissions of methyl and ethyl nitrates they found that the largest impact of oceanic alkyl nitrates is seen over the equatorial Western Pacific (Figure \ref{fig:Neu2008}).
+Atmospheric circulation (convergence of the trade winds) in this region isolates the marine boundary layer, resulting in sustained photochemical loss of $O_3$ (Figure \ref{fig:Neu2008M_O3contol}) and at the same time accumulation of alkyl nitrates (Figure \ref{fig:Neu2008M_MeONO2}), which drive up boundary layer $NO_x$ abundances by as much as 250\% (Figure \ref{fig:Neu2008M_NOxdiff}).
+Such combination of factors leads to increases in $O_3$ of up to 20\% (Figures \ref{fig:Neu2008M_O3contol} and \ref{fig:Neu2008M_O3diff}), with the largest increases occurring from May to October.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{figure}[t!]
 	\centering
@@
         \includegraphics[width=\linewidth]{{\picspath Neu2008_Fig.2_MeONO2_OCT_merge}.png}
-        \caption{}
+        \caption{}\label{fig:Neu2008M_MeONO2}
     \end{subfigure}%
@@
         \includegraphics[width=\linewidth]{{\picspath Neu2008_Fig.3_OCT_O3contol}.png}
-        \caption{}
+        \caption{}\label{fig:Neu2008M_O3contol}
     \end{subfigure}
@@
         \includegraphics[width=\linewidth]{{\picspath Neu2008_Fig.3_OCT_NOxdiff}.png}
-        \caption{}
+        \caption{}\label{fig:Neu2008M_NOxdiff}
     \end{subfigure}%
@@
         \includegraphics[width=\linewidth]{{\picspath Neu2008_Fig.3_OCT_O3diff}.png}
-        \caption{}
+        \caption{}\label{fig:Neu2008M_O3diff}
     \end{subfigure}
-\caption{October monthly mean mixing ratios of (a) methyl nitrate ($MeONO_2$, $CH_3ONO_2$) and (c) $O_3$, and percent difference (perturbation - control) in (b) $NO_x$, (d) $O_3$.
+\caption{October monthly mean mixing ratios of (a) methyl nitrate ($MeONO_2$, $CH_3ONO_2$) and (b) $O_3$, and percent difference (perturbation - control) in (c) $NO_x$, (d) $O_3$.
 All averaged over the boundary layer (below 850 mb).
-Adapted from \citep{Neu2008}}
+Adapted from \citep{Neu2008}}\label{fig:Neu2008}
 \end{figure}
@@
 \end{itemize}
-The full list of reactions can be found in Appendix ??
-\ref{sec:appendix1}.
+The full list of reactions can be found in Appendix \ref{sec:appendix1}.
 It is important to mention that due to purely theoretical purpose of this research and its focus on investigation of the impact of alkyl nitrate chemistry on ozone production some of the reactions that lead to formation of species that serve as a sink for $NO_x$ are not included into the chemical mechanism.
@@
 \end{table}
+
 \subsection{Model performance}
@@
 The concentration of $NO_x$ is the main factor which determines whether $O_3$ is produced or removed in the atmosphere, and theoretically there are three regimes of $O_3$ production that depend on $NO_x$ levels: at low $NO_x$ levels the regime is characterized by net $O_3$ removal, at intermediate $NO_x$ (also called $NO_x$-sensitive) - by net $O_3$ formation, and at high $NO_x$ - by net $O_3$ formation or removal depending more on $VOCs$ concentration (therefore called $VOC$-sensitive) (e.g., \citep{Fowler2008}).
-When the prevailing $NO_x$ levels are low, a net $O_3$ removal takes place because there is not enough $NO$ around to react with the peroxy radicals ($HO_2$ and $RO_2$) to produce $NO_2$, which photolysis in conjunction with the reaction (\ref{reac:O3P+O2}) is the main source of $O_3$ in the troposphere.
+When the prevailing $NO_x$ levels are low, a net $O_3$ removal takes place because there is not enough $NO$ around to react with the peroxy radicals ($HO_2$ and $RO_2$) to produce $NO_2$, which photolysis in conjunction with the reaction (\ref{reac:O3P+O2=O3}) is the main source of $O_3$ in the troposphere.
 Some additional $O_3$ removal in this regime occurs due to reaction of $HO_2$ radicals with $O_3$ leading to the regeneration of $OH$ radicals as part of an $O_3$ depleting $OH$-$HO_2$ ($HO_x$) cycle.
 In other words, the following reaction sequence predominates:
-\begin{equation}\label{reac:O3+hv}
-O_3 + hv \rightarrow O(^1D) + O_2
-\end{equation}
-\begin{equation}\label{reac:O1D+H2O}
-O(^1D) + H_2O \rightarrow OH + OH
-\end{equation}
-\begin{equation}\label{reac:O1D=O3P}
-O(^1D) \xrightarrow{M} O(^3P
-\end{equation}
-\begin{equation}\label{reac:O3P+O2}
-O(^3P) + O_2 \xrightarrow{M} O_3
-\end{equation}
-\begin{equation}\label{reac:OH+CO}
-OH + CO \rightarrow HO_2 + CO_2
-\end{equation}
-\begin{equation}\label{reac:OH+RH}
+\begin{equation}
+O_3 + hv \rightarrow O(^1D) + O_2\tag{\ref{reac:O3+hv=O1D+O2}}
+\end{equation}
+\begin{equation}
+O(^1D) + H_2O \rightarrow OH + OH\tag{\ref{reac:O1D+H2O=OH+OH}}
+\end{equation}
+\begin{equation}
+O(^1D) \xrightarrow{M} O(^3P)\tag{\ref{reac:O1D=O3P}}
+\end{equation}
+\begin{equation}
+O(^3P) + O_2 \xrightarrow{M} O_3\tag{\ref{reac:O3P+O2=O3}}
+\end{equation}
+\begin{equation}
+OH + CO \rightarrow HO_2 + CO_2\tag{\ref{reac:OH+CO=H+CO2}}
+\end{equation}
+\begin{equation}\label{reac:OH+RH=RO2+H2O}
 OH + RH \rightarrow RO_2 + H_2O
 \end{equation}
-\begin{equation}\label{reac:HO2+HO2}
-HO_2 + HO_2 \rightarrow H_2O_2 + O_2
-\end{equation}
-\begin{equation}\label{reac:RO2+HO2}
+\begin{equation}
+HO_2 + HO_2 \rightarrow H_2O_2 + O_2\tag{\ref{reac:HO2+HO2=H2O2+O2}}
+\end{equation}
+\begin{equation}\label{reac:RO2+HO2=ROOH+O2}
 RO_2 + HO_2 \rightarrow ROOH + O_2
 \end{equation}
-\begin{equation}\label{reac:O3+HO2}
-O_3 + HO_2 \rightarrow OH + 2O_2
+\begin{equation}
+O_3 + HO_2 \rightarrow OH + 2O_2\tag{\ref{reac:HO2+O3=OH+2O2}}
 \end{equation}
@@
 
-When the prevailing $NO_x$ levels are more than 0.5 ppb \citep{Sillman1999}, but less than a certain critical value, reactions (\ref{reac:HO2+NO}) and (\ref{reac:RO2+NO}) represent the dominant reaction pathways for $HO_2$ and $RO_2$ providing a link for catalytic generation of $O_3$:
-\begin{equation}\label{reac:HO2+NO}
-HO_2 + NO \rightarrow OH + NO_2
-\end{equation}
-\begin{equation}\label{reac:RO2+NO}
+When the prevailing $NO_x$ levels are more than 0.5 ppb \citep{Sillman1999}, but less than a certain critical value, reactions (\ref{reac:HO2+NO=NO2+OH}) and (\ref{reac:RO2+NO=RO+NO2}) represent the dominant reaction pathways for $HO_2$ and $RO_2$ providing a link for catalytic generation of $O_3$:
+\begin{equation}
+HO_2 + NO \rightarrow OH + NO_2\tag{\ref{reac:HO2+NO=NO2+OH}}
+\end{equation}
+\begin{equation}\label{reac:RO2+NO=RO+NO2}
 RO_2 + NO \rightarrow RO + NO_2
 \end{equation}
-\begin{equation}\label{eq:NO2+hv}
-NO_2 + hv \rightarrow NO + O(^3P)
-\end{equation}
-\begin{equation} %same
-!!!!! same O(^3P) + O_2 \xrightarrow{M} O_3
-\end{equation}
-The competition between reactions (\ref{reac:HO2+NO}) and (\ref{reac:HO2+HO2}) for $HO_2$, and reactions (\ref{reac:RO2+NO}) and (\ref{reac:RO2+HO2}) for $RO_2$, allows a greater number of free-radical propagated $O_3$ forming cycles to happen prior to termination (reactions (\ref{reac:HO2+HO2}) and (\ref{reac:RO2+HO2})), therefore resulting in a net $O_3$ formation.
-
-When the prevailing $NO_x$ levels are high, reactions (\ref{reac:HO2+NO}) and (\ref{reac:RO2+NO}) dominate for $HO_2$ and $RO_2$, making the rate of $O_3$ production being controlled by the availability of these radicals.
-The availability of $OH$ radical is also important at high $NO_x$ levels because $OH$ connects the $O_3$ production rate with rates of its reactions with $CO$ and $RH$ (reactions (\ref{reac:OH+CO}) and (\ref{reac:OH+RH})), therefore making this regime $VOC$-sensitive.
+\begin{equation}
+NO_2 + hv \rightarrow NO + O(^3P)\tag{\ref{reac:NO2+hv=NO+O3P}}
+\end{equation}
+\begin{equation}
+O(^3P) + O_2 \xrightarrow{M} O_3\tag{\ref{reac:O3P+O2=O3}}
+\end{equation}
+The competition between reactions (\ref{reac:HO2+NO=NO2+OH}) and (\ref{reac:HO2+HO2=H2O2+O2}) for $HO_2$, and reactions (\ref{reac:RO2+NO=RO+NO2}) and (\ref{reac:RO2+HO2=ROOH+O2}) for $RO_2$, allows a greater number of free-radical propagated $O_3$ forming cycles to happen prior to termination (reactions (\ref{reac:HO2+HO2=H2O2+O2}) and (\ref{reac:RO2+HO2=ROOH+O2})), therefore resulting in a net $O_3$ formation.
+
+When the prevailing $NO_x$ levels are high, reactions (\ref{reac:HO2+NO=NO2+OH}) and (\ref{reac:RO2+NO=RO+NO2}) dominate for $HO_2$ and $RO_2$, making the rate of $O_3$ production being controlled by the availability of these radicals.
+The availability of $OH$ radical is also important at high $NO_x$ levels because $OH$ connects the $O_3$ production rate with rates of its reactions with $CO$ and $RH$ (reactions (\ref{reac:OH+CO=H+CO2}) and (\ref{reac:OH+RH=RO2+H2O})), therefore making this regime $VOC$-sensitive.
 Plus at this point the reaction of $OH$ with $NO_2$ to form nitric acid ($HNO_3$) becomes the major termination reaction for free radicals:
-\begin{equation}\label{reac:OH+NO2}
-OH + NO_2 \xrightarrow{M} HNO_3
+\begin{equation}
+OH + NO_2 \xrightarrow{M} HNO_3\tag{\ref{reac:OH+NO2=HNO3}}
 \end{equation}
 Under these circumstances, the number of free-radical propagated $O_3$ forming cycles decreases with increasing $NO_x$, and the production rate of $O_3$ decreases.
-However, if the concentrations of $VOCs$ are high, reactions of $OH$ with $CO$ and $RH$ ((\ref{reac:OH+CO}), (\ref{reac:OH+RH})) can compete with reaction (\ref{reac:OH+NO2}), leading to an increase in the $O_3$ production rate.
+However, if the concentrations of $VOCs$ are high, reactions of $OH$ with $CO$ and $RH$ ((\ref{reac:OH+CO=H+CO2}), (\ref{reac:OH+RH=RO2+H2O})) can compete with reaction (\ref{reac:OH+NO2=HNO3}), leading to an increase in the $O_3$ production rate.
 
@@
 $NO$ is rapidly converted to $NO_2$, but slower than at low $NO_x$ levels because of their higher initial concentrations (Figure \ref{fig:NOvalid}, \ref{fig:NO2valid}).
-Due to a higher efficiency of reaction (\ref{reac:HO2+NO}) $OH$ mixing ratio is higher at higher $NO_x$ levels, resulting in a more pronounced loss of $CO$ (Figure \ref{fig:COvalid}).
+Due to a higher efficiency of reaction (\ref{reac:HO2+NO=NO2+OH}) $OH$ mixing ratio is higher at higher $NO_x$ levels, resulting in a more pronounced loss of $CO$ (Figure \ref{fig:COvalid}).
 
@@
 Above the 'ridge line' the $VOC$/$NO_x$ ratios are low, and the $O_3$ production rate increases with increasing $VOC$ and decreases with increasing $NO_x$ \citep{Sillman2013}.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \begin{figure}[t!]
@@
 As it has been noted in Section \ref{sec:intro}, the relationship between $O_3$ and $RONO_2$ production rates is controlled by the branching ratio ($\alpha_3$) between reactions:
-\begin{equation}\label{reac:RO2+NO=RO+NO2}
-RO_2 + NO \xrightarrow{1-\alpha_3} RO + NO_2
-\end{equation}
-\begin{equation}\label{reac:RO2+NO=RONO2}
-RO_2 + NO \xrightarrow{\alpha_3} RONO_2
-\end{equation}
-Reaction (\ref{reac:RO2+NO=RO+NO2}) leads to $O_3$ production and $NO$ regeneration, resulting in continuing $O_3$ production, while reaction (\ref{reac:RO2+NO=RONO2}) terminates this process.
+\begin{equation}
+RO_2 + NO \xrightarrow{1-\alpha_3} RO + NO_2\tag{\ref{reac:RO2+NO=aphaRO+NO2}}
+\end{equation}
+\begin{equation}
+RO_2 + NO \xrightarrow{\alpha_3} RONO_2\tag{\ref{reac:RO2+NO=aphaRONO2}}
+\end{equation}
+Reaction (\ref{reac:RO2+NO=aphaRO+NO2}) leads to $O_3$ production and $NO$ regeneration, resulting in continuing $O_3$ production, while reaction (\ref{reac:RO2+NO=aphaRONO2}) terminates this process.
 Branching ratios used in the model are taken from the MCM v3.3, and their values (Table \ref{tab:ANbranching}) were kept constant in all model runs.
@@
 
+\appendix
 \section{Appendix I} \label{sec:appendix1}

### 2015-08-05 MSc_diss.tex
Updates in figures (all in .eps)

@@
 \usepackage{vmargin}
-\setmarginsrb{2cm}{2cm}{2cm}{2cm}{0cm}{0cm}{0cm}{1.5cm}
+\setmarginsrb{2.5cm}{2cm}{2.5cm}{2cm}{0cm}{0cm}{0cm}{1.5cm}
 \usepackage{pdflscape}
@@
 \usepackage{multirow}
+\usepackage{tabularx}
+\usepackage{array}
+\usepackage{ragged2e}
+\newcolumntype{P}[1]{>{\RaggedRight\hspace{0pt}}p{#1}}
 % Tikz styles
@@
 \begin{abstract}
-Abstract.
+Abstract.300 words.
 \end{abstract}
@@
 \caption{!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!}\label{tab:ANmean} % mean AN mixing ratios
-\begin{tabular}{cccc}
+\begin{tabularx}{\linewidth}{P{3cm}P{2cm}P{5cm}X}
 \hline
@@
 \hline
-remote marine   & 10 pptv    & methyl                   & \citep{Reeves2007} \\
-(North Atlantic)&            & ethyl                    & \\
-airborne        &            & 1-propyl                 & \\   
-                             && 2-propyl                & \\   
-                             && 2-butyl                 & \\                                                                                               
-                             && 2-methyl-3-butyl        & \\
-                             && sum of 2- and 3-pentyl  & \\
-\hline
-rural           & 14 pptv    & methyl                   & \citep{Roberts1998} \\
-(Nova Scotia, Canada)        && ethyl   & \\
-ground                       && n-propyl & \\
-                             && 2-propyl & \\
-                             && 2-butyl  & \\ 
-\hline
-suburban                     & 17.48 pptv  & 2-propyl    & \citep{DeKock1994} \\
-(Cape Retire, South Africa)  &(3.62-122.23)& 1-propyl & \\
-ground                                    && 2-butyl  & \\
-                                          && 1-butyl  & \\
-                                          && 3-pentyl & \\
-                                          && 2-pentyl & \\
-                                          && 2-methyl-1-butyl & \\
-                                          && i-pentyl & \\
-\hline
-suburban                     & 119.8 ppt  & $C_1$-$C_8$       & \citep{Flocke1998} \\
-(Black Forest, Germany)      & (30-630 ppt) && \\ 
-ground                                     &&& \\
-\hline
-polluted marine & approx 1 ppb & $NO_y$-($NO_x$+$HNO_3$+$PAN$) & \citep{Neuman2012} \\
-(Gulf of Mexico)&(plume center) && \\
-airborne &&& \\
-\hline
-\end{tabular}
+remote marine (North Atlantic), airborne   & 10 pptv    & methyl, ethyl, 1-propyl, 2-propyl, 2-butyl, 2-methyl-3-butyl, sum of 2- and 3-pentyl  & \citep{Reeves2007} \\
+\hline
+rural ground (Nova Scotia, Canada)        & 14 pptv    & methyl, ethyl, n-propyl, 2-propyl, 2-butyl  & \citep{Roberts1998} \\
+\hline
+suburban ground (Cape Retire, South Africa)  & 17.48 pptv (3.62-122.23) & 2-propyl, 1-propyl, 2-butyl, 1-butyl, 3-pentyl, 2-pentyl, 2-methyl-1-butyl, i-pentyl & \citep{DeKock1994} \\
+\hline
+suburban ground (Black Forest, Germany)    & 119.8 ppt (30-630 ppt) & $C_1-C_8$ & \citep{Flocke1998} \\
+\hline
+polluted marine airborne (Gulf of Mexico) & approx 1 ppb (plume center)& $NO_y-(NO_x+HNO_3+PAN)$ & \citep{Neuman2012} \\
+\hline
+\end{tabularx}
 \end{table}
@@
 Branching ratios used in the model are taken from the MCM v3.3, and their values (Table \ref{tab:ANbranching}) were kept constant in all model runs.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \begin{table} %  AN branching ratios
@@
 \end{table}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{figure} % netO3rate noAN withAN diff
-\centering
-\begin{minipage}{.3\textwidth}
-  \centering
-  \includegraphics[width=.99\linewidth]{D:/FACSIMILE/ANsCBmodel/matlab/ANsCB_pics/ANsNOxVOC/noAN/chem_noAN_netO3rate_noINORG_NOx_250ppt_100ppb.png}
-\end{minipage}
-\begin{minipage}{.3\textwidth}
-  \centering
-  \includegraphics[width=.99\linewidth]{D:/FACSIMILE/ANsCBmodel/matlab/ANsCB_pics/ANsNOxVOC/allAN/chem_allAN_netO3rate_noINORG_NOx_250ppt_100ppb.png}
-\end{minipage}
-\begin{minipage}{.3\textwidth}
-  \centering
-  \includegraphics[width=.99\linewidth]{D:/FACSIMILE/ANsCBmodel/matlab/ANsCB_pics/ANsNOxVOC/allAN/chem_allAN_netO3ratediff_noINORG_NOx_250ppt_100ppb.png}
-\end{minipage}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{figure}[h!] % netO3rate noAN withAN diff
+	\centering
+	\begin{subfigure}[t]{0.45\textwidth}
+        \centering
+        \includegraphics[width=\linewidth]{D:/FACSIMILE/ANsCBmodel/matlab/ANsCB_pics/ANsNOxVOC/noAN/chem_noAN_netO3rate_noINORG_NOx_250ppt_100ppb.eps}
+        \caption{}
+    \end{subfigure}%
+    \hfill
+    \begin{subfigure}[t]{0.45\textwidth}
+        \centering
+        \includegraphics[width=\linewidth]{D:/FACSIMILE/ANsCBmodel/matlab/ANsCB_pics/ANsNOxVOC/allAN/chem_allAN_netO3rate_noINORG_NOx_250ppt_100ppb.eps}
+        \caption{}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[t]{0.45\textwidth}
+        \centering
+        \includegraphics[width=\linewidth]{D:/FACSIMILE/ANsCBmodel/matlab/ANsCB_pics/ANsNOxVOC/allAN/chem_allAN_netO3ratediff_noINORG_NOx_250ppt_100ppb.eps}
+        \caption{}
+    \end{subfigure}%
+    \hfill
 \caption{Isopleths giving net rate of $O_3$ production (ppb/h, color) as a function of $VOC$ (ppbC) and $NO_x$ (ppb) with alkyl nitrate chemistry switched off (left) and switched on (middle).
@@
 \end{figure}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{figure} % netO3mixrat noAN allAN diff diffpercent
-\centering
-\begin{minipage}{.4\textwidth} % noAN
-  \centering
-  \includegraphics[width=.99\linewidth]{D:/FACSIMILE/ANsCBmodel/matlab/ANsCB_pics/ANsNOxVOC/noAN/chem_noAN_netO3mixrat_noINORG_NOx_250ppt_100ppb.png}
+\begin{figure}[h] % netO3mixrat noAN allAN diff diffpercent
+\centering
+\begin{minipage}{.45\textwidth} % noAN
+  \centering
+  \includegraphics[width=.99\linewidth]{D:/FACSIMILE/ANsCBmodel/matlab/ANsCB_pics/ANsNOxVOC/noAN/chem_noAN_netO3mixrat_noINORG_NOx_250ppt_100ppb.eps}
 \end{minipage}
-\begin{minipage}{.4\textwidth} % diff
-  \centering
-  \includegraphics[width=.99\linewidth]{D:/FACSIMILE/ANsCBmodel/matlab/ANsCB_pics/ANsNOxVOC/allAN/chem_allAN_netO3mixratdiff_noINORG_NOx_250ppt_100ppb.png}
+\begin{minipage}{.45\textwidth} % diff
+  \centering
+  \includegraphics[width=.99\linewidth]{D:/FACSIMILE/ANsCBmodel/matlab/ANsCB_pics/ANsNOxVOC/allAN/chem_allAN_netO3mixratdiff_noINORG_NOx_250ppt_100ppb.eps}
 \end{minipage}
-\begin{minipage}{.4\textwidth} % allAN
-  \centering
-  \includegraphics[width=.99\linewidth]{D:/FACSIMILE/ANsCBmodel/matlab/ANsCB_pics/ANsNOxVOC/allAN/chem_allAN_netO3mixrat_noINORG_NOx_250ppt_100ppb.png}
+\begin{minipage}{.45\textwidth} % allAN
+  \centering
+  \includegraphics[width=.99\linewidth]{D:/FACSIMILE/ANsCBmodel/matlab/ANsCB_pics/ANsNOxVOC/allAN/chem_allAN_netO3mixrat_noINORG_NOx_250ppt_100ppb.eps}
 \end{minipage}
-\begin{minipage}{.4\textwidth} % diffpercent
-  \centering
-  \includegraphics[width=.99\linewidth]{D:/FACSIMILE/ANsCBmodel/matlab/ANsCB_pics/ANsNOxVOC/allAN/chem_allAN_netO3mixratdiffpercent_noINORG_NOx_250ppt_100ppb.png}
+\begin{minipage}{.45\textwidth} % diffpercent
+  \centering
+  \includegraphics[width=.99\linewidth]{D:/FACSIMILE/ANsCBmodel/matlab/ANsCB_pics/ANsNOxVOC/allAN/chem_allAN_netO3mixratdiffpercent_noINORG_NOx_250ppt_100ppb.eps}
 \end{minipage}
@@
 \begin{figure} % endtotalANs
-\includegraphics[width=.5\linewidth]{D:/FACSIMILE/ANsCBmodel/matlab/ANsCB_pics/ANsNOxVOC/allAN/chem_allAN_endtotalANs_noINORG_NOx_250ppt_100ppb.png}
+\centering
+\includegraphics[width=.7\linewidth]{D:/FACSIMILE/ANsCBmodel/matlab/ANsCB_pics/ANsNOxVOC/allAN/chem_allAN_endtotalANs_noINORG_NOx_250ppt_100ppb.eps}
 \caption{Isopleths of net production of alkyl nitrates ($\Sigma RONO_2$) as a function of $NO_x$ and $VOCs$ levels.}\label{fig:endtotalANs}

### 2015-08-06 MSc_diss.tex
Finished objective and aims

@@
 
-Unlike many other air pollutants, tropospheric ozone is not directly emitted, but produced following the oxidation of carbon monoxide ($CO$), methane ($CH_4$), and nonmethane volatile compounds (nmVOCs) in the presence of nitrogen oxides (NOx) \citep{Crutzen1973,Myhre2013}.
+Unlike many other air pollutants, tropospheric ozone is not directly emitted, but produced following the oxidation of carbon monoxide ($CO$), methane ($CH_4$), and nonmethane volatile compounds (nmVOCs) in the presence of nitrogen oxides ($NO_x$) \citep{Crutzen1973,Myhre2013}.
 The emissions of these precursors arise from both natural and anthropogenic sources which have intensified since pre-industrial times \citep{Parrish2014,Volz1988} (Figure \ref{fig:O3observations}).
@@
 
-According to the Intergovernmental Panel on Climate Change (IPCC) Fifth Assessment report (AR5) for present conditions (around year 2000) the global mean tropospheric ozone budget is approximately 331 Tg \citep{Myhre2013}.
-However, despite general agreement on how the drivers impact ozone concentrations at global and regional scales, its budget varies considerably between different modelling and observational studies \citep{Stevenson2006,Wild2007,Young2012}.
-Figure \ref{fig:O3budget} demonstrates that the range of the global tropospheric ozone budget estimates is approximately 60 Tg which is 18\% of the mean.
+According to the Intergovernmental Panel on Climate Change (IPCC) Fifth Assessment report (AR5) in the present conditions (around year 2000) the global mean tropospheric ozone budget is approximately 331 Tg \citep{Myhre2013}.
+However, despite general agreement on how the drivers impact ozone concentrations at global and regional scales, there is considerable inconsistency in its budget between different modelling and observational studies \citep{Stevenson2006,Wild2007,Young2012}.
+Figure \ref{fig:O3budget} demonstrates that the range of global tropospheric ozone budget estimates is approximately 60 Tg which is 18\% of the mean.
 In terms of ozone production the variation is even higher and accounts for 28\% of the corresponding value.
@@
 \centering
-\includegraphics[width=0.7\linewidth]{o3_studies.eps}
+\includegraphics[width=0.8\linewidth]{o3_studies.eps}
 \caption{Summary of tropospheric ozone global budget model and observation estimates for present (about 2000) conditions.
@@
 
-A considerable effort has been made to tackle these problems, and generally there are two possible ways of doing it.
+A significant effort has been made to tackle these problems, and broadly speaking there are two possible ways of doing it.
 The first one implies the direct analysis of tropospheric chemical schemes used in climate-chemistry models.
-For example, Emmerson and Evans (2009) compared the state of the art Master Chemical Mechanism which contains approximately 5600 species and 13500 reactions \citep{Jenkin2002} with other six chemical schemes and found four significant variations between them.
-This approach is linked with big challenges since a modern global model is a very sophisticated representation of the climate-chemistry interactions, and therefore it is often difficult to separate influence of one reaction from another.
-However, even being so developed climate-chemistry models do not take into account all known chemical reactions taking place in the real troposphere.
+For example, \cite{Emmerson2009} compared the state of the art Master Chemical Mechanism which contains approximately 5600 species and 13500 reactions \citep{Jenkin2002} with other six chemical schemes and found significant variations between them.
+This approach is associated with serious challenges since a modern global model is a very sophisticated representation of the climate-chemistry interactions, and therefore it is often difficult to separate influence of one reaction from another.
+However, even advanced climate-chemistry models do not take into account all known chemical reactions in the real troposphere.
 A complete representation requires many thousands of species and tens of thousands of reactions, which is beyond the numerical capabilities currently available.
@@
 
-! add something from report!
-
 The second way to understanding the discrepancies between models and observations is to conduct an idealised theoretical study using a conceptual model.
-The advantage of this approach is that such kind of models can be quite easily constructed, and more importantly can be focused on representation of a certain chemical reactions which presumably could explain the discrepancy between models and observations.
+The advantage of this approach is that such kind of models can be quite easily constructed, and more importantly can be focused on representation of a certain chemical reactions which possibly could explain the discrepancy between models and observations.
 
 \subsubsection*{Ozone tropospheric chemistry}\label{intro_O3chem}
-%\input{O3CycleScheme}
-Having identified the role of tropospheric ozone, it is time to remind ourselves the chemistry that controls abundance of it.
+Keeping in mind the role of tropospheric ozone, a necessary step is to revisit the chemistry that controls its abundance.
 
 The key chemical processes that lead to $O_3$ formation and destruction in the troposphere are driven by reaction cycles involving free-radical intermediates, which are mainly formed from the photolysis of $O_3$ itself (\ref{reac:O3+hv=O1D+O2}) \citep{Fowler2008}.
-At UV wavelengths shorter than 320 nm, $O_3$ photolysis generates electronically excited oxygen atoms ($O(^1D)$), which than can either react with water vapour to form $OH$ radicals (\ref{reac:O1D+H2O=OH+OH}), or collide with an inert molecule (denoted '$M$') to reform $O_3$ (\ref{reac:O1D=O3P} and \ref{reac:O3P+O2=O3}):
+At UV wavelengths shorter than 320 nm, $O_3$ photolysis generates excited oxygen atoms ($O(^1D)$), which than can either react with water vapour to form $OH$ radicals (\ref{reac:O1D+H2O=OH+OH}), or collide with an inert molecule (denoted '$M$') to reform $O_3$ (\ref{reac:O1D=O3P} and \ref{reac:O3P+O2=O3}):
 \begin{equation}\label{reac:O3+hv=O1D+O2}
@@
 \end{equation}
-As shown in Figure \ref{fig:O3scheme} reactions (\ref{reac:HO2+NO=NO2+OH}) and (\ref{reac:CH3O2+NO=NO2+CH3O}) form parts of the free-radical propagated $O_3$ forming cycles, which may occur a number of times before being halted by radical termination reactions (\ref{reac:HO2+HO2=H2O2+O2} and \ref{reac:HO2+CH3O2=CH3OOH+O2}), and the higher $NO_x$ level is the greater number of these cycles can occur.
+As shown in Figure \ref{fig:O3scheme}, the reactions (\ref{reac:HO2+NO=NO2+OH}) and (\ref{reac:CH3O2+NO=NO2+CH3O}) produce parts of the free-radical propagated $O_3$ forming cycles, which may occur a number of times before being halted by radical termination reactions (\ref{reac:HO2+HO2=H2O2+O2} and \ref{reac:HO2+CH3O2=CH3OOH+O2}), and the higher $NO_x$ level is the greater number of these cycles can occur.
 Therefore, in intermediate $NO_x$ conditions a net $O_3$ formation takes place and it is sensitive to $NO_x$.
@@
 However, elevated emissions of $CO$ and $CH_4$ (and other nonmethane volatile organic compounds (nmVOCs)) may make reactions (\ref{reac:OH+CO=H+CO2})-(\ref{reac:OH+CH4=CH3+H2O}) more competitive with the reaction (\ref{reac:OH+NO2=HNO3}) and lead to an increase in the $O_3$ production rate.
-Therefore, in high-$NO_x$ conditions the $O_3$ production rate can either decrease or increase depending on $NO_x$ and $VOCs$ levels.
+Therefore, in high-$NO_x$ conditions the $O_3$ production rate can either decrease or increase depending on $NO_x$ and VOCs levels.
 
@@
 therefore also having an impact on the rates of $O_3$ production \citep{Fowler2008} as do some other species, which are not mentioned in this section.
-Their role is described in other sections as far as it is necessary.
+Their role is described with a required depth in the sections below.
 
@@
 Alkyl nitrates ($RONO_2$) are organic trace gases, which through one of their formation mechanisms are closely linked with the production of tropospheric ozone \citep{Reeves2007}.
-This mechanism is present in the following section.
+This mechanism is presented in the following section.
 \subsubsection*{Alkyl nitrates chemistry}\label{into_ANchem}
@@
 \end{subequations}
-The latter reaction can serve as a termination step of the propagated $O_3$ forming cycle (Figure \ref{fig:O3scheme}), and as \citep{Roberts1998} estimated it is likely to begin to occur when the $NO$ mixing ratio increases up to 100-200 ppt.
+The latter reaction can serve as a termination step of the propagated $O_3$ forming cycle (Figure \ref{fig:O3scheme}), and as \cite{Roberts1998} estimated it is likely to begin to occur when the $NO$ mixing ratio increases up to 100-200 ppt.
 The yield of alkyl nitrate formation is controlled by the branching ratio ($\alpha_3$), which is temperature and pressure dependent and increases with the number of carbons in the alkyl peroxy radical \citep{Roberts1990}.
@@
 The lifetimes of alkyl nitrates depend on the seasonality of the photolysis and $OH$ sink, which in turn vary with latitude and altitude.
-According to \citep{Clemitshaw1997} and \citep{Turberg1990} alkyl nitrates have atmospheric lifetimes of several days to several years, which makes them temporary reservoirs for the short lived $NO_x$ compounds.
+According to \cite{Clemitshaw1997} and \cite{Turberg1990} alkyl nitrates have atmospheric lifetimes of several days to several years, which makes them temporary reservoirs for the short lived $NO_x$ compounds.
 Being transported from polluted areas to remote background regions and eventually destroyed alkyl nitrates can 'release' $NO_2$ back to the atmosphere, thereby contributing to tropospheric ozone formation.
+
 \subsection{Ozone, alkyl nitrates and carbon bonds}\label{sec:intro_CB}
@@
 In other words, $OPE$ shows the amount of ozone created ($P_{O_3}$) per molecule of $NO_x$ removed ($L_{NO_x}$) from the system \citep{Jacob1999}.
-Due to the reasons described in section \ref{sec:method_chem} \footnote{No representation of the nitric acid ($HNO_3$) in the model.}, we use another expression for calculations of the $O_3$ production efficiency.
+Due to the reasons mentioned in section \ref{sec:method_chem} \footnote{No representation of the nitric acid ($HNO_3$) in the model.}, we use another expression for calculations of the $O_3$ production efficiency.
 This expression has been recently proposed by Professor Matthew Evans (University of York) amid the mounting concerns about the need of thorough intercomparison of chemical mechanisms of global atmospheric models (including those used in the IPCC reports), why it is more universal and applicable to any kind of model.
 
-According to \citep{Evans2014} (personal communication), the $O_3$ production efficiency can be estimated via calculation of the number of carbon bonds emitted into the system.
-For example, if the system contains one molecule of methane, $CH_4$, as 'a carbon bond reservoir', it means that this system has 4 carbon bonds (4 C-H), if  instead of methane it contains one molecule of ethane, $C_2H_6$, the system has 7 carbon bonds (6 C-H bonds and 1 C-C), and so on \citep{Edwards2013}.
+According to \cite{Evans2014} (personal communication), the $O_3$ production efficiency can be estimated via calculation of the number of carbon bonds emitted into the system\footnote{The same kind of approach is used in the Common Representative Intermediates (CRI) mechanism for VOC degradation \citep{Jenkin2002, Jenkin2008}.
+It is a reduced chemical mechanism that describes $O_3$ formation from the tropospheric degradation of methane and 115 emitted non-methane hydrocarbons and oxygenated volatile organic compounds developed using the Master Chemical Mechanism version 3.1 (MCM v3.1) as a reference benchmark.}.
+For example, if the system contains one molecule of methane, $CH_4$, as 'a carbon bond reservoir', it means that this system has 4 carbon bonds (4 $C-H$), if  instead of methane it contains one molecule of ethane, $C_2H_6$, the system has 7 carbon bonds (6 $C-H$ bonds and 1 $C-C$), and so on \citep{Edwards2013}.
 Assuming that every carbon bond is broken chemically and every organic peroxy radical ($RO_2$) reacts with $NO$, the overall efficiency of carbon bonds conversion to $O_3$, $\epsilon$, can be calculated as:
@@
 The fact that $\epsilon$ does not account for reactions where no $NO$-to-$NO_2$ conversion happens, as, for example, in case of the formation of the alkyl nitrates, $\epsilon$ provides a way to relate the rate of $O_3$ production to the alkyl nitrate chemistry.
-Knowing the rate of carbon bond emissions ($E_{C_{bons}}$) from emissions inventory data and knowing the rate of $O_3$ production from a model run ($P_{O_3}$), one can estimate the efficiency of $O_3$ production per carbon bond emitted in their model, and than compare it with other such values from different models potentially revealing the true cause of discrepancies between models and observations of tropospheric $O_3$ budgets.
-
-For link at the bottom?
-The same kind of approach is used in the Common Representative Intermediates (CRI) mechanism for VOC degradation \citep{Jenkin2002, Jenkin2008}.
-It is a reduced chemical mechanism that describes $O_3$ formation from the tropospheric degradation of methane and 115 emitted non-methane hydrocarbons and oxygenated volatile organic compounds developed using the Master Chemical Mechanism version 3.1 (MCM v3.1) as a reference benchmark
+Knowing the rate of carbon bond emissions ($E_{C_{bons}}$) from emissions inventory data and knowing the rate of $O_3$ production from a model run ($P_{O_3}$), one can estimate the efficiency of $O_3$ production per carbon bond emitted in their model, and than compare it with other such values from different models potentially revealing the true cause of discrepancies in tropospheric $O_3$ budgets between models and observations.
+
 \section{Overall objective and specific aims}\label{sec:objams}
-As the reader might anticipateit is mentioned in section \ref{sec:lit_review}
-\citep{Sommariva2008}
-[Typically, chemical mechanisms do not contain
-enough information about the various formation routes of secondary organic compounds, leading to the underestimation of their concentrations, with possible consequences on the estimates of ozone production in remote regions, or to ineffective ozone-control policies.
-In this work, a quasi-explicit chemical model based upon the Master Chemical Mechanism (MCM) has been used to study in detail the photochemistry of a number of alkyl nitrates and determine quantitatively their formation processes.
-This kind of study shows the potential of using mechanisms with a high level of chemical detail to study the formation of pollutants in the atmosphere (Madronich, 2006).]
-\subsection{Overall objective}\label{sec:objams_obj}
-The overall purpose of this research is to investigate the impact of alkyl nitrate chemistry on ozone production efficiency.
-\subsubsection*{Specific aims}\label{sec:objams_ams}
-Specific aims are to investigate:
+Despite of all of the aforesaid, typically chemical mechanisms of global atmospheric models do not contain enough information about some secondary organic compounds \citep{Sommariva2008}, and alkyl nitrates are among them.
+Underestimation of their concentrations could potentially lead to errors in the estimates of ozone production in remote regions, or to ineffective ozone-control policies.
+Therefore, the overall objective of this research is to investigate the impact of alkyl nitrate chemistry on tropospheric ozone production by using a chemical mechanism that represents the chemistry related to the problem.
+
+The specific aims are:
 \begin{itemize}
-\item[(1)] the relationship between ozone production and carbon bond emissions;
-\item[(2)] the impact of inclusion of alkyl nitrates chemistry on this relationship;
-\item[(3)] how (1) and (2) vary in different $NO_x$ and $VOC$ regimes;
-\item[(4)] build different version of the box model;
-\item[(5)] to evaluate the results.
+\item[(1)] to build a conceptual box model which chemical mechanism represents the basics of tropospheric chemistry of ozone and includes reactions of formation and loss of alkyl nitrates;
+\item[(2)] to estimate the sensitivity of overall tropospheric ozone production to alkyl nitrates chemistry;
+\item[(3)] to explore the the relationship between ozone production and carbon bonds emissions;
+\item[(3)] to estimate how (2) and (3) vary in different $NO_x$ and VOC regimes, which could be achieved by designing an appropriate set of experiments.
 \end{itemize}
 
-The brief description of the measurement sites and data is presented in section \ref{sec:data}.
-In section \ref{sec:method} the methodology used in this work is defined.
+The literature review on the problem is presented in section \ref{sec:lit_review}.
+In section \ref{sec:method} the conceptual model of global troposhere is described.
 Section \ref{sec:res} contains the results.
-Section \ref{sec:discuss} discusses the results, particularly focusing attention on the analysis of the discrepancy between the observational data and the results of theoretical modelling.
+Section \ref{sec:discuss} discusses the results and also includes an outline of future work.
 Conclusions constitute section \ref{sec:conclusion}.
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Literature review} \label{sec:lit_review}
-A lot of work is has been dedicated to measurements of alkyl nitrates in different environments and even more to measurements of ozone.
+A lot of work has been dedicated to measurements of alkyl nitrates in different environments and even more to measurements of ozone.
 Here we focus our attention on some major investigations that lead us to our current understanding of the relationship between tropospheric ozone and alkyl nitrates.
 
-During one of these investigations, namely the SAGA campaign, \citep{Atlas1993} first postulated that the short-chained alkyl nitrates ($C_1$-$C_3$) can be produced not only photochemically through the oxidation of their parent hydrocarbons \citep{Roberts1990}, but also emitted directly from equatorial oceans.
+During one of these investigations, namely the SAGA campaign, \cite{Atlas1993} first postulated that the short-chained alkyl nitrates ($C_1$-$C_3$) can be produced not only photochemically through the oxidation of their parent hydrocarbons \citep{Roberts1990}, but also emitted directly from equatorial oceans.
 This has since been confirmed by field measurements \citep{Chuck2002,Blake2003,Dahl2005}.
-\citep{Chuck2002} provided the first direct evidence of that based on detection of supersaturation of surface waters by methyl and ethyl nitrates and their high concentrations in the marine air over equatorial areas (0-40$^{\circ}$) of the Atlantic.
-Later during two extensive NASA airborne campaigns Pacific Exploratory Mission (PEM)-Tropics A (September-October 1996) and PEM-Tropics B (March-April 1999) \citep{Blake2003} observed high levels of methyl and ethyl nitrates (approximately 50 ppt) in the air at the band between 8$^{\circ}$N-13$^{\circ}$S stretching across the equatorial Pacific, and pointed out that oceanic source of $\geq C_4$ alkyl nitrates is insignificant compared to anthropogenic sources.
-
-Alkyl nitrates also have a biomass burning source which have been first reported by \citep{Simpson2002}.
-They found that in the immediate vicinity of savanna fires in Northern Australia the average alkyl nitrate mixing ratios were 47–122 times higher than local background mixing ratios, exceeding 3 ppbv for methyl nitrate.
+\cite{Chuck2002} provided the first direct evidence of that based on detection of supersaturation of surface waters by methyl and ethyl nitrates and their high concentrations in the marine air over equatorial areas (0-40$^{\circ}$) of the Atlantic.
+Later during two extensive NASA airborne campaigns: Pacific Exploratory Mission (PEM)-Tropics A (September-October 1996) and PEM-Tropics B (March-April 1999), \cite{Blake2003} observed high levels of methyl and ethyl nitrates (approximately 50 ppt) in the air at the band between 8$^{\circ}$N--13$^{\circ}$S stretching across the equatorial Pacific, and pointed out that oceanic source of $\geq C_4$ alkyl nitrates is insignificant compared to anthropogenic sources.
+
+Alkyl nitrates also have a biomass burning source which have been first reported by \cite{Simpson2002}.
+They found that in the immediate vicinity of savanna fires in Northern Australia the average alkyl nitrate mixing ratios were 47-–122 times higher than local background mixing ratios, exceeding 3 ppbv for methyl nitrate.
 Methyl nitrate was primarily emitted during the 'flaming' stage, while $C_2$-$C_4$ ethyl-butyl nitrates dominated during the 'smoldering' stage.
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \renewcommand{\arraystretch}{1}
@@
 Owing to the existence of the relationship between alkyl nitrates and their parent hydrocarbons, measurements of their concentrations can provide information on the extent of photochemical processing in the air mass.
-\cite{Bertman1995} developed a kinetic approach to describe this relationship through the evolution of the ratio of alkyl nitrate to its parent hydrocarbon as a function of time.
+\cite{Bertman1995} developed a kinetic approach to describe this relationship through the evolution of the ratio of alkyl nitrate to its parent hydrocarbon.
 This approach has been extensively used to study air mass photochemical ageing at several sites, in North America \citep{Bertman1995,Roberts1998}, China \citep{Simpson2006}, the British Isles \citep{Worton2010} and over the Pacific \citep{Simpson2003} and North Atlantic Ocean \citep{Reeves2007,Stroud2001}.
 
-A study that enabled us to focus on examination of $O_3$ and $RONO_2$ formation processes has been conducted by \citep{Neuman2012}.
-After an explosion at the Deepwater Horizon (DWH) offshore oil platform on 20 April 2010 large amounts of oil from the DWH spill reached the ocean surface.
+A study that enabled us to focus on examination of $O_3$ and $RONO_2$ formation processes has been conducted by \cite{Neuman2012}.
+After the Deepwater Horizon (DWH) offshore oil platform accident on 20 April 2010 large amounts of oil from the DWH spill reached the ocean surface.
 Subsequent oil evaporation caused a release of little methane and other light alkanes to the atmosphere forming a $VOC$ mixture atypical for the remote marine environment.
 At the same the $NO_x$ levels were enhanced in the vicinity of the oil spill due to spill response efforts (ship exhaust, surface oil burning, flaring of recovered gas).
-Airborne measurements of reactive nitrogen species ($NO_y$ = $NO_x$+$HNO_3$+$PAN$(peroxyacetul nitrate at the bottom)+$RONO_2$) and $VOCs$ in the marine boundary layer during daytime downwind from the DWH area showed that oxidation of alkanes with between 5 and 10 carbons per molecule ($C_5$–$C_{10}$) dominated $O_3$ formation chemistry in the plume.
-This circumstance led to a rapid and nearly complete oxidation of $NO_x$ to $RONO_2$ with mostly 6-10 carbons per molecule ($C_6$–$C_{10}$) effectively terminating $O_3$ production.
+Airborne measurements of reactive nitrogen species ($NO_y$ = $NO_x$+$HNO_3$+$PAN$\footnote{Peroxyacetyl nitrate.}+$RONO_2$) and VOCs in the marine boundary layer during daytime downwind from the DWH area showed that oxidation of alkanes with between 5 and 10 carbons per molecule ($C_5$–$C_{10}$) dominated $O_3$ formation chemistry in the plume.
+This circumstance led to a rapid and nearly complete oxidation of $NO_x$ to $RONO_2$ with mostly 6--10 carbons per molecule ($C_6$–$C_{10}$) effectively terminating $O_3$ production.
 On 10 June, $O_3$ enhancements in the DWH plumes were less than 15 ppbv, even after $NO_x$ was completely oxidized.
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \begin{figure}[h!]
@@
 \end{figure}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 These values are comparable with those measured in U.S.
-urban areas [Trainer et al., 1995; Neuman et al., 2009], but smaller than those measured in the petrochemical plumes with $OH$ loss rates dominated by alkenes (10-18 ppbv $O_3$ per ppbv of $NO_x$) [Ryerson et al., 2003].
+urban areas \textbf{[Trainer et al., 1995; Neuman et al., 2009] ADD THESE TO BIBTEX}, but smaller than those measured in the petrochemical plumes with $OH$ loss rates dominated by alkenes (10--18 ppbv $O_3$ per ppbv of $NO_x$) [Ryerson et al., 2003] AND THIS ONE!!!.
 
@@
 This result differs dramatically from previous studies, where the estimated contribution of $RONO_2$ to $NO_y$ was much lower: approximately 4\% in urban and petrochemical plumes [Neuman et al., 2002; Ryerson et al., 2003], approx.
-10\% in rural areas [Parrish et al., 1993], a few percent in urban air masses [O’Brien et al., 1997; Flocke et al.,1998; Schneider et al., 1998; Talbot et al., 2003; Ryerson et al., 2003], and up to 10–20\% of NOy in some studies [Day et al., 2003; Perring et al., 2010].
-
-Probably the most detailed investigation of the formation processes and photochemical evolution of alkyl nitrates (but not their impact on $O_3$) has been done by \citep{Sommariva2008}.
+10\% in rural areas [Parrish et al., 1993], a few percent in urban air masses [O’Brien et al., 1997; Flocke et al.,1998; Schneider et al., 1998; Talbot et al., 2003; Ryerson et al., 2003], and up to 10–20\% of NOy in some studies [Day et al., 2003; Perring et al., 2010] AND ALL OF THESE ABOVE!!!!!!!.
+
+Probably the most detailed investigation of the formation processes and photochemical evolution of alkyl nitrates (but not their impact on $O_3$) has been done by \cite{Sommariva2008}.
 Using a subset of the Master Chemical Mechanism v3.1 (MCM) they constructed a highly detailed chemical model and run it for 5 days under two scenarios (Table \ref{tab:Sommariva2008init}) to explore the sensitivity of the evolution of the chemical composition of an air mass emitted from an urban environment to the initial conditions.
 As it seen in Figure \ref{fig:Sommariva2008fig} according to the modelling results $O_3$ is produced and $NO_x$ is removed from the system with the consequent formation of $HNO_3$ via the reaction of $NO_2$ with $OH$, which broadly agrees with the observations [(Neuman et al., 2006).].
-In scenario A the concentration of $NO_x$ after the first day is about 0.5–1 ppb for $NO_2$ and approximately 100 ppt for $NO$, which is high enough to sustain the production of $O_3$ to up to 135 ppb and the formation of alkyl nitrates during the simulation.
+In scenario A the concentration of $NO_x$ after the first day is about 0.5–-1.0 ppb for $NO_2$ and approximately 100 ppt for $NO$, which is high enough to sustain the production of $O_3$ to up to 135 ppb and the formation of alkyl nitrates during the simulation.
 In scenario B (Figure \ref{fig:Sommariva2008fig} right) the increase in $O_3$ is slower (note that initial concentration is 0) and the maximum concentration after five days is about 100 ppb.
-The concentration of $NO_x$ remains relatively high during the second (10–15 ppb) and third day (3–5 ppb) and then decreases to $<$200 ppt of $NO$ and $<$1ppb of $NO_2$ on the fourth and fifth days, therefore allowing the formation of alkyl nitrates too.
+The concentration of $NO_x$ remains relatively high during the second (10–-15 ppb) and third day (3-–5 ppb) and then decreases to $<$200 ppt of $NO$ and $<$1ppb of $NO_2$ on the fourth and fifth days, therefore allowing the formation of alkyl nitrates too.
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-The first 3-D chemical transport  model simulations examining the impact of methyl and ethyl nitrates on tropospheric ozone chemistry in remote regions have been conducted by \citep{Neu2008}.
+The first 3-D chemical transport  model simulations examining the impact of methyl and ethyl nitrates on tropospheric ozone chemistry in remote regions have been conducted by \cite{Neu2008}.
 In their study, they incorporated methyl and ethyl nitrates in the chemical mechanism of the UCI CTM [Carver et al., 1997] and by matching their observed atmospheric abundances derived a total oceanic emission flux of these two nitrates (0.35 Tg of N per year).
@@
 A zero-dimensional (box) model is developed to investigate the impact of alkyl nitrate chemistry on ozone production efficiency.
-Its code is written using FACSIMILE for Windows.
-FACSIMILE is a special high-level programming language which is a powerful modelling tool designed to efficiently solve differential equations including those resulting from a set of chemical reactions.
-
-! \citep{Newland2013}
-[The model is written using the programming language FACSIMILE (Curtis and Sweetenham, 1987) and the differential equations are assembled and integrated using the Gear’s method (1971).
-The mixing ratios of the input species in the model are driven entirely by the emissions, the loss processes and the transport with no redistribution of the burden to ensure that mixing ratios agree with observations.]
+Its code is written using a new to the author program \href{http://www.mcpa-software.com/}{FACSIMILE for Windows}.
+FACSIMILE is a special high-level programming language which is a powerful modelling tool designed to solve differential equations including those resulting from a set of chemical reactions using the Gear’s method.
 
@@
 These include:
-
-! \citep{Newland2013}
-[The majority of the kinetic reaction rates and photochemical data used in the full chemistry model were taken from the IUPAC Subcommittee for Gas Kinetic Data Evaluation (http://www.iupac-kinetic.ch.cam.ac.uk; Atkinson et al., 2004; Atkinson et al., 2006; Atkinson et al., 2008).
-If the required data were not provided by IUPAC then data were taken from the NASA / JPL ‘Chemical Kinetics and Photochemical Data for Use in Atmospheric Studies: Evaluation No.
-17’ (Sander et al., 2011).
-Both these sources consider the full range of literature available on each reaction and provide a recommendation based on a synthesis of this data.
-There are sometimes minor differences between the recommendations of each organisation.]
-
 \begin{itemize}
@@
 When it is formed at lower temperatures (e.g., at high altitude) or transported into colder regions, PAN can act as an $NO_x$ reservoir.
-Observations of PAN and the sum of reactive nitrogen ($NO_y$) confirm it revealing that PAN can constitute up to 90\% of the total $NO_y$ budget at higher northern latitudes or higher altitudes (e.g., \citep{Jacobi2000}).
+Observations of PAN and the sum of reactive nitrogen ($NO_y$) confirm it revealing that PAN can constitute up to 90\% of the total $NO_y$ budget at higher northern latitudes or higher altitudes (e.g., \cite{Jacobi2000}).
 When an air mass containing PAN is transported into warmer regions, $NO_2$ tied up in PAN can be released back and cause the formation of $O_3$ and $HNO_3$ far away from the original $NO_x$ source.
@@
 
-\citep{Newland2013}
-[The alkanes and alkyl nitrates studied in this work were all explicitly described in the model with oxidation of the alkane giving a peroxy radical and reaction of the peroxy radical with NO giving an alkyl nitrate.]
-
 Lastly, the chemical mechanism of the model does not include halogen chemistry.
@@
 \subsection{Design of experiments}\label{sec:method_expdesign}
-To investigate the influence of alkyl nitrate chemistry on (potential changes in) tropospheric $O_3$ production, box model runs were performed at a series of different $NO_x$ and $VOCs$ initial concentrations, so that an $O_3$ isopleth plot could be constructed, similar to those found in \citep{Dodge1977} and \citep{Sillman1999}.
+To investigate the influence of alkyl nitrate chemistry on (potential changes in) tropospheric $O_3$ production, box model runs were performed at a series of different $NO_x$ and VOCs initial concentrations, so that an $O_3$ isopleth plot could be constructed, similar to those found in \citep{Dodge1977} and \citep{Sillman1999}.
 Table \ref{tab:setupO3CO} shows initial concentrations of 'independent'/global parameters and indicates which of them are fixed or allowed to vary during the model run.
-Table \ref{tab:setupNOx} and Table \ref{tab:setupVOCs} show the initial concentrations of $NO_x$ and $VOCs$ respectively that are used at a series of experiments.
-
-The lower limit of the range of $NO_x$ and $VOCs$ initial concentrations is based on observational data from Mauna Loa station (Hawaii, US) which is believed to be representative for conditions of remote troposphere.
+Table \ref{tab:setupNOx} and Table \ref{tab:setupVOCs} show the initial concentrations of $NO_x$ and VOCs respectively that are used at a series of experiments.
+
+The lower limit of the range of $NO_x$ and VOCs initial concentrations is based on observational data from Mauna Loa station (Hawaii, US) which is believed to be representative for conditions of remote troposphere.
 The upper limit is based on data from
@@
 For the first experiment the model was run under two configurations, (1) with all alkyl nitrate chemistry switched on, and (2) with all alkyl nitrate chemistry switched off.
-In every model run $VOCs$ concentrations were (changed according to the data from Table \ref{tab:setupVOCs} and) were fixed throughout the run, while the initial concentrations of $NO_x$ were (changed according to the data from Table \ref{tab:setupNOx} and were) allowed to vary throughout the run.
+In every model run VOCs concentrations were (changed according to the data from Table \ref{tab:setupVOCs} and) were fixed throughout the run, while the initial concentrations of $NO_x$ were (changed according to the data from Table \ref{tab:setupNOx} and were) allowed to vary throughout the run.
 As a result, the first experiment consists of 253 model runs including 11 runs (varying $NO_x$) with inorganic chemistry only.
 
-For the second experiment the model was run using the same approach concerning variations in $NO_x$ and $VOCs$ concentrations, but changes in alkyl nitrate chemistry were treated differently.
-In every model run reactions of individual alkyl nitrates or their groups (defined in Table \ref{tab:RHandANS}) were either switched on or switched off, resulting in other 847 additional model runs.
+For the second experiment the model was run using the same approach concerning variations in $NO_x$ and VOCs concentrations, but changes in alkyl nitrate chemistry were treated differently.
+In every model run reactions of individual alkyl nitrates or their groups (defined in Table \ref{tab:RHandANs}) were either switched on or switched off, resulting in other 847 additional model runs.
 
@@
 Due to imperfection of the chemical mechanism time evolution of species concentrations given by the model does not completely agree with the photochemical theory, but reproduces it to some extent.
-Figure \ref{fig:model_validation} shows an example of output from the model, namely the changes in mixing ratios of the main species within 24-hours of simulation time under contrasting $NO_x$ and $VOCs$ conditions.
+Figure \ref{fig:model_validation} shows an example of output from the model, namely the changes in mixing ratios of the main species within 24-hours of simulation time under contrasting $NO_x$ and VOCs conditions.
 The values shown in Figure \ref{fig:model_validation} have been justified from the theoretical point of view and also compared with those derived from the Atmospheric Chemistry Learner's Box Model (ACLBM, \href{http://guille.uea.ac.uk/ACLBM/}{http://guille.uea.ac.uk/ACLBM/}).
@@
 
-The concentration of $NO_x$ is the main factor which determines whether $O_3$ is produced or removed in the atmosphere, and theoretically there are three regimes of $O_3$ production that depend on $NO_x$ levels: at low $NO_x$ levels the regime is characterized by net $O_3$ removal, at intermediate $NO_x$ (also called $NO_x$-sensitive) - by net $O_3$ formation, and at high $NO_x$ - by net $O_3$ formation or removal depending more on $VOCs$ concentration (therefore called $VOC$-sensitive) (e.g., \citep{Fowler2008}).
+The concentration of $NO_x$ is the main factor which determines whether $O_3$ is produced or removed in the atmosphere, and theoretically there are three regimes of $O_3$ production that depend on $NO_x$ levels: at low $NO_x$ levels the regime is characterized by net $O_3$ removal, at intermediate $NO_x$ (also called $NO_x$-sensitive) - by net $O_3$ formation, and at high $NO_x$ - by net $O_3$ formation or removal depending more on VOCs concentration (therefore called $VOC$-sensitive) (e.g., \cite{Fowler2008}).
 When the prevailing $NO_x$ levels are low, a net $O_3$ removal takes place because there is not enough $NO$ around to react with the peroxy radicals ($HO_2$ and $RO_2$) to produce $NO_2$, which photolysis in conjunction with the reaction (\ref{reac:O3P+O2=O3}) is the main source of $O_3$ in the troposphere.
@@
 At low $NO_x$ regime (for example, 50 ppt) our modelling data agrees well with the theory (Figure \ref{fig:O3valid}).
-$O_3$ mixing ratio decreases with time, and this decrease is comparable to one derived from ACLBM (by 6 ppb and 4.5 ppb at low and high $VOCs$ concentrations respectively).
+$O_3$ mixing ratio decreases with time, and this decrease is comparable to one derived from ACLBM (by 6 ppb and 4.5 ppb at low and high VOCs concentrations respectively).
 $NO$ is rapidly (almost instantly) converted to $NO_2$ representing a state equilibrium between the two molecules in the troposphere (Figures \ref{fig:NOvalid}, \ref{fig:NO2valid}).
@@
 Under these circumstances, the number of free-radical propagated $O_3$ forming cycles decreases with increasing $NO_x$, and the production rate of $O_3$ decreases.
-However, if the concentrations of $VOCs$ are high, reactions of $OH$ with $CO$ and $RH$ ((\ref{reac:OH+CO=H+CO2}), (\ref{reac:OH+RH=RO2+H2O})) can compete with reaction (\ref{reac:OH+NO2=HNO3}), leading to an increase in the $O_3$ production rate.
+However, if the concentrations of VOCs are high, reactions of $OH$ with $CO$ and $RH$ ((\ref{reac:OH+CO=H+CO2}), (\ref{reac:OH+RH=RO2+H2O})) can compete with reaction (\ref{reac:OH+NO2=HNO3}), leading to an increase in the $O_3$ production rate.
 
@@
 At higher $NO_x$ regime (5 ppb) $O_3$ mixing ratio in the model increases with time (Figure \ref{fig:O3valid}) which agrees with the photochemical theory.
-The net increase is by 186 ppb and 171 ppb at low and high $VOCs$ concentrations respectively.
+The net increase is by 186 ppb and 171 ppb at low and high VOCs concentrations respectively.
 $NO$ is rapidly converted to $NO_2$, but slower than at low $NO_x$ levels because of their higher initial concentrations (Figure \ref{fig:NOvalid}, \ref{fig:NO2valid}).
@@
     \hfill
-    \begin{subfigure}[t]{0.45\linewidth}
-        \centering
-        \includegraphics[width=\linewidth]{D:/FACSIMILE/ANsCBmodel/matlab/ANsCB_pics/ANsNOxVOC/noAN/chem_noAN_netO3rate_withINORG_NOx_1ppb_100ppb.png}
+    \begin{subfigure}[t]{0.5\linewidth}
+        \centering
+        \includegraphics[width=\linewidth]{D:/FACSIMILE/ANsCBmodel/matlab/ANsCB_pics/ANsNOxVOC/noAN/chem_noAN_netO3rate_withINORG_NOx_1ppb_100ppb.eps}
         \caption{as a function of $VOC$ (ppbC) and $NO_x$ (ppb) in the first modelling experiment with alkyl nitrate chemistry switched off combined with an additional run with inorganic chemistry only (i.e.
@@
 
-As it has been noted in Section \ref{sec:intro}, the relationship between $O_3$ and $RONO_2$ production rates is controlled by the branching ratio ($\alpha_3$) between reactions:
+As it has been noted in Section \ref{sec:intro}, the relationship between $O_3$ and $RONO_2$ production rates is controlled by the branching ratio ($\alpha_3$) between two reactions:
 \begin{equation}
@@
 \end{equation}
-Reaction (\ref{reac:RO2+NO=aphaRO+NO2}) leads to $O_3$ production and $NO$ regeneration, resulting in continuing $O_3$ production, while reaction (\ref{reac:RO2+NO=aphaRONO2}) terminates this process.
+The reaction (\ref{reac:RO2+NO=aphaRO+NO2}) leads to $O_3$ production and $NO$ regeneration, resulting in continuing $O_3$ production, while reaction (\ref{reac:RO2+NO=aphaRONO2}) terminates this process.
 Branching ratios used in the model are taken from the MCM v3.3, and their values (Table \ref{tab:ANbranching}) were kept constant in all model runs.
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{table} %  AN branching ratios
+\begin{table}[h] %  AN branching ratios
 \caption{Branching ratios for the formation of alkyl nitrates from their precursor peroxy radicals and $NO$ (taken from the MCM v3.3 (2015)).}\label{tab:ANbranching}
@@
 
-Figure \ref{fig:netO3rate_noAN_withAN_diff} shows isopleths of net $O_3$ production rate as a function of $NO_x$ and sum of $VOCs$ initial concentrations from two series of model runs (with and without alkyl chemistry) and the difference between them.
-Since the actual formation of $O_3$ in our model starts at 250 ppt of $NO_x$, Figure \ref{fig:netO3rate_noAN_withAN_diff} summarises modelling data from experiments with $NO_x$ initial mixing ratios from 250 ppt till 100 ppb.
+Figure \ref{fig:netO3rate_noAN_withAN_diff} shows isopleths of net $O_3$ production rate as a function of $NO_x$ and sum of VOCs initial concentrations from two series of model runs (with and without alkyl chemistry) and the difference between them.
+Since the actual formation of $O_3$ in our model starts at 250 ppt of $NO_x$, Figure \ref{fig:netO3rate_noAN_withAN_diff} summarises the modelling data from experiments with $NO_x$ initial mixing ratios from 250 ppt to 100 ppb.
 
@@
 As expected, the inclusion of alkyl nitrate chemistry slows down the net $O_3$ production rate.
-At low $NO_x$ and $VOCs$ initial levels, this slowdown is -0.2...-0.6 ppb of $O_3$ per hour.
-As $NO_x$ and $VOCs$ levels increase the $O_3$ production rate decreases, reaching the maximum decrease of -1.94 ppb of $O_3$ per hour at 100 ppb of $NO_x$ and $VOCs$ level from case I (not the highest $VOCs$ level).
-At the range of $NO_x$ levels from 1 ppb to 10 ppb the difference in $O_3$ production rate is slightly more sensitive to $VOCs$ levels, while it is almost insensitive to $VOCs$ at $NO_x$ levels higher than 10 ppb and 2050 ppbC of $VOCs$.
+At low $NO_x$ and VOCs initial levels, this slowdown is -0.2...-0.6 ppb of $O_3$ per hour.
+As $NO_x$ and VOCs levels increase the $O_3$ production rate decreases, reaching the maximum decrease of -1.94 ppb of $O_3$ per hour at 100 ppb of $NO_x$ and VOCs level from case I (not the highest VOCs level).
+At the range of $NO_x$ levels from 1 ppb to 10 ppb the difference in $O_3$ production rate is slightly more sensitive to VOCs levels, while it is almost insensitive to VOCs at $NO_x$ levels higher than 10 ppb and 2050 ppbC of VOCs.
 
@@
 
-The change in net rate of $O_3$ production due to inclusion of alkyl nitrate chemistry into the chemical mechanism has a direct impact on absolute $O_3$ levels calculated by the model.
+The change in net rate of $O_3$ production due to inclusion of alkyl nitrate chemistry into the chemical mechanism has a direct impact on absolute $O_3$ levels.
 Figure \ref{fig:netO3mixrat_noAN_withAN_diff} shows that the net $O_3$ production (difference between final and initial $O_3$ mixing ratio) is smaller when alkyl nitrate chemistry is switched on.
-At low $NO_x$ and $VOCs$ levels the negative (!) change in the net $O_3$ production is 5-10 ppb, at high $NO_x$ and $VOCs$ levels it is 35-45 ppb (per 24 hours) with maximum value of 46.5 ppb.
-
-These changes in the net $O_3$ production should be treated with care, because the model due to reasons described in Section \ref{sec:method} gives unrealistically high $O_3$ mixing ratios.
-The maximum $O_3$ mixing ratio in the model is equal to 650 ppb (at the highest $NO_x$ level (100 ppb) and the lowest $VOCs$ level (case A)), while maximum typical urban (Los Angeles, Mexico City) $O_3$ mixing ratio is 490 ppb (\href{http://www-personal.umich.edu/~sillman/ozone.htm}{http://www-personal.umich.edu/~sillman/ozone.htm}).
-Therefore, a more useful estimate is a percentage difference in $O_3$ mixing ratios between two series of model runs shown in Figure \ref{fig:netO3rate_noAN_withAN_diff} DDDDD!.
-It appears that the largest absolute changes in $O_3$ mixing ratios (40-45 ppb) are changes by only 20-30\%, while changes up to 90\% take place the model is in low $NO_x$ and high $VOCs$ conditions.
-
-To see what caused these changes it is useful to look at the net production of the sum of the alkyl nitrates themselves as a function of $NO_x$ (250 ppt - 100 ppb) and $VOCs$ (Figure \ref{fig:endtotalANs}).
+At low $NO_x$ and VOCs levels the negative change in the net $O_3$ production is 5-10 ppb, at high $NO_x$ and VOCs levels it is 35--45 ppb (per 24 hours) with maximum value of 46.5 ppb.
+
+These changes in the net $O_3$ production should be treated with care, because the model gives unrealistically high $O_3$ mixing ratios due to the reasons described in section \ref{sec:method}.
+The maximum $O_3$ mixing ratio in the model is equal to 650 ppb (at the highest $NO_x$ level (100 ppb) and the lowest VOCs level (case A)), while typical urban (e.g., Los Angeles, Mexico City) $O_3$ mixing ratio reaches 490 ppb (\href{http://www-personal.umich.edu/~sillman/ozone.htm}{http://www-personal.umich.edu/~sillman/ozone.htm}).
+Therefore, a more useful estimate is a percentage difference in $O_3$ mixing ratios between two series of model runs shown in Figure \ref{fig:netO3rate_noAN_withAN_diff}.
+It appears that the largest absolute changes in $O_3$ mixing ratios (40--45 ppb) are changes by only 20--30\%, while changes up to 90\% take place the model is in low $NO_x$ and high VOCs conditions.
+
+To see what caused these changes it is useful to look at the net production of the sum of the alkyl nitrates themselves as a function of $NO_x$ (250 ppt -- 100 ppb) and VOCs (Figure \ref{fig:endtotalANs}).
 Since alkyl nitrates are the only sink for $NO_x$ in our model high production of these species was expected.
-The $NO_x$-$VOC$ field of total $RONO_2$ isopleths is mostly characterized by mixing ratios of ppb level (0.2-1.4 ppb), while the observations show that their mean mixing ratios lay at a level of tens-thousands of ppt (Table \ref{tab:ANmean}).
-Nevertheless, Figure \ref{fig:endtotalANs} will be of interest to other researches, because it is probably the first ever attempt to derive the distribution of $\Sigma RONO_2$ mixing ratios in $NO_x$-$VOC$ field.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+The $NO_x$-VOC space of total $RONO_2$ is mostly characterized by mixing ratios of ppb level (0.2--1.4 ppb), while the observations show that their mean mixing ratios lay at a level of tens to hundreds ppt (Table \ref{tab:ANmean}).
+Nevertheless, Figure \ref{fig:endtotalANs} will be of interest to other researchers, because to author's knowledge it is the first attempt to derive the distribution of $\Sigma RONO_2$ mixing ratios in $NO_x$-VOC space.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \begin{figure} % endtotalANs
 \centering
-\includegraphics[width=.7\linewidth]{D:/FACSIMILE/ANsCBmodel/matlab/ANsCB_pics/ANsNOxVOC/allAN/chem_allAN_endtotalANs_noINORG_NOx_250ppt_100ppb.eps}
-\caption{Isopleths of net production of alkyl nitrates ($\Sigma RONO_2$) as a function of $NO_x$ and $VOCs$ levels.}\label{fig:endtotalANs}
-\end{figure}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-!
-Dependence of alkyl nitrate yield on carbon number (paper about n-alkanes): bigger yield of ANs from larger carbon number means that the potential for contributing to photochemical air pollution may be less for the larger (C>6) n-alkanes that for the smaller ones.
+\includegraphics[width=.6\linewidth]{D:/FACSIMILE/ANsCBmodel/matlab/ANsCB_pics/ANsNOxVOC/allAN/chem_allAN_endtotalANs_noINORG_NOx_250ppt_100ppb.eps}
+\caption{Isopleths of net production of alkyl nitrates ($\Sigma RONO_2$) as a function of $NO_x$ and VOCs levels.}\label{fig:endtotalANs}
+\end{figure}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
-To estimate the impact of individual alkyl nitrates on $O_3$ production an additional series of experiments have been conducted.
+To estimate the impact of individual alkyl nitrates on $O_3$ production an additional series of experiments has been conducted.
 Prior to calculations some of the alkyl nitrates have been gathered into groups according to their parent alkanes as it shown in Table \ref{tab:RHandANs}.
-Setting initial $NO_x$ mixing ratio to 10 ppb and taking the middlemost $VOCs$ initial values (i.e., case F), the model has been run with certain alkyl nitrate chemistry switched on.
+Setting initial $NO_x$ mixing ratio to 10 ppb and taking the middlemost VOCs initial values (i.e., case F), the model has run with certain alkyl nitrate chemistry switched on.
 For example, to estimate the impact of $CH_3ONO_2$ the reactions involving this nitrate were switched on, while the other reactions with nitrates were switched off.
-Experiment with all alkyl nitrate chemistry is considered as a base run.
-
-To estimate the impact of certain alkyl nitrates on $O_3$ the difference between base run and other runs including one with all alkyl nitrate chemistry switched off has been calculated.
+The experiment with full alkyl nitrate chemistry is considered as a base run.
+
+The difference between base run and other runs (including one without alkyl nitrate chemistry) has been computed.
 Figure \ref{fig:timeO3diff10ppbNOx6VOC} shows the time evolution of this difference.
-This difference becomes more and more negative with time as the mixing ratio of alkyl nitrates themselves rises (Figure \ref{fig:timeANs10ppbNOx6VOC}).
-By the end of simulation the drop in $O_3$ mixing ratio due to the rise in mixing ratio of total alkyl nitrates is equal to -11.9 ppb (Table \ref{tab:timeO3diffANs10ppbNOx6VOC}.
-
-The contribution of different alkyl nitrates to drop in $O_3$ is not the same.
-By the end of simulation, the drop in $O_3$ mixing ratio due to ethyl nitrate only ($C_2H_5ONO_2$) is 11.6 ppb, and is almost the same drop is caused by methyl nitrate.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+The values become more negative with time while the mixing ratio of alkyl nitrates rises (Figure \ref{fig:timeANs10ppbNOx6VOC}).
+By the end of the simulation the drop in $O_3$ mixing ratio due to the rise in total alkyl nitrates concentration is equal to -11.9 ppb (Table \ref{tab:timeO3diffANs10ppbNOx6VOC}).
+%($\Delta O_3$ hereafter)
+
+The contribution of different alkyl nitrates to the $O_3$ decrease varies: the largest drop --- 11.6 ppb --- is caused by ethyl nitrate ($C_2H_5ONO_2$), and is almost the same for methyl nitrate.
+On the other hand, iso-pentyl nitrates account for the smallest $O_3$ decrease.
+
+The biggest change in $O_3$ (11.6 ppb) is related to ethyl nitrate, though its production by the end of the simulation is the smallest in comparison to other nitrates.
+Furthermore, when these is no conversion of methane to methyl nitrate, but conversion of ethane to ethyl nitrate is present even at a small initial value of ethane ($C_2H_6$ 3.8 ppb in comparison to 1915 ppb of methane) we observe almost the same change in $O_3$ mixing ratio.
+And surprisingly, in experiment with full alkyl nitrate chemistry we see almost the same $O_3$ change as in case of methyl or ethyl nitrate, but slightly higher.
+
+The smallest change in $O_3$ (by 8.0 ppb) is associated with production of iso-pentyl nitrates ($\Sigma iCH_5H_{11}ONO_2$), which final total mixing ratio in the highest in comparison to other nitrates.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \begin{figure}
@@
 Black line shows the difference between base experiment and an experiment with all alkyl nitrate chemistry switched off.
-$NO_x$ = 10 ppb, $VOCs$ = case F.}\label{fig:timeO3diff10ppbNOx6VOC}
-\end{figure}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+$NO_x$ = 10 ppb, VOCs = case F.}\label{fig:timeO3diff10ppbNOx6VOC}
+\end{figure}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \begin{figure}
@@
 \includegraphics[width=0.7\linewidth]{D:/FACSIMILE/ANsCBmodel/matlab/ANsCB_pics/ANsNOxVOC/allAN/chem_allAN_timeANs_10ppbNOx6VOC.eps}
-\caption{Time evolution of alkyl nitrates mixing ratios. $NO_x$ = 10 ppb, $VOCs$ = case F.}\label{fig:timeANs10ppbNOx6VOC}
-\end{figure}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\caption{Time evolution of alkyl nitrates mixing ratios. $NO_x$ = 10 ppb, VOCs = case F.}\label{fig:timeANs10ppbNOx6VOC}
+\end{figure}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \begin{table}
@@
 $CH_3ONO_2$              & 43  & -11.5 \\
-$C_2H_5ONO_2$            & 38  & -11.6 \\
+$C_2H_5ONO_2$            & \textbf{38}  & \textbf{-11.6} \\
 $\Sigma C_3H7ONO_2$      & 115 & -10.5 \\
@@
 $\Sigma nC_5H_{11}ONO_2$ & 108 & -10.1 \\
-$\Sigma iC_5H_{11}ONO_2$ & 290 & -8.0  \\
-$\Sigma RONO_2$          & 867 & -11.9 \\
+$\Sigma iC_5H_{11}ONO_2$ & \textbf{290} & \textbf{-8.0}  \\
+\textbf{$\Sigma RONO_2$} & \textbf{851} & \textbf{-11.9} \\
 \hline
@@
 \end{table}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
-As it has been noted in section \ref{sec:intro_CB}, there is a way to estimate the $O_3$ production per carbon bond emitted into the system and relate it to the alkyl nitrate chemistry.
-Such calculations have been first conducted by \citep{Evans2014} for the global 3D chemistry-climate model GEOS-Chem \citep{Bey2001}.
-They estimated that $\epsilon$ value for the GEOS-Chem model in simulation with full chemistry is equal to 0.2 meaning that to produce 1 molecule of $O_3$ 0.2 carbon bonds should have been emitted.
-They also estimated that in simulation with $CH_4$ only $\epsilon$ is equal to 0.3, meaning that more carbon bonds should have been emitted to produce the same amount of $O_3$.
-
-We tried to calculate the same value for our model, however since it does not include any emissions, the $\epsilon$'s denominator has been treated a little bit differently.
-To be exact the term $E_{C_{bons}}$ is calculated as a rate of change of the difference between final and initial sum of carbon bonds ($\Sigma Cbonds$) in the system at constant sum of carbon bonds contained in all of the alkanes and $CO$ ($\Sigma Cbonds_{alkanes, CO}$):
+As it has been noted in section \ref{sec:intro_CB}, a novel method was recently proposed to estimate the $O_3$ production per carbon bond emitted into the system.
+Applying this method to the present study, we try to associate ozone production efficiency to the alkyl nitrate chemistry.
+Such calculations have been first conducted by \cite{Evans2014} for the global 3D chemistry-climate model GEOS-Chem \citep{Bey2001}.
+They estimated that $\epsilon$ (see eq.
+\ref{eq:epsilon}) for the GEOS-Chem model in simulation with full chemistry is equal to 0.2 meaning that a production of 1 molecule of $O_3$ requires 0.2 carbon bonds to be emitted.
+In the mentioned study authors also stated that in the simulation with $CH_4$ only $\epsilon$ equals to 0.3, which implies that more carbon bonds need to be emitted to produce the same amount of $O_3$.
+
+In the present study an attempt was made to get the same value, however since the model does not include any emissions, the denominator in the equation \ref{eq:epsilon} has been treated differently.
+Namely, the term $E_{C_{bons}}$ is calculated as a rate of change of the difference between final and initial sum of carbon bonds ($\Sigma Cbonds$) in the system at a constant sum of carbon bonds contained in all of the alkanes and $CO$ ($\Sigma Cbonds_{alkanes, CO}$):
 \begin{matriz}\label{eq:ECbonds}
@@
 In other words, we calculate an increment in the amount of carbon bonds contained in all organic species except for alkanes and $CO$ (e.g., in $ROOH$, $RONO_2$) in the system by the end of simulation and divide it by simulation time (24 hours).
-Despite that we somewhat excluded the impact of alkanes and $CO$ on $O_3$ production efficiency, we have managed to save an opportunity to take into account the impact of alkyl nitrates, because the carbon bonds contained in them are still in this formula.
-
-Figure \ref{fig:epsilon250ppt100ppb} shows $\epsilon$ values calculated from our model as a function of $NO_x$ (250 ppt - 100 ppb) and $VOCs$ mixing ratios.
-It appears that in both cases with and without alkyl nitrates $\epsilon$ is insensitive to $NO_x$ in the range of 1-100 ppb of $NO_x$, but quickly drops within this $NO_x$ range with increasing $VOCs$.
-At the rest of $\epsilon$ isopleth field it is equally sensitive to $NO_x$ and $VOCs$.
-The difference in $\epsilon$ between two series of model runs is small.
-The average $\epsilon$ value in model runs without alkyl nitrates and $NO_x$ of 250 ppt - 100 ppb is equal to 0.44, in case with them it is equal to 0.41, which means that 2.3 carbon bonds should have been emitted to produce 1 molecule of $O_3$ in case without alkyl nitrates, while in case with alkyl nitrates 2.4 carbon bonds should have been emitted to produce the same amount of $O_3$.
-
-If we construct the same kind of plot for all $NO_x$ conditions considered in the model, from 5 ppt to 100 ppb, we find that $\epsilon$ could be negative (Figure \ref{fig:epsilon5ppt100ppb}), which could be explained by a net $O_3$ removal at $NO_x$ levels lower 250 ppt in our model.
-The lower $VOCs$ level is more negative $epsilon$ becomes.
-The average $\epsilon$ value in all model runs without alkyl nitrates is equal to 0.25, in case with them it is equal to 0.22, meaning that 4 carbon bonds should have been emitted to produce 1 molecule of $O_3$ in case without alkyl nitrates, while in case with alkyl nitrates 4.5 carbon bonds should have been emitted to produce the same amount of $O_3$.
-The difference between $\epsilon$ values between two series of model runs is small.
+Despite that we somewhat excluded the impact of alkanes and $CO$ on $O_3$ production efficiency, we have managed to take into account the impact of alkyl nitrates, because the carbon bonds contained in them are still in this formula.
+
+Figure \ref{fig:epsilon250ppt100ppb} shows $\epsilon$ values calculated from our model as a function of $NO_x$ (250 ppt -- 100 ppb) and VOCs mixing ratios.
+It appears that in both cases with and without alkyl nitrates $\epsilon$ is insensitive to $NO_x$ in the range of 1-100 ppb of $NO_x$, but quickly drops within this $NO_x$ range with increasing VOCs.
+The rest of $\epsilon$ isopleths plot is equally sensitive to $NO_x$ and VOCs.
+The difference in $\epsilon$ between two series of model runs is almost negligible.
+The average $\epsilon$ value in the model runs without alkyl nitrates and $NO_x$ of 250 ppt -- 100 ppb is equal to 0.44, while in the other case it is equal to 0.41, which means that 2.3 carbon bonds should have been emitted to produce 1 molecule of $O_3$ without alkyl nitrates, while in presence of alkyl nitrates the number of carbon bonds is 2.4.
+
+If we construct the same kind of plot for all $NO_x$ conditions considered in the model --- from 5 ppt to 100 ppb --- we find that $\epsilon$ could be negative (Figure \ref{fig:epsilon5ppt100ppb}), which could be explained by a net $O_3$ removal at $NO_x$ levels lower 250 ppt in our model.
+The lower VOCs level is the more negative $\epsilon$ becomes.
+The average $\epsilon$ value in all model runs with (without) alkyl nitrates is equal to 0.22 (0.25).
+In simpler terms, 4.5 (4) carbon bonds need to be emitted to produce a molecule of $O_3$ with (without) alkyl nitrates.
+The difference between $\epsilon$ values between the two series of model runs is relatively small too.
 
@@
 \end{figure}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \begin{figure} % epsilon noANs allANs diff 5 ppt 100 ppb
@@
 \end{figure}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Discussion} \label{sec:discuss}
-The experimental results presented in the previous section provide an opportunity to speculate about the behaviour of our simple chemistry box conceptual model and its sensitivity to alkyl nitrate production and loss.
-
-This research has limitations link with the disadvantages of the model, not enough experiments.
-
-Due to obvious limitations
+This purely theoretical study continues our efforts to understand the tropospheric ozone chemistry, in particular focusing on the impact of alkyl nitrates on $O_3$ production.
+However, it has many limitations.
+
+First of all, the chemical mechanism of the model is not sufficient enough to represent full complexity of ozone tropospheric chemistry.
+It does not have all possible sinks for $NO_x$, such as the nitric and nitrous acid, dinitrogen pentoxide and PAN.
+In addition, it does not account for dry and wet deposition of any species.
+
+Secondly, a greater number o experiments needs to be carried out to explore in depth the sensitivity of the model to individual alkyl nitrates.
+It could be done by extending the range of $NO_x$ and VOC conditions for every individual nitrate.
+
+Thirdly, the model was run with fixed meteorological parameters, namely temperature and relative humidity.
+
+Forthly, it could have been better to run the model to a steady state, however since our model does not have a full oxidation of the organics, it could have caused bigger errors that otherwise.
+
+Nevertheless, the model appeared to be sensitive to alkyl nitrate chemistry because $O_3$ yields
+
+add night chemistry
+photolysis variation during the day
+compare with observations
 
 Future plans include illumination of all weaknesses from our model, particularly adding other sinks for $NO_x$ into the chemical mechanism.
+preliminary study
 
@@
 \section{Conclusion} \label{sec:conclusion}
-The present study provides an assessment of of
-
-The present study is purely theoretical, investigates the impact of alkyl nitrates on tropospheric ozone production.
-We examine the sensitivity of the modelled ozone production to the alkyl nitrate chemistry.
-
-The research hopes to contribute to the atmospheric chemistry research b
+The present study provides an preliminary assessment of the sensitivity of tropospheric ozone production efficiency to alkyl nitrate chemistry using conceptual box modelling.
+It was found that the model is capable to capture some features of daytime tropospheric ozone chemistry, but due to various limitations does do it very well.
+
+Inclusion of $RONO_2$ chemistry slows down the net $O_3$ production rate leading to a smaller overall $O_3$ formation by the end of simulation.
+
+The distribution of the sum of $RONO_2$ considered in the model in $NO_x$--VOC space is derived.
+It shows that the total concentration of alkyl nitrates rises with rising $NO_x$ and VOC conditions, it is more sensitive to $NO_x$ at high VOC levels and more sensitive to VOC at high $NO_x$ levels.
 

### 2015-08-06 MSc_diss.tex
Updates in model description

@@
 They found that in the immediate vicinity of savanna fires in Northern Australia the average alkyl nitrate mixing ratios were 47-–122 times higher than local background mixing ratios, exceeding 3 ppbv for methyl nitrate.
-Methyl nitrate was primarily emitted during the 'flaming' stage, while $C_2$-$C_4$ ethyl-butyl nitrates dominated during the 'smoldering' stage.
+Methyl nitrate was primarily emitted during the 'flaming' stage, while $C_2$-$C_4$ ethyl--butyl nitrates dominated during the 'smoldering' stage.
 
@@
 \begin{table}[h!]
-\caption{!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!}\label{tab:ANmean} % mean AN mixing ratios
+\caption{Mean mixing ratios of the sum of alkyl nitrates ($\Sigma RONO_2$) measured in different environments.
+The measurement area is indicated in parentheses, type of measurements is in brackets.}\label{tab:ANmean} % mean AN mixing ratios
 \begin{tabularx}{\linewidth}{P{3cm}P{2cm}P{5cm}X}
@@
 \hline
-remote marine (North Atlantic), airborne   & 10 pptv    & methyl, ethyl, 1-propyl, 2-propyl, 2-butyl, 2-methyl-3-butyl, sum of 2- and 3-pentyl  & \citep{Reeves2007} \\
-\hline
-rural ground (Nova Scotia, Canada)        & 14 pptv    & methyl, ethyl, n-propyl, 2-propyl, 2-butyl  & \citep{Roberts1998} \\
-\hline
-suburban ground (Cape Retire, South Africa)  & 17.48 pptv (3.62-122.23) & 2-propyl, 1-propyl, 2-butyl, 1-butyl, 3-pentyl, 2-pentyl, 2-methyl-1-butyl, i-pentyl & \citep{DeKock1994} \\
-\hline
-suburban ground (Black Forest, Germany)    & 119.8 ppt (30-630 ppt) & $C_1-C_8$ & \citep{Flocke1998} \\
-\hline
-polluted marine airborne (Gulf of Mexico) & approx 1 ppb (plume center)& $NO_y-(NO_x+HNO_3+PAN)$ & \citep{Neuman2012} \\
+remote marine (North Atlantic) [airborne]   & 10 pptv    & methyl, ethyl, 1-propyl, 2-propyl, 2-butyl, 2-methyl-3-butyl, sum of 2- and 3-pentyl  & \citep{Reeves2007} \\
+\hline
+rural (Nova Scotia, Canada) [ground]        & 14 pptv    & methyl, ethyl, n-propyl, 2-propyl, 2-butyl  & \citep{Roberts1998} \\
+\hline
+suburban (Cape Retire, South Africa) [ground] & 17.48 pptv (3.62-122.23) & 2-propyl, 1-propyl, 2-butyl, 1-butyl, 3-pentyl, 2-pentyl, 2-methyl-1-butyl, i-pentyl & \citep{DeKock1994} \\
+\hline
+suburban (Black Forest, Germany) [ground]   & 119.8 ppt (30-630 ppt) & $C_1-C_8$ & \citep{Flocke1998} \\
+\hline
+polluted marine (Gulf of Mexico) [airborne] & approx 1 ppb (plume center)& $NO_y-(NO_x+HNO_3+PAN)$ & \citep{Neuman2012} \\
 \hline
@@
 
-A study that enabled us to focus on examination of $O_3$ and $RONO_2$ formation processes has been conducted by \cite{Neuman2012}.
+A study that enabled us to re-examine the theory of $O_3$ and $RONO_2$ formation processes in a quite unique environment has been conducted by \cite{Neuman2012}.
 After the Deepwater Horizon (DWH) offshore oil platform accident on 20 April 2010 large amounts of oil from the DWH spill reached the ocean surface.
-Subsequent oil evaporation caused a release of little methane and other light alkanes to the atmosphere forming a $VOC$ mixture atypical for the remote marine environment.
+Subsequent oil evaporation caused a release of methane and other light alkanes to the atmosphere forming a $VOC$ mixture atypical for the remote marine environment.
 At the same the $NO_x$ levels were enhanced in the vicinity of the oil spill due to spill response efforts (ship exhaust, surface oil burning, flaring of recovered gas).
-Airborne measurements of reactive nitrogen species ($NO_y$ = $NO_x$+$HNO_3$+$PAN$\footnote{Peroxyacetyl nitrate.}+$RONO_2$) and VOCs in the marine boundary layer during daytime downwind from the DWH area showed that oxidation of alkanes with between 5 and 10 carbons per molecule ($C_5$–$C_{10}$) dominated $O_3$ formation chemistry in the plume.
-This circumstance led to a rapid and nearly complete oxidation of $NO_x$ to $RONO_2$ with mostly 6--10 carbons per molecule ($C_6$–$C_{10}$) effectively terminating $O_3$ production.
-On 10 June, $O_3$ enhancements in the DWH plumes were less than 15 ppbv, even after $NO_x$ was completely oxidized.
+Daytime airborne measurements of reactive nitrogen species ($NO_y$ = $NO_x$+$HNO_3$+$PAN$\footnote{Peroxyacetyl nitrate.}+$RONO_2$) and VOCs in the marine boundary layer downwind from the DWH area showed that oxidation of alkanes with between 5 and 10 carbons per molecule ($C_5$–$C_{10}$) dominated ozone formation chemistry in the plume.
+This circumstance led to a rapid and nearly complete oxidation of $NO_x$ to $RONO_2$ with mostly 6--10 carbons per molecule ($C_6$–$C_{10}$) effectively terminating ozone production.
+As a result, on 10 June, enhancements ozone levels in the DWH plumes were less than 15 ppbv, even after $NO_x$ was completely oxidized.
 
@@
 $C_1$–$C_5$ alkyl nitrates were measured in whole air samples at locations indicated by the circles.
-The sum of alkyl nitrates (black lines) is inferred from the difference between $NO_y$ and ($NO_x$+$HNO_3$+$PAN$) \citep{Neuman2012}.}\label{fig:Neuman2012}
-\end{figure}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-$O_3$ production efficiencies in the DWH plume were 6 plus minus 0.5 ppbv $O_3$ per ppbv of $NO_x$ oxidized after 1 h of transport.
+The sum of alkyl nitrates (black lines) is inferred from the difference between $NO_y$ and ($NO_x$+$HNO_3$+$PAN$).
+Source: \citep{Neuman2012}.}\label{fig:Neuman2012}
+\end{figure}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+Estimated $O_3$ production efficiency in the DWH plume is equal to 6$\pm$0.5 ppbv of $O_3$ per ppbv of $NO_x$ oxidized after 1 h of transport.
 These values are comparable with those measured in U.S.
-urban areas \textbf{[Trainer et al., 1995; Neuman et al., 2009] ADD THESE TO BIBTEX}, but smaller than those measured in the petrochemical plumes with $OH$ loss rates dominated by alkenes (10--18 ppbv $O_3$ per ppbv of $NO_x$) [Ryerson et al., 2003] AND THIS ONE!!!.
-
-Temporal changes in reactive nitrogen partitioning in the DWH plume (Figure \ref{fig:Neuman2012}) nicely illustrate that in fresh part of the plume (Figure \ref{fig:Neuman2012}a) $NO_x$ accounted for over 80\% of plume $NO_y$, in aged plumes (Figure \ref{fig:Neuman2012}b), $NO_x$ oxidation was nearly complete, but NOx, PAN, HNO3, and the sum of eight $C_5$–$C_{10}$ $RONO_2$ accounted for only 22\% of plume $NO_y$.
-It means that $C_6$–$C_{10}$ $RONO_2$ account for 78\% of the aged plume $NO_y$ \citep{Neuman2012}.
-This result differs dramatically from previous studies, where the estimated contribution of $RONO_2$ to $NO_y$ was much lower: approximately 4\% in urban and petrochemical plumes [Neuman et al., 2002; Ryerson et al., 2003], approx.
-10\% in rural areas [Parrish et al., 1993], a few percent in urban air masses [O’Brien et al., 1997; Flocke et al.,1998; Schneider et al., 1998; Talbot et al., 2003; Ryerson et al., 2003], and up to 10–20\% of NOy in some studies [Day et al., 2003; Perring et al., 2010] AND ALL OF THESE ABOVE!!!!!!!.
-
-Probably the most detailed investigation of the formation processes and photochemical evolution of alkyl nitrates (but not their impact on $O_3$) has been done by \cite{Sommariva2008}.
+urban areas \citep{Trainer1995,Neuman2009}, but smaller than those measured in the petrochemical plumes with $OH$ loss rates dominated by alkenes (10--18 ppbv $O_3$ per ppbv of $NO_x$) \citep{Ryerson2003}.
+
+Temporal changes in reactive nitrogen partitioning in the DWH plume (Figure \ref{fig:Neuman2012}) nicely illustrate that a quite large concentration of alkyl nitrates in the air can be build up under certain conditions.
+While in fresh part of the plume (Figure \ref{fig:Neuman2012}a) $NO_x$ accounted for over 80\% of plume $NO_y$, in aged plumes (Figure \ref{fig:Neuman2012}b) $NO_x$ oxidation was nearly complete, but $NO_x$, $PAN$, $HNO_3$, and the sum of eight $C_5$–$C_{10}$ $RONO_2$ accounted for only 22\% of plume $NO_y$.
+It means that $C_6$–$C_{10}$ $RONO_2$ accounted for 78\% of the aged plume $NO_y$ \citep{Neuman2012}.
+This result differs dramatically from other studies, where the estimated contribution of $RONO_2$ to $NO_y$ was much lower: approximately 4\% in urban and petrochemical plumes \citep{Ryerson2003}, 10\% in rural areas \citep{Parrish1993}, a few percent in urban air masses \citep{Ryerson2003}, and up to 10–20\% of $NO_y$ in some studies \citep{Day2003}.
+
+Probably the most detailed investigation of the formation processes and photochemical evolution of alkyl nitrates (but not their impact on ozone) has been done by \cite{Sommariva2008}.
 Using a subset of the Master Chemical Mechanism v3.1 (MCM) they constructed a highly detailed chemical model and run it for 5 days under two scenarios (Table \ref{tab:Sommariva2008init}) to explore the sensitivity of the evolution of the chemical composition of an air mass emitted from an urban environment to the initial conditions.
-As it seen in Figure \ref{fig:Sommariva2008fig} according to the modelling results $O_3$ is produced and $NO_x$ is removed from the system with the consequent formation of $HNO_3$ via the reaction of $NO_2$ with $OH$, which broadly agrees with the observations [(Neuman et al., 2006).].
-In scenario A the concentration of $NO_x$ after the first day is about 0.5–-1.0 ppb for $NO_2$ and approximately 100 ppt for $NO$, which is high enough to sustain the production of $O_3$ to up to 135 ppb and the formation of alkyl nitrates during the simulation.
-In scenario B (Figure \ref{fig:Sommariva2008fig} right) the increase in $O_3$ is slower (note that initial concentration is 0) and the maximum concentration after five days is about 100 ppb.
+As it seen in Figure \ref{fig:Sommariva2008fig}, the modelling shows that $O_3$ is produced and $NO_x$ is removed from the air mass with the consequent formation of $HNO_3$.
+In scenario A (Figure \ref{fig:Sommariva2008fig} left) the concentration of $NO_x$ after the first day is about 0.5–-1.0 ppb for $NO_2$ and approximately 100 ppt for $NO$, which is high enough to sustain the production of $O_3$ to up to 135 ppb and the formation of alkyl nitrates during the simulation.
+In scenario B (Figure \ref{fig:Sommariva2008fig} right) the increase in $O_3$ is slower (note that initial concentration is zero) and the maximum concentration after five days is about 100 ppb.
 The concentration of $NO_x$ remains relatively high during the second (10–-15 ppb) and third day (3-–5 ppb) and then decreases to $<$200 ppt of $NO$ and $<$1ppb of $NO_2$ on the fourth and fifth days, therefore allowing the formation of alkyl nitrates too.
@@
 \begin{table}
-\caption{Initial values of the some model parameters (in ppb) for two scenarios used in \citep{Sommariva2008}}\label{tab:Sommariva2008init}
+\caption{Initial values of some model parameters (in ppb) for two scenarios used in \citep{Sommariva2008}}\label{tab:Sommariva2008init}
 \centering
@@
 \hline
-Parameter & Initial value Scenario A & Scenario B \\
+Parameter & Scenario A & Scenario B \\
 \hline
@@
 \end{figure}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-Comparison of modelling results with the values observed inside a typical urban plume sampled using the NOAA WP-3D aircraft from the New York City area during the New England Air Quality Study (NEAQS) 2004 campaign enabled \citep{Sommariva2008} to conclude that the chemistry of alkyl nitrates is well represented in the MCM model.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+Comparison of modelling results with the values observed inside a urban plume sampled using the NOAA WP-3D aircraft from the New York City area during the New England Air Quality Study (NEAQS) 2004 campaign enabled \cite{Sommariva2008} to conclude that the chemistry of alkyl nitrates is well represented in the MCM model.
 They found that short-chain $<C_4$ alkyl nitrates can be formed from several precursors, while long-chain ($C_5$) alkyl nitrates are produced for 90\% or more from the oxidation of a single parent alkane.
@@
 The first 3-D chemical transport  model simulations examining the impact of methyl and ethyl nitrates on tropospheric ozone chemistry in remote regions have been conducted by \cite{Neu2008}.
-In their study, they incorporated methyl and ethyl nitrates in the chemical mechanism of the UCI CTM [Carver et al., 1997] and by matching their observed atmospheric abundances derived a total oceanic emission flux of these two nitrates (0.35 Tg of N per year).
+In their study, they incorporated methyl and ethyl nitrates in the chemical mechanism of the UCI CTM model \citep{Carver1997}, and by matching their observed atmospheric abundances derived a total oceanic emission flux of these two nitrates (0.35 Tg of N per year).
 Comparing two simulations with and without spatially uniform emissions of methyl and ethyl nitrates they found that the largest impact of oceanic alkyl nitrates is seen over the equatorial Western Pacific (Figure \ref{fig:Neu2008}).
 Atmospheric circulation (convergence of the trade winds) in this region isolates the marine boundary layer, resulting in sustained photochemical loss of $O_3$ (Figure \ref{fig:Neu2008M_O3contol}) and at the same time accumulation of alkyl nitrates (Figure \ref{fig:Neu2008M_MeONO2}), which drive up boundary layer $NO_x$ abundances by as much as 250\% (Figure \ref{fig:Neu2008M_NOxdiff}).
-Such combination of factors leads to increases in $O_3$ of up to 20\% (Figures \ref{fig:Neu2008M_O3contol} and \ref{fig:Neu2008M_O3diff}), with the largest increases occurring from May to October.
+Such combination of factors leads to increases in $O_3$ of up to 20\% (Figure \ref{fig:Neu2008M_O3diff}), with the largest increases occurring from May to October.
 
@@
 Since these compounds are removed from the troposphere by photolysis and OH oxidation they can provide information on photochemical evolution of an air mass through relationships with their parent hydrocarbons \citep{Bertman1995}.
-Acting as reservoir species for reactive nitrogen (NOy) and having a relatively long atmospheric lifetimes alkyl nitrates play an important role in long range transport of NOx.
-Furthermore, alkyl nitrates link to ozone and carbonyl production making them useful tracers for both species \citep{Worton2010}.
-Despite their relative importance alkyl nitrates are usually not included in the climate-chemistry models, for example in UKCA (Reeves, personal communication), which potentially could lead to discrepancies between models and observations described before.
+Acting as reservoir species for reactive nitrogen and having relatively long atmospheric lifetimes alkyl nitrates play an important role in long range transport of NOx.
+Despite their importance and quite large build up under certain conditions alkyl nitrates are usually not included in climate-chemistry models, for example in UKCA (Reeves, personal communication), which potentially could lead to discrepancies between models and observations mentioned before.
 
@@
 \subsection{Model development and description}\label{sec:method_develop}
-A zero-dimensional (box) model is developed to investigate the impact of alkyl nitrate chemistry on ozone production efficiency.
+A zero-dimensional (box) model is developed to investigate the impact of alkyl nitrate chemistry on ozone production.
 Its code is written using a new to the author program \href{http://www.mcpa-software.com/}{FACSIMILE for Windows}.
@@
 
-The model is intended to represent the steady state of the troposphere under mean summer daytime meteorology and clear skies.
+The model is intended to represent the global troposphere under mean summer daytime meteorology and clear skies.
 To achieve that a number of assumptions have been made.
@@
 \item the loss of ozone through its photolysis and reactions with $O(^3P)$, $OH$, $HO_2$ and $NO$, among which the last is of particular interest because it gives back $NO_2$;
-\item the formation of alkylperoxy radicals ($RO_2$) from oxidation of the hydrocarbons;
+\item the formation of alkyl peroxy radicals ($RO_2$) from oxidation of the hydrocarbons;
 \item the competition of $NO$ and $HO_2$ for reaction with $HO_2$ and $RO_2$;
@@
 \end{itemize}
-The full list of reactions can be found in Appendix \ref{sec:appendix1}.
-It is important to mention that due to purely theoretical purpose of this research and its focus on investigation of the impact of alkyl nitrate chemistry on ozone production some of the reactions that lead to formation of species that serve as a sink for $NO_x$ are not included into the chemical mechanism.
+The full list of reactions incorporated into the code of the model can be found in Appendix \ref{sec:appendix1} and also on this on-line repository: \href{https://github.com/justmeteomary4/ENV-MB4X_model}{\textbf{link}}.
+It is important to mention that due to purely theoretical purpose of this research and its focus on alkyl nitrate chemistry and ozone some of the reactions that lead to formation of species that serve as a sink for $NO_x$ are not included into the chemical mechanism.
 However, these reactions play an important role in the chemistry of the troposphere, and below here is a description of their role and anticipated effect of their exclusion.
@@
 Being a 'sticky' molecule $HNO_3$ readily adsorbs to surfaces, particularly if there is water on the surface, and deposits either in precipitation (wet deposition) or in dry form (dry deposition).
-Because of the high deposition velocity, dry deposition of $HNO_3$ can be responsible for much of the removal of inorganic nitrogen from the troposphere \citep{Finlayson-Pitts2000}, which means that exclusion of $HNO_3$ (formation and removal)/(chemistry) from the chemical mechanism potentially leads to overproduction of ozone in the model, especially at high $NO_2$ concentrations because the competition between $NO_2$ photolysis and its reaction with $OH$ radical ($NO_2 + OH \xrightarrow{M} HNO_3$) is ignored.
-\item Since this research is focused on the daytime chemistry some reactions that are believed to be significant in nighttime are not included into the chemical mechanism of the model.
+Because of the high deposition velocity, dry deposition of $HNO_3$ can be responsible for much of the removal of inorganic nitrogen from the troposphere \citep{Finlayson-Pitts2000}, which means that exclusion of $HNO_3$ chemistry from the chemical mechanism potentially leads to overproduction of ozone in the model, especially at high $NO_2$ concentrations because the competition between $NO_2$ photolysis and its reaction with $OH$ radical ($NO_2 + OH \xrightarrow{M} HNO_3$) is ignored.
+\item Since this research is focused on the daytime chemistry some reactions that are believed to be significant in nighttime are not included into the chemical mechanism.
 Among them are reactions that involve the nitrate radical ($NO_3$) and dinitrogen pentoxide ($N_2O_5$).
@@
 This increase in concentration and a lesser availability of the OH radical (which sources are mostly photolytic in nature) make $NO_3$ radical a major contributor to the night chemistry of organics in the troposphere \citep{Finlayson-Pitts2000}.
-However, exclusion of $NO_3$ radical reactions from our chemical mechanism should not have a huge impact on the results anyway since the mechanism considers mainly decomposition of alkanes with which $NO_3$ radical reacts to a generally minor extent \citep{Atkinson2008}???.
-\item ! the nighttime NO3 radical reaction is of minor significance in term of the overall removal of the alkanes (Atkinson, 1990a)
+However, exclusion of $NO_3$ radical reactions from our chemical mechanism should not have had a huge impact on the results anyway since the mechanism considers mainly decomposition of alkanes with which $NO_3$ radical reacts to a generally minor extent \citep{Atkinson2008}.
 \item In addition to reacting with organics, $NO_3$ also reacts with $NO_2$, forming dinitrogen pentoxide ($NO_3 + NO_2 \xleftrightarrow{M} N_2O_5$) which in turn is an important nighttime source of nitric acid through its rapid hydrolysis on wet surfaces and aerosol particles ($N_2O_5 + H_2O \xrightarrow{surf., aerosol} 2HNO_3$) \citep{Finlayson-Pitts2000}.
-Exclusion of $N_2O_5$ from the chemical mechanism assumes no cycling of nitrogen oxides within the troposphere through $NO_3$ and $N_2O_5$, eventually forming $HNO_3$, which are not included into the mechanism anyway.
-\item ! In any case, because of this equilibrium, sink for N2O5 such as hydrolysis are, in essence, also sinks for NO3 as well.
+Exclusion of $N_2O_5$ from the chemical mechanism assumes no cycling of nitrogen oxides within the troposphere through $NO_3$ and $N_2O_5$, eventually forming $HNO_3$, which is not included into the mechanism anyway.
 \end{itemize}
@@
 However, since $HONO$ strongly absorbs UV radiation, it rapidly photolyzes, and under particular conditions (when the nighttime $HONO$ concentration reaches almost 15 ppb \citep{Finlayson-Pitts2000}), photolysis of nighttime-generated $HONO$ can become a major source of $OH$ radicals in the early morning hours ($HONO + hv \rightarrow OH + NO$) \citep{Lammel1996}.
-Therefore exclusion of $HONO$ from the chemical mechanism potentially leads to underestimation of $OH$ concentrations, but having neither formation nor removal of $HONO$ should balance this effect out???.
+Therefore exclusion of $HONO$ from the chemical mechanism potentially leads to underestimation of $OH$ concentrations, but having neither formation nor removal of $HONO$ should balance this effect out.
 \item Peroxyacetyl nitrate ($CH_3C(O)OONO_2$, PAN) is the most abundant organic nitrate in the troposphere \citep{Hewitt1994}.
@@
 However, since the focus of this research is on the impact of alkyl nitrate chemistry on ozone such a miscount is acceptable due to purely theoretical purpose of the research.
-\item Inclusion of all chemistry listed above into the model is a subject for future work.
 \end{itemize}
@@
 For the same reason all self-reactions of organic peroxy radicals ($RO_2$) are omitted as well.
-To avoid a huge impact of incomplete oxidation of alkanes on the results the model's simulation time has been restricted to 24 hours.
+To avoid a huge impact of incomplete oxidation of alkanes on the results the simulation time has been restricted to 24 hours.
 
@@
 \subsection{Design of experiments}\label{sec:method_expdesign}
-To investigate the influence of alkyl nitrate chemistry on (potential changes in) tropospheric $O_3$ production, box model runs were performed at a series of different $NO_x$ and VOCs initial concentrations, so that an $O_3$ isopleth plot could be constructed, similar to those found in \citep{Dodge1977} and \citep{Sillman1999}.
-Table \ref{tab:setupO3CO} shows initial concentrations of 'independent'/global parameters and indicates which of them are fixed or allowed to vary during the model run.
+To investigate the influence of alkyl nitrate chemistry on tropospheric $O_3$ production, box model runs were performed at a series of different $NO_x$ and VOCs initial concentrations, so that an $O_3$ isopleth plot could be constructed, similar to those found in \citep{Dodge1977} and \citep{Sillman1999}.
+Table \ref{tab:setupO3CO} shows initial concentrations of main parameters and indicates which of them are fixed or allowed to vary during the model run.
 Table \ref{tab:setupNOx} and Table \ref{tab:setupVOCs} show the initial concentrations of $NO_x$ and VOCs respectively that are used at a series of experiments.
@@
 
-Two experiments have been performed.
-The aim of the first one is to reveal the total impact of alkyl nitrate chemistry on $O_3$ production, the aim of the second one is to estimate the contribution of individual alkyl nitrates to changes in $O_3$ production (sensitivity!).
-
-For the first experiment the model was run under two configurations, (1) with all alkyl nitrate chemistry switched on, and (2) with all alkyl nitrate chemistry switched off.
-In every model run VOCs concentrations were (changed according to the data from Table \ref{tab:setupVOCs} and) were fixed throughout the run, while the initial concentrations of $NO_x$ were (changed according to the data from Table \ref{tab:setupNOx} and were) allowed to vary throughout the run.
-As a result, the first experiment consists of 253 model runs including 11 runs (varying $NO_x$) with inorganic chemistry only.
-
-For the second experiment the model was run using the same approach concerning variations in $NO_x$ and VOCs concentrations, but changes in alkyl nitrate chemistry were treated differently.
-In every model run reactions of individual alkyl nitrates or their groups (defined in Table \ref{tab:RHandANs}) were either switched on or switched off, resulting in other 847 additional model runs.
-
-!! C2H5NO3 ethylnitrate, IC3H7NO3 i-propylnitrate, NC3H7NO3  n-propylnitrate, SC4H9NO3 2-butylnitrate, PEBNO3 2-pentylnitrate;PECNO3 3-pentylnitrate, IPEBNO3 3-methyl-2-butylnitrate
-
-\begin{table} % RH and ANs
-\caption{Alkyl nitrates and their parent alkanes}\label{tab:RHandANs}
-\centering
-\begin{tabular}{llll}
-\hline
-Parent alkane& Structural formula          & Name                     & Group\\
-\hline
-$CH_4$       & $CH_3ONO_2$                 & methyl nitrate           & $CH_3ONO_2$ \\
-\hline
-$C_2H_6$     & $CH_3CH_2ONO_2$             & ethyl nitrate            & $C_2H_5ONO_2$ \\
-\hline
-$C_3H_8$     & $(CH_3)_2CHONO_2$           & iso-propyl nitrate       & $\Sigma C_3H_7ONO_2$\\
-             & $CH_3(CH_2)_2ONO_2$         & n-propyl nitrate         & \\
-\hline
-$nC_4H_{10}$ & $CH_3(CH_2)_3ONO_2$         & n-butyl nitrate          & $\Sigma nC_4H_9ONO_2$\\
-             & $CH_3CH_2CH(CH_3)ONO_2$     & sec-butyl nitrate        & \\
-\hline
-$iC_4H_{10}$ & $CH_3CH(CH_3)CH_2ONO_2$     & iso-butyl nitrate        & $\Sigma iC_4H_9ONO_2$\\
-             & $CH_2C(CH_3)CH_2ONO_2$      & tert-butyl nitrate       & \\
-\hline
-$nC_5H_{12}$ & $CH_3(CH_2)_4ONO_2$         & n-pentyl nitrate         & $\Sigma nC_5H_{11}ONO_2$\\
-             & $CH_3(CH_2)_2CH(CH_3)ONO_2$ & 2-pentyl nitrate         & \\
-             & $???$                       & 3-pentyl nitrate         & \\
-\hline
-$iC_5H_{12}$ & $???$                       & 2-methylbutyl nitrate    & $\Sigma iC_5H_{11}ONO_2$\\
-             & $(CH_3)_2CH(CH_2)_2ONO_2$   & 3-methyl-2-butyl nitrate & \\
-             & $(CH_3)_2C(CH_2CH_3)ONO_2$  & 2-methyl-2-butyl nitrate & \\
-\hline
-\end{tabular}
-\end{table}
+Two series of experiments have been performed.
+The aim of the first one is to reveal the total impact of alkyl nitrate chemistry on $O_3$ production, the aim of the second one is to estimate the contribution of individual alkyl nitrates to changes in $O_3$ production.
+
+For the first series of experiments the model was run under two configurations, (1) with all alkyl nitrate chemistry switched on, and (2) with all alkyl nitrate chemistry switched off.
+In every model run VOCs concentrations were (changed according to the data from Table \ref{tab:setupVOCs}) and were fixed throughout the run, while the initial concentrations of $NO_x$ (Table \ref{tab:setupNOx}) were allowed to vary throughout the run.
+As a result, the first experiment consists of 253 model runs including 11 runs (with varying $NO_x$) with inorganic chemistry only.
+
+For the second series of experiments the model was run with $NO_x$ initial concentration equal to 10 ppb and VOCs concentrations from case F.
+In every model run reactions of individual alkyl nitrates or their groups (defined in Table \ref{tab:RHandANs}) were either switched on or switched off, resulting in other 7 additional model runs.
 
 \begin{table} % model setup: O3, CO
-\caption{Fixed and variable 'independent' parameters used in box model runs}\label{tab:setupO3CO}
+\caption{Fixed and variable main parameters used in all model runs}\label{tab:setupO3CO}
 \centering
@@
 3         & 7           & 10    \\
+15        & 35          & 50    \\
+30        & 70          & 100   \\
 \hline
@@
 K  & 2050 &	7.000 &	3.000 &	2.000 &	1.000 &	0.500 &	2.000 \\
+\hline
+\end{tabular}
+\end{table}
+
+\begin{table} % RH and ANs
+\caption{Alkyl nitrates and their parent alkanes}\label{tab:RHandANs}
+\centering
+\begin{tabular}{llll}
+\hline
+Parent alkane& Structural formula          & Name                     & Group\\
+\hline
+$CH_4$       & $CH_3ONO_2$                 & methyl nitrate           & $CH_3ONO_2$ \\
+\hline
+$C_2H_6$     & $CH_3CH_2ONO_2$             & ethyl nitrate            & $C_2H_5ONO_2$ \\
+\hline
+$C_3H_8$     & $(CH_3)_2CHONO_2$           & iso-propyl nitrate       & $\Sigma C_3H_7ONO_2$\\
+             & $CH_3(CH_2)_2ONO_2$         & n-propyl nitrate         & \\
+\hline
+$nC_4H_{10}$ & $CH_3(CH_2)_3ONO_2$         & n-butyl nitrate          & $\Sigma nC_4H_9ONO_2$\\
+             & $CH_3CH_2CH(CH_3)ONO_2$     & 2-butyl nitrate        & \\
+\hline
+$iC_4H_{10}$ & $CH_3CH(CH_3)CH_2ONO_2$     & iso-butyl nitrate        & $\Sigma iC_4H_9ONO_2$\\
+             & $CH_2C(CH_3)CH_2ONO_2$      & tert-butyl nitrate       & \\
+\hline
+$nC_5H_{12}$ & $CH_3(CH_2)_4ONO_2$         & n-pentyl nitrate         & $\Sigma nC_5H_{11}ONO_2$\\
+             & $CH_3(CH_2)_2CH(CH_3)ONO_2$ & 2-pentyl nitrate         & \\
+             &                        & 3-pentyl nitrate         & \\
+\hline
+$iC_5H_{12}$ &                        & 2-methylbutyl nitrate    & $\Sigma iC_5H_{11}ONO_2$\\
+             & $(CH_3)_2CH(CH_2)_2ONO_2$   & 3-methyl-2-butyl nitrate & \\
+             & $(CH_3)_2C(CH_2CH_3)ONO_2$  & 2-methyl-2-butyl nitrate & \\
 \hline

### 2015-08-06 MSc_diss.tex
Almost done

@@
 \begin{abstract}
-Abstract.300 words.
+This study attempts to investigate the role of alkyl nitrates in tropospheric ozone chemistry.
+A conceptual box model representing the chemistry of mean conditions of global troposphere was developed.
+The sensitivity of ozone production rate to inclusion of different alkyl nitrates into the set of reactions was analysed in various regimes of ozone precursors.
+For the first time the ozone production efficiency and its relation to alkyl nitrate presence was estimated using the so-called carbon bonds method.
+The model behaviour stays in accordance with previous research, while the major findings of the present study offer an insight into alkyl nitrate distribution in the $NO_x$-VOC space.
 \end{abstract}
 
+\newpage
 \tableofcontents
@@
 After the Deepwater Horizon (DWH) offshore oil platform accident on 20 April 2010 large amounts of oil from the DWH spill reached the ocean surface.
-Subsequent oil evaporation caused a release of methane and other light alkanes to the atmosphere forming a $VOC$ mixture atypical for the remote marine environment.
+Subsequent oil evaporation caused a release of methane and other light alkanes to the atmosphere forming a VOC mixture atypical for the remote marine environment.
 At the same the $NO_x$ levels were enhanced in the vicinity of the oil spill due to spill response efforts (ship exhaust, surface oil burning, flaring of recovered gas).
@@
 
-The concentration of $NO_x$ is the main factor which determines whether $O_3$ is produced or removed in the atmosphere, and theoretically there are three regimes of $O_3$ production that depend on $NO_x$ levels: at low $NO_x$ levels the regime is characterized by net $O_3$ removal, at intermediate $NO_x$ (also called $NO_x$-sensitive) - by net $O_3$ formation, and at high $NO_x$ - by net $O_3$ formation or removal depending more on VOCs concentration (therefore called $VOC$-sensitive) (e.g., \cite{Fowler2008}).
+The concentration of $NO_x$ is the main factor which determines whether $O_3$ is produced or removed in the atmosphere, and theoretically there are three regimes of $O_3$ production that depend on $NO_x$ levels: at low $NO_x$ levels the regime is characterized by net $O_3$ removal, at intermediate $NO_x$ (also called $NO_x$-sensitive) - by net $O_3$ formation, and at high $NO_x$ - by net $O_3$ formation or removal depending more on VOCs concentration (therefore called VOC-sensitive) (e.g., \cite{Fowler2008}).
 When the prevailing $NO_x$ levels are low, a net $O_3$ removal takes place because there is not enough $NO$ around to react with the peroxy radicals ($HO_2$ and $RO_2$) to produce $NO_2$, which photolysis in conjunction with the reaction (\ref{reac:O3P+O2=O3}) is the main source of $O_3$ in the troposphere.
@@
 When the prevailing $NO_x$ levels are high, reactions (\ref{reac:HO2+NO=NO2+OH}) and (\ref{reac:RO2+NO=RO+NO2}) dominate for $HO_2$ and $RO_2$, making the rate of $O_3$ production being controlled by the availability of these radicals.
-The availability of $OH$ radical is also important at high $NO_x$ levels because $OH$ connects the $O_3$ production rate with rates of its reactions with $CO$ and $RH$ (reactions (\ref{reac:OH+CO=H+CO2}) and (\ref{reac:OH+RH=RO2+H2O})), therefore making this regime $VOC$-sensitive.
+The availability of $OH$ radical is also important at high $NO_x$ levels because $OH$ connects the $O_3$ production rate with rates of its reactions with $CO$ and $RH$ (reactions (\ref{reac:OH+CO=H+CO2}) and (\ref{reac:OH+RH=RO2+H2O})), therefore making this regime VOC-sensitive.
 Plus at this point the reaction of $OH$ with $NO_2$ to form nitric acid ($HNO_3$) becomes the major termination reaction for free radicals:
@@
 
-Since the chemical mechanism of our model does not include the reaction of $HNO_3$ formation, a good representation of $VOC$-sensitive regime by our model is not expected.
-It also makes it difficult to split $NO_x$-sensitive and $VOC$-sensitive regimes in our model, because this split is determined by the recurrence (size) of the peroxide- and nitric-acid forming reactions \citep{Sillman1999}, while our mechanism lacks one of these reactions.
+Since the chemical mechanism of our model does not include the reaction of $HNO_3$ formation, a good representation of VOC-sensitive regime by our model is not expected.
+It also makes it difficult to split $NO_x$-sensitive and VOC-sensitive regimes in our model, because this split is determined by the recurrence (size) of the peroxide- and nitric-acid forming reactions \citep{Sillman1999}, while our mechanism lacks one of these reactions.
 However, needing to evaluate performance of our model under higher $NO_x$ conditions anyway below we consider just a contrasting case of $NO_x$ levels, for example, 5 ppb (Figure \ref{fig:model_validation}).
@@
 
-The best way to illustrate the central features of the relationship between $O_3$, $NO_x$ and $VOC$ mentioned above is to construct a ozone isopleth plot in a form shown in Figure \ref{fig:Sillman1999}.
-The isopleth plot shows that the rate of $O_3$ production is a nonlinear function of the $NO_x$ and $VOC$ concentrations.
-When $VOC$/$NO_x$ ratios are high, the $O_3$ production rate increases with increasing $NO_x$ eventually reaching a local maximum, and decreasing afterwards as $NO_x$ concentrations continue to increase.
-The local maximum (the 'ridge line') defines the split into $NO_x$-sensitive and $VOC$-sensitive regimes.
-Under the 'ridge line' the $VOC$/$NO_x$ ratios are high, and the $O_3$ production rate increases with increasing $NO_x$ and almost insensitive to $VOC$.
-Above the 'ridge line' the $VOC$/$NO_x$ ratios are low, and the $O_3$ production rate increases with increasing $VOC$ and decreases with increasing $NO_x$ \citep{Sillman2013}.
+The best way to illustrate the central features of the relationship between $O_3$, $NO_x$ and VOC mentioned above is to construct an ozone isopleth plot in a form shown in Figure \ref{fig:Sillman1999}.
+The isopleth plot shows that the rate of $O_3$ production is a nonlinear function of the $NO_x$ and VOC concentrations.
+When VOC/$NO_x$ ratios are high, the $O_3$ production rate increases with increasing $NO_x$ eventually reaching a local maximum, and decreasing afterwards as $NO_x$ concentrations continue to increase.
+The local maximum (the 'ridge line') defines the split into $NO_x$-sensitive and VOC-sensitive regimes.
+Under the 'ridge line' the VOC/$NO_x$ ratios are high, and the $O_3$ production rate increases with increasing $NO_x$ and almost insensitive to VOC.
+Above the 'ridge line' the VOC/$NO_x$ ratios are low, and the $O_3$ production rate increases with increasing VOC and decreases with increasing $NO_x$ \citep{Sillman2013}.
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
         \includegraphics[width=\linewidth]{{./pics/Sillman1999_Fig1_adapted_withdotsrectangle}.png}
-        \caption{(ppb/h, solid lines) as a function of $VOC$ (ppbC) and $NO_x$ (ppb) for mean summer daytime meteorology and clear skies.
+        \caption{for mean summer daytime meteorology and clear skies.
 The solid lines represent production rates of 1, 2.5, 5, 10, 15, 20 and 30 ppb/h.
@@
         \includegraphics[width=\linewidth]{D:/FACSIMILE/ANsCBmodel/matlab/ANsCB_pics/ANsNOxVOC/noAN/chem_noAN_netO3rate_withINORG_NOx_1ppb_100ppb.eps}
-        \caption{as a function of $VOC$ (ppbC) and $NO_x$ (ppb) in the first modelling experiment with alkyl nitrate chemistry switched off combined with an additional run with inorganic chemistry only (i.e.
-the only $VOC$ is $CO$).} \label{fig:noAN_netO3rate}
+        \caption{from the first series of experiments with alkyl nitrate chemistry switched off combined with an additional run with inorganic chemistry only (i.e.
+the only VOC is $CO$).} \label{fig:noAN_netO3rate}
     \end{subfigure}
-    \caption{Isopleths giving net rate of $O_3$ production (ppb/h, color)}\label{fig:netO3rate}
+    \caption{Isopleths of net rate of $O_3$ production (ppb/h) as a function of VOC (ppbC) and $NO_x$ (ppb).}\label{fig:netO3rate}
 \end{figure}
@@
 
-Rewrite!
-
-The same kind of plot can be found in Figure \ref{fig:noAN_netO3rate}, which is produced using data from our model from the first experiment with alkyl nitrate chemistry switched off.
-Taking into account the fact that the maximum $NO_x$ mixing ratio considered in the experiment is 10 ppb, our results are directly comparable only with a part of the data from Figure \ref{fig:Sillman1999}, namely with those inside the green dashed square.
-Moreover, since the minimum $NO_x$ mixing ratio considered in the experiment is 5 ppt, while the corresponding value in Figure \ref{fig:Sillman1999} is 1 ppb, a negative $O_3$ production rate is seen in our results, whereas it stays out of range of Figure \ref{fig:Sillman1999}.
+The same kind of plot can be found in Figure \ref{fig:noAN_netO3rate}, which is produced using data from our model from the first series of experiments with alkyl nitrate chemistry switched off.
+Taking into account the fact that the minimum sum of VOCs mixing ratios considered in this work is 1782 ppb, our results are directly comparable only with a part of the data from Figure \ref{fig:Sillman1999}, namely with those inside the green dashed rectangle.
+Moreover, since the minimum $NO_x$ mixing ratio considered in this work is 5 ppt, while the corresponding value in Figure \ref{fig:Sillman1999} is 1 ppb, a negative $O_3$ production rate occurs in our results, whereas it stays out of range of Figure \ref{fig:Sillman1999} (i.e., not all of our modelling data is shown in Figure \ref{fig:noAN_netO3rate}).
 
 One can see the same pattern in these two figures.
-At high $VOC$/$NO_x$ ratios, the rate of $O_3$ production increases with increasing $NO_x$ and reaches its maximum at the highest $NO_x$ and $VOC$ levels (which corresponds to the top right corner of the green dashed square).
-The change in $O_3$ production rate at high $VOC$/$NO_x$ ratios in our model is largely insensitive to $VOC$, which fits into the theory.
-At low $VOC$/$NO_x$ ratios the $O_3$ production rate starts to depend on $VOC$ concentration also resembling the behaviour of ozone isopleths at the top left corner of the green dashed square.
-The overall tendency of $O_3$ production rate to increase with increasing $VOC$ and $NO_x$ is present in both figures showing that the model reproduces this feature according to the theory.
-However, the modelled values of $O_3$ production rates are much lower than those in Figure \ref{fig:Sillman1999}: in the model the maximum production rate is 9 ppb/h, while in theory it is 30 ppb/h, meaning that our model underestimates the $O_3$ production rate, but reproduces some features of the relationship between $O_3$, $NO_x$ and $VOC$.
+At high VOC/$NO_x$ ratios, the rate of $O_3$ production increases with increasing $NO_x$ and reaches its maximum at the highest $NO_x$ and VOC levels.
+In theory at high $NO_x$ and VOC levels the $O_3$ production rate decreases (top right corner of green rectangle), due to impact of $HNO_3$ chemistry.
+Therefore, this figure confirms once again that the model does not account for impact of $HNO_3$ chemistry.
+
+The change in $O_3$ production rate at high VOC/$NO_x$ ratios in our model is somewhat insensitive to VOC, which fits into the theory.
+At low VOC/$NO_x$ ratios the $O_3$ production rate starts to depend on VOC concentration also resembling the behaviour of ozone isopleths at the middle of green rectangle.
+The overall tendency of $O_3$ production rate to increase with increasing VOC and $NO_x$ is present in both figures showing that the model reproduces this feature according to the theory.
+The modelled values of $O_3$ production rates are similar than those in Figure \ref{fig:Sillman1999}: in the model the maximum production rate is 25 ppb/h, in theory it is 30 ppb/h, meaning that our model shows the same $O_3$ production rates, but does not show the same distribution of $O_3$ production rate in $NO_x$-VOC space as in the theory.
 
@@
 
-To evaluate the sensitivity of $O_3$ production to alkyl nitrate chemistry under different $NO_x$ and $VOC$ conditions a number of simulations were carried out.
+To evaluate the sensitivity of $O_3$ production to alkyl nitrate chemistry under different $NO_x$ and VOC conditions a number of simulations were carried out.
 In these simulations the chemical mechanism of the model was adjusted to either include or exclude the formation and loss of alkyl nitrates, which was accomplished by switching on and off all reactions that involve $RONO_2$ while keeping the rest of the mechanism unchanged.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{table}[h] %  AN branching ratios
+\begin{table}[h!] %  AN branching ratios
 \caption{Branching ratios for the formation of alkyl nitrates from their precursor peroxy radicals and $NO$ (taken from the MCM v3.3 (2015)).}\label{tab:ANbranching}
@@
     \hfill
-\caption{Isopleths giving net rate of $O_3$ production (ppb/h, color) as a function of $VOC$ (ppbC) and $NO_x$ (ppb) with alkyl nitrate chemistry switched off (left) and switched on (middle).
-Difference between them is shown on the right.}\label{fig:netO3rate_noAN_withAN_diff}
+\caption{Isopleths giving net rate of $O_3$ production (ppb/h, color) as a function of VOC (ppbC) and $NO_x$ (ppb) with alkyl nitrate chemistry switched off (top left) and switched on (top right).
+Difference between them is shown at the bottom.}\label{fig:netO3rate_noAN_withAN_diff}
 \end{figure}
@@
   \centering
-  \includegraphics[width=.99\linewidth]{D:/FACSIMILE/ANsCBmodel/matlab/ANsCB_pics/ANsNOxVOC/allAN/chem_allAN_netO3mixratdiffpercent_noINORG_NOx_250ppt_100ppb.eps}
+  \includegraphics[width=.99\linewidth]{D:/FACSIMILE/ANsCBmodel/matlab/ANsCB_pics/ANsNOxVOC/allAN/chem_allAN_netO3mixratdiffpercent_noINORG_NOx_250ppt_100ppb_new.eps}
 \end{minipage}
-\caption{Isopleths giving net $O_3$ production (ppb, color) as a function of $VOC$ (ppbC) and $NO_x$ (ppb, range 250 ppt - 100 ppb) with alkyl nitrate chemistry switched off (a) and switched on (b).
-Absolute difference (with alkyl nitrates minus without them) between two series of model runs in ppb (c), same in \% (d).
+\caption{Isopleths of net $O_3$ production (ppb, color) as a function of VOC (ppbC) and $NO_x$ (ppb, range 250 ppt - 100 ppb) with alkyl nitrate chemistry switched off (top left) and switched on (top right).
+Absolute difference between two series of model runs in ppb (bottom left), same in \% (bottom right).
 Net production = final - initial value of $O_3$ mixing ratio.}\label{fig:netO3mixrat_noAN_withAN_diff}
@@
 Therefore, a more useful estimate is a percentage difference in $O_3$ mixing ratios between two series of model runs shown in Figure \ref{fig:netO3rate_noAN_withAN_diff}.
-It appears that the largest absolute changes in $O_3$ mixing ratios (40--45 ppb) are changes by only 20--30\%, while changes up to 90\% take place the model is in low $NO_x$ and high VOCs conditions.
+It appears that the largest absolute changes in $O_3$ mixing ratios (40--45 ppb) are changes by only 20--30\%, while changes up to 90\% take place the model in low $NO_x$ and high VOCs conditions.
 
@@
 
-The contribution of different alkyl nitrates to the $O_3$ decrease varies: the largest drop --- 11.6 ppb --- is caused by ethyl nitrate ($C_2H_5ONO_2$), and is almost the same for methyl nitrate.
-On the other hand, iso-pentyl nitrates account for the smallest $O_3$ decrease.
-
 The biggest change in $O_3$ (11.6 ppb) is related to ethyl nitrate, though its production by the end of the simulation is the smallest in comparison to other nitrates.
 Furthermore, when these is no conversion of methane to methyl nitrate, but conversion of ethane to ethyl nitrate is present even at a small initial value of ethane ($C_2H_6$ 3.8 ppb in comparison to 1915 ppb of methane) we observe almost the same change in $O_3$ mixing ratio.
-And surprisingly, in experiment with full alkyl nitrate chemistry we see almost the same $O_3$ change as in case of methyl or ethyl nitrate, but slightly higher.
+Surprisingly, in experiment with full alkyl nitrate chemistry we see almost the same $O_3$ change as in case of methyl or ethyl nitrate.
 
@@
 \centering
-\begin{minipage}{.3\textwidth}
+\begin{minipage}{.5\textwidth}
   \centering
@@
 \end{minipage}
-\begin{minipage}{.3\textwidth}
+\begin{minipage}{.5\textwidth}
   \centering
@@
 \end{minipage}
-\caption{Isopleths giving net $O_3$ production efficiency (or $\epsilon$, molecules of $O_3$ per carbon bond, color) as a function of $VOC$ (ppbC) and $NO_x$ (250ppt - 100 ppb) with alkyl nitrate chemistry switched off (left) and switched on (middle).
+\caption{Isopleths giving net $O_3$ production efficiency (or $\epsilon$, molecules of $O_3$ per carbon bond, color) as a function of VOC (ppbC) and $NO_x$ (250ppt - 100 ppb) with alkyl nitrate chemistry switched off (left) and switched on (middle).
 Difference between them is shown on the right.}\label{fig:epsilon250ppt100ppb}
@@
 \end{minipage}
-\caption{Isopleths giving net $O_3$ production efficiency (or $\epsilon$, molecules of $O_3$ per carbon bond, color) as a function of $VOC$ (ppbC) and $NO_x$ (5 ppt - 100 ppb) with alkyl nitrate chemistry switched off (left) and switched on (middle).
-Difference between them is shown on the right.}\label{fig:epsilon5ppt100ppb}
+\caption{Isopleths giving net $O_3$ production efficiency (or $\epsilon$, molecules of $O_3$ per carbon bond, color) as a function of VOC (ppbC) and $NO_x$ (5 ppt - 100 ppb) with alkyl nitrate chemistry switched off (top) and switched on (middle).
+Difference between them is shown at the bottom.}\label{fig:epsilon5ppt100ppb}
 \end{figure}
@@
 
-Forthly, it could have been better to run the model to a steady state, however since our model does not have a full oxidation of the organics, it could have caused bigger errors that otherwise.
-
-Nevertheless, the model appeared to be sensitive to alkyl nitrate chemistry because $O_3$ yields
-
-add night chemistry
-photolysis variation during the day
-compare with observations
-
-Future plans include illumination of all weaknesses from our model, particularly adding other sinks for $NO_x$ into the chemical mechanism.
-preliminary study
-
-Probably it is reasonable to conduct a similar kind of research using the Common Representative Mechanism (CRI) which includes representation of alkyl nitrates and is traceable to the MCM.
-
-Future plans and possible improvements in the model.
-
-\subsubsection*{Future research}
-A greater depth of information may have been obtained by conducting more experiments.
-Another possible improvement to the study could have been
+Last but not least, it could have been better to run the model to a steady state, however since our model does not have complete oxidation of the organics, running it for longer than 24 hours could cause larger errors.
+
+Nevertheless, the model appeared to be sensitive to alkyl nitrate chemistry, since the final ozone mixing ratios and average production rates were observed to be different if nitrate chemistry is switched on or off.
+
+All of the above said, this preliminary study needs to be extended into further research to produce more robust results.
+
+It seems reasonable to conduct a similar kind of investigation using the Common Representative Mechanism (CRI) which includes the representation of alkyl nitrates and is traceable to the MCM.
+
+Moreover, future plans include gradual elimination of all weaknesses from our model, particularly adding other sinks for $NO_x$ into the chemical mechanism.
+Another possible improvement to the study can be achieved if night chemistry is added, accounting for photolysis variation during the day.
+The latter will allow us to make a comparison with observations.
 
@@
 The present study provides an preliminary assessment of the sensitivity of tropospheric ozone production efficiency to alkyl nitrate chemistry using conceptual box modelling.
-It was found that the model is capable to capture some features of daytime tropospheric ozone chemistry, but due to various limitations does do it very well.
+It was found that the model is capable to capture some features of daytime tropospheric ozone chemistry, but due to various limitations does not do it perfectly.
 
 Inclusion of $RONO_2$ chemistry slows down the net $O_3$ production rate leading to a smaller overall $O_3$ formation by the end of simulation.
+At 100 ppb of $NO_x$ and high VOC level we observe the biggest decrease in $O_3$ production rate, which is almost 2 ppb/h.
 
@@
 
+The contribution of different alkyl nitrates to the $O_3$ decrease varies: the largest drop is caused by ethyl nitrate and is almost the same for methyl nitrate and the sum of all nitrates.
+On the other hand, iso-pentyl nitrates account for the smallest $O_3$ decrease.
+
+The sensitivity of our model to alkyl nitrate chemistry can be expressed in terms of amount of carbon bonds in the system: on average, 4.5 (4) carbon bonds need to be emitted to produce one molecule of $O_3$ with (without) alkyl nitrates.
+Although the difference between these values is too small, it is consistent with the results obtained with the model GEOS-Chem.
+
 We hope that our results concerning the sensitivity of ozone calculations to alkyl nitrate chemistry will be interesting to other researches, as well as the carbon bond efficiency calculations.
@@
 \bibliography{diss_refs}
-
+\newpage
 \appendix
-\section{Appendix I} \label{sec:appendix1}
+\section{Model code} \label{sec:appendix1}
+
+\input{model_code}
 

### 2020-04-06 code.tex
re-structure msc thesis code

\begin{verbatim}
*-------------------------------REACTIONS------------------------------;
*-------------------------------INORGANIC------------------------------;
*---Inorganic (0_00);
RI1A  % 6.0D-34*O2*(T/300)@-2.6*O2   : O = O3;
RI1B  % 5.6D-34*N2*(T/300)@-2.6*O2   : O = O3;
RI2   % 8.0D-12*EXP(-2060/T)         : O+O3 = ;
RI3   % 3.2D-11*EXP(67/T)*O2         : O1D = O;
RI12A % J<1>                         : O3 = O1D; 
RI12B % J<2>                         : O3 = O;
RI19  % 2.2D-10*H2O                  : O1D = OH+OH;
RI20  % 7.7D-12*EXP(-2100/T)         : OH+H2 = HO2;
RI23  % 4.8D-11*EXP(250/T)           : OH+HO2 = ;
RI24  % 2.9D-12*EXP(-160/T)          : OH+H2O2 = HO2;
RI25  % 1.7D-12*EXP(-940/T)          : OH+O3 = HO2;
RI26  % 2.2D-13*EXP(600/T)*KI2627    : HO2+HO2 = H2O2;
RI27  % 1.9D-33*N2*EXP(980/T)*KI2627 : HO2+HO2 = H2O2;
RI28  % 2.03D-16*(T/300)@4.57*EXP(693/T): HO2+O3 = OH;
RI30  % J<3>                         : H2O2 = OH+OH;
RI36  % 1.8D-11*EXP(107/T)*N2        : O1D = O;
RI45  % 3.6D-12*EXP(270/T)           : HO2+NO = OH+NO2;
RI54  % 1.4D-12*EXP(-1310/T)         : NO+O3 = NO2;
RI68  % J<4>                         : NO2 = NO+O;
RCH42 % 1.44D-13*(1+(M/4.2D+19))     : OH+CO = HO2; 
* RI69  % KMT08                        : OH+NO2 = HNO3;
* RI70  % J<8>                         : HNO3 = OH+NO2;
*--------------------------------ORGANIC-------------------------------;
*---Methane (0_01);
RCH41 % 1.85D-12*EXP(-1690/T)        : OH+CH4 = CH3O2; 
RCH43 % 5.4D-12*EXP(135/T)           : OH+HCHO = HO2+CO; 
RCH44 % 5.3D-12*EXP(190/T)*0.6       : OH+CH3OOH = CH3O2; 
RCH45 % 5.3D-12*EXP(190/T)*0.4       : OH+CH3OOH = HCHO+OH;    
RCH46 % 3.8D-13*EXP(780/T)*(1-1/(1+498*EXP(-1160/T))):HO2+CH3O2=CH3OOH; 
RCH47 % 3.8D-13*EXP(780/T)*(1/(1+498*EXP(-1160/T)))  :HO2+CH3O2=HCHO; 
RCH48 % 7.2D-14*EXP(-1080/T)*O2      : CH3O = HCHO+HO2;
RCH49 % 2.3D-12*EXP(360/T)*0.999     : CH3O2+NO = CH3O+NO2; 
RCH410% J<11>                        : HCHO = CO+HO2+HO2; 
RCH411% J<12>                        : HCHO = CO+H2; 
RCH412% J<41>                        : CH3OOH = CH3O+OH; 
*---Ethane (0_02);
RC2H61 % 6.9D-12*EXP(-1000/T)        : OH+C2H6 = C2H5O2;
RC2H62 % 3.8D-13*EXP(900/T)          : HO2+C2H5O2 = C2H5OOH;
RC2H63 % 2.55D-12*EXP(380/T)*0.991   : C2H5O2+NO = C2H5O+NO2;
RC2H64 % J<41>                       : C2H5OOH = C2H5O+OH;
RC2H65 % 1.90D-12*EXP(190/T)         : OH+C2H5OOH = C2H5O2;
RC2H66 % 8.01D-12                    : OH+C2H5OOH = CH3CHO+OH;
RC2H67 % 2.4D-14*EXP(-325/T)*O2      : C2H5O = CH3CHO+HO2;
RC2H68 % 3.0D-12*EXP(20/T)*0.05      : OH+C2H5OH = C2H5O;
RC2H69 % 3.0D-12*EXP(20/T)*0.90      : OH+C2H5OH = CH3CHO+HO2;
RC2H610% J<13>                       : CH3CHO = CH3O2+HO2+CO;
RC2H611% 4.7D-12*EXP(345/T)*0.95     : OH+CH3CHO = CH3CO3;
RC2H612% 4.7D-12*EXP(345/T)*0.05     : OH+CH3CHO = HCOCH2O2;
*---Propane (0_03);
RC3H81 % 7.6D-12*EXP(-585/T)*0.736   : OH+C3H8 = IC3H7O2;
RC3H82 % 7.6D-12*EXP(-585/T)*0.264   : OH+C3H8 = NC3H7O2;
RC3H83 % KRO2HO2*0.520               : IC3H7O2+HO2 = IC3H7OOH;
RC3H84 % 2.7D-12*EXP(360/T)*0.958    : IC3H7O2+NO = IC3H7O+NO2;
RC3H85 % KRO2HO2*0.520               : NC3H7O2+HO2 = NC3H7OOH;
RC3H86 % 2.9D-12*EXP(350/T)*0.980    : NC3H7O2+NO = NC3H7O+NO2;
RC3H87 % J<41>                       : IC3H7OOH = IC3H7O+OH;
RC3H88 % 1.66D-11                    : OH+IC3H7OOH = CH3COCH3+OH;
RC3H89 % 1.90D-12*EXP(190/T)         : OH+IC3H7OOH = IC3H7O2;
RC3H810% 1.5D-14*EXP(-230/T)*O2      : IC3H7O = CH3COCH3+HO2;
RC3H811% 8.8D-12*EXP(-1320/T)+1.7D-14*EXP(423/T):CH3COCH3+OH=CH3COCH2O2; 
RC3H812% J<21>                       : CH3COCH3 = CH3CO3+CH3O2;
RC3H813% J<41>                       : NC3H7OOH = NC3H7O+OH;
RC3H814% 1.10D-11                    : OH+NC3H7OOH = C2H5CHO+OH;
RC3H815% 1.90D-12*EXP(190/T)         : OH+NC3H7OOH = NC3H7O2;
RC3H816% 2.6D-14*EXP(-255/T)*O2      : NC3H7O = C2H5CHO+HO2;
RC3H817% 4.9D-12*EXP(405/T)          : C2H5CHO+OH = C2H5CO3;
RC3H818% J<14>                       : C2H5CHO = C2H5O2+HO2+CO;
*---n-butane (0_04);
RNC4H101 % 9.8D-12*EXP(-425/T)*0.127 : OH+NC4H10 = NC4H9O2;
RNC4H102 % 9.8D-12*EXP(-425/T)*0.873 : OH+NC4H10 = SC4H9O2;
RNC4H103 % KRO2HO2*0.625             : NC4H9O2+HO2 = NC4H9OOH;
RNC4H104 % KRO2NO*0.967              : NC4H9O2+NO = NC4H9O+NO2;
RNC4H105 % KRO2HO2*0.625             : SC4H9O2+HO2 = SC4H9OOH;
RNC4H106 % KRO2NO*0.910              : SC4H9O2+NO = SC4H9O+NO2;
RNC4H107 % J<41>                     : NC4H9OOH = NC4H9O+OH;	
RNC4H108 % 1.24D-11                  : OH+NC4H9OOH = C3H7CHO+OH;
RNC4H109 % 1.90D-12*EXP(190/T)       : OH+NC4H9OOH = NC4H9O2;
RNC4H1010% 8.9D-14*EXP(-550/T)*O2    : NC4H9O = C3H7CHO+HO2;
RNC4H1011% 4.6D10*EXP(-3570/T)       : NC4H9O = HO1C4O2;
RNC4H1012% J<41>                     : SC4H9OOH = SC4H9O+OH;
RNC4H1013% 5.6D12*EXP(-5780/T)       : SC4H9O = CH3CHO+C2H5O2;
RNC4H1014% 1.5D-14*EXP(-200/T)*O2    : SC4H9O = MEK+HO2 ;
RNC4H1015% J<22>                     : MEK = CH3CO3+C2H5O2;
*---i-butane;
RIC4H101 % 1.16D-17*T@2*EXP(225/T)*0.206 : OH+IC4H10 = IC4H9O2;
RIC4H102 % 1.16D-17*T@2*EXP(225/T)*0.794 : OH+IC4H10 = TC4H9O2;
RIC4H103 % KRO2HO2*0.625                 : IC4H9O2+HO2 = IC4H9OOH;
RIC4H104 % KRO2NO*0.967                  : IC4H9O2+NO = IC4H9O+NO2;
RIC4H105 % KRO2HO2*0.625                 : TC4H9O2+HO2 = TC4H9OOH;
RIC4H106 % KRO2NO*0.975                  : TC4H9O2+NO = TC4H9O+NO2;
RIC4H107 % J<41>                         : IC4H9OOH = IC4H9O+OH;
RIC4H108 % 1.90D-12*EXP(190/T)           : OH+IC4H9OOH = IC4H9O2;
RIC4H109 % 1.24D-11                      : OH+IC4H9OOH = IPRCHO+OH;
RIC4H1010% KROPRIM*O2                    : IC4H9O = IPRCHO+HO2;
RIC4H1011% 2.17D-12*EXP(190/T)           : OH+TC4H9OOH = TC4H9O2;
RIC4H1012% J<41>                         : TC4H9OOH = TC4H9O+OH;
RIC4H1013% 6.00D+14*EXP(-8153/T)         : TC4H9O = CH3COCH3+CH3O2;
*---n-pentane (0_05);
RNC5H121 % 2.44D-17*T@2*EXP(183/T)*0.083 : OH+NC5H12 = PEAO2;
RNC5H122 % 2.44D-17*T@2*EXP(183/T)*0.568 : OH+NC5H12 = PEBO2;
RNC5H123 % 2.44D-17*T@2*EXP(183/T)*0.349 : OH+NC5H12 = PECO2;
RNC5H124 % KRO2HO2*0.706                 : PEAO2+HO2 = PEAOOH;
RNC5H125 % KRO2NO*0.948                  : PEAO2+NO = PEAO+NO2;
RNC5H126 % KRO2HO2*0.706                 : PEBO2+HO2 = PEBOOH;
RNC5H127 % KRO2NO*0.871                  : PEBO2+NO = PEBO+NO2;
RNC5H128 % KRO2HO2*0.706                 : PECO2+HO2 = PECOOH;
RNC5H129 % KRO2NO*0.869                  : PECO2+NO = PECO+NO2;
RNC5H1210% 1.41D-11                      : OH+PEAOOH = C4H9CHO+OH;
RNC5H1211% 1.90D-12*EXP(190/T)           : OH+PEAOOH = PEAO2;
RNC5H1212% J<41>                         : PEAOOH = PEAO+OH;
RNC5H1213% 2.30D-11                      : OH+PEBOOH = MPRK+OH;
RNC5H1214% 1.90D-12*EXP(190/T)           : OH+PEBOOH = PEBO2;
RNC5H1215% J<41>                         : PEBOOH = PEBO+OH;
RNC5H1216% 2.73D-11                      : OH+PECOOH = DIEK+OH;
RNC5H1217% J<41>                         : PECOOH = PECO+OH;
*---i-pentane;
RIC5H121 % 3.70D-12*0.087                : OH+IC5H12 = IPEAO2;
RIC5H122 % 3.70D-12*0.297                : OH+IC5H12 = IPEBO2;
RIC5H123 % 3.70D-12*0.616                : OH+IC5H12 = IPECO2;
RIC5H124 % KRO2HO2*0.706                 : IPEAO2+HO2 = IPEAOOH;
RIC5H125 % KRO2NO*0.948                  : IPEAO2+NO = IPEAO+NO2;
RIC5H126 % KRO2HO2*0.706                 : IPEBO2+HO2 = IPEBOOH;
RIC5H127 % KRO2NO*0.859                  : IPEBO2+NO = IPEBO+NO2;
RIC5H128 % KRO2HO2*0.706                 : IPECO2+HO2 = IPECOOH;
RIC5H129 % KRO2NO*0.953                  : IPECO2+NO = IPECO+NO2;
RIC5H1210% J<41>                         : IPEAOOH = IPEAO+OH;
RIC5H1211% 1.41D-11                      : OH+IPEAOOH = BUT2CHO+OH;
RIC5H1212% 1.90D-12*EXP(190/T)           : OH+IPEAOOH = IPEAO2;
RIC5H1213% J<41>                         : IPEBOOH = IPEBO+OH;
RIC5H1214% 1.90D-12*EXP(190/T)           : OH+IPEBOOH = IPEBO2;
RIC5H1215% 2.29D-11                      : OH+IPEBOOH = MIPK+OH;
RIC5H1216% J<41>                         : IPECOOH = IPECO+OH;
RIC5H1217% 5.25D-12                      : OH+IPECOOH = IPECO2;
*-------------------------------ALKYL NITRATES-------------------------;
*---Methyl nitrate (1_01);
* RCH4NO31 % 2.3D-12*EXP(360/T)*0.001  : CH3O2+NO = CH3NO3;
* RCH4NO32 % 4.0D-13*EXP(-845/T)       : OH+CH3NO3 = HCHO+NO2;
* RCH4NO33 % J<51>                     : CH3NO3 = CH3O+NO2; 
*---Ethyl nitrate (1_02);
* RC2H6NO31 % 2.55D-12*EXP(380/T)*0.009: C2H5O2+NO = C2H5NO3;
* RC2H6NO32 % 6.7D-13*EXP(-395/T)      : OH+C2H5NO3 = CH3CHO+NO2;
* RC2H6NO33 % J<52>                    : C2H5NO3 = C2H5O+NO2;
*---Propyl nitrates (1_03);
* RC3H7NO31 % 2.7D-12*EXP(360/T)*0.042 : IC3H7O2+NO = IC3H7NO3;
* RC3H7NO32 % 6.2D-13*EXP(-230/T)      : OH+IC3H7NO3 = CH3COCH3+NO2;
* RC3H7NO33 % J<54>                    : IC3H7NO3 = IC3H7O+NO2;
* RC3H7NO34 % 2.9D-12*EXP(350/T)*0.020 : NC3H7O2+NO = NC3H7NO3;
* RC3H7NO35 % 5.8D-13                  : OH+NC3H7NO3 = C2H5CHO+NO2;
* RC3H7NO36 % J<53>                    : NC3H7NO3 = NC3H7O+NO2;
*---n-butyl nitrates (1_04);
* RNC4H10N1 % KRO2NO*0.033             : NC4H9O2+NO = NC4H9NO3;
* RNC4H10N2 % KRO2NO*0.090             : SC4H9O2+NO = SC4H9NO3;
* RNC4H10N3 % 1.6D-12                  : OH+NC4H9NO3 = C3H7CHO+NO2;
* RNC4H10N4 % 8.6D-13                  : OH+SC4H9NO3 = MEK+NO2;
* RNC4H10N5 % J<53>                    : NC4H9NO3 = NC4H9O+NO2;
* RNC4H10N6 % J<54>                    : SC4H9NO3 = SC4H9O+NO2;
*---i-butyl nitrates;
* RIC4H10N1 % KRO2NO*0.033             : IC4H9O2+NO = IC4H9NO3;
* RIC4H10N2 % KRO2NO*0.025             : TC4H9O2+NO = TC4H9NO3;
* RIC4H10N3 % 7.69D-13                 : OH+IC4H9NO3 = IPRCHO+NO2;
* RIC4H10N4 % 8.16D-14                 : OH+TC4H9NO3 = CH3COCH3+HCHO+NO2;
* RIC4H10N5 % J<53>                    : IC4H9NO3 = IC4H9O+NO2;
* RIC4H10N6 % J<55>                    : TC4H9NO3 = TC4H9O+NO2;
*---n-pentyl nitrates (1_05);
* RNC5H12N1 % KRO2NO*0.052             : PEAO2+NO = PEANO3;
* RNC5H12N2 % KRO2NO*0.129             : PEBO2+NO = PEBNO3;
* RNC5H12N3 % KRO2NO*0.131             : PECO2+NO = PECNO3;
* RNC5H12N4 % 3.00D-12                 : OH+PEANO3 = C4H9CHO+NO2;
* RNC5H12N5 % 1.85D-12                 : OH+PEBNO3 = MPRK+NO2;
* RNC5H12N6 % 1.12D-12                 : OH+PECNO3 = DIEK+NO2;
* RNC5H12N7 % J<53>                    : PEANO3 = PEAO+NO2;
* RNC5H12N8 % J<54>                    : PEBNO3 = PEBO+NO2;
* RNC5H12N9 % J<54>                    : PECNO3 = PECO+NO2;
*---i-pentyl nitrates;
RIC5H12N1 % KRO2NO*0.052             : IPEAO2+NO = IPEANO3;
RIC5H12N2 % KRO2NO*0.141             : IPEBO2+NO = IPEBNO3;
RIC5H12N3 % KRO2NO*0.047             : IPECO2+NO = IPECNO3;
RIC5H12N4 % 2.01D-12                 : OH+IPEANO3 = BUT2CHO+NO2;
RIC5H12N5 % 1.82D-12                 : OH+IPEBNO3 = MIPK+NO2;
RIC5H12N6 % 4.08D-13                 : OH+IPECNO3 = CH3COCH3+CH3CHO+NO2;
RIC5H12N7 % J<53>                    : IPEANO3 = IPEAO+NO2;
RIC5H12N8 % J<54>                    : IPEBNO3 = IPEBO+NO2;
RIC5H12N9 % J<55>                    : IPECNO3 = IPECO+NO2;
\end{verbatim}

### 2020-04-06 figures/o3_chemistry_scheme/o3_chemistry_scheme_full.tex
re-structure msc thesis code

%\documentclass[]{standalone}
%\usepackage[utf8]{inputenc}
%\usepackage[english]{babel}
%\usepackage{amsmath}
%\usepackage{amsfonts}
%\usepackage{amssymb}
%\usepackage{eulervm}
%
%\usepackage{graphicx}
%
%% Figures
%\usepackage{tikz}
%\usetikzlibrary{shapes,arrows}
%\usetikzlibrary{positioning}
%\usetikzlibrary{calc}
%
%\begin{document}

% Comment out everything above before including this .tex file via \input
% and \end{document} at the end

% Tikz styles
\tikzstyle{ar} = [->,draw=black!70, line width=1]
\tikzstyle{ar_decis} = [-|,draw=black!70, line width=1]
\tikzstyle{ar_red} = [->,draw=red, line width=1]
\tikzstyle{ln} = [-,draw=black!70, line width=1]
\tikzstyle{ln_dash} = [-,dashed,draw=black!70, line width=1]
\tikzstyle{bord} = [rounded rectangle, draw=black]

% Set a length that is relative to \linewidth
\newlength{\scl} \setlength{\scl}{0.08\linewidth}

% Begin Tikz picture
\begin{tikzpicture}[node distance = 4cm, auto]

% Nodes
\node (o3hv) at (0,0) {$O_3 + h\nu$};
\node[bord] (oh_1) at (2\scl, 0) {$OH$};
\node (h2o_1) at ($(o3hv)!.5!(oh_1)+(0,-0.25\scl)$) {$H_2O$};
\node [align=center] (plus_rh) at (2\scl, -0.7\scl) {$+$\\$RH$};
\node (r) at (2\scl, -2\scl) {$R$};
\node (plus_o2) at (2.5\scl, -2.5\scl) {$+O_2$};
\node (ro2) at (2\scl, -3\scl) {$RO_2$};
\node (ro) at (2\scl, -5\scl) {$RO$};
\node (ho2) at (0, -3\scl) {$HO_2$};
\node (rooh) at (0, -5\scl) {$ROOH$};
\node (foo1) at (1\scl, -4\scl) {};
\node (foo2) at (3\scl, -4\scl) {};
\node (foo3) at (3.5\scl, -5\scl) {};
\node (alpha3) at (3.5\scl, -4.75\scl) {$\alpha_3$};
\node (no_1) at (4.25\scl, -3\scl) {$NO$};
\node (no2_1) at (4.25\scl, -5\scl) {$NO_2$};
\node (rono2) at (4.5\scl, -6\scl) {$RONO_2$};
\node (foo4) at (5.5\scl, -4\scl) {};
\node (o3_1) at (6.5\scl, -3\scl) {$O_3$};
\node (hvo2_1) at (6.25\scl, -3.5\scl) {$h\nu + O_2$};

\node (ho2_2) at (2\scl, -7\scl) {$HO_2$};
\node[bord] (oh_2) at (2\scl, -9\scl) {$OH$};
\node (o3_2) at (0, -7\scl) {$O_3$};
\node (o2) at (0, -9\scl) {$O_2$};
\node (foo5) at (1\scl, -8\scl) {};
\node (foo9) at (3\scl, -8\scl) {};
\node (no_2) at (4.25\scl, -7\scl) {$NO$};
\node (no2_2) at (4.25\scl, -9\scl) {$NO_2$};
\node (foo10) at (5.5\scl, -8\scl) {};
\node (o3_3) at (6.5\scl, -7\scl) {$O_3$};
\node (hvo2_2) at (6.25\scl, -7.5\scl) {$h\nu + O_2$};

\node (plus_o2_2) at (2.5\scl, -6\scl) {$+O_2$};
\node (foo6) at (1.95\scl, -6\scl) {};
\node (foo7) at (1.5\scl, -6.25\scl) {};
\node (foo8) at (1.5\scl, -9.25\scl) {};
\node (rcho) at (2\scl, -10\scl) {$RCHO$};
\node (hv) at (2.25\scl, -10.5\scl) {$h\nu$};
\node (co) at (2\scl, -11\scl) {$CO~(+2HO_2)$};
\node (co2) at (2\scl, -12\scl) {$CO_2$};

\node (oz_loss) at (0\scl, -12\scl) {NET OZONE LOSS};
\node (oz_gain) at (4\scl, -12\scl) {NET OZONE GAIN};

% Connections
\draw[ar] (o3hv.east) to [out=0,in=180](oh_1.west);
\draw[ar] (plus_rh.south) to [out=270,in=90](r.north);
\draw[ar] (r.south) to [out=270,in=90](ro2.north);
\draw[ln] (ho2.east) to [out=0,in=90](foo1.center);
\draw[ln] (ro2.west) to [out=180,in=90](foo1.center);
\draw[ar] (foo1.center) to [out=270,in=45](rooh.north east);
\draw[ar] (rooh.east) to [out=0,in=180](ro.west);
\draw[ln] (ro2.east) to [out=0,in=90](foo2.center);
\draw[ar] (foo2.center) to [out=270,in=0](ro.east);
\draw[ln] (no_1.west) to [out=180,in=90](foo2.center);
\draw[ar_decis] (foo2.center) to [out=270,in=180](foo3.center);
\draw[ar_red] (foo3.center) to [out=0,in=180](no2_1.west);
\draw[ar_red] (foo3.center) to [out=0,in=180](rono2.west);
\draw[ln,red, dashed] (no2_1.east) to [out=0,in=270](foo4.center);
\draw[ar_red, dashed] (foo4.center) to [out=90,in=0](no_1.east);
\draw[ar_red, dashed] (foo4.center) to [out=90,in=180](o3_1.west);

\draw[ar] (ro.south) to [out=270,in=90](ho2_2.north);
\draw[ln] (o3_2.east) to [out=0,in=90](foo5.center);
\draw[ln] (ho2_2.west) to [out=180,in=90](foo5.center);
\draw[ar] (foo5.center) to [out=270,in=0](o2.east);
\draw[ar] (foo5.center) to [out=270,in=180](oh_2.west);
\draw[ln] (ho2_2.east) to [out=0,in=90](foo9.center);
\draw[ar] (foo9.center) to [out=270,in=0](oh_2.east);
\draw[ln] (no_2.west) to [out=180,in=90](foo9.center);
\draw[ar] (foo9.center) to [out=270,in=180](no2_2.west);
\draw[ln] (no2_2.east) to [out=0,in=270](foo10.center);
\draw[ln] (foo10.center) to [out=90,in=0](no_2.east);
\draw[ar] (foo10.center) to [out=90,in=180](o3_3.west);

\draw[ln] (foo6.center) to [out=180,in=90](foo7.center);
\draw[ln_dash] (foo7.center) to [out=270,in=90](foo8.center);
\draw[ar] (foo8.center) to [out=270,in=90](rcho.north);
\draw[ar] (rcho.south) to [out=270,in=90](co.north);
\draw[ar] (co.south) to [out=270,in=90](co2.north);
\draw[ar] (co.east) to [out=0,in=315](oh_2.south east);

\end{tikzpicture}

%\end{document}

### 2020-04-06 msc_thesis.tex
re-structure msc thesis code

--- 2015-08-06 MSc_diss.tex
@@
 % Paper size and margins
-%\usepackage{rotating}
 \usepackage{pdflscape}
-%\usepackage[a4paper,showframe]{geometry}
 \usepackage{vmargin}
@@
 \linespread{1.5}
-%\usepackage{setspace}
 \usepackage{indentfirst}
-%\setlength{\parindent}{1cm}
 \setlength{\parskip}{0.5cm}
@@
 % Equations
+\usepackage{mathtools} % for arrows
 \usepackage{eulervm} % upright font in eqns (Palatino)
 \usepackage{cool}
-\usepackage{mathtools} % for arrows
-
 \newcounter{matriz}
@@
 \renewcommand{\theequation}{R\arabic{equation}}
-
 % References
@@
 % Hyperlinks and labels
-%\usepackage{showkeys}
 \usepackage[linktocpage=true,plainpages=false,pdfpagelabels=false]{hyperref}
@@
 \newcommand{\depname}{School of Environmental Sciences}
-\newcommand{\mydate}{\the\year}
-
-\newcommand{\picspath}{D:/UEA/ENV-MB4X_Dissertation/ENV-MB4X_Dissertation_thesis/ENV-MB4X_Dissertation_thesis_latex/pics/}
-%% Remove author and date from the maketitle command
-%\makeatletter
-%\renewcommand{\@maketitle}{
-%\newpage
-% \null
-% \vskip 2em%
-% \begin{center}%
-%  {\LARGE \@title \par}%
-% \end{center}%
-% \par} \makeatother
-
+\newcommand{\mydate}{2015}
+\newcommand{\figspath}{figures/}
 % PDF meta-data
@@
 \title{\mytitle}
-% =========================================================================
+
 \begin{document}
 
-\input{diss_titlepage}
-%\onehalfspacing
+\input{titlepage}
 
@@
 \tableofcontents
+
 \newpage
-
 \section{Introduction} \label{sec:intro}
@@
 This capacity depends mainly on two factors: climate and the global burden of the principal oxidants such as ozone ($O_3$), the hydroxyl radical ($OH$) and hydrogen peroxide ($H_2O_2$) \citep{Prinn2003,Thompson1992}.
-\subsection{Tropospheric ozone}\label{sec:into_O3}
+
+\subsection{Tropospheric ozone} \label{sec:intro_O3}
 Tropospheric ozone ($O_3$) plays an important role in the chemical and thermal balance of the atmosphere.
@@
 Unlike many other air pollutants, tropospheric ozone is not directly emitted, but produced following the oxidation of carbon monoxide ($CO$), methane ($CH_4$), and nonmethane volatile compounds (nmVOCs) in the presence of nitrogen oxides ($NO_x$) \citep{Crutzen1973,Myhre2013}.
-The emissions of these precursors arise from both natural and anthropogenic sources which have intensified since pre-industrial times \citep{Parrish2014,Volz1988} (Figure \ref{fig:O3observations}).
+The emissions of these precursors arise from both natural and anthropogenic sources which have intensified since pre-industrial times \citep{Parrish2014,Volz1988} (Figure \ref{fig:Isaksen2009_fig15}).
 Presumably it has led to roughly a doubling in background tropospheric ozone concentration \citep{Guicherit2000,Hartmann2013,Tarasick2008,Vingarzan2004}, although contribution from another major source, transport from the stratosphere, may also took place \citep{Fowler2008}.
@@
 \centering
-\includegraphics[width=\linewidth]{{./pics/Isaksen2009_O3}.png}
+\includegraphics[width=\linewidth]{\figspath Isaksen2009_fig15.png}
 \caption{Observed surface ozone at different Northern Hemispheric surface stations \citep{Isaksen2009}.}
-\label{fig:O3observations}
+\label{fig:Isaksen2009_fig15}
 \end{figure}
@@
 However, despite general agreement on how the drivers impact ozone concentrations at global and regional scales, there is considerable inconsistency in its budget between different modelling and observational studies \citep{Stevenson2006,Wild2007,Young2012}.
-Figure \ref{fig:O3budget} demonstrates that the range of global tropospheric ozone budget estimates is approximately 60 Tg which is 18\% of the mean.
+Figure \ref{fig:o3_budget_studies} demonstrates that the range of global tropospheric ozone budget estimates is approximately 60 Tg which is 18\% of the mean.
 In terms of ozone production the variation is even higher and accounts for 28\% of the corresponding value.
@@
 
-\begin{figure}[t] % O3 budget
-\centering
-\includegraphics[width=0.8\linewidth]{o3_studies.eps}
+\begin{figure}[t]
+\centering
+\includegraphics[width=0.8\linewidth]{\figspath /o3_budget_studies/o3_budget_studies.pdf}
 \caption{Summary of tropospheric ozone global budget model and observation estimates for present (about 2000) conditions.
 All uncertainties quoted as 1 standard deviation (68\% confidence interval).
-Adapted from \citep{Myhre2013}}\label{fig:O3budget}
+Adapted from \citep{Myhre2013}.}
+\label{fig:o3_budget_studies}
 \end{figure}
@@
 
-\subsubsection*{Ozone tropospheric chemistry}\label{intro_O3chem}
+\subsubsection*{Ozone tropospheric chemistry} \label{intro_O3chem}
 Keeping in mind the role of tropospheric ozone, a necessary step is to revisit the chemistry that controls its abundance.
@@
 At UV wavelengths shorter than 320 nm, $O_3$ photolysis generates excited oxygen atoms ($O(^1D)$), which than can either react with water vapour to form $OH$ radicals (\ref{reac:O1D+H2O=OH+OH}), or collide with an inert molecule (denoted '$M$') to reform $O_3$ (\ref{reac:O1D=O3P} and \ref{reac:O3P+O2=O3}):
-\begin{equation}\label{reac:O3+hv=O1D+O2}
+\begin{equation} \label{reac:O3+hv=O1D+O2}
 O_3 + hv \rightarrow O(^1D) + O_2
 \end{equation}
-\begin{equation}\label{reac:O1D+H2O=OH+OH}
+\begin{equation} \label{reac:O1D+H2O=OH+OH}
 O(^1D) + H_2O \rightarrow OH + OH
 \end{equation}
-\begin{equation}\label{reac:O1D=O3P}
+\begin{equation} \label{reac:O1D=O3P}
 O(^1D) \xrightarrow{M} O(^3P)
 \end{equation}
-\begin{equation}\label{reac:O3P+O2=O3}
+\begin{equation} \label{reac:O3P+O2=O3}
 O(^3P) + O_2 \xrightarrow{M} O_3
@@
 Once formed the OH radicals mainly react with $CO$ and $CH_4$ and initiate the $O_3$ formation or removal cycles by producing the peroxy radicals ($HO_2$, $CH_3O_2$):
-\begin{equation}\label{reac:OH+CO=H+CO2}
+\begin{equation} \label{reac:OH+CO=H+CO2}
 OH + CO \rightarrow H + CO_2
 \end{equation}
-\begin{equation}\label{reac:H+O2=HO2}
+\begin{equation} \label{reac:H+O2=HO2}
 H + O_2 \xrightarrow{M} HO_2
 \end{equation}
-\begin{equation}\label{reac:OH+CH4=CH3+H2O}
+\begin{equation} \label{reac:OH+CH4=CH3+H2O}
 OH + CH_4 \rightarrow CH_3 + H_2O
 \end{equation}
-\begin{equation}\label{reac:CH3O+O2=CH3O2}
+\begin{equation} \label{reac:CH3O+O2=CH3O2}
 CH_3 + O_2 \xrightarrow{M} CH_3O_2
@@
 Plus in the same conditions some additional $O_3$ removal can occur because of the reaction of the $HO_2$ radical with $O_3$ leading to further destruction of ozone in a chain sequence involving formation of $OH$ radicals:
-\begin{equation}\label{reac:HO2+O3=OH+2O2}
+\begin{equation} \label{reac:HO2+O3=OH+2O2}
 HO_2 + O_3 \rightarrow OH + 2O_2
 \end{equation}
-\begin{equation}\label{reac:OH+O3=HO2+O2}
+\begin{equation} \label{reac:OH+O3=HO2+O2}
 OH + O_3 \rightarrow HO_2 + O_2
@@
 In intermediate $NO_x$ conditions (rural areas of most industrialised countries), the peroxy radicals react with nitrogen monoxide ($NO$) to form nitrogen dioxide ($NO_2$), which subsequent photolysis leads to $O_3$ generation:
-\begin{equation}\label{reac:HO2+NO=NO2+OH}
+\begin{equation} \label{reac:HO2+NO=NO2+OH}
 HO_2 + NO \rightarrow NO_2 + OH
 \end{equation}
-\begin{equation}\label{reac:CH3O2+NO=NO2+CH3O}
+\begin{equation} \label{reac:CH3O2+NO=NO2+CH3O}
 CH_3O_2 + NO \rightarrow NO_2 + CH_3O
 \end{equation}
-\begin{equation}\label{reac:NO2+hv=NO+O3P}
+\begin{equation} \label{reac:NO2+hv=NO+O3P}
 NO_2 + hv \rightarrow NO + O(^3P)
@@
 \end{equation}
-As shown in Figure \ref{fig:O3scheme}, the reactions (\ref{reac:HO2+NO=NO2+OH}) and (\ref{reac:CH3O2+NO=NO2+CH3O}) produce parts of the free-radical propagated $O_3$ forming cycles, which may occur a number of times before being halted by radical termination reactions (\ref{reac:HO2+HO2=H2O2+O2} and \ref{reac:HO2+CH3O2=CH3OOH+O2}), and the higher $NO_x$ level is the greater number of these cycles can occur.
+As shown in Figure \ref{fig:o3_chemistry_scheme}, the reactions (\ref{reac:HO2+NO=NO2+OH}) and (\ref{reac:CH3O2+NO=NO2+CH3O}) produce parts of the free-radical propagated $O_3$ forming cycles, which may occur a number of times before being halted by radical termination reactions (\ref{reac:HO2+HO2=H2O2+O2} and \ref{reac:HO2+CH3O2=CH3OOH+O2}), and the higher $NO_x$ level is the greater number of these cycles can occur.
 Therefore, in intermediate $NO_x$ conditions a net $O_3$ formation takes place and it is sensitive to $NO_x$.
@@
 \centering
-\input{diss_NewO3scheme}
-\caption{$O_3$ chemistry scheme}\label{fig:O3scheme}
+\includegraphics{\figspath /o3_chemistry_scheme/o3_chemistry_scheme.pdf}
+\caption{$O_3$ chemistry scheme.}
+\label{fig:o3_chemistry_scheme}
 \end{figure}
@@
 Depending on the concentration of $NO_2$ it can react with it to form nitric acid ($HNO_3$):
-\begin{equation}\label{reac:OH+NO2=HNO3}
+\begin{equation} \label{reac:OH+NO2=HNO3}
 OH + NO_2 \xrightarrow{M} HNO_3
@@
 Its photolysis and further oxidation contributes to radical generation:
-\begin{equation}\label{reac:HCHO+hv=H+HCO}
+\begin{equation} \label{reac:HCHO+hv=H+HCO}
 HCHO + hv \rightarrow H + HCO
 \end{equation}
-\begin{equation}\label{reac:HCO+O2=HO2+CO}
+\begin{equation} \label{reac:HCO+O2=HO2+CO}
 HCO + O_2 \rightarrow HO_2 + CO
@@
 \end{equation}
-
 therefore also having an impact on the rates of $O_3$ production \citep{Fowler2008} as do some other species, which are not mentioned in this section.
@@
 
-\subsection{Alkyl nitrates}\label{sec:intro_AN}
+\subsection{Alkyl nitrates} \label{sec:intro_AN}
 Alkyl nitrates ($RONO_2$) are organic trace gases, which through one of their formation mechanisms are closely linked with the production of tropospheric ozone \citep{Reeves2007}.
 This mechanism is presented in the following section.
-\subsubsection*{Alkyl nitrates chemistry}\label{into_ANchem}
+
+\subsubsection*{Alkyl nitrates chemistry} \label{into_ANchem}
 The photochemical formation of the alkyl nitrates starts when chemistry of the troposphere reaches the regime of $O_3$ production, i.e.
@@
 \end{subequations}
-The latter reaction can serve as a termination step of the propagated $O_3$ forming cycle (Figure \ref{fig:O3scheme}), and as \cite{Roberts1998} estimated it is likely to begin to occur when the $NO$ mixing ratio increases up to 100-200 ppt.
+The latter reaction can serve as a termination step of the propagated $O_3$ forming cycle (Figure \ref{fig:o3_chemistry_scheme}), and as \cite{Roberts1998} estimated it is likely to begin to occur when the $NO$ mixing ratio increases up to 100-200 ppt.
 The yield of alkyl nitrate formation is controlled by the branching ratio ($\alpha_3$), which is temperature and pressure dependent and increases with the number of carbons in the alkyl peroxy radical \citep{Roberts1990}.
@@
 An alternative photochemical pathway of the formation of alkyl nitrates is the reaction of the alkoxy radical ($RO$) with $NO_2$:
-\begin{equation}\label{reac:RO+NO2=RONO2}
+\begin{equation} \label{reac:RO+NO2=RONO2}
 RO + NO_2 \rightarrow RONO_2
@@
 
-\subsection{Ozone, alkyl nitrates and carbon bonds}\label{sec:intro_CB}
+\subsection{Ozone, alkyl nitrates and carbon bonds} \label{sec:intro_CB}
 In the previous sections we covered the chemistry that controls the production of $O_3$ in the troposphere. A traditional way to assess the efficiency of this production is to calculate the ratio:
-\begin{matriz}\label{eq:OPE}
+\begin{matriz} \label{eq:OPE}
 OPE = \dfrac{k_{HO_2+NO}[HO_2][NO]}{k_{OH+NO_2}[OH][NO_2]} = \dfrac{P_{O_3}}{L_{NO_x}}
@@
 Assuming that every carbon bond is broken chemically and every organic peroxy radical ($RO_2$) reacts with $NO$, the overall efficiency of carbon bonds conversion to $O_3$, $\epsilon$, can be calculated as:
-\begin{matriz}\label{eq:epsilon}
+\begin{matriz} \label{eq:epsilon}
 \epsilon = \dfrac{P_{O_3}}{E_{C_{bons}}} = \chi\cdot\upsilon
@@
 
-\section{Overall objective and specific aims}\label{sec:objams}
+\section{Overall objective and specific aims} \label{sec:objams}
 Despite of all of the aforesaid, typically chemical mechanisms of global atmospheric models do not contain enough information about some secondary organic compounds \citep{Sommariva2008}, and alkyl nitrates are among them.
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Literature review} \label{sec:lit_review}
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \renewcommand{\arraystretch}{1}
 \begin{table}[h!]
-\caption{Mean mixing ratios of the sum of alkyl nitrates ($\Sigma RONO_2$) measured in different environments.
-The measurement area is indicated in parentheses, type of measurements is in brackets.}\label{tab:ANmean} % mean AN mixing ratios
+\caption{Mean mixing ratios of the sum of alkyl nitrates ($\Sigma RONO_2$) measured in different environments. The measurement area is indicated in parentheses, type of measurements is in brackets.}
+\label{tab:ANmean} % mean AN mixing ratios
 \begin{tabularx}{\linewidth}{P{3cm}P{2cm}P{5cm}X}
@@
 \end{table}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \begin{figure}[h!]
 \centering
-\includegraphics[width=0.5\linewidth]{{\picspath Neuman2012_Fig.4}.png}
+\includegraphics[width=0.5\linewidth]{\figspath Neuman2012_fig4.png}
 \caption{Time series of reactive nitrogen species measured on 10 June 2010 in the MBL in plumes (a) 10 km and (b) 50 km downwind from DWH.
@@
 The sum of alkyl nitrates (black lines) is inferred from the difference between $NO_y$ and ($NO_x$+$HNO_3$+$PAN$).
-Source: \citep{Neuman2012}.}\label{fig:Neuman2012}
-\end{figure}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+Source: \citep{Neuman2012}.}
+\label{fig:Neuman2012_fig4}
+\end{figure}
 
@@
 
-Temporal changes in reactive nitrogen partitioning in the DWH plume (Figure \ref{fig:Neuman2012}) nicely illustrate that a quite large concentration of alkyl nitrates in the air can be build up under certain conditions.
-While in fresh part of the plume (Figure \ref{fig:Neuman2012}a) $NO_x$ accounted for over 80\% of plume $NO_y$, in aged plumes (Figure \ref{fig:Neuman2012}b) $NO_x$ oxidation was nearly complete, but $NO_x$, $PAN$, $HNO_3$, and the sum of eight $C_5$–$C_{10}$ $RONO_2$ accounted for only 22\% of plume $NO_y$.
+Temporal changes in reactive nitrogen partitioning in the DWH plume (Figure \ref{fig:Neuman2012_fig4}) nicely illustrate that a quite large concentration of alkyl nitrates in the air can be build up under certain conditions.
+While in fresh part of the plume (Figure \ref{fig:Neuman2012_fig4}a) $NO_x$ accounted for over 80\% of plume $NO_y$, in aged plumes (Figure \ref{fig:Neuman2012_fig4}b) $NO_x$ oxidation was nearly complete, but $NO_x$, $PAN$, $HNO_3$, and the sum of eight $C_5$–$C_{10}$ $RONO_2$ accounted for only 22\% of plume $NO_y$.
 It means that $C_6$–$C_{10}$ $RONO_2$ accounted for 78\% of the aged plume $NO_y$ \citep{Neuman2012}.
@@
 The concentration of $NO_x$ remains relatively high during the second (10–-15 ppb) and third day (3-–5 ppb) and then decreases to $<$200 ppt of $NO$ and $<$1ppb of $NO_2$ on the fourth and fifth days, therefore allowing the formation of alkyl nitrates too.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \begin{table}
-\caption{Initial values of some model parameters (in ppb) for two scenarios used in \citep{Sommariva2008}}\label{tab:Sommariva2008init}
+\caption{Initial values of some model parameters (in ppb) for two scenarios used in \citep{Sommariva2008}}
+\label{tab:Sommariva2008init}
 \centering
@@
 \hline
-Parameter & Scenario A & Scenario B \\
-\hline
-$CO$      & 332.5 & 232.5 \\
-$NO_x$    & 34.9  & 46.5  \\
-$O_3$     & 90    & 0   \\
+Parameter & Scenario A & Scenario B\\
+\hline
+$CO$      & 332.5      & 232.5\\
+$NO_x$    & 34.9       & 46.5\\
+$O_3$     & 90         & 0\\
 \hline
@@
 \begin{figure}
-\includegraphics[width=\linewidth]{{D:/UEA/ENV-MB4X_Dissertation/ENV-MB4X_Dissertation_thesis/ENV-MB4X_Dissertation_thesis_latex/pics/Sommariva2008_Fig.1}.png}
-\caption{Modelled concentrations of $O_3$, $NO_x$ and $HNO_3$ for scenario A (left) and B (right). Source: \citep{Sommariva2008}.}\label{fig:Sommariva2008fig}
-\end{figure}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\includegraphics[width=\linewidth]{\figspath Sommariva2008_fig1.png}
+\caption{Modelled concentrations of $O_3$, $NO_x$ and $HNO_3$ for scenario A (left) and B (right). Source: \citep{Sommariva2008}.}
+\label{fig:Sommariva2008fig}
+\end{figure}
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \begin{figure}[t!]
@@
         \centering
-        \includegraphics[width=\linewidth]{{\picspath Neu2008_Fig.2_MeONO2_OCT_merge}.png}
-        \caption{}\label{fig:Neu2008M_MeONO2}
+        \includegraphics[width=\linewidth]{\figspath Neu2008_fig2_oct_MeONO2.png}
+        \caption{}
+        \label{fig:Neu2008M_MeONO2}
     \end{subfigure}%
@@
         \centering
-        \includegraphics[width=\linewidth]{{\picspath Neu2008_Fig.3_OCT_O3contol}.png}
-        \caption{}\label{fig:Neu2008M_O3contol}
+        \includegraphics[width=\linewidth]{\figspath Neu2008_fig3_oct_O3contol.png}
+        \caption{}
+        \label{fig:Neu2008M_O3contol}
     \end{subfigure}
@@
         \centering
-        \includegraphics[width=\linewidth]{{\picspath Neu2008_Fig.3_OCT_NOxdiff}.png}
-        \caption{}\label{fig:Neu2008M_NOxdiff}
+        \includegraphics[width=\linewidth]{\figspath Neu2008_fig3_oct_NOxdiff.png}
+        \caption{}
+        \label{fig:Neu2008M_NOxdiff}
     \end{subfigure}%
@@
         \centering
-        \includegraphics[width=\linewidth]{{\picspath Neu2008_Fig.3_OCT_O3diff}.png}
-        \caption{}\label{fig:Neu2008M_O3diff}
+        \includegraphics[width=\linewidth]{\figspath Neu2008_fig3_oct_O3diff.png}
+        \caption{}
+        \label{fig:Neu2008M_O3diff}
     \end{subfigure}
@@
 All averaged over the boundary layer (below 850 mb).
-Adapted from \citep{Neu2008}}\label{fig:Neu2008}
-\end{figure}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+Adapted from \citep{Neu2008}}
+\label{fig:Neu2008}
+\end{figure}
 
@@
 \section{Methodology} \label{sec:method}
-\subsection{Model development and description}\label{sec:method_develop}
+
+\subsection{Model development and description} \label{sec:method_develop}
 A zero-dimensional (box) model is developed to investigate the impact of alkyl nitrate chemistry on ozone production.
@@
 
-\subsection{Chemical mechanism}\label{sec:method_chem}
+\subsection{Chemical mechanism} \label{sec:method_chem}
 The chemical mechanism of the model is based on recommendations from \citep{Atkinson2004} and information from the Master Chemical Mechanism, MCM v3.3 \citep{Jenkin1997,Saunders2003}, retrieved via website: \href{http://mcm.leeds.ac.uk/MCM}{http://mcm.leeds.ac.uk/MCM}.
@@
 Despite that halogens (chlorine ($Cl$), bromine ($Br$) and iodine ($I$)) are very reactive chemical compounds mostly known by their role in anthropogenic stratospheric ozone-depletion chemistry, and that they are directly (for example, reaction of the $Cl$ radical with hydrocarbons (e.g., $CH_4$)) and indirectly (coupling with the cycles of sulfur and nitrogen that eventually influence $O_3$, $OH$ and $NO_x$ concentrations) involved in the tropospheric chemistry \citep{VonGlasow2014,Platt2003}, they are excluded from the chemical mechanism to keep it more transparent.
-\subsection{Photolysis}\label{sec:method_photolysis}
+
+\subsection{Photolysis} \label{sec:method_photolysis}
 To estimate photolysis rates for a given geographical location, one must take into account the latitude and season, as well as the time of the day.
@@
 
-\subsection{Design of experiments}\label{sec:method_expdesign}
+\subsection{Design of experiments} \label{sec:method_expdesign}
 To investigate the influence of alkyl nitrate chemistry on tropospheric $O_3$ production, box model runs were performed at a series of different $NO_x$ and VOCs initial concentrations, so that an $O_3$ isopleth plot could be constructed, similar to those found in \citep{Dodge1977} and \citep{Sillman1999}.
@@
 \begin{table} % model setup: O3, CO
-\caption{Fixed and variable main parameters used in all model runs}\label{tab:setupO3CO}
+\caption{Fixed and variable main parameters used in all model runs}
+\label{tab:setupO3CO}
 \centering
@@
 \begin{table} % model setup: NOx
-\caption{Model setup: NOx initial concentrations (ppb)}\label{tab:setupNOx}
+\caption{Model setup: NOx initial concentrations (ppb)}
+\label{tab:setupNOx}
 \centering
@@
 \begin{table} % model setup: VOCs
-\caption{Model setup: VOCs concentrations (ppb)}\label{tab:setupVOCs}
+\caption{Model setup: VOCs concentrations (ppb)}
+\label{tab:setupVOCs}
 \centering
@@
 \begin{table} % RH and ANs
-\caption{Alkyl nitrates and their parent alkanes}\label{tab:RHandANs}
+\caption{Alkyl nitrates and their parent alkanes}
+\label{tab:RHandANs}
 \centering
@@
 \end{equation}
-\begin{equation}\label{reac:OH+RH=RO2+H2O}
+\begin{equation} \label{reac:OH+RH=RO2+H2O}
 OH + RH \rightarrow RO_2 + H_2O
@@
 \end{equation}
-\begin{equation}\label{reac:RO2+HO2=ROOH+O2}
+\begin{equation} \label{reac:RO2+HO2=ROOH+O2}
 RO_2 + HO_2 \rightarrow ROOH + O_2
@@
 
-\input{diss_valid_vs_time}
+\input{figures/model_validation/model_validation}
 
@@
 \end{equation}
-\begin{equation}\label{reac:RO2+NO=RO+NO2}
+\begin{equation} \label{reac:RO2+NO=RO+NO2}
 RO_2 + NO \rightarrow RO + NO_2
@@
 Above the 'ridge line' the VOC/$NO_x$ ratios are low, and the $O_3$ production rate increases with increasing VOC and decreases with increasing $NO_x$ \citep{Sillman2013}.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \begin{figure}[t!]
@@
         \centering
-        \includegraphics[width=\linewidth]{{./pics/Sillman1999_Fig1_adapted_withdotsrectangle}.png}
-        \caption{for mean summer daytime meteorology and clear skies.
-The solid lines represent production rates of 1, 2.5, 5, 10, 15, 20 and 30 ppb/h.
-Adapted from \citep{Sillman1999}.} \label{fig:Sillman1999}
+        \includegraphics[width=\linewidth]{\figspath Sillman1999_fig1_adapted_withdotsrectangle.png}
+        \caption{for mean summer daytime meteorology and clear skies. The solid lines represent production rates of 1, 2.5, 5, 10, 15, 20 and 30 ppb/h. Adapted from \citep{Sillman1999}.}
+        \label{fig:Sillman1999}
     \end{subfigure}%
@@
         \centering
-        \includegraphics[width=\linewidth]{D:/FACSIMILE/ANsCBmodel/matlab/ANsCB_pics/ANsNOxVOC/noAN/chem_noAN_netO3rate_withINORG_NOx_1ppb_100ppb.eps}
-        \caption{from the first series of experiments with alkyl nitrate chemistry switched off combined with an additional run with inorganic chemistry only (i.e.
-the only VOC is $CO$).} \label{fig:noAN_netO3rate}
+        \includegraphics[width=\linewidth]{\figspath facsimile_results/chem_noAN_netO3rate_withINORG_NOx_1ppb_100ppb.png}
+        \caption{from the first series of experiments with alkyl nitrate chemistry switched off combined with an additional run with inorganic chemistry only (i.e. the only VOC is $CO$).}
+        \label{fig:noAN_netO3rate}
     \end{subfigure}
-    \caption{Isopleths of net rate of $O_3$ production (ppb/h) as a function of VOC (ppbC) and $NO_x$ (ppb).}\label{fig:netO3rate}
-\end{figure}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+    \caption{Isopleths of net rate of $O_3$ production (ppb/h) as a function of VOC (ppbC) and $NO_x$ (ppb).}
+    \label{fig:netO3rate}
+\end{figure}
 
@@
 \section{Results} \label{sec:res}
-\subsection{Sensitivity of $O_3$ production to alkyl nitrate chemistry}\label{sec:res_O3ANsensetivity}
+
+\subsection{Sensitivity of $O_3$ production to alkyl nitrate chemistry} \label{sec:res_O3ANsensetivity}
+
 \subsubsection{$O_3$ and total alkyl nitrates}
@@
 Branching ratios used in the model are taken from the MCM v3.3, and their values (Table \ref{tab:ANbranching}) were kept constant in all model runs.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \begin{table}[h!] %  AN branching ratios
-\caption{Branching ratios for the formation of alkyl nitrates from their precursor peroxy radicals and $NO$ (taken from the MCM v3.3 (2015)).}\label{tab:ANbranching}
+\caption{Branching ratios for the formation of alkyl nitrates from their precursor peroxy radicals and $NO$ (taken from the MCM v3.3 (2015)).}
+\label{tab:ANbranching}
 \centering
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \begin{figure}[h!] % netO3rate noAN withAN diff
@@
         \centering
-        \includegraphics[width=\linewidth]{D:/FACSIMILE/ANsCBmodel/matlab/ANsCB_pics/ANsNOxVOC/noAN/chem_noAN_netO3rate_noINORG_NOx_250ppt_100ppb.eps}
+        \includegraphics[width=\linewidth]{\figspath facsimile_results/chem_noAN_netO3rate_noINORG_NOx_250ppt_100ppb.png}
         \caption{}
@@
         \centering
-        \includegraphics[width=\linewidth]{D:/FACSIMILE/ANsCBmodel/matlab/ANsCB_pics/ANsNOxVOC/allAN/chem_allAN_netO3rate_noINORG_NOx_250ppt_100ppb.eps}
+        \includegraphics[width=\linewidth]{\figspath facsimile_results/chem_allAN_netO3rate_noINORG_NOx_250ppt_100ppb.png}
         \caption{}
@@
         \centering
-        \includegraphics[width=\linewidth]{D:/FACSIMILE/ANsCBmodel/matlab/ANsCB_pics/ANsNOxVOC/allAN/chem_allAN_netO3ratediff_noINORG_NOx_250ppt_100ppb.eps}
+        \includegraphics[width=\linewidth]{\figspath facsimile_results/chem_allAN_netO3ratediff_noINORG_NOx_250ppt_100ppb.png}
         \caption{}
@@
 \caption{Isopleths giving net rate of $O_3$ production (ppb/h, color) as a function of VOC (ppbC) and $NO_x$ (ppb) with alkyl nitrate chemistry switched off (top left) and switched on (top right).
-Difference between them is shown at the bottom.}\label{fig:netO3rate_noAN_withAN_diff}
-\end{figure}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+Difference between them is shown at the bottom.}
+\label{fig:netO3rate_noAN_withAN_diff}
+\end{figure}
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \begin{figure}[h] % netO3mixrat noAN allAN diff diffpercent
@@
   \centering
-  \includegraphics[width=.99\linewidth]{D:/FACSIMILE/ANsCBmodel/matlab/ANsCB_pics/ANsNOxVOC/noAN/chem_noAN_netO3mixrat_noINORG_NOx_250ppt_100ppb.eps}
+  \includegraphics[width=.99\linewidth]{\figspath facsimile_results/chem_noAN_netO3mixrat_noINORG_NOx_250ppt_100ppb.png}
 \end{minipage}
@@
   \centering
-  \includegraphics[width=.99\linewidth]{D:/FACSIMILE/ANsCBmodel/matlab/ANsCB_pics/ANsNOxVOC/allAN/chem_allAN_netO3mixratdiff_noINORG_NOx_250ppt_100ppb.eps}
+  \includegraphics[width=.99\linewidth]{\figspath facsimile_results/chem_allAN_netO3mixratdiff_noINORG_NOx_250ppt_100ppb.png}
 \end{minipage}
@@
   \centering
-  \includegraphics[width=.99\linewidth]{D:/FACSIMILE/ANsCBmodel/matlab/ANsCB_pics/ANsNOxVOC/allAN/chem_allAN_netO3mixrat_noINORG_NOx_250ppt_100ppb.eps}
+  \includegraphics[width=.99\linewidth]{\figspath facsimile_results/chem_allAN_netO3mixrat_noINORG_NOx_250ppt_100ppb.png}
 \end{minipage}
@@
   \centering
-  \includegraphics[width=.99\linewidth]{D:/FACSIMILE/ANsCBmodel/matlab/ANsCB_pics/ANsNOxVOC/allAN/chem_allAN_netO3mixratdiffpercent_noINORG_NOx_250ppt_100ppb_new.eps}
+  \includegraphics[width=.99\linewidth]{\figspath facsimile_results/chem_allAN_netO3mixratdiffpercent_noINORG_NOx_250ppt_100ppb_new.png}
 \end{minipage}
@@
 Absolute difference between two series of model runs in ppb (bottom left), same in \% (bottom right).
-Net production = final - initial value of $O_3$ mixing ratio.}\label{fig:netO3mixrat_noAN_withAN_diff}
-\end{figure}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+Net production = final - initial value of $O_3$ mixing ratio.}
+\label{fig:netO3mixrat_noAN_withAN_diff}
+\end{figure}
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \begin{figure} % endtotalANs
 \centering
-\includegraphics[width=.6\linewidth]{D:/FACSIMILE/ANsCBmodel/matlab/ANsCB_pics/ANsNOxVOC/allAN/chem_allAN_endtotalANs_noINORG_NOx_250ppt_100ppb.eps}
-\caption{Isopleths of net production of alkyl nitrates ($\Sigma RONO_2$) as a function of $NO_x$ and VOCs levels.}\label{fig:endtotalANs}
-\end{figure}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\includegraphics[width=.6\linewidth]{\figspath facsimile_results/chem_allAN_endtotalANs_noINORG_NOx_250ppt_100ppb.png}
+\caption{Isopleths of net production of alkyl nitrates ($\Sigma RONO_2$) as a function of $NO_x$ and VOCs levels.}
+\label{fig:endtotalANs}
+\end{figure}
 
 \subsubsection{$O_3$ and individual alkyl nitrates}
-
 To estimate the impact of individual alkyl nitrates on $O_3$ production an additional series of experiments has been conducted.
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \begin{figure}
 \centering
-\includegraphics[width=0.7\linewidth]{D:/FACSIMILE/ANsCBmodel/matlab/ANsCB_pics/ANsNOxVOC/allAN/chem_allAN_timeO3diff_10ppbNOx6VOC.eps}
+\includegraphics[width=0.7\linewidth]{\figspath facsimile_results/chem_allAN_timeO3diff_10ppbNOx6VOC.png}
 \caption{Time evolution of the difference in $O_3$ mixing ratio between base experiment (with all alkyl nitrate chemistry switched on) and experiments where chemistry of individual alkyl nitrates or their groups is switch on.
 Black line shows the difference between base experiment and an experiment with all alkyl nitrate chemistry switched off.
-$NO_x$ = 10 ppb, VOCs = case F.}\label{fig:timeO3diff10ppbNOx6VOC}
-\end{figure}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+$NO_x$ = 10 ppb, VOCs = case F.}
+\label{fig:timeO3diff10ppbNOx6VOC}
+\end{figure}
+
 \begin{figure}
 \centering
-\includegraphics[width=0.7\linewidth]{D:/FACSIMILE/ANsCBmodel/matlab/ANsCB_pics/ANsNOxVOC/allAN/chem_allAN_timeANs_10ppbNOx6VOC.eps}
-\caption{Time evolution of alkyl nitrates mixing ratios. $NO_x$ = 10 ppb, VOCs = case F.}\label{fig:timeANs10ppbNOx6VOC}
-\end{figure}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\includegraphics[width=0.7\linewidth]{\figspath facsimile_results/chem_allAN_timeANs_10ppbNOx6VOC.png}
+\caption{Time evolution of alkyl nitrates mixing ratios. $NO_x$ = 10 ppb, VOCs = case F.}
+\label{fig:timeANs10ppbNOx6VOC}
+\end{figure}
+
 \begin{table}
-\caption{Difference in $O_3$ mixing ratio by the end of simulation depending alkyl nitrate chemistry.}\label{tab:timeO3diffANs10ppbNOx6VOC}
+\caption{Difference in $O_3$ mixing ratio by the end of simulation depending alkyl nitrate chemistry.}
+\label{tab:timeO3diffANs10ppbNOx6VOC}
 \centering
@@
 \end{table}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\subsection{$O_3$ production and carbon bonds}\label{sec:res_O3ANCB}
-
+
+\subsection{$O_3$ production and carbon bonds} \label{sec:res_O3ANCB}
 As it has been noted in section \ref{sec:intro_CB}, a novel method was recently proposed to estimate the $O_3$ production per carbon bond emitted into the system.
@@
 Namely, the term $E_{C_{bons}}$ is calculated as a rate of change of the difference between final and initial sum of carbon bonds ($\Sigma Cbonds$) in the system at a constant sum of carbon bonds contained in all of the alkanes and $CO$ ($\Sigma Cbonds_{alkanes, CO}$):
-\begin{matriz}\label{eq:ECbonds}
+\begin{matriz} \label{eq:ECbonds}
 E_{C_{bons}} = {\dfrac{\Sigma Cbonds|_{final} - \Sigma Cbonds|_{initial}}{24 hours}}\bigg|_{\Sigma Cbonds_{alkanes, CO} =const}
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \begin{figure} % epsilon noANs allANs diff 250 ppt 100 ppb
@@
   \centering
-  \includegraphics[width=.99\linewidth]{D:/FACSIMILE/ANsCBmodel/matlab/ANsCB_pics/ANsNOxVOC/noAN/chem_noAN_epsilon_noINORG_NOx_250ppt_100ppb.png}
+  \includegraphics[width=.99\linewidth]{\figspath facsimile_results/chem_noAN_epsilon_noINORG_NOx_250ppt_100ppb.png}
 \end{minipage}
@@
   \centering
-  \includegraphics[width=.99\linewidth]{D:/FACSIMILE/ANsCBmodel/matlab/ANsCB_pics/ANsNOxVOC/allAN/chem_allAN_epsilon_noINORG_NOx_250ppt_100ppb.png}
+  \includegraphics[width=.99\linewidth]{\figspath facsimile_results/chem_allAN_epsilon_noINORG_NOx_250ppt_100ppb.png}
 \end{minipage}
@@
   \centering
-  \includegraphics[width=.99\linewidth]{D:/FACSIMILE/ANsCBmodel/matlab/ANsCB_pics/ANsNOxVOC/allAN/chem_allAN_epsilondiff_noINORG_NOx_250ppt_100ppb.png}
+  \includegraphics[width=.99\linewidth]{\figspath facsimile_results/chem_allAN_epsilondiff_noINORG_NOx_250ppt_100ppb.png}
 \end{minipage}
 \caption{Isopleths giving net $O_3$ production efficiency (or $\epsilon$, molecules of $O_3$ per carbon bond, color) as a function of VOC (ppbC) and $NO_x$ (250ppt - 100 ppb) with alkyl nitrate chemistry switched off (left) and switched on (middle).
-Difference between them is shown on the right.}\label{fig:epsilon250ppt100ppb}
-\end{figure}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+Difference between them is shown on the right.}
+\label{fig:epsilon250ppt100ppb}
+\end{figure}
+
 \begin{figure} % epsilon noANs allANs diff 5 ppt 100 ppb
@@
   \centering
-  \includegraphics[width=.99\linewidth]{D:/FACSIMILE/ANsCBmodel/matlab/ANsCB_pics/ANsNOxVOC/noAN/chem_noAN_epsilon_noINORG_NOx_5ppt_100ppb.png}
+  \includegraphics[width=.99\linewidth]{\figspath facsimile_results/chem_noAN_epsilon_noINORG_NOx_5ppt_100ppb.png}
 \end{minipage}
@@
   \centering
-  \includegraphics[width=.99\linewidth]{D:/FACSIMILE/ANsCBmodel/matlab/ANsCB_pics/ANsNOxVOC/allAN/chem_allAN_epsilon_noINORG_NOx_5ppt_100ppb.png}
+  \includegraphics[width=.99\linewidth]{\figspath facsimile_results/chem_allAN_epsilon_noINORG_NOx_5ppt_100ppb.png}
 \end{minipage}
@@
   \centering
-  \includegraphics[width=.99\linewidth]{D:/FACSIMILE/ANsCBmodel/matlab/ANsCB_pics/ANsNOxVOC/allAN/chem_allAN_epsilondiff_noINORG_NOx_5ppt_100ppb.png}
+  \includegraphics[width=.99\linewidth]{\figspath facsimile_results/chem_allAN_epsilondiff_noINORG_NOx_5ppt_100ppb.png}
 \end{minipage}
 \caption{Isopleths giving net $O_3$ production efficiency (or $\epsilon$, molecules of $O_3$ per carbon bond, color) as a function of VOC (ppbC) and $NO_x$ (5 ppt - 100 ppb) with alkyl nitrate chemistry switched off (top) and switched on (middle).
-Difference between them is shown at the bottom.}\label{fig:epsilon5ppt100ppb}
-\end{figure}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+Difference between them is shown at the bottom.}
+\label{fig:epsilon5ppt100ppb}
+\end{figure}
+
 \section{Discussion} \label{sec:discuss}
@@
 First of all, the chemical mechanism of the model is not sufficient enough to represent full complexity of ozone tropospheric chemistry.
-It does not have all possible sinks for $NO_x$, such as the nitric and nitrous acid, dinitrogen pentoxide and PAN.
+It does not possess $NO_x$ removal by the nitric and nitrous acid, dinitrogen pentoxide and PAN.
 In addition, it does not account for dry and wet deposition of any species.
@@
 
-\bibliography{diss_refs}
+\bibliography{refs}
 \newpage
@@
 
-\input{model_code}
+\input{code}
 
